### 2019-02-07 README.Rmd
statistics after first glance

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# venomanaphylaxiscompendium

This repository contains the data and code for our paper:

> Francuzik et al. , (2019). _Insect Venom Anaphylaxis, observetional study of the European Anaphylaxis Registry_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _Title of paper_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for 'Insect Venom Anaphylaxis, observetional study of the European Anaphylaxis Registry'_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: <https://github.com/wolass/venomanaphylaxiscompendium/archive/master.zip>

Or you can install this compendium as an R package, venomanaphylaxiscompendium, from GitHub with:
```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("wolass/venomanaphylaxiscompendium")
```
### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2019-02-07 analysis/paper/paper.Rmd
statistics after first glance

---
title: "Insect venom anaphylaxis: a case - control study of the European Anaphylaxis Registry"
author:
  - Wojciech Francuzik^1^
  - Sabine Dölle-Bierke^1^
  - Franziska Ruëff^2^ 
  - Claudia Pföhler^3^
  - Kathrin Scherer Hofmeier^4^
  - Margitta Worm^1^
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

^1^ Department of Dermatology, Venerology and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,

^2^ Department of Dermatology and Allergology, Klinikum der Universität München, Germany

^3^ Department of Dermatology, Saarland University Hospital, Homburg/Saar, Germany

^4^ Department of Dermatology, University Hospital Basel, Switzerland

**Corresponding author: **
  Prof. Dr. med. M. Worm margitta.worm@charite.de
Phone: +49 30 450 529 005;	  Fax: +49 30 450 529 902

**Keywords:**
  anaphylaxis, adrenaline (epinephrine), beta-blockers, insect venom allergy, yellow-jacket

**Highlights:**
  These are the highlights. 
**Document statistics:**
  Word count, figures, tables, references

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r setup, echo = FALSE,warning=FALSE,include=F}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

devtools::load_all('.',quiet = T) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
###### SETUP

```

# Abstract
Insect-venom elicited anaphylaxis is a common hypersensitivity reaction which may be life-threatening. Using the data from the European Anaphylaxis Registry (11596 cases in total) we identified insect-venom elicited anaphylaxis cases (n = 4482) and analyzed these in comparison to anaphylaxis elicited by other elicitors (n = 7114).

The data show that 68.57% of all insect elicited cases were elicited by yellow jackets, followed by bees (21.86%). The insect venom elicited cases occurred mostly in outdoor places (44.65%) patients' homes (12.87%) or urban places (9.616%).

Skin, gastrointestinal and respiratory symptoms occurred less frequently in insect elicited cases of anaphylaxis, whereas cardiologic symptoms (with hypotension, collapse, and loss of consciousness) were more frequent.  Intramuscular adrenaline (as a first-line therapy) was administered significantly less often in insect venom elicited cases (4.04%, p < 0.0001). The mortality rate in insect anaphylaxis was comparable (0.156%) to other cases (0.295%, p = 0.174).

Patients who experienced insect-venom anaphylaxis were older (p < 0.0001), more often had concomitant mastocytosis (p < 0.0001) and cardiologic conditions (p < 0.0001) and females more often had concomitant thyroid diseases and less often suffered from a food allergy or atopic dermatitis.

Symptoms of insect venom anaphylaxis are distinctively different from other reactions, indicating that the therapy of insect elicited cases of anaphylaxis should be considered separately. Indeed we observed different therapeutic patterns in insect elicited cases of anaphylaxis (more antihistaminics but fewer corticosteroids, bronchodilators, and surprisingly - adrenaline). This indicates that the management of insect-venom induced anaphylaxis may be improved and is especially required in patients with concomitant cardiologic conditions and these with hyperreactive mast cells.

# Introduction

# Methods

The European Anaphylaxis Registry [@Grabenhenrich2016] database from March 2018 wassearched for anaphylaxis cases elicited by insect venom. The flowchart in figure \@ref(fig:flowchart) represents the detailed case-selection process.

The final database consisted of `r countries %>% summarise(sum(n))` cases of insect elicited anaphylaxis from `r countries %>% summarize(length(n))` countries. Severe reactions were identified based on the definition by NIAID/FAAN [@Sampson2006] and presented with significant hypoxia, hypotension, confusion, collapse and loss of consciousness, or incontinence. 
We compared the frequency of various elicitors, symptoms, and factors known to increase the risk of severe anaphylaxis [@Worm2018] in both groups. We evaluated symptoms, managment and risk factors of insect elicited cases in comparison to other known triggers of anaphylaxis.

<!-- ## Data handling and statisitcs -->
The statistical analysis was performed in the R Statistical Package [@R]. Simple comparison of categorical variables was performed using Fisher's exact test, continuous variables were analyzed using Mann-Whitney U test. We defined statistical significance as α = 0.05. Data along with the analysis script can be accessed at online at....

# Results

## Demographics

```{r}
demoTab
```

```{r, results='asis'}
st_options('bootstrap.css', FALSE)
rdb %>% select()
ctable(rdb$d_centres_country %>% sapply(function(x){
  as.character(x) %>% factor()
}), rdb$q_340_insects%>% sapply(function(x){
  as.character(x) %>% factor()
}), style = 'rmarkdown',
       useNA = "no",
       prop = "r",
       totals = F,
       round.digits = 1,
dnn = c("Countries","Insects"))

```

```{r}
cdplot(grouping~d_age,data = rdb)
```

```{r}
mosaicplot(formula = d_severity_rm~d_insect_gr4,data = rdb)
```

## Symptoms
```{r}
testInsectsbinomialTab$symptoms[,-7] %>% knitr::kable()
plotSympt
```

```{r}
crammerElicitorBinvsSympt[1:15,] %>% arrange(desc(CV)) %>% knitr::kable()
```

Patients with Insect Venom Allergy had a lot more cardiologic symptoms than anaphylaxis cases that were elicited through other triggers.  
Dizziness, Loss of consciousness hypotension, collapse, Throat tightness and nausea were a lot more often associated with the insect cases.

Forest plot ! Risk that the symptom will occure given that the reaction was elicited by Insects or not

## Co-factors
```{r}
testInsectsbinomialTab$cofactors[,-7] %>% knitr::kable()
```

## Managment
```{r}
testInsectsbinomialTab$management[,-7] %>% knitr::kable()
```

```{r}
testInsectsbinomialTab$prophylaxis[,-7] %>% knitr::kable()
```

## Repeated reactions

```{r}
treact <- table(rdb$grouping,rdb$q_160_ever_react)[,1:2] %>% prop.table(1)*100
p <- table(rdb$grouping,rdb$q_160_ever_react)[,1:2] %>% summary() %>% {.$p.value}
```
In general `r treact[1,2]`% patient with insect allergy had experienced anaphylaxis in the past which is less than if the reaction is elicited by other elcitors (`r treact[2,2]`%, p = `r p`).  We documented `r sum(repeated_reaction_in_Reg)` patients with two documented reactions in our registry. 
Out of these `r sum(greter_same)` (`r sum(greter_same)/sum(repeated_reaction_in_Reg)*100`%) had Insect elicited anaphylaxis and in `r greter_same[1]` (`r greter_same[1]/sum(greter_same)*100`) The follwoeing reaction was more severe than before.

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

# Figures

```{r flow}
knitr::include_graphics("../figures/flow.png")
```

```{r figforestcofactors}
knitr::include_graphics("../../figFOREST.png")
```

```{r}
mosaic(d_organ_sum~d_elicitor_gr5,data=rdb)
```

```{r}
mosaic(d_severity_rm~d_elicitor_gr5,data=rdb)

```

### 2019-02-07 R/stat.R

@@ -1,0 +1,814 @@
+#Setup
+require(dplyr)
+require(magrittr)
+require(ggplot2)
+require(forestplot)
+require(summarytools)
+st_options('bootstrap.css', FALSE)
+#### GET THE DATA ########
+# Note the path that we need to use to access our data files when rendering this document
+load('../RefractoryAnaOrg/analysis/data/raw_data/data4.R')
+
+
+####### Functions #############
+rcalc <- function(var,level="yes"){
+  tab <- table(data3[,var],data3$rANA)
+  if(level %in% (data3[,var] %>% levels)){
+    pn <- tab[level,"no"]
+    pp <- tab[level,"yes"]
+    l <- level
+  } else {
+    pn <- tab[2,"no"]
+    pp <- tab[2,"yes"]
+    l <- levels(data3[,var])[2]
+  }
+  nn <- data3$rANA %>% factor %>% summary %>% {.[1]}-pn
+  np <- data3$rANA %>% factor %>% summary %>% {.[2]}-pp
+  eval <- c(pn,pp,nn,np) %>% matrix(byrow = T,ncol = 2)
+  prt <- (prop.table(eval,2)*100) %>% {signif(.,3)}
+  p <- eval %>% fisher.test() %>% .$p.value %>% roP
+  return(c(var,prt[1,1],prt[1,2],p,l))
+}
+
+roP<- function(x){
+  ifelse(x > 0.001,
+         round(x,3),ifelse(x==0,0,0.0001)
+  )
+}
+
+f1 <- function(x){
+  names(x) <- NULL
+  return(do.call(t.test,x)$p.value %>% roP)
+}
+f2 <- function(x){
+  x %>% table(data3$rANA) %>% {.[1:2,]} %>% fisher.test() %>% {.$p.value} %>% roP
+}
+f3 <- function(x){
+  x %>%
+    lapply(function(x){(length(which(x=="yes"))/length(x)*100) %>% roP})
+}
+
+
+
+#### cases selection ####
+# Define a new variable - reaction type to exclude non ANA cases
+
+correctLabels1 <- function(data) {
+  data$q_120_time_between_v4 %<>% factor(levels = c("00 – 10 Minutes",
+                                                    "11 - 30 Minutes",
+                                                    "31 - 60 Minutes",
+                                                    "61 - 120 Minutes",
+                                                    "121 – 240 Minutes (2 – 4 hours)",
+                                                    "more than 240 Minutes" ,
+                                                    "more than 120 min (V4)"),
[…]

### 2019-02-07 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(venomanaphylaxiscompendium)
+
+test_check("venomanaphylaxiscompendium")

### 2019-02-07 tests/testthat/test-stat.R

@@ -1,0 +1,11 @@
+context("test-stat")
+
+test_that("testing the proper renaming function", {
+  s <- data.frame(q_120_time_between_v4=c("unknown","-9","00 – 10 Minutes"))
+  expect_equal(
+    s %>% correctLabels1() %>% {.$q_120_time_between_v4[3]},
+   factor("10", levels = c("10","30",
+                           "60","120","4h","more","4h-5"))
+  )
+  expect_true(s %>% correctLabels1() %>% {.$q_120_time_between_v4[1]} %>% is.na)
+})

### 2019-02-07 analysis/paper/paper.Rmd
age adaptation done

@@
 testInsectsbinomialTab$symptoms[,-7] %>% knitr::kable()
+```
+
+```{r}
 plotSympt
 ```
+
+Patients who underent anaphylaxis due to insect venom more often experienced 
+cardiologic symptoms. Especially collapsing due to hypotension was prominently 
+more often seen in patient who underwent venom anaphylaxis. This difference was especially visible in children under the age of 18 years.
 
@@
 
+## Diagnostics
+
 ## Repeated reactions
@@
 # Discussion
+
+Cardiologic symptoms and hypotensive collapse might be associated with vanom anaphylaxis due to the vaso-vagal reflex.
+As seen in patients undrgoing blood sampling needles (and probably also insect stings) may elicit a hypotensive response due to extreme emotional distress with bradycardia, limiting the bloodflow to the brain and causing a syncope.
+This mechanism may contribute to the ongoing histamin-induced vasodialation and may worsen the symptoms of an allergic reaction.
 
@@
 mosaic(d_severity_rm~d_elicitor_gr5,data=rdb)
+```
 
+```{r}
+hypotensionPlot
 ```
+
+```{r}
+cramerPlot
+```

### 2019-02-07 R/stat.R

@@ -4,7 +4,7 @@ require(magrittr)
 require(ggplot2)
 require(forestplot)
 require(summarytools)
-st_options('bootstrap.css', FALSE)
+
 #### GET THE DATA ########
 # Note the path that we need to use to access our data files when rendering this document
 load('../RefractoryAnaOrg/analysis/data/raw_data/data4.R')
@@ -294,7 +294,12 @@ funF2 <- function(var, grouping_var){
   #group_by(grouping_var) %>%
   #summarise(count=n())
 }
-
+funC <- function(var, grouping_var){
+  temp <- var %>%
+    data.frame(grouping_var) %>%
+    table %>% {.[1:2,]} %>%
+    assocstats() %>% {.$cramer}
+}
 #' This function takes grouping vector and a DATA FRame ONLY OF  variables to test
 #'
 ChiF <- function(variablesDF,grouping){
@@ -303,7 +308,8 @@ ChiF <- function(variablesDF,grouping){
              counts_1 = lapply(variablesDF,funN1,grouping) %>%unlist,
              counts_2 = lapply(variablesDF,funN2,grouping) %>%unlist,
              fraq_1 = lapply(variablesDF,funF1,grouping) %>%unlist,
-             fraq_2 = lapply(variablesDF,funF2,grouping) %>%unlist
+             fraq_2 = lapply(variablesDF,funF2,grouping) %>%unlist,
+             Crammer = lapply(variablesDF,funC,grouping) %>% unlist
   )
 }
 
@@ -650,25 +656,8 @@ FigForestCofactors <- makeForestPlot(rdb,
 
 
 
-#### MEasures of Association ############
-rdb$q_114_hypotension_collapse_v5
-vcd::assocstats(table(rdb$q_114_hypotension_collapse_v5,
-                      grouping))
-require(vcd)
 
-# Tu skonczyles - zrób assocstats lapply na wielu zmiennych z groupingiem
-crammerElicitorBinvsSympt <-
-lapply(variableSelectionTab(rdb) %>% filter(section=="symptoms",level==3) %>%
-         pull(variableName),function(x){
-           assocstats(rdb[,x] %>%
-                      table(grouping))[[4]]
-         }) %>%
-  unlist %>%
-  {data.frame(CV=.,
-              var=variableSelectionTab(rdb) %>%
-                filter(section=="symptoms",level==3) %>%
-                pull(variableName))}
-
+##### Gonvert to function ######
 crammerElicitorBinvsSympt2 <-
[…]

### 2019-02-08 analysis/paper/paper.Rmd
plot crammer with confidence intervals

@@
 st_options('bootstrap.css', FALSE)
-rdb %>% select()
+#rdb %>% select()
 ctable(rdb$d_centres_country %>% sapply(function(x){
@@
 dnn = c("Countries","Insects"))
+
+```
+```{r}
+plot.proportions(rdb, "d_centres_country","q_340_insects")
 
@@
 cardiologic symptoms. Especially collapsing due to hypotension was prominently 
-more often seen in patient who underwent venom anaphylaxis. This difference was especially visible in children under the age of 18 years.
+more often seen in patient who underwent venom anaphylaxis. This difference was especially visible in children under the age of 18 years. (See the ade adapted symptoms plot). 
+We saw prominent differences in association of specific symptoms with eliocitor in cerain age groups.
+Children under 13 yoears of age showed association with IVIA and vomiting - whereas adults did not.
+Hypotension collapse was more prominently associated with IVIA in children whereas loss of consciousness in adults.
 
@@
 ```
+
+```{r}
+plot.Cramer(supraCramerFun(data =rdbp,
+               vars =c( "q_114",
+               "q_112_incontinence",
+               "q_114_dizziness",
+               "q_112_abdominal_pain",
+               "q_114_loss_of_consciousness",
+               "q_114_hypotension_collapse_v5",
+               "q_114_reductions_of_alertness",
+               "q_111_angioedema",
+               "q_113_wheezing_expiratory_distress_v5"),
+               grouping = "grouping",
+               subset = "d_age_gr2"))
+
+```
+
+```{r}
+plot.Cramer(cramerFun(data =rdbp,
+               vars=c("q_114",
+               "q_112_incontinence",
+               "q_114_dizziness",
+               "q_112_abdominal_pain",
+               "q_114_loss_of_consciousness",
+               "q_114_hypotension_collapse_v5",
+               "q_114_reductions_of_alertness",
+               "q_111_angioedema",
+               "q_113_wheezing_expiratory_distress_v5"),
+               grouping = "grouping"))
+```
+
+```{r}
+tryptase_plot(rdb)+
+  scale_x_log10()
+
+```

### 2019-02-08 R/stat.R

@@ -4,6 +4,7 @@ require(magrittr)
 require(ggplot2)
 require(forestplot)
 require(summarytools)
+require(vcd)
 
 #### GET THE DATA ########
 # Note the path that we need to use to access our data files when rendering this document
@@ -295,21 +296,42 @@ funF2 <- function(var, grouping_var){
   #summarise(count=n())
 }
 funC <- function(var, grouping_var){
-  temp <- var %>%
+  var %>%
     data.frame(grouping_var) %>%
     table %>% {.[1:2,]} %>%
     assocstats() %>% {.$cramer}
 }
+
+# funC <- function(var, grouping_var){
+#   temp <- var %>%
+#     data.frame(grouping_var) %>%
+#     table %>% {.[1:2,]} %>%
+#     DescTools::CramerV(conf.level = T,method = "fisher")
+# }
+# funCl <- function(var, grouping_var){
+#   temp <- var %>%
+#     data.frame(grouping_var) %>%
+#     table %>% {.[1:2,]} %>%
+#     DescTools::CramerV(conf.level = T) %>% .[2]
+# }
+# funCu <- function(var, grouping_var){
+#   temp <- var %>%
+#     data.frame(grouping_var) %>%
+#     table %>% {.[1:2,]} %>%
+#     DescTools::CramerV(conf.level = T)
+# }
+
 #' This function takes grouping vector and a DATA FRame ONLY OF  variables to test
 #'
 ChiF <- function(variablesDF,grouping){
+  # C <- lapply(variablesDF,funC,grouping)
   data.frame(variables = names(variablesDF),
              pval = lapply(variablesDF,funChi,grouping) %>% unlist,
              counts_1 = lapply(variablesDF,funN1,grouping) %>%unlist,
              counts_2 = lapply(variablesDF,funN2,grouping) %>%unlist,
              fraq_1 = lapply(variablesDF,funF1,grouping) %>%unlist,
              fraq_2 = lapply(variablesDF,funF2,grouping) %>%unlist,
-             Crammer = lapply(variablesDF,funC,grouping) %>% unlist
+             Cramer = lapply(variablesDF,funC,grouping) %>% unlist
   )
 }
 
@@ -658,30 +680,84 @@ FigForestCofactors <- makeForestPlot(rdb,
 
 
 ##### Gonvert to function ######
-crammerElicitorBinvsSympt2 <-
[…]

### 2019-02-11 analysis/paper/paper.Rmd
Forest plots with subsets as age

@@
 ```{r figforestcofactors}
-knitr::include_graphics("../../figFOREST.png")
+knitr::include_graphics("../figures/figForest.png")
+```
+
+```{r figforestcofactorskids}
+knitr::include_graphics("../figures/kidsForest.png")
+```
+
+```{r figforestcofactorsadults}
+knitr::include_graphics("../figures/adultsForest.png")
 ```

### 2019-02-11 R/stat.R

@@ -372,12 +372,16 @@ countries <-rdb %>%
   group_by(d_centres_country) %>%
   summarize(n=n())%>% arrange(desc(n))
 
-grouping <- ifelse(rdb$d_elicitor_gr5=="insects","insects",
-                   ifelse(is.na(rdb$d_elicitor_gr5)|rdb$d_elicitor_gr5=="unkown",NA,"other")) %>%
+grouping <- ifelse(rdb$d_elicitor_gr5=="insects",
+                   "insects",
+                   ifelse(is.na(rdb$d_elicitor_gr5)|rdb$d_elicitor_gr5=="unkown",
+                          NA,
+                          "other")
+                   ) %>%
   factor
 
-group_triggerBinomial <- ifelse(rdb$d_elicitor_gr5=="insects","insects",
-                                ifelse(rdb$d_elicitor_gr5=="unkown",NA,"other"))
+rdbp <-rdb
+rdbp$grouping<- grouping
 
 
 variableSelectionTab <- function(data){
@@ -620,6 +624,7 @@ library(forestplot)
 #
 
 ## Creat a formatted data frame
+
 makeOdds <- function(data,var1,var2){ # var1 to predyktor, var2 - outcome
   if(var1=="b_sex"){
     or <- questionr::odds.ratio(data[,var1],data[,var2])
@@ -635,6 +640,7 @@ makeOdds <- function(data,var1,var2){ # var1 to predyktor, var2 - outcome
 }
 makeDF <- function(data,variables,grouping,outcome){
   o <-lapply(variables,function(x){
+    grouping <- data[,grouping]
     grA <- which(grouping==levels(grouping)[1])
     grB <- which(grouping==levels(grouping)[2])
     rbind(
@@ -654,7 +660,7 @@ makeTableText <- function(makeDF){
         makeDF[-c(1),c(1,6,2)])
 }
 
-makeForestPlot <- function(data,variables,grouping,outcome){
+makeForestPlot <- function(data,variables,grouping,outcome,cLow=0.2,cHigh = 4){
   df <- makeDF(data,variables,grouping,outcome)
   forestplot(labeltext = makeTableText(df),
              mean = df[,"mean"] %>% unlist,
@@ -662,22 +668,44 @@ makeForestPlot <- function(data,variables,grouping,outcome){
              upper = df[,"upper"] %>% unlist,
              new_page = TRUE,
              is.summary=c(T,rep(F,length(df[,1])-1)),
[…]

### 2019-02-13 README.Rmd
month of reaction analysis done

@@
 )
+Title <- "Case-control study of Insect Venom Anaphylaxis from the European Anaphylaxis Registry."
 ```
@@
 
-> Francuzik et al. , (2019). _Insect Venom Anaphylaxis, observetional study of the European Anaphylaxis Registry_. Name of journal/book <https://doi.org/xxx/xxx>
-
-Our pre-print is online here:
-
-> Authors, (YYYY). _Title of paper_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+> Francuzik et al., (2019).  Name of journal/book <https://doi.org/xxx/xxx>
 
@@
 
-> Authors, (`r format(Sys.Date(), "%Y")`).
-_Compendium of R code and data for 'Insect Venom Anaphylaxis, observetional study of the European Anaphylaxis Registry'_.
-Accessed `r format(Sys.Date(), "%d %b %Y")`.
-Online at <https://doi.org/xxx/xxx>
+> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_ Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
 
@@
 
-**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse
+**Data :** [CC-BY-4.0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse
 

### 2019-02-13 analysis/paper/paper.Rmd
month of reaction analysis done

@@
 
-The European Anaphylaxis Registry [@Grabenhenrich2016] database from March 2018 wassearched for anaphylaxis cases elicited by insect venom.
+The European Anaphylaxis Registry [@Grabenhenrich2016] database from March 2018 was searched for anaphylaxis cases elicited by insect's venom.
 The flowchart in figure \@ref(fig:flowchart) represents the detailed case-selection process.
@@
 The statistical analysis was performed in the R Statistical Package [@R].
-Simple comparison of categorical variables was performed using Fisher's exact test, continuous variables were analyzed using Mann-Whitney U test.
+Simple comparison of categorical variables was performed using either Chi^2^ test or Fisher's exact test (where the number of observation in a bin was less then 10), continuous variables were analyzed using Mann-Whitney U test.
 We defined statistical significance as α = 0.05.
-Data along with the analysis script can be accessed at online at....
+Data along with the analysis script can be accessed at online at https://github.com/wolass/venomanaphylaxiscompendium.
 
@@
 
+We saw a significant difference in the clinical features of insect anaphylaxis in children and adults.
+The younger popoulation significantly less often had concomitant conditions (i.e.
+DM, HT, malignant diseases and mastocytosis), and more often presented with atopic dermatitis, rhinitis and asthma.
+Therefore, we decided to analyse these groups separately.
+
 ```{r}
-demoTab
+demoTab %>% rbind(demo_age_tab) %>% t() %>%  
+  {data.frame(` `=rownames(.),.)} %>%
+#   flextable(col_keys = c("X.","female","male","children.adolescents","adults")) 
+# topology <- data.frame(col_keys = c("X.","female","male","children.adolescents","adults"),
+#                        colB = c("","Sex","Sex","Age","Age"),
+#                        ColA = c("","Female","Male","<18",">=18"))
+# tempT %>%  
+#   set_header_df(mapping = topology,key = "col_keys") %>% 
+#   merge_h(part="header") %>% 
+#   merge_v() %>% 
+#   empty_blanks() %>% 
+#   autofit()
+knitr::kable(col.names = c("","female","male","<18","adults"),row.names = F)
 ```
@@
 ```{r, results='asis'}
-st_options('bootstrap.css', FALSE)
-#rdb %>% select()
-ctable(rdb$d_centres_country %>% sapply(function(x){
-  as.character(x) %>% factor()
-}), rdb$q_340_insects%>% sapply(function(x){
-  as.character(x) %>% factor()
-}), style = 'rmarkdown',
-       useNA = "no",
-       prop = "r",
-       totals = F,
-       round.digits = 1,
-dnn = c("Countries","Insects"))
-
-```
-```{r}
-plot.proportions(rdb, "d_centres_country","q_340_insects")
+# st_options('bootstrap.css', FALSE)
+# #rdb %>% select()
+# ctable(rdb$d_centres_country %>% sapply(function(x){
+#   as.character(x) %>% factor()
+# }), rdb$q_340_insects%>% sapply(function(x){
+#   as.character(x) %>% factor()
+# }), style = 'rmarkdown',
+#        useNA = "no",
+#        prop = "r",
+#        totals = F,
+#        round.digits = 1,
+# dnn = c("Countries","Insects"))
 
@@
 
+```{r insectsgeo, fig.cap="Proportion of anaphylaxis cases elicited by specific insects in reporting countries. Less frequent insects were groupeped together as 'other'."}
+plot.proportions(rdb, "d_centres_country","q_340_insects")
+```
+
+The Figure \@ref(fig:insectsgeo) shows the distribution of insects according to the geographic area (countries).
+
 ```{r}
-cdplot(grouping~d_age,data = rdb)
+plot_MOR
 ```
@@
 testInsectsbinomialTab$management[,-7] %>% knitr::kable()
-```
-
-```{r}
-testInsectsbinomialTab$prophylaxis[,-7] %>% knitr::kable()
 ```
@@
 ```
+
+```{r reactionmonthincountries, fig.cap="Proportional insect venom induced anaphylaxis to specific insects during a year according to countries.", fig.height=8,fig.width=5}
+plot_MOR+
+  facet_grid(d_centres_country~.)
+```
+
+```{r}
+plot_MOR
+```

### 2019-02-13 R/stat.R

@@ -263,21 +263,21 @@ funChi <-function(var,grouping_var){
     {chisq.test(.[,1],.[,2])$p.value}
 }
 funN1 <- function(var, grouping_var){
-  temp <- var %>%
+  var %>%
     data.frame(grouping_var) %>% table %>%
     {.[2,1]}
     #group_by(grouping_var) %>%
     #summarise(count=n())
 }
 funN2 <- function(var, grouping_var){
-  temp <- var %>%
+  var %>%
     data.frame(grouping_var) %>% table %>%
     {.[2,2]}
   #group_by(grouping_var) %>%
   #summarise(count=n())
 }
 funF1 <- function(var, grouping_var){
-  temp <- var %>%
+  var %>%
     data.frame(grouping_var) %>%
     table %>% {.[1:2,]} %>%
     prop.table(2) %>%
@@ -287,7 +287,7 @@ funF1 <- function(var, grouping_var){
 }
 
 funF2 <- function(var, grouping_var){
-  temp <- var %>%
+  var %>%
     data.frame(grouping_var) %>%
     table %>% {.[1:2,]} %>%
     prop.table(2) %>%
@@ -379,7 +379,7 @@ grouping <- ifelse(rdb$d_elicitor_gr5=="insects",
                           "other")
                    ) %>%
   factor
-
+rdb$grouping <- grouping
 rdbp <-rdb
 rdbp$grouping<- grouping
 
@@ -787,28 +787,54 @@ plot.Cramer <- function(x){
 
 #### Demographics####
 
+f4 <- function(x){
+  x[rdb$d_elicitor_gr5=="insects"]%>%
+  split(rdb$b_sex[rdb$d_elicitor_gr5=="insects"]) %>%
+  f3
+}
+f5 <- function(x){
+  x[rdb$d_elicitor_gr5=="insects"]%>%
+    split(rdb$d_age_gr2[rdb$d_elicitor_gr5=="insects"]) %>%
+    f3
+}
 demoTab <- cbind(n = rdb$b_sex[rdb$d_elicitor_gr5=="insects"] %>% summary(),
                  Age = rdb$d_age[rdb$d_elicitor_gr5=="insects"] %>%
                    split(rdb$b_sex[rdb$d_elicitor_gr5=="insects"]) %>%
                    lapply(.,function(x){mean(x) %>% signif(3)}),
[…]

### 2019-02-27 analysis/paper/paper.Rmd
add repeated reaction severity

@@
 
-```{r insectsgeo, fig.cap="Proportion of anaphylaxis cases elicited by specific insects in reporting countries. Less frequent insects were groupeped together as 'other'."}
-plot.proportions(rdb, "d_centres_country","q_340_insects")
+```{r insectsgeo, fig.cap="Proportion of anaphylaxis cases elicited by specific insects in reporting countries.
+Less frequent insects were groupeped together as 'other'.
+Countries which reported less than 20 cases of anaphylaxis due to insect venom were not illustrated in this figure."}
+plot.proportions(rdb, "d_centres_country","q_340_insects",20)
 ```
@@
 The Figure \@ref(fig:insectsgeo) shows the distribution of insects according to the geographic area (countries).
-
-```{r}
-plot_MOR
-```
+The most common elicitor among most of the countries were yellow-jackets wasps.
+Switzerland had significantly more reactions to bees than Germany.
+Centers from Italy reported more cases elicited by hornets in comparison to other countries.
 
@@
 tryptase_plot(rdb)+
-  scale_x_log10()
+  scale_x_log10()+
+  facet_grid(~d_age_gr2)
 

### 2019-02-27 R/findingOrientedResearch.R

@@ -1,0 +1,17 @@
+# Define a new class - finding
+finding <- setClass("finding", slots=c(rationale="character",
+                               libs="character",
+                               data="list",
+                               vis = "list",
+                               conclusion ="character",
+                               discussion = "character"))
+
+# Example
+# F2<- finding(rationale = "Testing the new class",
+#              libs = "purrr",
+#              data = list(test="a"),
+#              vis = list(plot="1"),
+#              conclusion = "This is dumb test",
+#              discussion = "test could be ok?")
+#
+# F2@rationale <- "This finding is obwiously going to be based on pure hypothetical shooting"

### 2019-02-27 R/stat.R

@@ -1074,8 +1074,123 @@ plot_MOR <- rdbp %>%
 
 
 # Plot countries proportions
-plot.proportions <- function(data,varx,vary){
-  ggplot(data[!is.na(data[,vary]),], aes(get(varx), fill = get(vary)))+
+plot.proportions <- function(data,varx,vary,minN){
+  ns <- data %>% filter(grouping == "insects") %>% group_by(get(varx)) %>%
+    summarize(n=n()) %>% filter(n>minN)
+  l=length(ns$n)
+  ggplot(data[!is.na(data[,vary])&
+                data[,varx]%in%ns$`get(varx)`,],
+         aes(get(varx),
+             fill = get(vary)))+
     geom_bar(position = "fill")+
-    theme(axis.text.x = element_text(angle = 45,hjust=1))
+    theme(axis.text.x = element_text(angle = 45,hjust=1))+
+    annotate(geom = "text",
+             x = 1:l,
+             y = rep(0.1,l),
+             label = paste("n =",ns%>% {as.character(.$n)}),
+             angle = 90)
 }
+
+
+##### reaction severity after SIT ######
+### rationale
+# Do patients with previous SIT had less severe reactions than these without previous SIT?
+### data and variables
+dt <- rdb %>%
+  filter(d_elicitor_gr5=="insects",
+         q_610_sit_prior_v5 %in% c("no","yes")) %>%
+  group_by(q_610_sit_prior_v5,d_severity_rm) %>%
+  summarize(n = n())
+
+### Plot
+dt %>% ggplot(aes(q_610_sit_prior_v5,fill=factor(d_severity_rm),y = n))+
+  geom_bar(stat="identity",position="fill")
+### stat test
+fisher.test(matrix(dt$n,ncol=4,byrow=T)[,2:4])
+### coonclusion
+# There were no statisitcally significant differences in these groups
+### discussion
+# Probably there is a potential difference in the tests when we look into
+# a previous reaction severity and next reaction severity
+
+
+##### F1 repeate reaction severity after SIT ######
+F1 <- list()
+### rationale
+F1$rationale<- "Do patients with previous SIT had less severe reactions than these without previous SIT? we compare these reactions in the same patients."
+### libraries
+F1$libs <- "purrr"
+### data and variables
+# patients with at leat two documented reactions in our database
+# split-apply-combine using dplyr,tidyr,purrr
+F1$data[["pat_ids"]]<-
+  data4 %>%
[…]

### 2019-02-27 R/stat.R
add severity-adrenaline analysis

@@ -372,6 +372,7 @@ countries <-rdb %>%
   group_by(d_centres_country) %>%
   summarize(n=n())%>% arrange(desc(n))
 
+#### Grouping variable insects vs no insects #####
 grouping <- ifelse(rdb$d_elicitor_gr5=="insects",
                    "insects",
                    ifelse(is.na(rdb$d_elicitor_gr5)|rdb$d_elicitor_gr5=="unkown",
@@ -1113,7 +1114,7 @@ fisher.test(matrix(dt$n,ncol=4,byrow=T)[,2:4])
 # Probably there is a potential difference in the tests when we look into
 # a previous reaction severity and next reaction severity
 
-
+require(purrr)
 ##### F1 repeate reaction severity after SIT ######
 F1 <- list()
 ### rationale
@@ -1151,7 +1152,7 @@ F1$data[["dt"]]  <-
   group_by(b_patient_code) %>%
   #do(ifelse(.$Rdate==max(.$Rdate),"last","prev"))
   tidyr::nest() %>%
-  mutate(reaction = map(data,function(df){
+  mutate(reaction = purrr::map(data,function(df){
     ifelse(df$Rdate==max(df$Rdate),
            "last",
            "prev")
@@ -1183,14 +1184,65 @@ F1$conclusion <-  "More patients showed no reduction in severity after in repeat
 ### discussion
 F1$discussion <- "Low patient numbers could have prevented an adequate analysis"
 
+source("R/findingOrientedResearch.R")
+
 ### Adrenaline managment in insect cases ###
 # F2 Adrenaline use Corellation with severity? #####
-F2<- finding(rationale = "We saw that patients who were treated for anaphylaxis after insect sting less often recived adrenaline than patients who were treated for anaphylaxis due to other triggers. We want to evaluate the cause of this observation",
-             libs = "",
-             data = list(test=
-                           rdb$d_522_adren_agg
-                           ),
-             vis = list(plot="1"),
-             conclusion = "This is dumb test",
-             discussion = "test could be ok?")
[…]

### 2019-03-25 analysis/paper/paper.Rmd
add severity-adrenaline analysis
add MOR plot+description

@@
 DM, HT, malignant diseases and mastocytosis), and more often presented with atopic dermatitis, rhinitis and asthma.
-Therefore, we decided to analyse these groups separately.
+Therefore, we decided to adjust the analysis for age and sex.
 
@@
 
+## Temporal distribution of anaphylaxis cases by elicitor.
+
+```{r insectstime, fig.cap="Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occured.
+Less frequent insects were groupeped together as 'other'.
+The lowest panel shows the proportion of insect elicited cases to ceses elicited by other triggers."}
+plot.proportions(rdb, "d_centres_country","q_340_insects",20)
+```
+
+Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May till October.
+Their proportion to other cases during the summer seasons reached 60 % and was as under 1% of cases during winter.
+Yellow-jacket was the most prominent causing-insect reported in these cases foolowed by bee (Fig \@ref(fig:insectstime)).
+
 ```{r insectsgeo, fig.cap="Proportion of anaphylaxis cases elicited by specific insects in reporting countries.
@@
 The Figure \@ref(fig:insectsgeo) shows the distribution of insects according to the geographic area (countries).
-The most common elicitor among most of the countries were yellow-jackets wasps.
+The most common elicitor among most of the countries were yellow-jackets.
 Switzerland had significantly more reactions to bees than Germany.
@@
 
+```{r age}
+cdplot(rdb$grouping~as.numeric(rdb$d_age))
+```
+
 ```{r}
 mosaicplot(formula = d_severity_rm~d_insect_gr4,data = rdb)
+rdb %>% filter(!is.na(d_insect_gr4)) %>% 
+ggplot(aes(d_insect_gr4,fill=factor(d_severity_rm)))+
+  geom_bar(position="fill")+
+  facet_grid(.~d_age_gr2)
 ```
@@
 crammerElicitorBinvsSympt[1:15,] %>% arrange(desc(CV)) %>% knitr::kable()
+
+crammerElicitorBinvsSympt[1:15,] %>% arrange(desc(CV)) %>%
+  filter(CV<0.3,
+         CV>0.15) %>% 
+  ggplot(aes(reorder(var,-CV),CV))+
+  geom_col()+
+  theme(axis.text.x = element_text(angle=90))
 ```
@@
 testInsectsbinomialTab$cofactors[,-7] %>% knitr::kable()
+
+testInsectsbinomialTab$cofactors[,-7] %>%
+  tidyr::gather(value ="Proportion",
+         key="Group",
+         4:5) %>% 
+  ggplot(aes(reorder(variableName,-Proportion),Proportion,fill=Group))+
+  geom_bar(stat="identity",position="dodge")+
+  theme(axis.text.x=element_text(angle=90))
 ```
@@
 testInsectsbinomialTab$management[,-7] %>% knitr::kable()
+testInsectsbinomialTab$management[,-7]%>%
+  tidyr::gather(value ="Proportion",
+         key="Group",
+         4:5) %>% 
+  ggplot(aes(reorder(variableName,-Proportion),Proportion,fill=Group))+
+  geom_bar(stat="identity",position="dodge")+
+  theme(axis.text.x=element_text(angle=90))+
+  geom_text(aes(y=0.5,label=reorder(signif(pval,2),-Proportion)),angle=90)
 ```
@@
 ```
+
+```{R adrenalineuse, fig.cap = "The proportion of insect venom elicited anaphylaxis cases that were treated with adrenaline in comparison to other elicitors."}
+F2$plot2
+
+```

### 2019-03-25 R/stat.R

@@ -700,7 +700,7 @@ makeForestPlot(rdbp[rdbp$d_age>18,],
                testInsectsbinomial %>%
                  filter(section=="cofactors") %>%
                  arrange(pval) %>%
-                 select(variableName) %>% pull() %>% {c(.[c(1,6,16,17,18,19
+                 select(variableName) %>% pull() %>% {c(.[c(1,6,16,41,18,27
                                                             )])},
                "grouping",
                "severity_brown",
@@ -852,11 +852,11 @@ t2 <- t1 %>% prop.table(1) %>% {round(.*100,1)}
 
 #### Previous ANA ####
 rdb$grouping <- grouping
-rdb$q_160_ever_react %>% table(rdb$grouping) %>% assocstats()
+rdb$q_160_ever_react %>% table(rdb$grouping) %>% {.[1:2,]} %>% assocstats()
 
-rdb$q_1622_ever_mild_v4 %>% table(grouping) %>% assocstats()
+rdb$q_1622_ever_mild_v4 %>% table(grouping) %>% {.[1:2,]}%>% assocstats()
 
-rdb$q_1621_ever_severe_v4 %>% table(grouping) %>% assocstats()
+rdb$q_1621_ever_severe_v4 %>% table(grouping) %>% {.[1:2,]} %>% assocstats()
 
 #### Previous Reacktions ########
 # Check for previous
@@ -908,12 +908,9 @@ iCodeDF <- tempDF %>% filter(b_patient_code%in%iCode) %>%
 greter_same <- iCodeDF %>%
   group_by(difference) %>% summarize(n = n()) %>% pull() %>% as.numeric()
 
-### Check what was the elicitor in the repeated cases
-
+iCodeDF %>% filter(difference=="greater")
 
 ### PLOT Szmptom differences ####
-
-
 
 plotSympt <- ggplot(testInsectsbinomialTab$symptoms %>%
                       tidyr::gather(key = "group",value = "Fraction", 4:5) %>%
@@ -1034,17 +1031,21 @@ AssociatedVars <- testInsectsbinomial %>%
 #                            "q_114_hypotension_collapse_v5",
 #                            "q_112_abdominal_pain",
 #                            "q_112_vomiting"))
-# ggplot(ex %>%
-#          tidyr::gather(key = "group",value = "Fraction", c("fraq_1","fraq_2")) %>%
-#          arrange(pval),
[…]

### 2019-04-03 analysis/paper/paper.Rmd
make multiple outcome plotting fun

@@
 output: 
-    bookdown::html_document2:
+    #bookdown::html_document2:
+    bookdown::word_document2:
       fig_caption: yes
@@
 Less frequent insects were groupeped together as 'other'.
-The lowest panel shows the proportion of insect elicited cases to ceses elicited by other triggers."}
-plot.proportions(rdb, "d_centres_country","q_340_insects",20)
+The lowest panel shows the proportion of insect elicited cases to ceses elicited by other triggers.", fig.height=6,fig.width=4}
+#### plot MOR#####
+plot_MOR <- gridExtra::grid.arrange(
+  #cowplot::plot_grid(
+  rdbp %>%
+  select(b_reactiondate,grouping,q_340_insects,d_centres_country) %>%
+  mutate(MOR = substr(b_reactiondate,4,5)) %>%
+  filter(!is.na(q_340_insects),MOR!="00") %>%
+  ggplot(aes(MOR,fill=q_340_insects))+
+  geom_bar(position = "fill")+
+    theme_classic()+
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        legend.position = "none"
+        )+
+  ylab("Proportion"),
+rdbp %>%
+  select(b_reactiondate,grouping,q_340_insects,d_centres_country) %>%
+  mutate(MOR = substr(b_reactiondate,4,5)) %>%
+  filter(!is.na(q_340_insects),MOR!="00") %>%
+  ggplot(aes(MOR,fill=q_340_insects))+
+  geom_bar()+
+  theme_classic()+
+  labs(fill="Insect",x = "Month of the year")+
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        legend.position = c(0.01,.98),
+        legend.justification = c(0,1)),
+rdbp %>%
+  select(b_reactiondate,grouping,d_elicitor_gr5,d_centres_country) %>%
+  mutate(MOR = substr(b_reactiondate,4,5),
+         d_elicitor_gr5 = relevel(d_elicitor_gr5, "insects")) %>%
+  filter(!is.na(grouping),MOR!="00") %>%
+  group_by(MOR,grouping) %>%
+  summarize(n = n()) %>%
+  group_by(MOR) %>%
+  summarise(prop = n[1]/sum(n)) %>%
+  ggplot(aes(MOR,prop))+
+  geom_bar(stat="identity")+
+  theme_classic()+
+  labs(x = "Month of the year",y = "Fraction of insect elicited ANA"),
+  #theme(legend.position = c(0.01,.98),
+  #      legend.justification = c(0,1)),
+heights = c(0.4,1,0.5),
+ncol = 1
+)
 ```
@@
 Countries which reported less than 20 cases of anaphylaxis due to insect venom were not illustrated in this figure."}
-plot.proportions(rdb, "d_centres_country","q_340_insects",20)
+plot.proportions(rdb, "d_centres_country","q_340_insects",5)+
+  labs(x = "Country",y="Proportion of IVA cases",fill = "Species")
 ```
@@
 ```
+
+Cases of IVA were more requent in oder patients. Children suffered mainly form food elicited anaphylaxis. Therefore we adjusted the anlysis apropriately for age.
 
@@
 
+### ACE-I
+Ruef wrote in 2009 that the use of ACE I increases the risk of a severe reaction.
+
+```{r}
+data$q_423_ace <- yesonly(data$q_423_ace)
+
+wilcox.test(q_116_VAS_v7~q_423_ace,data = data)
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_423_ace,q_116_VAS_v7))+
+  geom_violin()
+
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+group_by(q_423_ace) %>% 
+  summarise(medianVAS=median(q_116_VAS_v7,na.rm=T),
+            IQRVAS = IQR(q_116_VAS_v7,na.rm=T))
+
+wilcox.test(ANAscore~q_423_ace,data = data %>% 
+  filter(d_elicitor_gr5=="insects") 
+)
+
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_423_ace,ANAscore))+
+  geom_violin()
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ group_by(q_423_ace) %>% 
+  summarise(medianANAscore=median(ANAscore,na.rm=T),
+            IQRANAscore = IQR(ANAscore,na.rm=T))
+
+cowplot::plot_grid(
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_423_ace,q_116_VAS_v7))+
+  geom_boxplot(),
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_423_ace,ANAscore))+
+  geom_boxplot()
+)
+
+```
+
+### Beta blockers 
+cardioselectivity??????
+```{r}
+data$q_423_beta <- yesonly(data$q_423_beta)
+
+wilcox.test(q_116_VAS_v7~q_423_beta,data = data)
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_423_beta,q_116_VAS_v7))+
+  geom_violin()
+
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+group_by(q_423_beta) %>% 
+  summarise(medianVAS=median(q_116_VAS_v7,na.rm=T),
+            IQRVAS = IQR(q_116_VAS_v7,na.rm=T))
+
+wilcox.test(ANAscore~q_423_beta,data = data %>% 
+  filter(d_elicitor_gr5=="insects") 
+)
+
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_423_beta,ANAscore))+
+  geom_violin()
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ group_by(q_423_beta) %>% 
+  summarise(medianANAscore=median(ANAscore,na.rm=T),
+            IQRANAscore = IQR(ANAscore,na.rm=T))
+
+cowplot::plot_grid(
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_423_beta,q_116_VAS_v7))+
+  geom_boxplot(),
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_423_beta,ANAscore))+
+  geom_boxplot()
+)
+
+```
+
+### mast cell burden - mastocytosis
+
+```{r}
+data$q_410_masto_cur %<>%  yesonly()
+
+wilcox.test(q_116_VAS_v7~q_410_masto_cur,data = data)
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_410_masto_cur,q_116_VAS_v7))+
+  geom_violin()
+
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+group_by(q_410_masto_cur) %>% 
+  summarise(medianVAS=median(q_116_VAS_v7,na.rm=T),
+            IQRVAS = IQR(q_116_VAS_v7,na.rm=T))
+
+wilcox.test(ANAscore~q_410_masto_cur,data = data %>% 
+  filter(d_elicitor_gr5=="insects") 
+)
+
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_410_masto_cur,ANAscore))+
+  geom_violin()
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ group_by(q_410_masto_cur) %>% 
+  summarise(medianANAscore=median(ANAscore,na.rm=T),
+            IQRANAscore = IQR(ANAscore,na.rm=T))
+
+cowplot::plot_grid(
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_410_masto_cur,q_116_VAS_v7))+
+  geom_boxplot(),
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_410_masto_cur,ANAscore))+
+  geom_boxplot()
+)
+
+```
+
+```{r}
+cowplot::plot_grid(
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_212_tryptase_value_v5,q_116_VAS_v7))+
+  geom_point()+
+  scale_x_log10(),
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+ggplot(aes(q_212_tryptase_value_v5,ANAscore))+
+  geom_point()+
+  scale_x_log10()
+)
+
+cowplot::plot_grid(
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                  "low",
+                  "high")) %>% 
+  filter(!is.na(typ_cat)) %>% 
+  ggplot(aes(typ_cat,q_116_VAS_v7))+
+  geom_boxplot(),
+data %>% 
+  filter(d_elicitor_gr5=="insects") %>% 
+  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                  "low",
+                  "high")) %>% 
+  filter(!is.na(typ_cat)) %>% 
+  ggplot(aes(typ_cat,ANAscore))+
+  geom_boxplot(),
+data %>% 
+  filter(d_elicitor_gr5=="insects",
+         !is.na(d_severity_rm)) %>% 
+  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                  "low",
+                  "high")) %>% 
+  group_by(d_severity_rm,typ_cat) %>% 
+  filter(!is.na(typ_cat)) %>% 
+    summarize(n = n()) %>% 
+  ggplot(aes(typ_cat,n,fill=d_severity_rm))+
+  geom_bar(stat="identity",position= "fill"),
+data %>% 
+  filter(d_elicitor_gr5!="insects") %>% 
+  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                  "low",
+                  "high")) %>% 
+  filter(!is.na(typ_cat)) %>% 
+  ggplot(aes(typ_cat,q_116_VAS_v7))+
+  geom_boxplot(),
+data %>% 
+  filter(d_elicitor_gr5!="insects") %>% 
+  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                  "low",
+                  "high")) %>% 
+  filter(!is.na(typ_cat)) %>% 
+  ggplot(aes(typ_cat,ANAscore))+
+  geom_boxplot(),
+data %>% 
+  filter(d_elicitor_gr5!="insects",
+         !is.na(d_severity_rm)) %>% 
+  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                  "low",
+                  "high")) %>% 
+  group_by(d_severity_rm,typ_cat) %>% 
+  filter(!is.na(typ_cat)) %>% 
+    summarize(n = n()) %>% 
+  ggplot(aes(typ_cat,n,fill=d_severity_rm))+
+  geom_bar(stat="identity",position= "fill")
+)
+
+```
+
+Here put the mastocytosis and baseline serum tryptase as the evaluation of severe risk
+Here it is crucial to compare to even these reactions that had nearly no reactions
+
+### Model the risk of severity of a severe reaction in IVA
+
+USe the model from our previous paper
+```
+```
+
 ## Managment
@@
 Out of these `r sum(greter_same)` (`r sum(greter_same)/sum(repeated_reaction_in_Reg)*100`%) had Insect elicited anaphylaxis and in `r greter_same[1]` (`r greter_same[1]/sum(greter_same)*100`) The follwoeing reaction was more severe than before.
+
+### Do we know how mnay beekeepers were in our dataset?
+We were unable to estimate the fraction of cases that were elicited in bee-keepers as this information is not asked in the questionnaire.
+
+### Hypothesis: Wasp stings are more severe than bee stings
+We previously saw that refractory anaphylaxis cases come from bees. Here we see no difference in the severity of cases elicited by bees and wasps (in the RM).
+
+```{r}
+ana_outcome_matched <- function(data = data4,
+         outcome = "d_severity_rmr",
+         matching_var = "q_340_insects",
+         confounders_cat = c("b_sex", "cc_cardio"),
+         confounders_int = c("d_age"),
+         selected_predictors = c(1:4)
+         ){
+  possible_outcomes <- c("d_severity_rm",
+                         "d_severity_rmr",
+                         "severity_brown",
+                         "q_116_VAS_v7",
+                         "ANAscore")
+  predictors <- c(matching_var, confounders_cat,confounders_int)
+  df <- data %>% # first get the data frame
+   filter(!!as.name(matching_var) %in% c("yellow jacket", "bee")) %>% 
+   select(!!!matching_var,
+          !!!outcome,
+          !!!confounders_cat,
+          !!!confounders_int
+          ) %>%
+          {.[complete.cases(.),]}
+  ### we need to match the cases now to the first level of theoutcome variable
+  ref_level <- levels(df[,matching_var] %>% droplevels())
+  matched <- df %>% 
+    mutate(match = ifelse(!!as.name(matching_var)==ref_level[1],T,F),
+           out=ifelse(!!as.name(outcome)=="severe",1,0)) %>% 
+    matchit(formula = as.formula(
+      paste("match ~ ",paste0(confounders_cat),"+", paste0(confounders_int,collapse = " + ")))) %>% 
+    match.data() 
+  ## IFELSE
+  if(outcome %in% possible_outcomes[c(2,3)]){
+    plot <- matched %>% 
+      ggplot(aes(!!as.name(matching_var),
+               fill = !!as.name(outcome)))+
+      geom_bar()+
+      facet_grid(reformulate(confounders_cat[1],confounders_cat[2]))
+
+    fit <- glm(as.formula(paste0(
+      "out ~ ",paste0(predictors[selected_predictors], collapse = " + ")
+    )),data= matched,family= "binomial")
+  } else if (outcome %in% possible_outcomes[c(1,4,5)]){
+     plot <- matched %>% 
+      ggplot(aes(!!as.name(matching_var),
+               y = !!as.name(outcome)))+
+      geom_boxplot()+
+      facet_grid(reformulate(confounders_cat[1],confounders_cat[2]))
+
+    fit <- glm(as.formula(paste0(
+      outcome, " ~ ", paste0(predictors[selected_predictors], collapse = " + "))),
+             data = matched)
+  }
+
+  return(list(mtched_df = matched,
+              ref_level = ref_level,
+              fit = fit,
+              plot = plot)) 
+}
+
+possible_outcomes <- c("d_severity_rm",
+                         "d_severity_rmr",
+                         "severity_brown",
+                         "q_116_VAS_v7",
+                         "ANAscore")
+
+output_severity <- lapply(possible_outcomes,
+                         function(x){
+                         ana_outcome_matched(data = data4,
+                            outcome = x,
+                            matching_var = "q_340_insects",
+                            confounders_cat = c("b_sex", "cc_cardio"),
+                            confounders_int = c("d_age"),
+                            selected_predictors = c(1:4))
+                           }
+)
+
+do_cof_tab <- function(x){
+  output_severity[[x]]$fit %>% summary() %>% {.$coefficients}%>% tidy() %>% as.matrix()
+}
+
+temp <- lapply(1:5,do_cof_tab) %>% do.call(what = rbind) %>%
+                          as_tibble() %>%
+  {mutate(.,
+         Estimate = as.numeric(Estimate),
+         Pr...t.. = as.numeric(Pr...t..)
+  )} %>%
+          {cbind(.,outcome = lapply(possible_outcomes,
+                          rep,
+                          nrow(.)/5) %>% unlist)
+          } %>% 
+           filter(.rownames!="(Intercept)")
+gridExtra::grid.arrange(output_severity[[2]]$plot,
+                        output_severity[[3]]$plot,
+                        output_severity[[1]]$plot,
+                        output_severity[[4]]$plot,
+                        output_severity[[5]]$plot,
+                        temp %>% 
+  ggplot(aes(.rownames,Pr...t.., color=outcome))+
+  geom_segment(x=0.5,xend=4.5,y=log10(0.05),yend=log10(0.05))+
+  scale_y_log10()+
+  geom_jitter()+
+  theme(axis.text.x = element_text(angle=40, hjust=1))
+  )#+
+  #geom_line(aes(group=outcome))
+```
+
+Based on the above graphs of patients matched according to sex and age we cannot say that the severity of their reaction differ between yellow-jackets and bees in any of the models except ANAscore - where yellow jacket were linked to  a slightly more severe reaction.
+
+But in the ANAscore variable it shows up to be significant !
+Yellowjackets stings are more severe than the bee stings.
+This is however very hard to compare to refractory cases.
+Maybe we can use the refractory case identifier to say in which group we see more refractory cases?
+
+### Increased BTSC develops in the normal range <8
+
+```{r}
+output_severity <- lapply(possible_outcomes,
+                         function(x){
+                         ana_outcome_matched(data = data4,
+                            outcome = x,
+                            matching_var = "q_340_insects",
+                            confounders_cat = c("tryp_cat", "cc_cardio"),
+                            confounders_int = c("d_age"),
+                            selected_predictors = c(1,2,4))
+                           }
+)
+
+do_cof_tab <- function(x){
+  output_severity[[x]]$fit %>% summary() %>% {.$coefficients}%>% tidy() %>% as.matrix()
+}
+
+temp <- lapply(1:5,do_cof_tab) %>% do.call(what = rbind) %>%
+                          as_tibble() %>%
+  {mutate(.,
+         Estimate = as.numeric(Estimate),
+         Pr...t.. = as.numeric(Pr...t..)
+  )} %>%
+          {cbind(.,outcome = lapply(possible_outcomes,
+                          rep,
+                          nrow(.)/5) %>% unlist)
+          } %>% 
+           filter(.rownames!="(Intercept)")
+gridExtra::grid.arrange(output_severity[[2]]$plot,
+                        output_severity[[3]]$plot,
+                        output_severity[[1]]$plot,
+                        output_severity[[4]]$plot,
+                        output_severity[[5]]$plot,
+                        temp %>% 
+  ggplot(aes(.rownames,Pr...t.., color=outcome))+
+  geom_segment(x=0.5,xend=4.5,y=log10(0.05),yend=log10(0.05))+
+  scale_y_log10()+
+  geom_jitter()+
+  theme(axis.text.x = element_text(angle=40, hjust=1))
+  )
+```
+
+Based on the above we see that the lower tryptase significantly decreased the severity of
+
+### Mastocytosis
+
+### Older Age
+
+### ACE - I
+
+### previous reactions (severe)
+
+### Male sex
+
+### Beta blockers and CARDIOSELECTIVE BLOCKERS
+
+### Evaluate the pediatric population separately
+
+### Number of SAR during VIT compare to other hypos
+#### Compare bee sit to wasp sit for AEs
+#### correlate the tryptase value with SIT types and severity
+### 
+they used generalized aditive models (package GAM?)
+Also ROC curves for confirmation
 
@@
 F2$plot2
-
 ```

### 2019-04-03 R/stat.R

@@ -8,7 +8,15 @@ require(vcd)
 
 #### GET THE DATA ########
 # Note the path that we need to use to access our data files when rendering this document
-load('../RefractoryAnaOrg/analysis/data/raw_data/data4.R')
+# data5 <-foreign::read.spss("analysis/data/raw_data/anaphylaxis_registry_15_mar_2019_mergedSD.sav",
+#                            use.value.labels = TRUE,
+#                            to.data.frame = T,
+#                            trim.factor.names = T,
+#                            trim_values = T)
+
+#load('../RefractoryAnaOrg/analysis/data/raw_data/data4.R')
+load("data.R") # here the anascore is done using the severity_analysis R script
+data4 <-data
 
 
 ####### Functions #############
@@ -363,7 +371,50 @@ levels(data4$q_561_hospital_admission_v6) %<>% {c(NA,.[2:3],NA)}
 levels(data4$q_562_intensive_care_v6)%<>% {c(NA,.[2:3],NA)}
 levels(data4$q_632_autoinj_number_v7) %<>% {c(.[1:3],NA)}
 levels(data4$q_422_stress) <- c("no","yes")
+levels(data4$q_340_insects) <-c(levels(data4$q_340_insects)[1:7],"other")
 
+### REstricted ring and messmer (conformat with the RUEFF work)
+data4$d_severity_rmr <- ifelse(data4$d_severity_rm%in%c(3,4),
+                               "severe",
+                               "mild")
+### ATOPIC VARIABLE (posible many NAs)
+data4$atopy <- ifelse(data4$q_410_rhinitis_cur=="yes"|
+                        data4$q_410_rhinitis_prev_v5=="yes"|
+                        data4$q_410_asthma_cur=="yes"|
+                        data4$q_410_asthma_prev_v5=="yes"|
+                        data4$q_410_ad_cur=="yes"|
+                        data4$q_410_ad_prev_v5=="yes",
+                      "yes",
+                      "no") %>% factor()
+summary(data4$q_410_ad_cur)
+
+#### Cardiologic grouping var
+data4$cc_cardio <- ifelse(data4$q_410_cardio_cur=="yes"|
+                        data4$q_410_cardio_prev_v5=="yes"
+                        ,
+                      "yes",
+                      "no") %>% factor()
+data4$cc_cardio %>% summary()
+data4$q_410_cardio_cur %>% summary()
+data4$q_410_cardio_prev_v5 %>% summary()
+
[…]

### 2019-04-08 analysis/paper/paper.Rmd
add abstract r code

@@
 output: 
-    #bookdown::html_document2:
-    bookdown::word_document2:
+    bookdown::html_document2:
+    #bookdown::word_document2:
       fig_caption: yes
@@
 Insect-venom elicited anaphylaxis is a common hypersensitivity reaction which may be life-threatening.
-Using the data from the European Anaphylaxis Registry (11596 cases in total) we identified insect-venom elicited anaphylaxis cases (n = 4482) and analyzed these in comparison to anaphylaxis elicited by other elicitors (n = 7114).
+Using the data from the European Anaphylaxis Registry (`r nrow(data)` cases in total) we identified insect-venom elicited anaphylaxis cases (n = `r {data$d_elicitor_gr5=="insects"} %>% sum()`) and analyzed these in comparison to anaphylaxis elicited by other elicitors (n = `r {data$d_elicitor_gr5!="insects"} %>% sum()`).
 
-The data show that 68.57% of all insect elicited cases were elicited by yellow jackets, followed by bees (21.86%).
-The insect venom elicited cases occurred mostly in outdoor places (44.65%) patients' homes (12.87%) or urban places (9.616%).
+The data show that `r {{rdb$q_340_insects=="yellow jacket"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`% of all insect elicited cases were elicited by yellow jackets, followed by bees (`r {{rdb$q_340_insects=="bee"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`%).
+The insect venom elicited cases occurred mostly in outdoor places (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(4),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%) patients' homes (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(8),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%) or urban places (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(5),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%).
 
 Skin, gastrointestinal and respiratory symptoms occurred less frequently in insect elicited cases of anaphylaxis, whereas cardiologic symptoms (with hypotension, collapse, and loss of consciousness) were more frequent.
-Intramuscular adrenaline (as a first-line therapy) was administered significantly less often in insect venom elicited cases (4.04%, p < 0.0001).
-The mortality rate in insect anaphylaxis was comparable (0.156%) to other cases (0.295%, p = 0.174).
+Intramuscular adrenaline (as a first-line therapy) was administered significantly less often in insect venom elicited cases (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(d_522_adren_agg) %>% summarize(n=n()) %>% {.[2,2]/(.[1,2])*100} %>% round(1)`% vs `r rdb %>% filter(is.na(q_340_insects)) %>% group_by(d_522_adren_agg) %>% summarize(n=n()) %>% {.[2,2]/(.[1,2])*100} %>% round(1)`, p < 0.0001).
+The mortality rate in insect anaphylaxis was comparable (`r `%) to other cases (0.295%, p = 0.174).
 
@@
 Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May till October.
-Their proportion to other cases during the summer seasons reached 60 % and was as under 1% of cases during winter.
+Their proportion to other cases during the summer seasons reached 60% and was as under 1% of cases during winter.
 Yellow-jacket was the most prominent causing-insect reported in these cases foolowed by bee (Fig \@ref(fig:insectstime)).
+
+```{r MORaltogether}
+
+```
 
@@
 Countries which reported less than 20 cases of anaphylaxis due to insect venom were not illustrated in this figure."}
-plot.proportions(rdb, "d_centres_country","q_340_insects",5)+
-  labs(x = "Country",y="Proportion of IVA cases",fill = "Species")
 ```
@@
 
-```{r}
+```{r,fig.height=8}
+require(broom)
+
 ana_outcome_matched <- function(data = data4,
@@
   geom_jitter()+
-  theme(axis.text.x = element_text(angle=40, hjust=1))
+  theme(axis.text.x = element_text(angle=40, hjust=1))+
+    annotate("text",x=4,y = c(log10(c(0.0001,0.00001,.000001,.0000001,.00000001))),
+             label = c(1776,1780,1206,234,1780))
   )#+
   #geom_line(aes(group=outcome))
+
+ggplot(temp,aes(.rownames,Pr...t.., color=outcome))+
+  geom_segment(x=0.5,xend=4.5,y=log10(0.05),yend=log10(0.05))+
+  scale_y_log10()+
+  geom_jitter(width = 0.2)+
+  theme(axis.text.x = element_text(angle=40, hjust=1))+
+    annotate("text",x=4,y = c(0.0001,0.00001,.000001,.0000001,.00000001)*10,
+             label = c("n=1776","n=1780","n=1206","n=234","n=1780"),size = 3)+
+  labs(x = "",y="p value")
 ```
@@
 
-```{r}
+```{r,fig.height=8}
 output_severity <- lapply(possible_outcomes,
@@
 
-Based on the above we see that the lower tryptase significantly decreased the severity of
+Based on the above we see that the severity of anaphylaxis significantly depends on tryptase concentration (here under 8 µg/ml). Only in the VAS scale this did not sho as significant.
+There might be problems with the vas scale...
 

### 2019-04-08 R/stat.R

@@ -270,6 +270,7 @@ funChi <-function(var,grouping_var){
     data.frame(gr = grouping_var) %>%
     {chisq.test(.[,1],.[,2])$p.value}
 }
+
 funN1 <- function(var, grouping_var){
   var %>%
     data.frame(grouping_var) %>% table %>%
@@ -469,7 +470,7 @@ makeTests <- function(grouping,rdb){
     rowwise() %>%
     mutate(fun = if(level==3|level==2){
       "chi"
-    } else if(variableName%in%c("q_212_tryptase_value_v5","d_age")) {
+    } else if(variableName%in%c("q_212_tryptase_value_v5","d_age","ANAscore")) {
       "t.test"
     } else if (variableName%in%c("d_severity_rm","d_organ_sum",
                                  "q_116_VAS_v7", "q_120_time_between_v4",
@@ -562,7 +563,7 @@ testInsectsbinomial <- makeTests(grouping,rdb=rdb) %>% arrange(pval)
 #   geom_point()+
 #   scale_y_log10()
 
-testInsectsbinomialTab <- testInsectsbinomial %>% filter(pval<1e-30) %>%
+testInsectsbinomialTab <- testInsectsbinomial %>% #filter(pval<1e-30) %>%
   select(variableName,counts_1,counts_2,fraq_1,fraq_2,
          pval,
          section) %>%
@@ -697,13 +698,13 @@ makeDF <- function(data,variables,grouping,outcome){
     grA <- which(grouping==levels(grouping)[1])
     grB <- which(grouping==levels(grouping)[2])
     rbind(
-      c(makeOdds(data =data, var1 = x, var2 = outcome),S="all"),
-      c(makeOdds(data =data[grA,], var1 = x, var2 = outcome),S="Lev1"),
-      c(makeOdds(data =data[grB,], var1 = x, var2 = outcome),S="Lev2"),
+      #c(makeOdds(data =data, var1 = x, var2 = outcome),S="all"),
+      c(makeOdds(data =data[grA,], var1 = x, var2 = outcome),S="IVA"),
+      c(makeOdds(data =data[grB,], var1 = x, var2 = outcome),S="non-IVA"),
       NA
     )
   }) %>% do.call(what=rbind) %>% data.frame()
-  return(rbind(rep(NA,length(o[1,])),o))
+  return(rbind(rep(NA,length(o[1,])),o) %>% data.frame())
 }
 
 
[…]

### 2019-04-08 R/stat.R
matching function done!

@@ -170,7 +170,14 @@ other_insects <- c("horse fly",
 data4$d_insect_gr4 <- as.character(data4$q_340_insects)
 data4$d_insect_gr4[which(data4$q_340_insects%in%other_insects)] <- "other"
 data4$d_insect_gr4 %<>% factor
-
+data4%<>%
+  mutate(d_AAI_prescribed = ifelse(q_540_why_autoinj_v5 %in% c("prescribed, available, but not used",
+                                                               "prescribed, but not available"),
+                                   "yes",
+                                   ifelse(q_540_why_autoinj_v5 =="not prescribed",
+                                          "no",
+                                          NA)
+  ))
 
 ##### DIAGRAM ##############
 library(DiagrammeRsvg)
@@ -1342,18 +1349,25 @@ F3$plot
 #rdb$
 #ggplot(aes())
 
-####Propensity Score MAtching #####
+#### Propensity Score MAtching for Adrenaline use #####
 #' The functition to match samples based on the minimal set of predictors
 #' provide data, grouping variables and predictor variables as well as others..
 require(MatchIt)
 prop_fun <- function(data, grouping_var, predictor_vars, other_vars=NULL,
                      method = "optimal",
                      samplesize = 200) {
+  if(is.logical(grouping_var)){
+    temp <- data %>%
+      select(!!!predictor_vars,!!!other_vars) %>%
+      mutate(grouping = grouping_var) %>%
+      filter(complete.cases(.))
+  } else {
   temp <- data %>%
     select(!!!grouping_var,!!!predictor_vars,!!!other_vars) %>%
     filter(complete.cases(.)) %>%
     mutate(grouping = ifelse(get(grouping_var) == "insects", T,
                            F))
+  }
   match_stats <- matchit(formula = paste0(grouping_var,
                                           "~",
                                           paste0(predictor_vars,
@@ -1386,6 +1400,78 @@ rdb %>%
 
 ggplot(temp$post_data, aes(grouping, q_116_VAS_v7))+
   geom_boxplot()
+
+
+####3 Here is the adrenalin data analysis
+#temp <-
+rdb %>%
+  #filter(d_AAI_prescribed!="no") %>%
+  group_by(grouping, d_AAI_prescribed,d_522_adren_agg) %>%
+  summarize(n = n()) %>%
+  ggplot(aes(y=n, x = d_AAI_prescribed,fill=d_522_adren_agg))+
[…]

### 2019-04-09 analysis/paper/paper.Rmd
makeTests function rewrite to use 'grouping' var

@@
 
+### Adrenaline use is lower in patients with IVA compared to other anaphylaxis cases
+
+```{r}
+grid.draw(fig_adrenuse1)
+```
+
+I matched patients with IVA to a group of similarly severe cases of non-IVA anaphylaxis and evaluated the use of Adrenaline in both of these groups.
+
 ## Diagnostics
@@
 But in the ANAscore variable it shows up to be significant !
-Yellowjackets stings are more severe than the bee stings.
+Yellow-jackets stings are more severe than the bee stings.
 This is however very hard to compare to refractory cases.

### 2019-04-09 R/stat.R

@@ -5,6 +5,8 @@ require(ggplot2)
 require(forestplot)
 require(summarytools)
 require(vcd)
+require(MatchIt)
+
 
 #### GET THE DATA ########
 # Note the path that we need to use to access our data files when rendering this document
@@ -57,6 +59,39 @@ f3 <- function(x){
     lapply(function(x){(length(which(x=="yes"))/length(x)*100) %>% roP})
 }
 
+
+#' Match patients
+#'
+#' Function match_patients takes a data frame from ANAreg and otuputs matched
+#' case_id according to a given matching variable. The various grouping variables
+#' have to still be implemented. Right now only the "grouping" variable is recognized
+#'
+#' @param grouping_var has to be a binomial var  of 1s and 0s
+#' @param data is a data frame derived from ANAreg
+#' @param matching_vars variable name to which we shouldmatch the two groups
+#' @return vector of case ids that can be used for further analysis
+#'
+#'
+match_patients <- function(data,grouping_var,matching_vars,df=F){
+  o <- data %>%
+    select(b_case_id,!!!grouping_var,!!!matching_vars) %>%
+    {filter(.,complete.cases(.))} %>%
+    mutate(grouping = ifelse(grouping=="insects",1,0)) %>%
+    matchit(formula = as.formula(paste0(grouping_var,
+                                        "~",
+                                        paste0(matching_vars,
+                                               collapse = "+"))
+    ),
+    method = "nearest",
+    ratio = 1) %>%
+    match.data() %>%
+    select(b_case_id) %>% pull()
+
+  if(df == T){
+    o <- data[data$b_case_id %in% o,]
+  }
+  return(o)
+}
 
 
 #### cases selection ####
@@ -471,7 +506,8 @@ variableSelectionTab <- function(data){
 #' @param rdb database on which to perform the test (you can restrict it)
 #' @return Tbl
 #' @export
-makeTests <- function(grouping,rdb){
+makeTests <- function(groups,rdb){
+  grouping <- rdb[,groups]
   variableSelectionTab <- variableSelectionTab(rdb)
   variableSelectionTab %<>% # Make division for tests
     rowwise() %>%
@@ -565,23 +601,36 @@ makeTests <- function(grouping,rdb){
 }
 
[…]

### 2019-05-13 analysis/paper/paper.Rmd
po konferencji w Groningen

@@
          4:5) %>% 
+  filter(variableName %in% c("q_522_cortico_iv","q_522_antih_iv","q_522_adren_im",
+                             "q_522_adren_iv","q_522_antih_oral","q_522_beta2_inhal")) %>% 
+  mutate(variableName = car::recode(
+    variableName,recodes = "'q_522_cortico_iv'='corticoids iv.';
+    'q_522_antih_iv'='antihistamines iv.';
+    'q_522_adren_im'='adrenaline im.';
+    'q_522_adren_iv'='adrenaline iv.';
+    'q_522_antih_oral'='antihistamines po.';
+    'q_522_beta2_inhal'='beta-2 mimetics inh.'"),
+    Group = car::recode(Group,"'fraq_1'='IVA';
+                        'fraq_2'='non-IVA'")) %>% 
   ggplot(aes(reorder(variableName,-Proportion),Proportion,fill=Group))+
   geom_bar(stat="identity",position="dodge")+
-  theme(axis.text.x=element_text(angle=90))+
-  geom_text(aes(y=0.5,label=reorder(signif(pval,2),-Proportion)),angle=90)
+  theme_classic()+
+  theme(axis.text.x = element_text(angle = 30,hjust =1),
+        legend.position = c(1,1),
+        legend.justification = c(1,1))+
+  labs(x = "Symptoms",y = "Proportion of cases",fill="")+
+  scale_fill_manual(values = rev(c("#E5F5E0", "#74C476","#005A32")))
+  #geom_text(aes(y=0.5,label=reorder(signif(pval,2),-Proportion)),angle=90)
 ```

### 2019-05-13 R/VAS.R

@@ -1,0 +1,48 @@
+require(tidyverse)
+#require(dplyr)
+d <- readxl::read_xlsx("analysis/data/raw_data/NORA Konferenz Auswertung Fragebogen.xlsx")
+d %>% str()
+names(d)
+clean <- d %>%
+  select(1:44) %>%
+  tidyr::gather(key = "symptom",value = "VAS",6:44) %>%
+  mutate(VAS = as.numeric(VAS),
+         experience = as.numeric(`Erfahrung in jahren`),
+         symptom = factor(symptom))
+clean$VAS[667] <- 55.6
+clean[which(clean$symptom == "death"&clean$VAS <70),]
+clean <- clean %>%
+  filter(ID != 23)
+
+clean %>%
+  ggplot(aes(reorder(symptom,VAS),VAS))+
+  geom_boxplot()+
+  theme(axis.text.x = element_text(angle = 70,hjust = 1))
+
+MEAN <- clean %>%
+  filter(complete.cases(.)) %>%
+  #split(.$symptom)
+  group_by(symptom) %>%
+  #summarize()
+  summarize(mean_VAS = mean(VAS,na.rm = T),
+            weighted_mean_VAS = weighted.mean(x = VAS, w=experience,na.rm = T),
+            sd_VAS = sd(VAS))
+MEAN %>%
+  ggplot(aes(reorder(symptom,weighted_mean_VAS),weighted_mean_VAS))+
+           geom_point()+
+  theme(axis.text.x = element_text(angle = 70,hjust = 1))
+
+MEAN %>%
+  ggplot(aes(reorder(symptom,sd_VAS),sd_VAS))+
+  geom_point()+
+  theme(axis.text.x = element_text(angle = 70,hjust = 1))
+
+clean %>%
+  filter(symptom =="angioedema") %>%
+  ggplot(aes(Land,experience))+
+  geom_jitter()+
+  theme(axis.text.x = element_text(angle = 70,hjust = 1))
+
+
+
+

### 2019-05-13 R/stat.R

@@ -279,8 +279,10 @@ create_graph() %>%
   add_global_graph_attrs(value="black",
                          attr = "fontcolor",
                          attr_type = "node") %>% #render_graph()
-  export_graph(file_name = "analysis/figures/flow.png", file_type = "png", title = NULL,
-               width = NULL, height = NULL)
+  export_graph(file_name = "analysis/figures/flow.png", file_type = "png",
+               title = NULL,
+               width = NULL,
+               height = NULL)
 
 
 ###Outcomes############
@@ -618,7 +620,7 @@ age_sex_matched <- rdb %>%
                  df=T)
 tests_matched <- makeTests(groups ="grouping",
                            rdb =age_sex_matched)
-tests_matchedTab <- testInsectsbinomial %>% filter(pval<1e-30) %>%
+tests_matchedTab <- tests_matched %>% filter(pval<1e-30) %>%
   select(variableName,counts_1,counts_2,fraq_1,fraq_2,
          pval,
          section) %>%
@@ -766,7 +768,7 @@ makeDF <- function(data,variables,grouping,outcome){
 
 #makeDF(rdb,"q_114",grouping,"severity_brown")
 makeTableText <- function(makeDF){
-  rbind(c("Symptoms","Elicitors","p"),
+  rbind(c("Cofactor","Elicitor","p"),
         makeDF[-c(1),c(1,6,2)])
 }
 
@@ -790,19 +792,20 @@ test1 <- makeDF(rdbp,
                    "q_410_masto_cur",
                    "q_410_cardio_cur",
                    "q_423_beta",
+                   "q_423_ace",
                    "q_4211_exercise",
                    "q_422_stress"),
                 "grouping",
                 "severity_brown")
 
 test1[1] <- c(NA,"Concomitant disease",NA,NA,"Asthma",NA,NA,"Mastocytosis",NA,NA,
-              "Cardiologic disese",NA,NA,"Beta-blockers",NA,NA,"Exercise",NA,NA,"Stress",NA,NA)
-test1 <- test1[-c(2:7,20:22),]
+              "Cardiologic disese",NA,NA,"Beta-blockers",NA,NA,"ACE-I",NA,NA,"Exercise",NA,NA,"Stress",NA,NA)
+test1 <- test1[-c(2:7,23:25),]
 png("analysis/figures/figForestfinal.png",
-    height = 400,
[…]

### 2019-05-15 analysis/paper/paper.Rmd
figure about ana managment

--- 2019-04-09 analysis/paper/paper.Rmd
@@
 USe the model from our previous paper
-```
-```
 
 ## Managment
-```{r}
-testInsectsbinomialTab$management[,-7] %>% knitr::kable()
-testInsectsbinomialTab$management[,-7]%>%
-  tidyr::gather(value ="Proportion",
-         key="Group",
-         4:5) %>% 
-  ggplot(aes(reorder(variableName,-Proportion),Proportion,fill=Group))+
-  geom_bar(stat="identity",position="dodge")+
-  theme(axis.text.x=element_text(angle=90))+
-  geom_text(aes(y=0.5,label=reorder(signif(pval,2),-Proportion)),angle=90)
-```
 
@@
 
-```{r}
-grid.draw(fig_adrenuse1)
+```{r adrenuse, fig.cap="**Therapy of IVA cases.** A: General managment methods in IVA and non-IVA.
+B: Reasons for not administering adrenaline in both IVA and non-IVA cases.
+C,D,E: Age, sex and severity distribution was matched in cases in both groups to allow for comparable results between IVA and non_IVA cases"}
+fig_adrenuse1
 ```
 
-I matched patients with IVA to a group of similarly severe cases of non-IVA anaphylaxis and evaluated the use of Adrenaline in both of these groups.
+Patients who underwent IVA significantly less often recived adrenaline treatment than in other adrenaline cases.
+After adjusting both groups for simmilar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the route of administration.
+Antihistaminic drugs on the other had were given more often in cases of IVA when compared to non-IVA.
+
+IVA was most often treated with corticosteroids and antihistamines (significantly more frequent than in other anaphylaxis cases).
+On the other hand adrenaline, beta-2 mimetics and oxygen were given more often in non-IVA.
+
+```{r}
+ANAscore_matched %>%
+        mutate(grouping = recode(grouping,
+                                 'insects'='IVA',
+                                 'other'='non-IVA'),
+               d_522_adren_agg = recode(d_522_adren_agg,
+                                       'no'='no adrenaline',
+                                       'yes'='adrenaline given') ) %>%
+                         group_by(grouping,
+                                  #d_AAI_prescribed,
+                                  d_522_adren_agg) %>%
+                         summarize(n = n()) %>%
+  pull(n) %>% 
+  matrix(nrow=2) %>% 
+  chisq.test() %>% 
+```
 

### 2019-05-15 R/stat.R

@@ -6,6 +6,7 @@ require(forestplot)
 require(summarytools)
 require(vcd)
 require(MatchIt)
+require(tidyr)
 
 
 #### GET THE DATA ########
@@ -431,7 +432,6 @@ data4$atopy <- ifelse(data4$q_410_rhinitis_cur=="yes"|
                         data4$q_410_ad_prev_v5=="yes",
                       "yes",
                       "no") %>% factor()
-summary(data4$q_410_ad_cur)
 
 #### Cardiologic grouping var
 data4$cc_cardio <- ifelse(data4$q_410_cardio_cur=="yes"|
@@ -439,9 +439,6 @@ data4$cc_cardio <- ifelse(data4$q_410_cardio_cur=="yes"|
                         ,
                       "yes",
                       "no") %>% factor()
-data4$cc_cardio %>% summary()
-data4$q_410_cardio_cur %>% summary()
-data4$q_410_cardio_prev_v5 %>% summary()
 
 ### Add tryptase coategory based on RUEFF (cut off 8)
 data4 %<>%
@@ -1502,105 +1499,230 @@ ANAscore_matched <-
                  "grouping",
                  c("ANAscore","d_age"),
                  T)
+testANAscoreMatched <- makeTests(groups = "grouping",
+                                 rdb=ANAscore_matched) %>%
+  #arrange(pval) %>%
+  #filter(pval<1e-30) %>%
+  select(variableName,counts_1,counts_2,fraq_1,fraq_2,
+         pval,
+         section) %>%
+         {split(.,.$section)}
 
 
+countYesManagment <- function(var){
+  tbl <- ANAscore_matched %>%
+    group_by(grouping,!! sym(var))  %>%
+    summarize(n = n()) %>%
+    filter(!!sym(var) == "yes")
+  data.frame(variable = names(tbl)[2],
+             IVA = tbl[1,3],
+             nonIVA = tbl[2,3])
+}
 
-fig_adrenuse1 <-
-  gridExtra::grid.arrange(
-    ANAscore_matched %>%
-  group_by(grouping,
-           d_AAI_prescribed,
-           d_522_adren_agg) %>%
-  summarize(n = n()) %>%
-  ggplot(aes(y = n, x = grouping, fill =
-               d_522_adren_agg)) +
-  geom_bar(stat = "identity")+
-    theme_classic()+
-    theme(legend.position = "right")+
-    scale_fill_brewer(palette = "Greens")+
[…]

### 2019-05-21 R/stat.R
RADAR PLOT

@@ -48,6 +48,12 @@ roP<- function(x){
   )
 }
 
+pval<- function(x){
+  ifelse(x > 0.001,
+         paste0("p = ",round(x,3)),
+         "p < 0.001")
+}
+
 f1 <- function(x){
   names(x) <- NULL
   return(do.call(t.test,x)$p.value %>% roP)
@@ -93,6 +99,39 @@ match_patients <- function(data,grouping_var,matching_vars,df=F){
   }
   return(o)
 }
+
+
+unknown_to_na <- function(vect){
+  ifelse(vect == "unknown",
+         NA,
+         ifelse(grepl(vect,pattern = "no"),
+                0,
+                vect)
+  )
+}
+
+levels(data$d_114_sum) %>%
+  unknown_to_na() %>%
+  {which(!is.na(.)==T)}-1
+
+char_to_num_levels <- function(variable){
+  levels(variable) <-  levels(variable) %>%
+           car::recode("'one' = 1;
+                  'two' = 2;
+                  'three' = 3;
+                  'four' = 4;
+                  'five' = 5;
+                  'six' = 6;
+                  'seven' = 7;
+                  'eight' = 8;
+                  'nine' = 9") %>%
+           unknown_to_na()
+  return(variable %>%
+           as.character() %>%
+           as.numeric())
+}
+
+
 
 
 #### cases selection ####
@@ -392,6 +431,11 @@ ChiF <- function(variablesDF,grouping){
 
 ### Data Cleanin #####
 
+data4$d_111_sum_wf <- data$d_111_sum %>% char_to_num_levels()
+data4$d_112_sum_wf <- data$d_112_sum %>% char_to_num_levels()
+data4$d_113_sum_wf <- data$d_113_sum %>% char_to_num_levels()
+data4$d_114_sum_wf <- data$d_114_sum %>% char_to_num_levels()
+
 data4$q_200_height_v7 %<>% as.character %>% as.numeric
 data4$q_200_weight_v7 %<>% as.character %>% as.numeric
 
@@ -605,7 +649,7 @@ testInsectsbinomial <- makeTests(groups = "grouping",rdb=rdb) %>% arrange(pval)
 #   geom_point()+
 #   scale_y_log10()
 
-testInsectsbinomialTab <- testInsectsbinomial %>% filter(pval<1e-30) %>%
+testInsectsbinomialTab <- testInsectsbinomial %>% #filter(pval<1e-30) %>%
   select(variableName,counts_1,counts_2,fraq_1,fraq_2,
          pval,
[…]

### 2019-05-23 analysis/paper/paper.Rmd
figure symptoms

@@
 
+We saw a significant difference in the clinical features of insect anaphylaxis in children and adults.
+The younger popoulation significantly less often had concomitant conditions (i.e.
+DM, HT, malignant diseases and mastocytosis), and more often presented with atopic dermatitis, rhinitis and asthma.
+Therefore, we decided to adjust the analysis for age and sex to reduce the comparison bias.
+When comparing the management of both types of anaphylaxis we also matched the control group according to severity.
+
 <!-- ## Data handling and statisitcs -->
@@
 
-## Demographics
-
-We saw a significant difference in the clinical features of insect anaphylaxis in children and adults.
-The younger popoulation significantly less often had concomitant conditions (i.e.
-DM, HT, malignant diseases and mastocytosis), and more often presented with atopic dermatitis, rhinitis and asthma.
-Therefore, we decided to adjust the analysis for age and sex.
-
-```{r}
-demoTab %>% rbind(demo_age_tab) %>% t() %>%  
-  {data.frame(` `=rownames(.),.)} %>%
-#   flextable(col_keys = c("X.","female","male","children.adolescents","adults")) 
-# topology <- data.frame(col_keys = c("X.","female","male","children.adolescents","adults"),
-#                        colB = c("","Sex","Sex","Age","Age"),
-#                        ColA = c("","Female","Male","<18",">=18"))
-# tempT %>%  
-#   set_header_df(mapping = topology,key = "col_keys") %>% 
-#   merge_h(part="header") %>% 
-#   merge_v() %>% 
-#   empty_blanks() %>% 
-#   autofit()
-knitr::kable(col.names = c("","female","male","<18","adults"),row.names = F)
-```
-
-```{r, results='asis'}
-# st_options('bootstrap.css', FALSE)
-# #rdb %>% select()
-# ctable(rdb$d_centres_country %>% sapply(function(x){
-#   as.character(x) %>% factor()
-# }), rdb$q_340_insects%>% sapply(function(x){
-#   as.character(x) %>% factor()
-# }), style = 'rmarkdown',
-#        useNA = "no",
-#        prop = "r",
-#        totals = F,
-#        round.digits = 1,
-# dnn = c("Countries","Insects"))
-
-```
-
 ## Temporal distribution of anaphylaxis cases by elicitor.
 
-```{r insectstime, fig.cap="Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occured.
+```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occured.
 Less frequent insects were groupeped together as 'other'.
-The lowest panel shows the proportion of insect elicited cases to ceses elicited by other triggers.", fig.height=6,fig.width=4}
-#### plot MOR#####
-plot_MOR <- gridExtra::grid.arrange(
-  #cowplot::plot_grid(
-  rdbp %>%
-  select(b_reactiondate,grouping,q_340_insects,d_centres_country) %>%
-  mutate(MOR = substr(b_reactiondate,4,5)) %>%
-  filter(!is.na(q_340_insects),MOR!="00") %>%
-  ggplot(aes(MOR,fill=q_340_insects))+
-  geom_bar(position = "fill")+
-    theme_classic()+
-  theme(axis.text.x = element_blank(),
-        axis.title.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        legend.position = "none"
-        )+
-  ylab("Proportion"),
-rdbp %>%
-  select(b_reactiondate,grouping,q_340_insects,d_centres_country) %>%
-  mutate(MOR = substr(b_reactiondate,4,5)) %>%
-  filter(!is.na(q_340_insects),MOR!="00") %>%
-  ggplot(aes(MOR,fill=q_340_insects))+
-  geom_bar()+
-  theme_classic()+
-  labs(fill="Insect",x = "Month of the year")+
-  theme(axis.text.x = element_blank(),
-        axis.title.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        legend.position = c(0.01,.98),
-        legend.justification = c(0,1)),
-rdbp %>%
-  select(b_reactiondate,grouping,d_elicitor_gr5,d_centres_country) %>%
-  mutate(MOR = substr(b_reactiondate,4,5),
-         d_elicitor_gr5 = relevel(d_elicitor_gr5, "insects")) %>%
-  filter(!is.na(grouping),MOR!="00") %>%
-  group_by(MOR,grouping) %>%
-  summarize(n = n()) %>%
-  group_by(MOR) %>%
-  summarise(prop = n[1]/sum(n)) %>%
-  ggplot(aes(MOR,prop))+
-  geom_bar(stat="identity")+
-  theme_classic()+
-  labs(x = "Month of the year",y = "Fraction of insect elicited ANA"),
-  #theme(legend.position = c(0.01,.98),
-  #      legend.justification = c(0,1)),
-heights = c(0.4,1,0.5),
-ncol = 1
-)
+B: The proportion of insect elicited cases to ceses elicited by other triggers considering patient's age.
+C: Geographical differences in the most common eliciotors of IVA.
+Countries which reported less than 10 cases of anaphylaxis due to insect venom were not illustrated in this figure."}
+fig_basic
 ```
@@
 Their proportion to other cases during the summer seasons reached 60% and was as under 1% of cases during winter.
-Yellow-jacket was the most prominent causing-insect reported in these cases foolowed by bee (Fig \@ref(fig:insectstime)).
+Yellow-jacket was the most prominent IVA-causing insect reported in these cases followed by bees.
+The IVA causing insects differed in European countries with hornets being more prominient on the south of europe.
+IVA was more frequent in adults and seniors when compared to children and youg adults.
+(Fig \@ref(fig:insectstime)).
 
-```{r MORaltogether}
-
-```
-
-```{r insectsgeo, fig.cap="Proportion of anaphylaxis cases elicited by specific insects in reporting countries.
-Less frequent insects were groupeped together as 'other'.
-Countries which reported less than 20 cases of anaphylaxis due to insect venom were not illustrated in this figure."}
-```
-
-The Figure \@ref(fig:insectsgeo) shows the distribution of insects according to the geographic area (countries).
-The most common elicitor among most of the countries were yellow-jackets.
-Switzerland had significantly more reactions to bees than Germany.
-Centers from Italy reported more cases elicited by hornets in comparison to other countries.
-
-```{r age}
-cdplot(rdb$grouping~as.numeric(rdb$d_age))
-```
-
-Cases of IVA were more requent in oder patients. Children suffered mainly form food elicited anaphylaxis. Therefore we adjusted the anlysis apropriately for age.
-
-```{r}
-mosaicplot(formula = d_severity_rm~d_insect_gr4,data = rdb)
-rdb %>% filter(!is.na(d_insect_gr4)) %>% 
-ggplot(aes(d_insect_gr4,fill=factor(d_severity_rm)))+
-  geom_bar(position="fill")+
-  facet_grid(.~d_age_gr2)
-```
+Cases of IVA were more frequent in older patients. Children suffered mainly from food elicited anaphylaxis. Therefore we adjusted the anlysis apropriately for age.
 
 ## Symptoms
-```{r}
-testInsectsbinomialTab$symptoms[,-7] %>% knitr::kable()
-```
-
-```{r}
-plotSympt
-```
 
@@
 
-### Adrenaline use is lower in patients with IVA compared to other anaphylaxis cases
+### Adrenaline is less often used in patients with IVA compared to other anaphylaxis cases
 
@@
 
-Patients who underwent IVA significantly less often recived adrenaline treatment than in other adrenaline cases.
+Patients who underwent IVA significantly less often recived adrenaline treatment than in other adrenaline cases (`r test_adrenuse1$observed %>% prop.table()`, `r test_adrenuse1$p.value %>% pval()`).
 After adjusting both groups for simmilar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the route of administration.
@@
 On the other hand adrenaline, beta-2 mimetics and oxygen were given more often in non-IVA.
-
-```{r}
-ANAscore_matched %>%
-        mutate(grouping = recode(grouping,
-                                 'insects'='IVA',
-                                 'other'='non-IVA'),
-               d_522_adren_agg = recode(d_522_adren_agg,
-                                       'no'='no adrenaline',
-                                       'yes'='adrenaline given') ) %>%
-                         group_by(grouping,
-                                  #d_AAI_prescribed,
-                                  d_522_adren_agg) %>%
-                         summarize(n = n()) %>%
-  pull(n) %>% 
-  matrix(nrow=2) %>% 
-  chisq.test() %>% 
-```
 
@@
 ```{R adrenalineuse, fig.cap = "The proportion of insect venom elicited anaphylaxis cases that were treated with adrenaline in comparison to other elicitors."}
-F2$plot2
+fig_adrenuse1
 ```
+
+```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors",fig.width=8}
+fig_symptoms
+```

### 2019-05-24 analysis/paper/paper.Rmd
Figure cofactors

@@
 output: 
-    bookdown::html_document2:
-    #bookdown::word_document2:
+    #bookdown::html_document2:
+    bookdown::word_document2:
       fig_caption: yes
@@
 
-```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occured.
-Less frequent insects were groupeped together as 'other'.
-B: The proportion of insect elicited cases to ceses elicited by other triggers considering patient's age.
-C: Geographical differences in the most common eliciotors of IVA.
-Countries which reported less than 10 cases of anaphylaxis due to insect venom were not illustrated in this figure."}
-fig_basic
-```
-
 Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May till October.
@@
 
-```{r}
-crammerElicitorBinvsSympt[1:15,] %>% arrange(desc(CV)) %>% knitr::kable()
+Patients with Insect Venom Allergy had a lot more cardiologic symptoms than anaphylaxis cases that were elicited through other triggers.  
+Dizziness, Loss of consciousness hypotension, collapse, Throat tightness and nausea were more often associated with the insect cases.
 
-crammerElicitorBinvsSympt[1:15,] %>% arrange(desc(CV)) %>%
-  filter(CV<0.3,
-         CV>0.15) %>% 
-  ggplot(aes(reorder(var,-CV),CV))+
-  geom_col()+
-  theme(axis.text.x = element_text(angle=90))
-```
-
-Patients with Insect Venom Allergy had a lot more cardiologic symptoms than anaphylaxis cases that were elicited through other triggers.  
-Dizziness, Loss of consciousness hypotension, collapse, Throat tightness and nausea were a lot more often associated with the insect cases.
-
-Forest plot ! Risk that the symptom will occure given that the reaction was elicited by Insects or not
+<!-- Forest plot ! Risk that the symptom will occure given that the reaction was elicited by Insects or not -->
 
@@
 
-```{r flow}
+```{r flow,fig.width=6,fig.height=4}
 knitr::include_graphics("../figures/flow.png")
@@
 
-```{r figforestcofactors}
-knitr::include_graphics("../figures/figForest.png")
+##### pagebreak
+
+```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occured.
+Less frequent insects were groupeped together as 'other'.
+B: The proportion of insect elicited cases to ceses elicited by other triggers considering patient's age.
+C: Geographical differences in the most common eliciotors of IVA.
+Countries which reported less than 10 cases of anaphylaxis due to insect venom were not illustrated in this figure.",fig.height=4,fig.width = 7}
+fig_basic
 ```
 
-```{r figforestcofactorskids}
-knitr::include_graphics("../figures/kidsForest.png")
-```
+##### pagebreak
 
-```{r figforestcofactorsadults}
-knitr::include_graphics("../figures/adultsForest.png")
-```
-
-```{r}
-mosaic(d_organ_sum~d_elicitor_gr5,data=rdb)
-```
-
-```{r}
-mosaic(d_severity_rm~d_elicitor_gr5,data=rdb)
-```
-
-```{r}
-hypotensionPlot
-```
-
-```{r}
-cramerPlot
-```
-
-```{r}
-plot.Cramer(supraCramerFun(data =rdbp,
-               vars =c( "q_114",
-               "q_112_incontinence",
-               "q_114_dizziness",
-               "q_112_abdominal_pain",
-               "q_114_loss_of_consciousness",
-               "q_114_hypotension_collapse_v5",
-               "q_114_reductions_of_alertness",
-               "q_111_angioedema",
-               "q_113_wheezing_expiratory_distress_v5"),
-               grouping = "grouping",
-               subset = "d_age_gr2"))
-
-```
-
-```{r}
-plot.Cramer(cramerFun(data =rdbp,
-               vars=c("q_114",
-               "q_112_incontinence",
-               "q_114_dizziness",
-               "q_112_abdominal_pain",
-               "q_114_loss_of_consciousness",
-               "q_114_hypotension_collapse_v5",
-               "q_114_reductions_of_alertness",
-               "q_111_angioedema",
-               "q_113_wheezing_expiratory_distress_v5"),
-               grouping = "grouping"))
-```
-
-```{r}
-tryptase_plot(rdb)+
-  scale_x_log10()+
-  facet_grid(~d_age_gr2)
-
-```
-
-```{r reactionmonthincountries, fig.cap="Proportional insect venom induced anaphylaxis to specific insects during a year according to countries.", fig.height=8,fig.width=5}
-plot_MOR+
-  facet_grid(d_centres_country~.)
-```
-
-```{r}
-plot_MOR
-```
-
-```{R adrenalineuse, fig.cap = "The proportion of insect venom elicited anaphylaxis cases that were treated with adrenaline in comparison to other elicitors."}
+```{R adrenalineuse, fig.cap = "The proportion of insect venom elicited anaphylaxis cases that were treated with adrenaline in comparison to other elicitors.",fig.height=4,fig.width = 6}
 fig_adrenuse1
@@
 
-```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors",fig.width=8}
+##### pagebreak
+
+```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors",fig.width=9.5,fig.height = 6}
 fig_symptoms
 ```
+
+##### pagebreak
+
+```{r figcofactors, fig.cap="Cofactors of insect venom anaphylaxis ad their relationship to severity of a reaction",fig.width=4.2,fig.height = 6}
+cof_fig
+```

### 2019-05-24 R/stat.R

@@ -836,8 +836,51 @@ makeForestPlot <- function(data,variables,grouping,outcome,cLow=0.2,cHigh = 4){
              col=fpColors(box="royalblue",line="darkblue", summary="royalblue"))
 }
 
+library(forcats)
+or_plot <- function(my_result){
+  my_result %>%
+    mutate_all(unlist) %>%
+    mutate(variable = forcats::fct_inorder(V)) %>% #make variable into factor (otherwise ggplot will make alphabetically)
+    ##start ggplot()
+    ggplot(aes(x = forcats::fct_rev(V),
+               y = mean))+
+    geom_hline(yintercept = 1,
+               colour = 'grey')+ #1.0 reference line
+    geom_errorbar(aes(ymin = lower, ymax = upper),
+                  width  = 0.2,
+                  colour = '#41ae76',
+                  size   = 0.9) + #OR lower-upper errorbars
+    ##geom_linerange(aes(ymin=or_lower, ymax=or_upper), width=0.2, colour='#41ae76', size=0.9) + #alternative to errorbars
+    geom_point(shape  = 18,
+               size   = 2.5,
+               colour = '#41ae76') +
+    ##explanatory labels
+    geom_text(aes(label = S),
+              vjust = -0.5) +
+    ##reference level labels
+    #geom_text(data = my_result %>%
+    #            filter(is.na(or)) %>%
+    #            mutate(or_reference = 1.0),
+    #          aes(label = value, x = variable, y = or_reference), angle = 90, vjust = -0.2, colour = '#737373') +
+    coord_flip() +
+    scale_y_log10() + #log scale breaks=c(0.2, 0.5, 1, 1.5, 2.0, 3.0), breaks=c(0.01, 0.1, 0.5, 1, 2.0)
+    theme_bw() +                          #theme
+    theme(axis.text.y  = element_blank(),
+          axis.ticks.y = element_blank(),
+          axis.title   = element_blank(),
+          #panel.border = element_blank(),
+          axis.text.x  = element_text(size=12, colour='black')
+          #,panel.grid.major.y = element_blank() #optinally only remove horizontal grid lines (keeping odds ratio ones)
+          ,panel.grid.major = element_blank() #these remove all grid lines
+          ,panel.grid.minor = element_blank()
+    )
+
+  #or_plot
+
+}
+
 ### testing
-test1 <- makeDF(rdbp,
[…]

### 2019-07-03 R/VAS.R
VAS analysis from the NORA conference

@@ -1,5 +1,51 @@
+#library(plyr)
 require(tidyverse)
 #require(dplyr)
+
+## Summarizes data.
+## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
+##   data: a data frame.
+##   measurevar: the name of a column that contains the variable to be summariezed
+##   groupvars: a vector containing names of columns that contain grouping variables
+##   na.rm: a boolean that indicates whether to ignore NA's
+##   conf.interval: the percent range of the confidence interval (default is 95%)
+summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
+                      conf.interval=.95, .drop=TRUE) {
+  #library(plyr)
+
+  # New version of length which can handle NA's: if na.rm==T, don't count them
+  length2 <- function (x, na.rm=FALSE) {
+    if (na.rm) sum(!is.na(x))
+    else       length(x)
+  }
+
+  # This does the summary. For each group's data frame, return a vector with
+  # N, mean, and sd
+  datac <- ddply(data, groupvars, .drop=.drop,
+                 .fun = function(xx, col) {
+                   c(N    = length2(xx[[col]], na.rm=na.rm),
+                     mean = mean   (xx[[col]], na.rm=na.rm),
+                     sd   = sd     (xx[[col]], na.rm=na.rm)
+                   )
+                 },
+                 measurevar
+  )
+
+  # Rename the "mean" column
+  datac <- rename(datac, c("mean" = measurevar))
+
+  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
+
+  # Confidence interval multiplier for standard error
+  # Calculate t-statistic for confidence interval:
+  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
+  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
+  datac$ci <- datac$se * ciMult
+
+  return(datac)
+}
+
+
 d <- readxl::read_xlsx("analysis/data/raw_data/NORA Konferenz Auswertung Fragebogen.xlsx")
 d %>% str()
 names(d)
@@ -24,13 +70,16 @@ MEAN <- clean %>%
   #split(.$symptom)
   group_by(symptom) %>%
   #summarize()
[…]

### 2019-07-03 R/stat.R

@@ -10,6 +10,8 @@ require(tidyr)
 library(ggradar)
 library(tibble)
 library(scales)
+require(ggpubr)
+require(rlang)
 
 #### GET THE DATA ########
 # Note the path that we need to use to access our data files when rendering this document
@@ -88,11 +90,16 @@ f3 <- function(x){
 #' @return vector of case ids that can be used for further analysis
 #'
 #'
-match_patients <- function(data,grouping_var,matching_vars,df=F){
+match_patients <- function(data,
+                           grouping_var,
+                           matching_vars,
+                           df=F,
+                           grouping_var_level = "insects"
+){
   o <- data %>%
     select(b_case_id,!!!grouping_var,!!!matching_vars) %>%
     {filter(.,complete.cases(.))} %>%
-    mutate(grouping = ifelse(grouping=="insects",1,0)) %>%
+    mutate(grouping = ifelse(grouping==grouping_var_level,1,0)) %>%
     matchit(formula = as.formula(paste0(grouping_var,
                                         "~",
                                         paste0(matching_vars,
@@ -996,19 +1003,19 @@ makeForestPlot(age_sex_matched,
                "d_severity_rmr")
 dev.off()
 
-makeForestPlot(age_sex_matched %>%
-                 mutate(tryp_cat %>% factor),
-               c( "q_410_masto_cur",
-                  "q_410_asthma_cur",
-                  "q_410_cardio_cur",
-                  "q_423_beta",
-                  "q_423_ace",
-                  #"tryp_cat",
-                  "q_423_asa"
-                  ),
-               "grouping",
-               "severity_brown"
-               )
+# makeForestPlot(age_sex_matched %>%
+#                  mutate(tryp_cat %>% factor),
+#                c( "q_410_masto_cur",
+#                   "q_410_asthma_cur",
+#                   "q_410_cardio_cur",
+#                   "q_423_beta",
+#                   "q_423_ace",
+#                   #"tryp_cat",
+#                   "q_423_asa"
+#                   ),
[…]

### 2019-07-03 R/test_diagramm.R

@@ -1,0 +1,56 @@
+nodes <-
+  create_node_df(
+    n = 6,
+    type = c("a","b","b","c","d","d"),
+    label= c("insect sting reaction in the ER",
+             "anaphylaxis",
+             "local reaction",
+             "treatment",
+             "acute serum samples\nwithin 2h of reaction",
+             "baseline serum samples\nafter > 2 weeks"
+             ),
+
+    #color = c("red", "green",
+    #          "grey", "blue"),
+    #value = c(3.5, 2.6, 9.4, 2.7),
+    shape = "rectangle",
+    width = c(2,1,1,1,2,2),
+    x = c(0, -.6, .6, -0.6,-0, -0),
+    y = c(0,-0.5,-0.5,  -1,-1.5,-2),
+    height = c(0.3,0.3,0.3,0.3,0.4,0.4))
+
+
+edges <- create_edge_df(from = c(1,1,2,4,5,3),
+                        to =   c(2,3,4,5,6,5),
+                        headport = c(5,0,0,0,0,0))
+# Add the node data frame to the
+# graph object to create a graph
+# with nodes
+gr <- create_graph() %>%
+  add_node_df(node_df = nodes) %>%
+  add_edge_df(edge_df = edges) %>%
+  add_global_graph_attrs(
+    attr = "ortho",
+    value = 1,
+    attr_type = "graph") %>%
+  add_global_graph_attrs(value="black",
+                         attr = "fontcolor",
+                         attr_type = "node") #%>% render_graph()
+  export_graph(file_name = "analysis/figures/flow.png", file_type = "png",
+               title = NULL,
+               width = NULL,
+               height = NULL)
+
+  library(network)
+  library(sna)
+  n <- network(rgraph(6, tprob = 0.2), directed = FALSE)
+
+library(ggnetwork)
+
+  df <- ggnetwork(n, layout = "fruchtermanreingold", cell.jitter = 0.75)
+  df$x <- c(0, -.6, .6, -0.6,-0, -0)
+  ggplot(df,
+         aes(x = x, y = y, xend = xend, yend = yend))+
+   #ggplot(n,aes(x = x, y = y, xend = xend, yend = yend))+
+    geom_edges(color = "grey50") +
+    theme_blank()

### 2019-07-03 helpers/recode_variables.R

@@ -1,0 +1,72 @@
+recode_variables <- function(x){
+  car::recode(x,
+  recodes =
+    "'q_111'='cutaneous symptoms';
+     'q_111_angioedema'= 'angioedema';
+     'q_111_erythema_flush_v5' = 'flush';
+     'q_111_pruritus' = 'pruritus';
+     'q_111_urticaria' = 'urticaria';
+     'q_111_conjunctivitis_v5' = 'cojunctivitis';
+     'q_112' = 'gastrointestinal symptoms';
+     'q_114_dizziness'='dizziness';
+     'q_114_loss_of_consciousness'='loss of consciousness';
+     'q_112_abdominal_pain'='abdominal pain';
+     'q_113_wheezing_expiratory_distress_v5'='expiratory distress';
+      'q_522_volume'='volume iv.';
+      'q_561_hospital_admission_v6'='hospital admission';
+      'q_562_intensive_care_v6'='intensive care';
+  'q_112_nausea' = 'nausea';
+                                             'q_113_throat_tightness_v5' = 'throat tightness';
+                                             'q_114_reductions_of_alertness' = 'reduced allertness';
+                                             'q_112_incontinence' = 'incontinence';
+                                             'q_112_abdominal_pain' = 'abdominal pain';
+                                             'q_112_diarrhoea' = 'diarrhoea';
+                                             'q_113_wheezing_expiratory_distress_v5' = 'expiratory distress';
+                                             'q_113_rhinitis_v5' = 'rhinitis';
+                                             'q_114_dizziness' = 'dizziness';
+                                             'q_114_hypotension_collapse_v5' = 'hypotension';
+  'q_112_dysphagia_v5' = 'dysphagia';
+  'q_112_vomiting' = 'vomiting';
+  'q_113' = 'respiratory symptoms';
+  'q_113_respiratory_arrest' = 'respiratory arrest';
+  'q_113_dyspnea' = 'dyspnea';
+  'q_113_chest_tightness_v5' = 'chest tightness';
[…]

### 2019-07-04 additional analyses/VAS.Rmd
figures corrected

---
output:
  word_document: default
  html_document: default
---
# Expert opinion on the severity of individual symptoms in anaphylaxis

```{r init,include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F)
#library(plyr)
require(tidyverse,quietly = T)
```

```{r fig.height=6}
d <- readxl::read_xlsx("../analysis/data/raw_data/NORA Konferenz Auswertung Fragebogen.xlsx")
#d %>% str()
#names(d)
clean <- d %>%
  select(1:44) %>%
  tidyr::gather(key = "symptom",value = "VAS",6:44) %>%
  mutate(VAS = as.numeric(VAS),
         experience = as.numeric(`Erfahrung in jahren`)#,
         #symptom = factor(symptom)
         )
clean$VAS[667] <- 55.6
#clean[which(clean$symptom == "death"&clean$VAS <70),]
clean <- clean %>%
  filter(ID != 23)
clean$symptom[clean$symptom == "pruritus/ ich"] <- "pruritus / itch"

MEAN <- clean %>%
  filter(complete.cases(.)) %>%
  #split(.$symptom)
  group_by(symptom) %>%
  #summarize()
  summarise(mean_VAS = mean(VAS,na.rm = T),
            weighted_mean_VAS = weighted.mean(x = VAS, w=experience,na.rm = T),
            sd_VAS = sd(VAS, na.rm=T))
MEAN %>%
  ggplot(aes(reorder(symptom,weighted_mean_VAS),weighted_mean_VAS))+
           geom_point()+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 70,hjust = 1))+
  geom_errorbar(aes(ymin = weighted_mean_VAS-sd_VAS, ymax =weighted_mean_VAS+ sd_VAS))+
  labs(x = "Symptoms", y = "severity according to weighted mean VAS")
```

We recived answers from `r clean$ID %>% factor()%>% levels() %>% length()` experts with a mean experience of `r clean$experience %>% mean(na.rm = T) %>% round (2)` years (SD = `r clean$experience %>% sd(na.rm = T) %>% round (2)`) from `r clean$Land %>% factor() %>% levels() %>% length()` countries.

The table below summarizes the expert group.

```{r}
clean %>% 
  mutate(Country = Land) %>% 
  group_by(Country) %>% 
  summarize(n = n()/39,
            `mean experience [y]` = mean(experience, na.rm = T),
            `mean VAS across all questions` = mean(VAS, na.rm = T), 
            `SD VAS` = sd(VAS, na.rm = T)) %>% 
  knitr::kable(digits = 2,caption = "Summary statistics for the expert panel according to country of origin" )
```

### 2019-07-04 analysis/paper/paper.Rmd
VAS analysis from the NORA conference
figures corrected

@@
 
-```{r adrenuse, fig.cap="**Therapy of IVA cases.** A: General managment methods in IVA and non-IVA.
-B: Reasons for not administering adrenaline in both IVA and non-IVA cases.
-C,D,E: Age, sex and severity distribution was matched in cases in both groups to allow for comparable results between IVA and non_IVA cases"}
-fig_adrenuse1
+```{r adrenuse, fig.cap="**Therapy of IVA cases.** A: General managment methods in IVA and non-IVA. B: Reasons for not administering adrenaline in both IVA and non-IVA cases. C,D,E: "}
+#fig_adrenuse1
 ```
@@
 <!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
-##### pagebreak
+#####
 
@@
 
-##### pagebreak
+#####
 
@@
 
-```{r flow,fig.width=6,fig.height=4}
-knitr::include_graphics("../figures/flow.png")
+```{r flow,fig.width=6,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale to include cases in our analysis.
+B,C,D: Sex, age, and severity distribution was matched in cases in both groups to allow for comparable results between IVA and non_IVA cases"}
+ggarrange(ggplot()+
+            background_image(png::readPNG("../figures/flow.png")),
+          lower_panel,
+          ncol = 1,
+          nrow = 2,
+          heights = c(1,0.4),
+          labels = c("A","")
+)
 ```
 
-##### pagebreak
+#####
 
@@
 
-##### pagebreak
+#####
 
-```{R adrenalineuse, fig.cap = "The proportion of insect venom elicited anaphylaxis cases that were treated with adrenaline in comparison to other elicitors.",fig.height=4,fig.width = 6}
-fig_adrenuse1
-```
-
-##### pagebreak
-
-```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors",fig.width=9.5,fig.height = 6}
+```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors",fig.width=9.5,fig.height = 8}
 fig_symptoms
@@
 
-##### pagebreak
+#####
 
-```{r figcofactors, fig.cap="Cofactors of insect venom anaphylaxis ad their relationship to severity of a reaction",fig.width=4.2,fig.height = 6}
+```{r adrenalineuse, fig.cap = "Therapy modes in patients with IVA compared to oter elicitors. A)  B)",fig.height=10,fig.width = 7}
+
+ggpubr::ggarrange(
+  ggarrange(
+    plotManagement,
+    rf_ggplot_treatment,
+    labels = c("A","B")
+  ),
+  #lower_panel,
+  matrix_venom$p+
+    labs (x = "", y = "")+
+    geom_segment(aes(x = c(8.5),
+                     xend = c(8.5), 
+                     y = c(0.5),
+                     yend = c(39.5)),color = "red")+
+    geom_segment(aes(x = c(0.5),
+                     xend = c(21.5), 
+                     y = c(28.5),
+                     yend = c(28.5)), color = "red" ),
+  nrow = 2,
+  heights = c(1.5,3),
+  labels = c("","C")
+)
+```
+
+#####
+
+```{r figcofactors, fig.cap="Cofactors of insect venom anaphylaxis ad their relationship to severity of a reaction",fig.width=5,fig.height = 8}
 cof_fig
 ```
+
+# Supplementrary Figures
+
+```{r treatmentmatrixother}
+#matrix_other
+```

### 2019-07-04 R/stat.R

@@ -147,6 +147,83 @@ char_to_num_levels <- function(variable){
            as.numeric())
 }
 
+recode_variables <- function(x){
+  car::recode(x,
+              recodes =
+                "'q_111'='cutaneous symptoms';
+              'q_111_angioedema'= 'angioedema';
+              'q_111_erythema_flush_v5' = 'flush';
+              'q_111_pruritus' = 'pruritus';
+              'q_111_urticaria' = 'urticaria';
+              'q_111_conjunctivitis_v5' = 'cojunctivitis';
+              'q_112' = 'gastrointestinal symptoms';
+              'q_114_dizziness'='dizziness';
+              'q_114_loss_of_consciousness'='loss of consciousness';
+              'q_112_abdominal_pain'='abdominal pain';
+              'q_113_wheezing_expiratory_distress_v5'='expiratory distress';
+              'q_522_volume'='volume iv.';
+              'q_561_hospital_admission_v6'='hospital admission';
+              'q_562_intensive_care_v6'='intensive care';
+              'q_112_nausea' = 'nausea';
+              'q_113_throat_tightness_v5' = 'throat tightness';
+              'q_114_reductions_of_alertness' = 'reduced allertness';
+              'q_112_incontinence' = 'incontinence';
+              'q_112_abdominal_pain' = 'abdominal pain';
+              'q_112_diarrhoea' = 'diarrhoea';
+              'q_113_wheezing_expiratory_distress_v5' = 'expiratory distress';
+              'q_113_rhinitis_v5' = 'rhinitis';
+              'q_114_dizziness' = 'dizziness';
+              'q_114_hypotension_collapse_v5' = 'hypotension';
+              'q_112_dysphagia_v5' = 'dysphagia';
+              'q_112_vomiting' = 'vomiting';
+              'q_113' = 'respiratory symptoms';
+              'q_113_respiratory_arrest' = 'respiratory arrest';
+              'q_113_dyspnea' = 'dyspnea';
+              'q_113_chest_tightness_v5' = 'chest tightness';
+              'q_113_cough_v5' = 'cough';
+              'q_113_change_in_voice_v5' = 'change in voice';
[…]

### 2019-07-04 R/functions.R
new package structure test completed

@@ -1,0 +1,293 @@
+rcalc <- function(var,level="yes"){
+  tab <- table(data3[,var],data3$rANA)
+  if(level %in% (data3[,var] %>% levels)){
+    pn <- tab[level,"no"]
+    pp <- tab[level,"yes"]
+    l <- level
+  } else {
+    pn <- tab[2,"no"]
+    pp <- tab[2,"yes"]
+    l <- levels(data3[,var])[2]
+  }
+  nn <- data3$rANA %>% factor %>% summary %>% {.[1]}-pn
+  np <- data3$rANA %>% factor %>% summary %>% {.[2]}-pp
+  eval <- c(pn,pp,nn,np) %>% matrix(byrow = T,ncol = 2)
+  prt <- (prop.table(eval,2)*100) %>% {signif(.,3)}
+  p <- eval %>% fisher.test() %>% .$p.value %>% roP
+  return(c(var,prt[1,1],prt[1,2],p,l))
+}
+
+
+roP<- function(x){
+  ifelse(x > 0.001,
+         round(x,3),ifelse(x==0,0,0.0001)
+  )
+}
+
+pval<- function(x){
+  ifelse(x > 0.001,
+         paste0("p = ",round(x,3)),
+         "p < 0.001")
+}
+
+f1 <- function(x){
+  names(x) <- NULL
+  return(do.call(t.test,x)$p.value %>% roP)
+}
+f2 <- function(x){
+  x %>% table(data3$rANA) %>% {.[1:2,]} %>% fisher.test() %>% {.$p.value} %>% roP
+}
+f3 <- function(x){
+  x %>%
+    lapply(function(x){(length(which(x=="yes"))/length(x)*100) %>% roP})
+}
+
+
+
+
+#' Match patients
+#'
+#' Function match_patients takes a data frame from ANAreg and otuputs matched
+#' case_id according to a given matching variable. The various grouping variables
+#' have to still be implemented. Right now only the "grouping" variable is recognized
+#'
+#' @param grouping_var has to be a binomial var  of 1s and 0s
+#' @param data is a data frame derived from ANAreg
+#' @param matching_vars variable name to which we shouldmatch the two groups
+#' @return vector of case ids that can be used for further analysis
+#'
+#'
+match_patients <- function(data,
+                           grouping_var,
+                           matching_vars,
+                           df=F,
+                           grouping_var_level = "insects"
+){
+  o <- data %>%
[…]

### 2019-07-04 R/stat.R

@@ -33,302 +33,14 @@ manual_greens <- rev(c("#E5F5E0",
                        "#238B45",
                        "#005A32"))
 
-####### Functions #############
-rcalc <- function(var,level="yes"){
-  tab <- table(data3[,var],data3$rANA)
-  if(level %in% (data3[,var] %>% levels)){
-    pn <- tab[level,"no"]
-    pp <- tab[level,"yes"]
-    l <- level
-  } else {
-    pn <- tab[2,"no"]
-    pp <- tab[2,"yes"]
-    l <- levels(data3[,var])[2]
-  }
-  nn <- data3$rANA %>% factor %>% summary %>% {.[1]}-pn
-  np <- data3$rANA %>% factor %>% summary %>% {.[2]}-pp
-  eval <- c(pn,pp,nn,np) %>% matrix(byrow = T,ncol = 2)
-  prt <- (prop.table(eval,2)*100) %>% {signif(.,3)}
-  p <- eval %>% fisher.test() %>% .$p.value %>% roP
-  return(c(var,prt[1,1],prt[1,2],p,l))
-}
+####### General Functions #############
+source("R/functions.R")
 
-roP<- function(x){
-  ifelse(x > 0.001,
-         round(x,3),ifelse(x==0,0,0.0001)
-  )
-}
-
-pval<- function(x){
-  ifelse(x > 0.001,
-         paste0("p = ",round(x,3)),
-         "p < 0.001")
-}
-
-f1 <- function(x){
-  names(x) <- NULL
-  return(do.call(t.test,x)$p.value %>% roP)
-}
-f2 <- function(x){
-  x %>% table(data3$rANA) %>% {.[1:2,]} %>% fisher.test() %>% {.$p.value} %>% roP
-}
-f3 <- function(x){
-  x %>%
-    lapply(function(x){(length(which(x=="yes"))/length(x)*100) %>% roP})
-}
-
-
-#' Match patients
-#'
-#' Function match_patients takes a data frame from ANAreg and otuputs matched
-#' case_id according to a given matching variable. The various grouping variables
-#' have to still be implemented. Right now only the "grouping" variable is recognized
-#'
-#' @param grouping_var has to be a binomial var  of 1s and 0s
-#' @param data is a data frame derived from ANAreg
-#' @param matching_vars variable name to which we shouldmatch the two groups
-#' @return vector of case ids that can be used for further analysis
-#'
-#'
-match_patients <- function(data,
-                           grouping_var,
-                           matching_vars,
[…]

### 2019-07-05 analysis/paper/paper.Rmd
results first run

@@
 
-^1^ Department of Dermatology, Venerology and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,
+^1^ Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,
 
@@
 
-The European Anaphylaxis Registry [@Grabenhenrich2016] database from March 2018 was searched for anaphylaxis cases elicited by insect's venom.
+The European Anaphylaxis Registry [@Grabenhenrich2016] database from March 2019 was searched for anaphylaxis cases elicited by insect's venom.
 The flowchart in figure \@ref(fig:flowchart) represents the detailed case-selection process.
@@
 The final database consisted of `r countries %>% summarise(sum(n))` cases of insect elicited anaphylaxis from `r countries %>% summarize(length(n))` countries.
-Severe reactions were identified based on the definition by NIAID/FAAN [@Sampson2006] and presented with significant hypoxia, hypotension, confusion, collapse and loss of consciousness, or incontinence.
+Severe reactions were identified based on the definition by NIAID/FAAN [@Sampson2006] or as Ring and Messmer Scale - grades III and IV) and presented with significant hypoxia, hypotension, confusion, collapse and loss of consciousness, or incontinence.
 We compared the frequency of various elicitors, symptoms, and factors known to increase the risk of severe anaphylaxis [@Worm2018] in both groups.
-We evaluated symptoms, managment and risk factors of insect elicited cases in comparison to other known triggers of anaphylaxis.
+We evaluated symptoms, management and risk factors of insect elicited cases in comparison to other known triggers of anaphylaxis.
 
-We saw a significant difference in the clinical features of insect anaphylaxis in children and adults.
-The younger popoulation significantly less often had concomitant conditions (i.e.
-DM, HT, malignant diseases and mastocytosis), and more often presented with atopic dermatitis, rhinitis and asthma.
+We observed a significant difference in the clinical features of insect anaphylaxis in children and adults.
+The younger population significantly less often had concomitant conditions (i.e.
+DM, HT, malignant diseases and mastocytosis), and more often presented with atopic dermatitis, rhinitis, and asthma.
 Therefore, we decided to adjust the analysis for age and sex to reduce the comparison bias.
@@
 The statistical analysis was performed in the R Statistical Package [@R].
-Simple comparison of categorical variables was performed using either Chi^2^ test or Fisher's exact test (where the number of observation in a bin was less then 10), continuous variables were analyzed using Mann-Whitney U test.
+A simple comparison of categorical variables was performed using either Chi^2^ test or Fisher's exact test (where the number of observation in a bin was less than 10), continuous variables were analyzed using Mann-Whitney U test.
 We defined statistical significance as α = 0.05.
-Data along with the analysis script can be accessed at online at https://github.com/wolass/venomanaphylaxiscompendium.
+Data along with the analysis script can be accessed online at https://github.com/wolass/venomanaphylaxiscompendium.
 
@@
 Yellow-jacket was the most prominent IVA-causing insect reported in these cases followed by bees.
-The IVA causing insects differed in European countries with hornets being more prominient on the south of europe.
-IVA was more frequent in adults and seniors when compared to children and youg adults.
+The IVA causing insects differed in European countries with hornets being more prominent in the south of Europe.
+IVA was more frequent in adults and seniors when compared to children and young adults.
 (Fig \@ref(fig:insectstime)).
-
-Cases of IVA were more frequent in older patients. Children suffered mainly from food elicited anaphylaxis. Therefore we adjusted the anlysis apropriately for age.
 
@@
 
-Patients who underent anaphylaxis due to insect venom more often experienced 
-cardiologic symptoms. Especially collapsing due to hypotension was prominently 
-more often seen in patient who underwent venom anaphylaxis. This difference was especially visible in children under the age of 18 years. (See the ade adapted symptoms plot). 
-We saw prominent differences in association of specific symptoms with eliocitor in cerain age groups.
-Children under 13 yoears of age showed association with IVIA and vomiting - whereas adults did not.
-Hypotension collapse was more prominently associated with IVIA in children whereas loss of consciousness in adults.
+IVA had a specific symptom pattern.
+Patients, who underwent IVA, more often experienced cardiologic symptoms (dizziness, reduced alertness, unconsciousness) than in other elicitors of anaphylaxis and less often showed gastrointestinal symptoms  (association measured with Cramer's V, Fig \@ref(fig:figsymptoms)).
+The difference in severe hypotension frequency was especially prominent in children under 13 (Fig.
+S\@ref(fig:hypotensionkids)).
 
-Patients with Insect Venom Allergy had a lot more cardiologic symptoms than anaphylaxis cases that were elicited through other triggers.  
-Dizziness, Loss of consciousness hypotension, collapse, Throat tightness and nausea were more often associated with the insect cases.
+The pattern of organ involvement showed similarities in gastrologic, skin and respiratory symptoms, and did not differ in the fraction of elevated baseline serum tryptase.
+Although, less atopic diseases were seen in patients undergoing IVA (Fig \@ref(fig:figsymptoms)C).
 
-<!-- Forest plot ! Risk that the symptom will occure given that the reaction was elicited by Insects or not -->
+Severe reactions were more prevalent in adults in comparison to children under 18, and in IVA cases vs other elicitors (Fig.
+\@ref(fig:figsymptoms)D).
+There were no differences in severity between insect species.
+
+## Repeated reactions
+
+```{r repreactvars}
+treact <- table(rdb$grouping,rdb$q_160_ever_react)[,1:2] %>% prop.table(1)*100
+p <- table(rdb$grouping,rdb$q_160_ever_react)[,1:2] %>% summary() %>% {.$p.value}
+```
+
+In general `r treact[1,2]`% patient with insect allergy had experienced anaphylaxis in the past which is less than if the reaction is elicited by other elcitors (`r treact[2,2]`%, p = `r p`).
+We documented `r sum(repeated_reaction_in_Reg)` patients with two documented reactions in our registry.
+Out of these `r sum(greater_same[,2])` (`r sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100`%) had Insect elicited anaphylaxis and in `r greater_same[1,2]` (`r greater_same[1,2]/sum(greater_same[,2])*100`) The following reaction was more severe than before.
+In `r greater_same[3,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was graded exactly as before.
+
+## Management
+
+Patients who underwent IVA significantly less often received adrenaline treatment than in other adrenaline cases (`r test_adrenuse1$observed %>% prop.table()`, `r test_adrenuse1$p.value %>% pval()`).
+After adjusting both groups for similar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the route of administration.
+Antihistaminic drugs, on the other hand, were given more often in cases of IVA when compared to non-IVA.
+
+IVA was most often treated with corticosteroids and antihistamines (significantly more frequent than in other anaphylaxis cases).
+On the other hand adrenaline, beta-2 mimetics and oxygen were given more often in non-IVA.
+
+We discovered clusters of symptoms and therapy modes.
+In IVA patients urticaria, and cardiologic symptoms (cardiac arrest, hypotension, loss of consciousness), as well as respiratory distress, were treated more similarly than gastroenterologic symptoms (Fig.
+\@ref(fig:adrenalineuse)C).
+Therapy that was usually used together was dependent on the symptom profile and consisted of AAI, i.v.
+adrenaline in multiple doses, pure oxygen inhalation, an initial dose of antihistamines, and beta 2 agonists.
+Corticosteroids, volume replacement and i.v. beta2 agonists formed another therapy mode.
+
+The most differences in the therapy of IVA vs.
+other forms of anaphylaxis was observed in the frequency of inhaled beta2-agonists and antihistamines (probably because these are not available for the IVA patients, but atopic patients use them regularly) (Fig.
+\@ref(fig:adrenalineuse)B).
 
 ## Co-factors
+
+Although the physical exercise was more often related to IVA than other triggers of anaphylaxis - it was not predictive of the severity of a reaction in these patients.
+
+Concomitant use of ACE-I did not correlate with reaction severity measured in ANAscore and VAS scales.
+It was however more often associated with cardiac arrests in both IVA and other-triggers anaphylaxis (Fig.
+S\@ref(fig:acei)).
+
+Beta-blockers use was associated with higher severity scores in VAS and ANAscore and with cardiac arrest, chest pain.
+Especially arrhythmia was more frequently reported in patients undergoing IVA with concomitant beta-blockers.
+
+Mastocytosis was increasing the risk of cardiac arrest in patients undergoing IVA significantly more in patients undergoing anaphylaxis due to other triggers (Fig S\@ref(fig:cardiacmasto)).
+
+Baseline serum tryptase correlated with the severity of reactions and especially in the Ring and Messmer scale was more prominently increasing risk of severe anaphylaxis in IVA patients than in other cases (Fig.
+\@ref(fig:figcofactors)B and S\@ref(fig:tryp)).
+In the VAS scale, this did not show as significant, but it did in all the other severity scales.
+We decided to use Tryptase cut off value of 8 ng/ml (Fig.
+S\@ref(fig:tryp2)).
+We saw that cases with cardiac arrest were associated with increased tryptase over 8 ng/ml and this proportion was higher in IVA when compared to other elicitors.
+Loss of consciousness was also associated with increased tryptase levels but only in patients with IVA (Fig.
+\@ref(fig:figcofactors)C).
+
+We compared the Hymenoptera species which was responsible for triggering IVA according to the severity of the reaction.
+Patients matched according to sex and age has similar reactions to bees and yellow-jackets (in Brown and Ring and Messmer scales).
+In ANAscore - yellow jacket stings were linked to a slightly more severe reaction.
+This data, however, is not comparable to the refractory anaphylaxis cases as refractory anaphylaxis seems to be a distinct phenotype.
+
+<!-- ### Mastocytosis -->
+
+<!-- ### Older Age -->
+
+<!-- ### ACE - I -->
+
+<!-- ### previous reactions (severe) -->
+
+<!-- ### Male sex -->
+
+<!-- ### Beta blockers and CARDIOSELECTIVE BLOCKERS -->
+
+<!-- ### Evaluate the pediatric population separately -->
+
+<!-- ### Number of SAR during VIT compare to other hypos -->
+<!-- #### Compare bee sit to wasp sit for AEs -->
+<!-- #### correlate the tryptase value with SIT types and severity -->
+
+# Discussion
+
+Cardiologic symptoms and hypotensive collapse might be associated with venom anaphylaxis due to the vaso-vagal reflex.
+As seen in patients undergoing blood sampling needles (and probably also insect stings) may elicit a hypotensive response due to extreme emotional distress with bradycardia, limiting the blood flow to the brain and causing syncope.
+This mechanism may contribute to the ongoing histamine-induced vasodilatation and may worsen the symptoms of an allergic reaction.
+
+# Conclusion
+
+# Acknowledgements
+
+<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
+#####
+
+# References 
+<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
+<div id="refs"></div>
+
+#####
+
+# Figures
+
+```{r flow,fig.width=6,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale to include cases in our analysis.
+B,C,D: Sex, age, and severity distribution was matched in cases in both groups to allow for comparable results between IVA and non_IVA cases"}
+ggarrange(ggplot()+
+            background_image(png::readPNG("../figures/flow.png")),
+          lower_panel,
+          ncol = 1,
+          nrow = 2,
+          heights = c(1,0.4),
+          labels = c("A","")
+)
+```
+
+#####
+
+```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occurred.
+Less frequent insects were grouped together as 'other'.
+B: The proportion of insect elicited cases to cases elicited by other triggers considering the patient's age.
+C: Geographical differences in the most common elicitors of IVA.
+Countries which reported less than 10 cases of anaphylaxis due to insect venom were not illustrated in this figure.",fig.height=4,fig.width = 7}
+fig_basic
+```
+
+#####
+
+```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors",fig.width=9.5,fig.height = 8}
+fig_symptoms
+```
+
+#####
+
+```{r adrenalineuse, fig.cap = "Therapy modes in patients with IVA compared to oter elicitors. A)  B)",fig.height=10,fig.width = 7}
+
+ggpubr::ggarrange(
+  ggarrange(
+    plotManagement,
+    rf_ggplot_treatment,
+    labels = c("A","B")
+  ),
+  #lower_panel,
+  matrix_venom$p+
+    labs (x = "", y = "")+
+    geom_segment(aes(x = c(8.5),
+                     xend = c(8.5), 
+                     y = c(0.5),
+                     yend = c(39.5)),color = "red")+
+    geom_segment(aes(x = c(0.5),
+                     xend = c(21.5), 
+                     y = c(28.5),
+                     yend = c(28.5)), color = "red" ),
+  nrow = 2,
+  heights = c(1.5,3),
+  labels = c("","C")
+)
+```
+
+#####
+
+```{r figcofactors, fig.cap="Cofactors of insect venom anaphylaxis ad their relationship to severity of a reaction",fig.width=5,fig.height = 8}
+cof_fig
+```
+
+# Supplementary Figures
+
+```{r treatmentmatrixother}
+#matrix_other
+```
+
+```{r hypotensionkids, fig.cap="Hypotension in IVA of kids"}
+hypotensionPlot
+```
+
 ```{r}
+ggarrange(dt,
+          dt3 %>%
+ggbarplot(x = "q_610_sit_prior_v5",
+          y = "n",
+          fill = "severity_brown",
+          position = position_fill())+
+  labs(fill="FAAN"))
+```
+
+```{r cofactorscomparison}
 testInsectsbinomialTab$cofactors[,-7] %>% knitr::kable()
@@
 
-### ACE-I
-Ruef wrote in 2009 that the use of ACE I increases the risk of a severe reaction.
+```{r severityinsectspecies}
+yj_bee %>%
+  ggbarplot(x = "grouping",
+            y = "n",
+            fill = "d_severity_rm",
+            position = position_fill())
 
-```{r}
-data$q_423_ace <- yesonly(data$q_423_ace)
+```
 
-wilcox.test(q_116_VAS_v7~q_423_ace,data = data)
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_423_ace,q_116_VAS_v7))+
-  geom_violin()
+```{r acei}
 
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-group_by(q_423_ace) %>% 
-  summarise(medianVAS=median(q_116_VAS_v7,na.rm=T),
-            IQRVAS = IQR(q_116_VAS_v7,na.rm=T))
-
-wilcox.test(ANAscore~q_423_ace,data = data %>% 
-  filter(d_elicitor_gr5=="insects") 
-)
-
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_423_ace,ANAscore))+
-  geom_violin()
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
- group_by(q_423_ace) %>% 
-  summarise(medianANAscore=median(ANAscore,na.rm=T),
-            IQRANAscore = IQR(ANAscore,na.rm=T))
-
-cowplot::plot_grid(
-data %>% 
+ggarrange(
+  plot_ace_cardiacs,
+ggarrange(rdb %>% 
   filter(d_elicitor_gr5=="insects") %>% 
@@
   geom_boxplot(),
-data %>% 
+rdb %>% 
   filter(d_elicitor_gr5=="insects") %>% 
@@
   geom_boxplot()
+),
+nrow = 2,
+ncol = 1
 )
@@
 
-### Beta blockers 
-cardioselectivity??????
-```{r}
-data$q_423_beta <- yesonly(data$q_423_beta)
+```{r betas}
+rdb_temp <- rdb
+rdb_temp$q_423_beta <- yesonly(rdb_temp$q_423_beta)
 
-wilcox.test(q_116_VAS_v7~q_423_beta,data = data)
-data %>% 
+wilcox.test(q_116_VAS_v7~q_423_beta,data = rdb_temp)
+rdb_temp %>% 
   filter(d_elicitor_gr5=="insects") %>% 
@@
 
-data %>% 
+rdb_temp %>% 
   filter(d_elicitor_gr5=="insects") %>% 
@@
 
-wilcox.test(ANAscore~q_423_beta,data = data %>% 
+wilcox.test(ANAscore~q_423_beta,data = rdb_temp %>% 
   filter(d_elicitor_gr5=="insects") 
@@
   geom_violin()
-data %>% 
+rdb_temp %>% 
   filter(d_elicitor_gr5=="insects") %>% 
@@
 
-cowplot::plot_grid(
-data %>% 
+ggarrange(
+  plot_beta_cardiacs,
+  ggarrange(
+    rdb_temp %>% 
   filter(d_elicitor_gr5=="insects") %>% 
@@
   geom_boxplot(),
-data %>% 
+rdb_temp %>% 
   filter(d_elicitor_gr5=="insects") %>% 
@@
   geom_boxplot()
+  ),
+nrow =2,
+ncol = 1
 )
@@
 
-### mast cell burden - mastocytosis
-
-```{r}
-data$q_410_masto_cur %<>%  yesonly()
-
-wilcox.test(q_116_VAS_v7~q_410_masto_cur,data = data)
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_410_masto_cur,q_116_VAS_v7))+
-  geom_violin()
-
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-group_by(q_410_masto_cur) %>% 
-  summarise(medianVAS=median(q_116_VAS_v7,na.rm=T),
-            IQRVAS = IQR(q_116_VAS_v7,na.rm=T))
-
-wilcox.test(ANAscore~q_410_masto_cur,data = data %>% 
-  filter(d_elicitor_gr5=="insects") 
-)
-
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_410_masto_cur,ANAscore))+
-  geom_violin()
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
- group_by(q_410_masto_cur) %>% 
-  summarise(medianANAscore=median(ANAscore,na.rm=T),
-            IQRANAscore = IQR(ANAscore,na.rm=T))
-
-cowplot::plot_grid(
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_410_masto_cur,q_116_VAS_v7))+
-  geom_boxplot(),
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_410_masto_cur,ANAscore))+
-  geom_boxplot()
-)
-
+```{r cardiacmasto}
+cardiac_beta_plots[[6]]+
+              labs(x = "cardiac arrest")
 ```
 
-```{r}
-cowplot::plot_grid(
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_212_tryptase_value_v5,q_116_VAS_v7))+
-  geom_point()+
-  scale_x_log10(),
-data %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_212_tryptase_value_v5,ANAscore))+
-  geom_point()+
-  scale_x_log10()
-)
-
-cowplot::plot_grid(
-data %>% 
+```{r tryp, fig.cap="Severity of anaphylaxis is influenced by baseline serum tryptase levels with the cut off of 8 ng/ml in IVA (top row) vs other anaphylaxis cases",fig.width = 7}
+ggarrange(
+rdb %>% 
   filter(d_elicitor_gr5=="insects") %>% 
@@
   geom_boxplot(),
-data %>% 
+rdb %>% 
   filter(d_elicitor_gr5=="insects",
@@
                   "low",
-                  "high")) %>% 
+                  "high"),
+         d_severity_rm = factor(d_severity_rm)) %>% 
   group_by(d_severity_rm,typ_cat) %>% 
@@
   geom_bar(stat="identity",position= "fill"),
-data %>% 
+rdb %>%
   filter(d_elicitor_gr5!="insects") %>%
@@
   geom_boxplot(),
-data %>% 
+rdb %>%
   filter(d_elicitor_gr5!="insects") %>%
@@
   geom_boxplot(),
-data %>% 
+rdb %>%
   filter(d_elicitor_gr5!="insects",
@@
                   "low",
-                  "high")) %>% 
+                  "high"),
+         d_severity_rm = factor(d_severity_rm)) %>%
   group_by(d_severity_rm,typ_cat) %>%
@@
   ggplot(aes(typ_cat,n,fill=d_severity_rm))+
-  geom_bar(stat="identity",position= "fill")
-)
+  geom_bar(stat="identity",position= "fill"),
+ncol =3 , nrow = 2,
+widths = c(1,1.1,2))
 
@@
 
-Here put the mastocytosis and baseline serum tryptase as the evaluation of severe risk
-Here it is crucial to compare to even these reactions that had nearly no reactions
-
-### Model the risk of severity of a severe reaction in IVA
-
-USe the model from our previous paper
-
-## Managment
-
-### Adrenaline is less often used in patients with IVA compared to other anaphylaxis cases
-
-```{r adrenuse, fig.cap="**Therapy of IVA cases.** A: General managment methods in IVA and non-IVA. B: Reasons for not administering adrenaline in both IVA and non-IVA cases. C,D,E: "}
-#fig_adrenuse1
-```
-
-Patients who underwent IVA significantly less often recived adrenaline treatment than in other adrenaline cases (`r test_adrenuse1$observed %>% prop.table()`, `r test_adrenuse1$p.value %>% pval()`).
-After adjusting both groups for simmilar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the route of administration.
-Antihistaminic drugs on the other had were given more often in cases of IVA when compared to non-IVA.
-
-IVA was most often treated with corticosteroids and antihistamines (significantly more frequent than in other anaphylaxis cases).
-On the other hand adrenaline, beta-2 mimetics and oxygen were given more often in non-IVA.
-
-## Diagnostics
-
-## Repeated reactions
-
-```{r}
-treact <- table(rdb$grouping,rdb$q_160_ever_react)[,1:2] %>% prop.table(1)*100
-p <- table(rdb$grouping,rdb$q_160_ever_react)[,1:2] %>% summary() %>% {.$p.value}
-```
-In general `r treact[1,2]`% patient with insect allergy had experienced anaphylaxis in the past which is less than if the reaction is elicited by other elcitors (`r treact[2,2]`%, p = `r p`).
-We documented `r sum(repeated_reaction_in_Reg)` patients with two documented reactions in our registry.
-Out of these `r sum(greter_same)` (`r sum(greter_same)/sum(repeated_reaction_in_Reg)*100`%) had Insect elicited anaphylaxis and in `r greter_same[1]` (`r greter_same[1]/sum(greter_same)*100`) The follwoeing reaction was more severe than before.
-
-### Do we know how mnay beekeepers were in our dataset?
-We were unable to estimate the fraction of cases that were elicited in bee-keepers as this information is not asked in the questionnaire.
-
-### Hypothesis: Wasp stings are more severe than bee stings
-We previously saw that refractory anaphylaxis cases come from bees. Here we see no difference in the severity of cases elicited by bees and wasps (in the RM).
-
-```{r,fig.height=8}
+```{r yjbee, fig.height=8, fig.width = 7, fig.cap="Models of severity and the influence of jellow jacket vs bees"}
 require(broom)
@@
 }
-
+require(broom)
 temp <- lapply(1:5,do_cof_tab) %>% do.call(what = rbind) %>%
@@
   #geom_line(aes(group=outcome))
+```
+
+```{r modelfitness, fig.cap="Comparison of modelling to different severity scales."}
 
@@
 
-Based on the above graphs of patients matched according to sex and age we cannot say that the severity of their reaction differ between yellow-jackets and bees in any of the models except ANAscore - where yellow jacket were linked to  a slightly more severe reaction.
-
-But in the ANAscore variable it shows up to be significant !
-Yellow-jackets stings are more severe than the bee stings.
-This is however very hard to compare to refractory cases.
-Maybe we can use the refractory case identifier to say in which group we see more refractory cases?
-
-### Increased BTSC develops in the normal range <8
-
-```{r,fig.height=8}
+```{r tryp2,fig.height=8, fig.cap = "tryptase cut off value plot"}
 output_severity <- lapply(possible_outcomes,
@@
 ```
-
-Based on the above we see that the severity of anaphylaxis significantly depends on tryptase concentration (here under 8 µg/ml). Only in the VAS scale this did not sho as significant.
-There might be problems with the vas scale...
-
-### Mastocytosis
-
-### Older Age
-
-### ACE - I
-
-### previous reactions (severe)
-
-### Male sex
-
-### Beta blockers and CARDIOSELECTIVE BLOCKERS
-
-### Evaluate the pediatric population separately
-
-### Number of SAR during VIT compare to other hypos
-#### Compare bee sit to wasp sit for AEs
-#### correlate the tryptase value with SIT types and severity
-### 
-they used generalized aditive models (package GAM?)
-Also ROC curves for confirmation
-
-# Discussion
-
-Cardiologic symptoms and hypotensive collapse might be associated with vanom anaphylaxis due to the vaso-vagal reflex.
-As seen in patients undrgoing blood sampling needles (and probably also insect stings) may elicit a hypotensive response due to extreme emotional distress with bradycardia, limiting the bloodflow to the brain and causing a syncope.
-This mechanism may contribute to the ongoing histamin-induced vasodialation and may worsen the symptoms of an allergic reaction.
-
-# Conclusion
-
-# Acknowledgements
-
-<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
-#####
-
-# References 
-<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
-<div id="refs"></div>
-
-#####
-
-# Figures
-
-```{r flow,fig.width=6,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale to include cases in our analysis.
-B,C,D: Sex, age, and severity distribution was matched in cases in both groups to allow for comparable results between IVA and non_IVA cases"}
-ggarrange(ggplot()+
-            background_image(png::readPNG("../figures/flow.png")),
-          lower_panel,
-          ncol = 1,
-          nrow = 2,
-          heights = c(1,0.4),
-          labels = c("A","")
-)
-```
-
-#####
-
-```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occured.
-Less frequent insects were groupeped together as 'other'.
-B: The proportion of insect elicited cases to ceses elicited by other triggers considering patient's age.
-C: Geographical differences in the most common eliciotors of IVA.
-Countries which reported less than 10 cases of anaphylaxis due to insect venom were not illustrated in this figure.",fig.height=4,fig.width = 7}
-fig_basic
-```
-
-#####
-
-```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors",fig.width=9.5,fig.height = 8}
-fig_symptoms
-```
-
-#####
-
-```{r adrenalineuse, fig.cap = "Therapy modes in patients with IVA compared to oter elicitors. A)  B)",fig.height=10,fig.width = 7}
-
-ggpubr::ggarrange(
-  ggarrange(
-    plotManagement,
-    rf_ggplot_treatment,
-    labels = c("A","B")
-  ),
-  #lower_panel,
-  matrix_venom$p+
-    labs (x = "", y = "")+
-    geom_segment(aes(x = c(8.5),
-                     xend = c(8.5), 
-                     y = c(0.5),
-                     yend = c(39.5)),color = "red")+
-    geom_segment(aes(x = c(0.5),
-                     xend = c(21.5), 
-                     y = c(28.5),
-                     yend = c(28.5)), color = "red" ),
-  nrow = 2,
-  heights = c(1.5,3),
-  labels = c("","C")
-)
-```
-
-#####
-
-```{r figcofactors, fig.cap="Cofactors of insect venom anaphylaxis ad their relationship to severity of a reaction",fig.width=5,fig.height = 8}
-cof_fig
-```
-
-# Supplementrary Figures
-
-```{r treatmentmatrixother}
-#matrix_other
-```

### 2019-07-05 R/data_cleaning.R

@@ -1,0 +1,3 @@
+data_cleaninig <- function(){
+
+}

### 2019-07-05 R/functions.R

@@ -291,3 +291,461 @@ severity_brown <- function(data){
   severity_brown <- as.factor(severity_brown)
   return(severity_brown)
 }
+
+#'This function makes a chisq test on selected variables given a grouping variable.
+#'It outputs a p.value
+#'
+#'
+#'
+funChi <-function(var,grouping_var){
+  var %>%
+    data.frame(gr = grouping_var) %>%
+    {chisq.test(.[,1],.[,2])$p.value}
+}
+
+funN1 <- function(var, grouping_var){
+  var %>%
+    data.frame(grouping_var) %>% table %>%
+    {.[2,1]}
+  #group_by(grouping_var) %>%
+  #summarise(count=n())
+}
+funN2 <- function(var, grouping_var){
+  var %>%
+    data.frame(grouping_var) %>% table %>%
+    {.[2,2]}
+  #group_by(grouping_var) %>%
+  #summarise(count=n())
+}
+funF1 <- function(var, grouping_var){
+  var %>%
+    data.frame(grouping_var) %>%
+    table %>% {.[1:2,]} %>%
+    prop.table(2) %>%
+    {.[2,1]}
+  #group_by(grouping_var) %>%
+  #summarise(count=n())
+}
+
+funF2 <- function(var, grouping_var){
+  var %>%
+    data.frame(grouping_var) %>%
+    table %>% {.[1:2,]} %>%
+    prop.table(2) %>%
+    {.[2,2]}
+  #group_by(grouping_var) %>%
+  #summarise(count=n())
+}
+funC <- function(var, grouping_var){
+  var %>%
+    data.frame(grouping_var) %>%
+    table %>% {.[1:2,]} %>%
+    assocstats() %>% {.$cramer}
+}
+
+#' This function takes grouping vector and a DATA FRame ONLY OF  variables to test
+#'
+ChiF <- function(variablesDF,grouping){
+  # C <- lapply(variablesDF,funC,grouping)
+  data.frame(variables = names(variablesDF),
+             pval = lapply(variablesDF,funChi,grouping) %>% unlist,
+             counts_1 = lapply(variablesDF,funN1,grouping) %>%unlist,
+             counts_2 = lapply(variablesDF,funN2,grouping) %>%unlist,
+             fraq_1 = lapply(variablesDF,funF1,grouping) %>%unlist,
+             fraq_2 = lapply(variablesDF,funF2,grouping) %>%unlist,
[…]

### 2019-07-05 R/make_all_forest_plots.R

@@ -1,0 +1,141 @@
+### make all forest plots
+make_all_forest_plots <- function(){
+
+
+test1 <- makeDF(age_sex_matched,
+                c( "q_410_cur",
+                   "q_410_asthma_cur",
+                   "q_410_masto_cur",
+                   "q_410_cardio_cur",
+                   "q_423_beta",
+                   "q_423_ace",
+                   "q_4211_exercise",
+                   "q_422_stress",
+                   "q_423_asa"),
+                "grouping",
+                "severity_brown")
+
+test1[1] <- c(NA,"Concomitant disease",
+              NA,NA,"Asthma",
+              NA,NA,"Mastocytosis",
+              NA,NA,"Cardiologic disease",
+              NA,NA,"β-blockers",
+              NA,NA,"ACE-I",
+              NA,NA,"Exercise",
+              NA,NA,"Stress",
+              NA,NA,"ASA",NA,NA)
+test1 <- test1[-c(2:4,20:25),]
+
+
+test2 <- makeDF(age_sex_matched,
+                c( "q_410_cur",
+                   "q_410_masto_cur",
+                   "q_410_asthma_cur",
+                   "q_410_cardio_cur",
+                   "q_423_beta",
+                   "q_423_ace",
+                   "q_4211_exercise",
+                   "q_422_stress",
+                   "q_423_asa"),
+                "grouping",
+                "d_severity_rmr")
+
+test2[1] <- c(NA,"Concomitant disease",
+              NA,NA,"Mastocytosis",
+              NA,NA,"Asthma",
+              NA,NA,"Cardiologic disease",
+              NA,NA,"β-blockers",
+              NA,NA,"ACE-I",
+              NA,NA,"Exercise",
+              NA,NA,"Stress",
+              NA,NA,"ASA",NA,NA)
+test2 <- test2[-c(2:4,20:25),]
+
+
+#or_plot(test1)
+
+
+png("analysis/figures/figForestfinal.png",
+    height = 800,
+    width = 1000,
+    #res = 300,
+    pointsize = 40,
+    units="px")
+forestplot(labeltext = makeTableText(test1),
+           mean = test1[,"mean"] %>% unlist,
+           lower = test1[,"lower"] %>% unlist,
+           upper = test1[,"upper"] %>% unlist,
+           new_page = TRUE,
+           is.summary=c(T,rep(F,length(test1[,1])-1)),
+           clip=c(0.2,4),
+           xlog=TRUE,
[…]

### 2019-07-05 R/make_flow.R

@@ -1,0 +1,72 @@
+### mAKE THE DIAGRAMM FOR THE fLOW CHART
+
+make_flow <- function(){
+require(DiagrammeRsvg)
+require(DiagrammeR)
+
+# Create a node data frame
+nodes <-
+  create_node_df(
+    n = 7,
+    type = "a",
+    label= c(paste0("All cases in the\nEuropean Anaphylaxis Registry\n",length(data4$b_submitdate)),
+             paste0("Formal anaphylaxis definition met\n",
+                    length(which(data4$reaction_type_brown=="anaphylaxis"))),
+             paste0("Insects as elicitors\n",
+                    length(which(data4$d_elicitor_gr5== "insects" &
+                                   data4$reaction_type_brown=="anaphylaxis"))),
+             paste0("Other elicitors\n",
+                    length(which(data4$d_elicitor_gr5!= "insects"&
+                                   data4$reaction_type_brown=="anaphylaxis"))),
+             paste(paste0("Yellow-jackets = ",
+                          length(which(data4$q_340_insects== "yellow jacket"&
+                                         data4$reaction_type_brown=="anaphylaxis"))),
+                   paste0("Bees = ",
+                          length(which(data4$q_340_insects== "bee"&
+                                         data4$reaction_type_brown=="anaphylaxis"))),
+                   paste0("Hornets = ",
+                          length(which(data4$q_340_insects== "hornet"&
+                                         data4$reaction_type_brown=="anaphylaxis"))),
+                   paste0("Other insects = ",
+                          length(which(data4$d_insect_gr4=="other" &
+                                         data4$reaction_type_brown=="anaphylaxis"))),
+                   sep="\n"
+             ),
+             paste0("Elicitor unknown\n",
+                    length(which(data4$d_elicitor_gr5=="unkown" &
+                                   data4$reaction_type_brown=="anaphylaxis"))),
+             paste0("Other known\nelicitors\n",
+                    length(which(data4$d_elicitor_gr5!="insects"&
+                                   data4$d_elicitor_gr5!="unkown" &
+                                   data4$reaction_type_brown=="anaphylaxis")))
+    ),
+
+    #color = c("red", "green",
+    #          "grey", "blue"),
+    #value = c(3.5, 2.6, 9.4, 2.7),
+    shape= "rectangle",
[…]

### 2019-07-05 R/stat.R

@@ -1,4 +1,4 @@
-#Setup
+#####Setup####
 require(dplyr)
 require(magrittr)
 require(ggplot2)
@@ -12,6 +12,10 @@ library(tibble)
 library(scales)
 require(ggpubr)
 require(rlang)
+library(forcats)
+require(fmsb)
+require(purrr)
+
 
 #### GET THE DATA ########
 # Note the path that we need to use to access our data files when rendering this document
@@ -23,6 +27,10 @@ require(rlang)
 
 #load('../RefractoryAnaOrg/analysis/data/raw_data/data4.R')
 load("data.R") # here the anascore is done using the severity_analysis R script
+
+# update the anascore according to nora consensus
+#source("R/anascore_vas.R")
+
 data4 <-data
 
 manual_greens <- rev(c("#E5F5E0",
@@ -47,6 +55,7 @@ other_insects <- c("horse fly",
                    "bumble-bee",
                    "mosquito",
                    "insects")
+
 data4$d_insect_gr4 <- as.character(data4$q_340_insects)
 data4$d_insect_gr4[which(data4$q_340_insects%in%other_insects)] <- "other"
 data4$d_insect_gr4 %<>% factor
@@ -59,183 +68,6 @@ data4%<>%
                                           NA)
   ))
 
-##### DIAGRAM ##############
-library(DiagrammeRsvg)
-require(DiagrammeR)
-
-# Create a node data frame
-nodes <-
-  create_node_df(
-    n = 7,
-    type = "a",
-    label= c(paste0("All cases in the\nEuropean Anaphylaxis Registry\n",length(data4$b_submitdate)),
-             paste0("Formal anaphylaxis definition met\n",
-                    length(which(data4$reaction_type_brown=="anaphylaxis"))),
-             paste0("Insects as elicitors\n",
-                    length(which(data4$d_elicitor_gr5== "insects" &
-                                   data4$reaction_type_brown=="anaphylaxis"))),
-             paste0("Other elicitors\n",
-                    length(which(data4$d_elicitor_gr5!= "insects"&
-                                   data4$reaction_type_brown=="anaphylaxis"))),
-             paste(paste0("Yellow-jackets = ",
-                    length(which(data4$q_340_insects== "yellow jacket"&
-                                   data4$reaction_type_brown=="anaphylaxis"))),
-                paste0("Bees = ",
-                    length(which(data4$q_340_insects== "bee"&
-                                   data4$reaction_type_brown=="anaphylaxis"))),
-                paste0("Hornets = ",
[…]

### 2019-07-05 analysis/paper/paper.Rmd
results hotfix

@@
 Intramuscular adrenaline (as a first-line therapy) was administered significantly less often in insect venom elicited cases (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(d_522_adren_agg) %>% summarize(n=n()) %>% {.[2,2]/(.[1,2])*100} %>% round(1)`% vs `r rdb %>% filter(is.na(q_340_insects)) %>% group_by(d_522_adren_agg) %>% summarize(n=n()) %>% {.[2,2]/(.[1,2])*100} %>% round(1)`, p < 0.0001).
-The mortality rate in insect anaphylaxis was comparable (`r `%) to other cases (0.295%, p = 0.174).
+The mortality rate in insect anaphylaxis was comparable to other cases (0.295%, p = 0.174).
 
@@
 IVA was more frequent in adults and seniors when compared to children and young adults.
-(Fig \@ref(fig:insectstime)).
+(Fig \@ref(fig:insectstime)B).
 
@@
 
-In general `r treact[1,2]`% patient with insect allergy had experienced anaphylaxis in the past which is less than if the reaction is elicited by other elcitors (`r treact[2,2]`%, p = `r p`).
+In general `r treact[1,2] %>% round(1)`% patient with insect allergy had experienced anaphylaxis in the past which is less than if the reaction is elicited by other elcitors (`r treact[2,2]%>% round(1)`%, `r p %>% pval()`).
 We documented `r sum(repeated_reaction_in_Reg)` patients with two documented reactions in our registry.
-Out of these `r sum(greater_same[,2])` (`r sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100`%) had Insect elicited anaphylaxis and in `r greater_same[1,2]` (`r greater_same[1,2]/sum(greater_same[,2])*100`) The following reaction was more severe than before.
+Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had Insect elicited anaphylaxis and in `r greater_same[1,2]` (`r greater_same[1,2]/sum(greater_same[,2])*100`) The following reaction was more severe than before.
 In `r greater_same[3,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was graded exactly as before.
@@
 
-Patients who underwent IVA significantly less often received adrenaline treatment than in other adrenaline cases (`r test_adrenuse1$observed %>% prop.table()`, `r test_adrenuse1$p.value %>% pval()`).
+Patients who underwent IVA significantly less often received adrenaline treatment than in other adrenaline cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
 After adjusting both groups for similar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the route of administration.
@@
 \@ref(fig:adrenalineuse)C).
-Therapy that was usually used together was dependent on the symptom profile and consisted of AAI, i.v.
+Therapy was dependent on the symptom profile and consisted of AAI, i.v.
 adrenaline in multiple doses, pure oxygen inhalation, an initial dose of antihistamines, and beta 2 agonists.
@@
 
-```{r acei}
+```{r acei, fig.cap="Ace inhibitors and cardiologic symptoms"}
 
@@
 
-```{r betas}
+```{r betas, fig.cap = "Beta blockers and cardiologic symptoms"}
 rdb_temp <- rdb
@@
 
-```{r cardiacmasto}
-cardiac_beta_plots[[6]]+
+```{r cardiacmasto, fig.cap = "MAsotcytosis and cardiologic symptoms" }
+cardiac_masto_plots[[6]]+
               labs(x = "cardiac arrest")
@@
 
-```{r tryp2,fig.height=8, fig.cap = "tryptase cut off value plot"}
+```{r tryp2,fig.height=8, fig.width = 7, fig.cap = "tryptase cut off value plot"}
 output_severity <- lapply(possible_outcomes,

### 2019-07-05 R/functions.R

@@ -163,7 +163,7 @@ recode_variables <- function(x){
               'q_522_adren_iv'   = 'adrenaline i.v. 1st';
               'q_552_adren_iv_v5'   = 'adrenaline i.v. 2nd';
               'q_521_other_v5'   = 'other drugs 1st';
-              'q_521_antih_v5'   = 'antihistmines 1st';
+              'q_521_antih_v5'   = 'antihistamines 1st';
               'q_522_adren_im'   = 'adrenaline i.m. 1st';
               'q_522_antih_oral'   = 'antihistamines p.o. 1st';
               'q_552_adren_im_v5'   = 'adrenaline i.m. 2nd';

### 2019-07-17 analysis/paper/paper.Rmd
figs 1-4 in B-and-W

@@
 
-```{r flow,fig.width=6,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale to include cases in our analysis.
-B,C,D: Sex, age, and severity distribution was matched in cases in both groups to allow for comparable results between IVA and non_IVA cases"}
+```{r flow,fig.width=6,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale for including registry cases in the final analysis.
+B,C,D: Age, sex, and severity distribution was matched in cases in both groups to allow for comparable results between IVA and non-IVA cases.
+Two age-subsets of patients could be recognized based on the density plot of age (B)."}
 ggarrange(ggplot()+
@@
 ```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occurred.
-Less frequent insects were grouped together as 'other'.
-B: The proportion of insect elicited cases to cases elicited by other triggers considering the patient's age.
+Less frequent insect species were grouped together as 'other'.
+B: The density distribution of IVA cases to cases elicited by other triggers considering the patient's age.
 C: Geographical differences in the most common elicitors of IVA.
-Countries which reported less than 10 cases of anaphylaxis due to insect venom were not illustrated in this figure.",fig.height=4,fig.width = 7}
+Countries which reported less than 10 IVA cases were not illustrated in this figure.", fig.height=4, fig.width = 7}
 fig_basic
@@
 
-```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors",fig.width=9.5,fig.height = 8}
+```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors.
+A: Proportional presentation of specific reaction symptoms.
+B: High-level overview of involved organ systems and selected cofactors in form of a radar plot.", fig.width=8.5, fig.height = 3.2}
 fig_symptoms
+```
+
+#####
+
+```{r figcofactors, fig.cap="Cofactors of insect venom anaphylaxis. A: Odds ratios of eliciting severe anaphylaxis. B: Proportional ",fig.width=5.6,fig.height = 6}
+cof_fig
 ```
@@
 
-#####
-
-```{r figcofactors, fig.cap="Cofactors of insect venom anaphylaxis ad their relationship to severity of a reaction",fig.width=5,fig.height = 8}
-cof_fig
-```
-
 # Supplementary Figures
@@
 #matrix_other
+```
+
+```{r crammersv}
+tests_matched %>%
+    mutate(organ = factor(ifelse(grepl(variableName,pattern = "111"),"skin",
+                          ifelse(grepl(variableName,pattern = "112"),"gastro",
+                                 ifelse(grepl(variableName,pattern = "113"),"respiratory",
+                                        ifelse(grepl(variableName,pattern = "114"),"cardio",NA)))),
+                          levels = c("respiratory","cardio","gastro","skin"))
+           ) %>%
+    filter(!is.na(Cramer),
+           section == "symptoms",
+           Cramer > 0.11,
+           !(variableName %in% c("q_114","q_143_fatal_adre_v5", "q_112_abdominal_pain")))%>%
+    mutate(variableName = car::recode(variableName,
+                                      "'q_112_nausea' = 'nausea';
+                                      'q_114_loss_of_consciousness' = 'unconsciousness';
+                                      'q_113_throat_tightness_v5' = 'throat tightness';
+                                      'q_114_reductions_of_alertness' = 'reduced allertness';
+                                      'q_112_incontinence' = 'incontinence';
+                                      'q_112_abdominal_pain' = 'abdominal pain';
+                                      'q_112_diarrhoea' = 'diarrhoea';
+                                      'q_113_wheezing_expiratory_distress_v5' = 'expiratory distress';
+                                      'q_113_rhinitis_v5' = 'rhinitis';
+                                      'q_114_dizziness' = 'dizziness';
+                                      'q_114_hypotension_collapse_v5' = 'hypotension';
+                                      'q_111' = 'any skin symptoms'"
+    )) %>% #{.[,1]}
+    ggdotchart(x = "variableName",
+               y = "Cramer",
+               color = "organ",
+               palette = "lancet",
+               sorting = "descending",
+               add = "segments",
+               add.params = list(color = "organ"),
+               rotate = T,
+               #group = "organ",
+               dot.size = 5,
+               size = 2,
+               #label = "Cramer"
+               ggtheme = theme_pubr())+
+    labs(x = "",y = "Cramer's V",color = "")+
+    theme(legend.justification = c(-3,0))
+
+```
+
+```{r}
+severity_joined_brown %>%
+    ggbarplot(x = "grouping",
+              y = "n",
+              fill = "d_severity_rm",
+              position = position_fill(reverse = T),
+              color = "d_severity_rm",
+              palette = "lancet")+
+    facet_grid(.~subset,scale = "free_x",space = "free_x")+
+    labs(fill = "Severity")+
+    theme(axis.title.x =  element_blank(),
+          axis.text.x = element_text(angle=20,hjust = 1),
+          legend.position = "right"
+    )+
+    annotate("text",x = 1.5, y = 0.5, label = c("*","*",""),
+             size =12)+
+    guides(color = F)
 ```

### 2019-07-17 R/make_all_forest_plots.R

@@ -43,7 +43,7 @@ test2 <- makeDF(age_sex_matched,
 test2[1] <- c(NA,"Concomitant disease",
               NA,NA,"Mastocytosis",
               NA,NA,"Asthma",
-              NA,NA,"Cardiologic disease",
+              NA,NA,"Cardiovascular dis.",
               NA,NA,"β-blockers",
               NA,NA,"ACE-I",
               NA,NA,"Exercise",
@@ -55,7 +55,7 @@ test2 <- test2[-c(2:4,20:25),]
 #or_plot(test1)
 
 
-png("analysis/figures/figForestfinal.png",
+png("analysis/figures/figForestfinalb.png",
     height = 800,
     width = 1000,
     #res = 300,
@@ -69,13 +69,13 @@ forestplot(labeltext = makeTableText(test1),
            is.summary=c(T,rep(F,length(test1[,1])-1)),
            clip=c(0.2,4),
            xlog=TRUE,
-           col=fpColors(box="#92c442",line="gray", summary="royalblue"),
+           col=fpColors(box="#1f1f1f",line="gray", summary="royalblue"),
            lwd.xaxis = 5,
            lwd.ci = 4,
            lwd.zero = 4)
 dev.off()
 
-png("analysis/figures/figForestfinal.png",
+png("analysis/figures/figForestfinalrmr.png",
     height = 640,
     width = 860,
     #res = 300,
@@ -90,7 +90,7 @@ forestplot(labeltext = makeTableText(test2)[,1:2],
            is.summary=c(T,rep(F,length(test2[,1])-1)),
            clip=c(0.2,4),
            xlog=TRUE,
-           col=fpColors(box="#92c442",line="gray", summary="royalblue"),
+           col=fpColors(box="#1f1f1f",line="gray", summary="royalblue"),
            lwd.xaxis = 5,
            lwd.ci = 4,
            lwd.zero = 4,

### 2019-07-17 R/make_flow.R

@@ -9,21 +9,28 @@ nodes <-
   create_node_df(
     n = 7,
     type = "a",
-    label= c(paste0("All cases in the\nEuropean Anaphylaxis Registry\n",length(data4$b_submitdate)),
+    label= c(paste0("All cases in the\nEuropean Anaphylaxis Registry\n",
+                    format(length(data4$b_submitdate),
+                           big.mark = ",")),
              paste0("Formal anaphylaxis definition met\n",
-                    length(which(data4$reaction_type_brown=="anaphylaxis"))),
+                    format(length(which(data4$reaction_type_brown=="anaphylaxis")),
+                           big.mark = ",")),
              paste0("Insects as elicitors\n",
-                    length(which(data4$d_elicitor_gr5== "insects" &
-                                   data4$reaction_type_brown=="anaphylaxis"))),
+                    format(length(which(data4$d_elicitor_gr5== "insects" &
+                                   data4$reaction_type_brown=="anaphylaxis")),
+                           big.mark = ",")),
              paste0("Other elicitors\n",
-                    length(which(data4$d_elicitor_gr5!= "insects"&
-                                   data4$reaction_type_brown=="anaphylaxis"))),
+                    format(length(which(data4$d_elicitor_gr5!= "insects"&
+                                   data4$reaction_type_brown=="anaphylaxis")),
+                           big.mark = ",")),
              paste(paste0("Yellow-jackets = ",
-                          length(which(data4$q_340_insects== "yellow jacket"&
-                                         data4$reaction_type_brown=="anaphylaxis"))),
+                          format(length(which(data4$q_340_insects== "yellow jacket"&
+                                         data4$reaction_type_brown=="anaphylaxis")),
+                                 big.mark = ",")),
                    paste0("Bees = ",
-                          length(which(data4$q_340_insects== "bee"&
-                                         data4$reaction_type_brown=="anaphylaxis"))),
+                          format(length(which(data4$q_340_insects== "bee"&
+                                         data4$reaction_type_brown=="anaphylaxis")),
+                                 big.mark = ",")),
                    paste0("Hornets = ",
                           length(which(data4$q_340_insects== "hornet"&
                                          data4$reaction_type_brown=="anaphylaxis"))),
@@ -33,12 +40,14 @@ nodes <-
                    sep="\n"
              ),
              paste0("Elicitor unknown\n",
-                    length(which(data4$d_elicitor_gr5=="unkown" &
[…]

### 2019-07-17 R/stat.R

@@ -1009,34 +1009,37 @@ align = "h",
 labels = c("A","B"))
 
 lower_panel <- ggpubr::ggarrange(
+  ANAscore_matched  %>%
+    ggplot(aes(x = d_age, fill = grouping)) +
+    geom_density(alpha = 0.5)+
+    labs(x = "Age [years]",fill = "elicitor")+
+    theme_classic()+
+    theme(legend.position = "none")+
+    scale_fill_manual(values = rev(c("#c4c4c4", "#000000")))+
+    labs(y ="density",x = "Age [years]",fill = "")+
+    geom_segment(aes(x = 22, xend =22, y = 0, yend = 0.02),
+                 linetype = 2),
 
-ANAscore_matched %>%
-  group_by(b_sex,grouping) %>%
-  ggplot(aes(fill=grouping,x=b_sex))+
-  geom_bar(position = "fill")+
-  theme_classic()+
-  theme(legend.position = "none")+
-  scale_fill_manual(values = rev(c("#E5F5E0", "#74C476","#005A32")))+
-  labs(y ="proportion",x = "Sex",fill = "elicitor"),
+  ANAscore_matched %>%
+    group_by(b_sex,grouping) %>%
+    ggplot(aes(fill=grouping,x=b_sex))+
+    geom_bar(position = "fill")+
+    theme_classic()+
+    theme(legend.position = "none")+
+    scale_fill_manual(values = rev(c("#c4c4c4", "#454545")))+
+    labs(y ="proportion",x = "Sex",fill = "elicitor"),
 
-ANAscore_matched  %>%
-  ggplot(aes(x = d_age, fill = grouping)) +
-  geom_density(alpha = 0.5)+
-  labs(x = "Age [years]",fill = "elicitor")+
-  theme_classic()+
-  theme(legend.position = "none")+
-  scale_fill_manual(values = rev(c("#E5F5E0", "#74C476","#005A32")))+
-  labs(y ="density",x = "Age [years]",fill = ""),
-ANAscore_matched  %>%
-  #select(grouping,d_severity_rm) %>%
-  group_by(grouping,d_severity_rm) %>%
-  summarize(n = n()) %>%
-  ggplot(aes(x = as.factor(d_severity_rm),y=n, fill = grouping)) +
-  geom_bar(stat="identity",position="dodge")+
-  theme_classic()+
-  labs(x = "Severity grade [R&M]",fill = "R&M",y="cases [n]")+
[…]

### 2019-07-22 analysis/paper/paper.Rmd
REsults done

@@
 
-Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May till October.
-Their proportion to other cases during the summer seasons reached 60% and was as under 1% of cases during winter.
-Yellow-jacket was the most prominent IVA-causing insect reported in these cases followed by bees.
-The IVA causing insects differed in European countries with hornets being more prominent in the south of Europe.
+Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was frequently reported from May till October.
+Their proportion of IVA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was as under 1% of cases during winter.
+Nevertheless, `r rare_spring_autumn$n %>% sum` cases of IVA (bee in spring; yellow jacket in autumn) anaphylaxis were also reported in March, April and November.
+Yellow-jacket was the most prominent IVA-causing insect followed by bees.
+The IVA-causing insects differed in European countries with hornets being more prominent in the south of Europe.
 IVA was more frequent in adults and seniors when compared to children and young adults.
@@
 
+When we looked at the density plot of cases according to age - we saw a bimodal distribution forming two subset of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B.
+
 ## Symptoms
 
-IVA had a specific symptom pattern.
-Patients, who underwent IVA, more often experienced cardiologic symptoms (dizziness, reduced alertness, unconsciousness) than in other elicitors of anaphylaxis and less often showed gastrointestinal symptoms  (association measured with Cramer's V, Fig \@ref(fig:figsymptoms)).
-The difference in severe hypotension frequency was especially prominent in children under 13 (Fig.
+IVA showed a specific symptom pattern.
+Patients, who underwent IVA, more often experienced cardiologic symptoms (dizziness, reduced alertness, unconsciousness) than in other elicitors of anaphylaxis and less often showed gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A).
+The difference in severe hypotension frequency was especially prominent in the younger group under 22 (Fig.
 S\@ref(fig:hypotensionkids)).
 
-The pattern of organ involvement showed similarities in gastrologic, skin and respiratory symptoms, and did not differ in the fraction of elevated baseline serum tryptase.
-Although, less atopic diseases were seen in patients undergoing IVA (Fig \@ref(fig:figsymptoms)C).
+The pattern of organ involvement showed similarities in gastrologic, skin and respiratory symptoms, and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
+Although, less concomittant atopic diseases were seen in patients undergoing IVA (Fig \@ref(fig:figsymptoms)B).
 
-Severe reactions were more prevalent in adults in comparison to children under 18, and in IVA cases vs other elicitors (Fig.
-\@ref(fig:figsymptoms)D).
-There were no differences in severity between insect species.
+Severe reactions were more prevalent in older patients in comparison to patients under 22, and in IVA cases vs other elicitors (Fig.
+\@ref(fig:sev)D).
+There were no differences in severity of reactions elicited by yellow-jackets and insect species.
 
@@
 
-In general `r treact[1,2] %>% round(1)`% patient with insect allergy had experienced anaphylaxis in the past which is less than if the reaction is elicited by other elcitors (`r treact[2,2]%>% round(1)`%, `r p %>% pval()`).
-We documented `r sum(repeated_reaction_in_Reg)` patients with two documented reactions in our registry.
-Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had Insect elicited anaphylaxis and in `r greater_same[1,2]` (`r greater_same[1,2]/sum(greater_same[,2])*100`) The following reaction was more severe than before.
+In general `r treact[1,2] %>% round(1)`% of patients with insect allergy had experienced anaphylaxis in the past.
+If the reaction was elicited by other elcitors previous reactions were more frequently seen (`r treact[2,2]%>% round(1)`%, `r p %>% pval()`).
+We observed `r sum(repeated_reaction_in_Reg)` patients with two documented reactions in our registry.
+Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had Insect elicited anaphylaxis and in `r greater_same[1,2]` (`r greater_same[1,2]/sum(greater_same[,2])*100`%) the following reaction was more severe than before.
 In `r greater_same[3,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was graded exactly as before.
+
+## Co-factors
+
+The factor most prominently associated with an increased risk of severe anaphylaxis was mastocytosis, and there were no differences in elicitor groups (Fig.
+\@ref(fig:figcofactors)).
+Mastocytosis was increasing the risk of cardiac arrest in patients undergoing IVA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
+S\@ref(fig:cardiacmasto)).
+
+Concomitant cardiovascular diseases were associated with higher risk of severe anaphylaxis when elicited by insects.
+They played insignificant role when anaphylaxis was triggered by other elicitors (Fig.
+\@ref(fig:figcofactors)).
+
+Risk of severe anaphylaxis in patients concomitantly using ACE-I (as well as beta-blockes) could not be independently measured due to coexsisting cardiovascular pathologies.
+ACE-I use was however more often associated with cardiac arrests in both IVA and anaphylaxis by other triggers (Fig.
+\@ref(fig:figcofactors)C).
+Whereas beta-blockers use was associated with higher severity of anaphylaxis and with cardiovascular symptoms (cardiac arrest, chest pain).
+Suprisingly, arrhythmia was more frequently reported in patients undergoing IVA with concomitant beta-blockers (Fig.
+\@ref(fig:figcofactors)D).
+
+Baseline serum tryptase correlated with the severity of reactions in the Ring and Messmer scale and was more prominently increasing risk of severe anaphylaxis in IVA patients than in other cases (Fig.
+\@ref(fig:figcofactors)B).
+Cases with cardiac arrest were associated with increased tryptase over 8 ng/ml and this proportion was higher in IVA when compared to other elicitors.
+Loss of consciousness was also associated with increased tryptase levels but only in patients with IVA (Fig.
+\@ref(fig:figcofactors)C).
+Based on the severity and symptom profile, we decided to use a tryptase cut off value of 8 ng/ml (Fig.
+S\@ref(fig:tryp2)) rather than the currently used 11.5 ng/ml.
+
+We compared the Hymenoptera species responsible for triggering IVA according to the severity of the reaction.
+Patients matched according to sex and age had similar reactions to bees and yellow-jackets when severity of a reaction was considered.
+<!-- This data, however, is not comparable to the refractory anaphylaxis cases as refractory anaphylaxis seems to be a distinct phenotype.  -->
+
+```{r physactivivty}
+ttexercise2 <- age_sex_matched %>% 
+  filter(!is.na(grouping),
+         q_4211_exercise != "unknown") %>% 
+  group_by(grouping,q_4211_exercise,d_severity_rmr) %>% 
+  summarize(n = n()) %>% 
+  filter(grouping == "insects") %>% 
+  spread(key = q_4211_exercise, value = n) %>% 
+  {chisq.test(.[,3:4])$p.value} %>% pval
+  # ggbarplot(x = "q_4211_exercise",
+  #           y = "n",
+  #           facet.by = "grouping",
+  #           fill = "d_severity_rmr",
+  #           position = position_fill())
+
+ttexercise1 <- age_sex_matched %>% 
+  filter(!is.na(grouping),
+         q_4211_exercise != "unknown") %>% 
+  group_by(grouping,q_4211_exercise) %>% 
+  summarize(n = n()) %>% 
+  spread(key = q_4211_exercise,value = n) %>% 
+  {.[,2:3]} %>% chisq.test() %>% {.$p.value} %>% pval()
+
+```
+
+Physical exercise in the fresh air (e.g. jogging in the park) predisposed to anaphylaxis due to insect stings (`r ttexercise1`).
+But even though the physical exercise was more often associated with IVA than other triggers of anaphylaxis - it was not predictive of the severity of a reaction in these patients, as there were no differences in relation to concomitant exercise (`r ttexercise2`).
 
@@
 
-Patients who underwent IVA significantly less often received adrenaline treatment than in other adrenaline cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
-After adjusting both groups for similar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the route of administration.
-Antihistaminic drugs, on the other hand, were given more often in cases of IVA when compared to non-IVA.
+```{r}
+ttadren1 <- ANAscore_matched %>% 
+  group_by(d_522_adren_agg,grouping) %>% 
+  summarize(n = n()) %>% 
+  spread(key = grouping, value = n) %>% 
+  {chisq.test(.[,2:3])$p.value} %>% 
+  pval()
+```
 
-IVA was most often treated with corticosteroids and antihistamines (significantly more frequent than in other anaphylaxis cases).
+Patients who underwent IVA significantly less often received adrenaline treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
+After adjusting both groups for similar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)).
+<!-- Antihistaminic drugs, on the other hand, were given more often in cases of IVA when compared to non-IVA. -->
+IVA was most often treated with corticosteroids and antihistamines (significantly more frequently than in other anaphylaxis cases).
 On the other hand adrenaline, beta-2 mimetics and oxygen were given more often in non-IVA.
 
-We discovered clusters of symptoms and therapy modes.
-In IVA patients urticaria, and cardiologic symptoms (cardiac arrest, hypotension, loss of consciousness), as well as respiratory distress, were treated more similarly than gastroenterologic symptoms (Fig.
+We discovered clusters of associated symptoms and therapy modes in IVA patients (association measured useing phi).
+Cardiologic symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria, were treated more similarly than  respiratory or gastroenteral symptoms (Fig.
 \@ref(fig:adrenalineuse)C).
-Therapy was dependent on the symptom profile and consisted of AAI, i.v.
-adrenaline in multiple doses, pure oxygen inhalation, an initial dose of antihistamines, and beta 2 agonists.
-Corticosteroids, volume replacement and i.v. beta2 agonists formed another therapy mode.
+Therapy of these symptoms consisted of adrenaline autoinjector (AAI), i.v.
+adrenaline in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists.
+Corticosteroids, i.v. volume replacement and i.v. β-2 agonists formed another therapy mode.
 
@@
 \@ref(fig:adrenalineuse)B).
-
-## Co-factors
-
-Although the physical exercise was more often related to IVA than other triggers of anaphylaxis - it was not predictive of the severity of a reaction in these patients.
-
-Concomitant use of ACE-I did not correlate with reaction severity measured in ANAscore and VAS scales.
-It was however more often associated with cardiac arrests in both IVA and other-triggers anaphylaxis (Fig.
-S\@ref(fig:acei)).
-
-Beta-blockers use was associated with higher severity scores in VAS and ANAscore and with cardiac arrest, chest pain.
-Especially arrhythmia was more frequently reported in patients undergoing IVA with concomitant beta-blockers.
-
-Mastocytosis was increasing the risk of cardiac arrest in patients undergoing IVA significantly more in patients undergoing anaphylaxis due to other triggers (Fig S\@ref(fig:cardiacmasto)).
-
-Baseline serum tryptase correlated with the severity of reactions and especially in the Ring and Messmer scale was more prominently increasing risk of severe anaphylaxis in IVA patients than in other cases (Fig.
-\@ref(fig:figcofactors)B and S\@ref(fig:tryp)).
-In the VAS scale, this did not show as significant, but it did in all the other severity scales.
-We decided to use Tryptase cut off value of 8 ng/ml (Fig.
-S\@ref(fig:tryp2)).
-We saw that cases with cardiac arrest were associated with increased tryptase over 8 ng/ml and this proportion was higher in IVA when compared to other elicitors.
-Loss of consciousness was also associated with increased tryptase levels but only in patients with IVA (Fig.
-\@ref(fig:figcofactors)C).
-
-We compared the Hymenoptera species which was responsible for triggering IVA according to the severity of the reaction.
-Patients matched according to sex and age has similar reactions to bees and yellow-jackets (in Brown and Ring and Messmer scales).
-In ANAscore - yellow jacket stings were linked to a slightly more severe reaction.
-This data, however, is not comparable to the refractory anaphylaxis cases as refractory anaphylaxis seems to be a distinct phenotype.
 
@@
 
-```{r figcofactors, fig.cap="Cofactors of insect venom anaphylaxis. A: Odds ratios of eliciting severe anaphylaxis. B: Proportional ",fig.width=5.6,fig.height = 6}
+```{r figcofactors, fig.cap="Cofactors of insect venom anaphylaxis.
+A: Odds ratios of eliciting severe anaphylaxis.
+B: Proportion of cases elicited by insects or other elicitors (upper panels) according to tryptase levels and cardiologic symptoms.
+", fig.width = 5.6, fig.height = 10}
 cof_fig
@@
 
-```{r adrenalineuse, fig.cap = "Therapy modes in patients with IVA compared to oter elicitors. A)  B)",fig.height=10,fig.width = 7}
+```{r adrenalineuse, fig.cap = "Therapy in patients with IVA compared to oter elicitors, cases matched according to sex, age and severity of a reaction.
+A: Proportional use of therapy measures in both anaphylaxis groups.
+B: Heatmap visualizing the associacion of symptoms and corresponding treatment -  presented as a scaled correlation coefficient (phi).", fig.height=10, fig.width = 7}
 
 ggpubr::ggarrange(
-  ggarrange(
     plotManagement,
-    rf_ggplot_treatment,
-    labels = c("A","B")
-  ),
+
   #lower_panel,
@@
     geom_segment(aes(x = c(0.5),
-                     xend = c(21.5), 
+                     xend = c(20.5), 
                      y = c(28.5),
-                     yend = c(28.5)), color = "red" ),
+                     yend = c(28.5)), color = "red" )+
+    geom_segment(aes(x = c(0.5),
+                     xend = c(20.5), 
+                     y = c(34.5),
+                     yend = c(34.5)), color = "red", linetype = 2  ),
   nrow = 2,
@@
 )
+
+# ggarrange(matrix_venom$py+
+#             theme(plot.margin=unit(c(0.1,9.8,0,9.5),"lines")),
+#           ggarrange(matrix_venom$p,
+#                     matrix_venom$px+
+#                       theme(plot.margin=unit(c(1,0,8,0),"lines")),
+#                     widths = c(1,0.2)),
+#           ncol = 1,
+#           nrow = 2,
+#           heights = c(0.2,1)
+# )
+
 ```
@@
 
-```{r}
+```{r sev, fig.cap="Severity of anaphylaxis in subgroups.
+Severity of patients with IVA in two age groups (left), according to elicitor type (center) and according to the responsible insect species (right)"}
 severity_joined_brown %>%
@@
               fill = "d_severity_rm",
-              position = position_fill(reverse = T),
+              position = position_fill(reverse = F),
               color = "d_severity_rm",
-              palette = "lancet")+
+              palette = c("#848484","#1f1f1f"))+
     facet_grid(.~subset,scale = "free_x",space = "free_x")+
@@
     )+
-    annotate("text",x = 1.5, y = 0.5, label = c("*","*",""),
-             size =12)+
+    #annotate("text",x = 1.5, y = 0.5, label = c("*"),
+     #        size =12)+
     guides(color = F)
@@
 ```{r cofactorscomparison}
-testInsectsbinomialTab$cofactors[,-7] %>% knitr::kable()
+#testInsectsbinomialTab$cofactors[,-7] %>% knitr::kable()
 
@@
 ```
+
+```{r rfplot}
+rf_ggplot_treatment
+```

### 2019-07-22 R/stat.R

@@ -446,15 +446,15 @@ tempdf <- age_sex_matched %>%
 # Split the plot into kids and grown ups AGE division - 12
 # Get the proper table
 
-symptTabKids <- makeTests(rdb=rdb[rdb$d_age<13,],
+symptTabKids <- makeTests(rdb=rdb[rdb$d_age<22,],
           groups = "grouping") %>%
           {split(.,.$section)} %>%
-  .$symptoms %>% mutate(subset = "children")
+  .$symptoms %>% mutate(subset = "under 22")
 
-symptTabAdults <- makeTests(rdb=rdb[rdb$d_age>13,],
+symptTabAdults <- makeTests(rdb=rdb[rdb$d_age>=22,],
                                   groups = "grouping") %>%
                                   {split(.,.$section)} %>%
-                          .$symptoms %>% mutate(subset = "adults")
+                          .$symptoms %>% mutate(subset = "over 22")
 ex <- full_join(symptTabKids,symptTabAdults)
 
 # ggplot(ex %>% filter(pval<1e-10) %>%
@@ -491,7 +491,8 @@ hypotensionPlot <-
   #      axis.ticks.x = element_blank())+
   labs(y = "proportion [%]", x = "hypotension")+
    theme(axis.text.x = element_text(angle = 50,hjust =1),
-         legend.position = "top")
+         legend.position = "top",
+         legend.title = element_blank())
 
 
 
@@ -885,8 +886,9 @@ plotManagement <- testANAscoreMatched$management$variableName %>%
     legend.justification = c(1, 1)
   ) +
   labs(x = "Therapy", y = "proportion", fill = "") +
-  scale_fill_manual(values = rev(c("#E5F5E0", "#74C476", "#005A32"))) +
-  theme(axis.title.x = element_blank())+ scale_y_continuous(labels = scales::percent_format(accuracy = 1))
+  scale_fill_manual(values = rev(c("#848484","#1f1f1f"))) +
+  theme(axis.title.x = element_blank())+
+  scale_y_continuous(labels = scales::percent_format(accuracy = 1))
 
 
 backup <- testANAscoreMatched$management[, -7] %>%
@@ -1053,8 +1055,9 @@ labels = c("B","C","D"))
 
 test_adrenuse1<- ANAscore_matched %>%
   group_by(grouping,d_522_adren_agg) %>%
-  summarize(n = n()) %>% {matrix(.$n,ncol = 2)} %>%
[…]

### 2019-07-23 analysis/paper/paper.Rmd
spellcheck gramarly 1

@@
 
-Symptoms of insect venom anaphylaxis are distinctively different from other reactions, indicating that the therapy of insect elicited cases of anaphylaxis should be considered separately.
+Symptoms of insect venom anaphylaxis are distinctively different from other anaphylactic reactions.
+What made us consider the therapy of insect elicited cases of anaphylaxis separately.
 Indeed we observed different therapeutic patterns in insect elicited cases of anaphylaxis (more antihistaminics but fewer corticosteroids, bronchodilators, and surprisingly - adrenaline).
-This indicates that the management of insect-venom induced anaphylaxis may be improved and is especially required in patients with concomitant cardiologic conditions and these with hyperreactive mast cells.
+The presented evidence indicates that the management of insect-venom anaphylaxis may be improved.
+Patients with concomitant cardiologic conditions and these with hyperreactive mast cells require intensified prophylactic measures.
 
@@
 
+Hypersensitivity to insect venom may manifest as a systemic reaction (anaphylaxis) in up to 0.3-7.5% of the adult population [@Bil2008].
+Insect venom anaphylaxis (IVA) can be fatal, and patients sometimes require lifelong systemic immunotherapy [@Sturm2017].
+
+Recent expert position on the future anaphylaxis research indicates the need for a more precise description of diagnosis, biomarkers, and phenotypes of anaphylaxis [@JimenezRodriguez2018].
+Nevertheless, in order to facilitate precision-medicine approach [@Muraro2017] in the treatment of anaphylaxis - we first need to investigate its subtypes and pathomechanisms better.
+
+Clinical controlled trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition.
+Therefore large registries, gathering clinical data as well as biological samples from patients with a well documented (recent) history of anaphylaxis are a crucial tool in investigating this entity.
+
+This study aimed to identify current patterns of Insect venom anaphylaxis (IVA) in symptoms, cofactors, and management and compare them to other types of anaphylaxis (non-IVA) in a case-control manner.
+
 # Methods
 
-The European Anaphylaxis Registry [@Grabenhenrich2016] database from March 2019 was searched for anaphylaxis cases elicited by insect's venom.
-The flowchart in figure \@ref(fig:flowchart) represents the detailed case-selection process.
+We searched the European Anaphylaxis Registry [@Grabenhenrich2016] database from March 2019 for anaphylaxis cases elicited by insect's venom.
+The flowchart in figure \@ref(fig:flow) represents the detailed case-selection process.
 
@@
 Severe reactions were identified based on the definition by NIAID/FAAN [@Sampson2006] or as Ring and Messmer Scale - grades III and IV) and presented with significant hypoxia, hypotension, confusion, collapse and loss of consciousness, or incontinence.
-We compared the frequency of various elicitors, symptoms, and factors known to increase the risk of severe anaphylaxis [@Worm2018] in both groups.
-We evaluated symptoms, management and risk factors of insect elicited cases in comparison to other known triggers of anaphylaxis.
 
-We observed a significant difference in the clinical features of insect anaphylaxis in children and adults.
-The younger population significantly less often had concomitant conditions (i.e.
-DM, HT, malignant diseases and mastocytosis), and more often presented with atopic dermatitis, rhinitis, and asthma.
-Therefore, we decided to adjust the analysis for age and sex to reduce the comparison bias.
-When comparing the management of both types of anaphylaxis we also matched the control group according to severity.
+The cases were either assigned to IVA group (if the triggering factor was an insect sting) or non-IVA group (other elicitors of anaphylaxis).
+We compared the frequency of various symptoms, cofactors, known to increase the risk of severe anaphylaxis, [@Worm2018] and management in both groups.
 
-<!-- ## Data handling and statisitcs -->
+Due to a large number of documented reactions in the European Anaphylaxis registry - we were able to match the IVA cases to selected non-IVA cases according to sex and age to reduce the comparison bias.
+Subsequently, we compared the management in both groups and matched the control group according to the severity of a reaction.
+Propensity score matching has been performed using the "MatchIt" package for R [@matchit].
+The results of the propensity score matching are in Fig.
+\ref(fig:flow)B-D.
+
+We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find which therapy approaches were best fit for classifying cases either to IVA or non-IVA group and presented the results as mean decrease in Gini.
+We also performed an association study of therapeutic interventions and symptoms.
+The resulting phi values are presented in a heatmap with automatic clustering using Ward's D and euclidean distancing method [@heatmaply].
+
+<!-- ## Data handling and statistics -->
 The statistical analysis was performed in the R Statistical Package [@R].
@@
 We defined statistical significance as α = 0.05.
-Data along with the analysis script can be accessed online at https://github.com/wolass/venomanaphylaxiscompendium.
+Data, along with the analysis script, can be accessed online at https://github.com/wolass/venomanaphylaxiscompendium.
 
@@
 
-## Temporal distribution of anaphylaxis cases by elicitor.
+## Perennial distribution of anaphylaxis cases by elicitor.
 
@@
 Their proportion of IVA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was as under 1% of cases during winter.
-Nevertheless, `r rare_spring_autumn$n %>% sum` cases of IVA (bee in spring; yellow jacket in autumn) anaphylaxis were also reported in March, April and November.
+Nevertheless, `r rare_spring_autumn$n %>% sum` cases of IVA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were also reported in March, April, and November.
 Yellow-jacket was the most prominent IVA-causing insect followed by bees.
-The IVA-causing insects differed in European countries with hornets being more prominent in the south of Europe.
+The IVA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in the south of Europe.
+
 IVA was more frequent in adults and seniors when compared to children and young adults.
 (Fig \@ref(fig:insectstime)B).
-
-When we looked at the density plot of cases according to age - we saw a bimodal distribution forming two subset of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B.
+When we looked at the density plot of IVA cases according to age - we saw a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B.
 
 ## Symptoms
-
 IVA showed a specific symptom pattern.
@@
 The difference in severe hypotension frequency was especially prominent in the younger group under 22 (Fig.
-S\@ref(fig:hypotensionkids)).
+S\@ref(fig:supplement)).
 
-The pattern of organ involvement showed similarities in gastrologic, skin and respiratory symptoms, and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
-Although, less concomittant atopic diseases were seen in patients undergoing IVA (Fig \@ref(fig:figsymptoms)B).
+The pattern of organ involvement showed similarities in gastrologic, skin, and respiratory symptoms, and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
+Although, patients undergoing IVA  less often had concomitant atopic diseases (Fig \@ref(fig:figsymptoms)B).
 
-Severe reactions were more prevalent in older patients in comparison to patients under 22, and in IVA cases vs other elicitors (Fig.
+Severe reactions were more prevalent in older patients in comparison to patients under 22, and in IVA cases vs.
+other elicitors (Fig.
 \@ref(fig:sev)D).
@@
 
-Concomitant cardiovascular diseases were associated with higher risk of severe anaphylaxis when elicited by insects.
-They played insignificant role when anaphylaxis was triggered by other elicitors (Fig.
+Concomitant cardiovascular diseases were associated with a higher risk of severe anaphylaxis when elicited by insects.
+They played an insignificant role in non-IVA cases (Fig.
 \@ref(fig:figcofactors)).
 
-Risk of severe anaphylaxis in patients concomitantly using ACE-I (as well as beta-blockes) could not be independently measured due to coexsisting cardiovascular pathologies.
-ACE-I use was however more often associated with cardiac arrests in both IVA and anaphylaxis by other triggers (Fig.
-\@ref(fig:figcofactors)C).
-Whereas beta-blockers use was associated with higher severity of anaphylaxis and with cardiovascular symptoms (cardiac arrest, chest pain).
-Suprisingly, arrhythmia was more frequently reported in patients undergoing IVA with concomitant beta-blockers (Fig.
+Risk of severe anaphylaxis in patients concomitantly using ACE-I (as well as beta-blockers) could not be independently measured due to coexisting cardiovascular pathologies.
+ACE-I use was, however, more often associated with cardiac arrests in both IVA and anaphylaxis by other triggers (Fig.
+\@ref(fig:figcofactors)C) whereas beta-blockers use was associated with higher severity of anaphylaxis and with cardiovascular symptoms (cardiac arrest, chest pain).
+Surprisingly, arrhythmia was more frequently reported in patients undergoing IVA with concomitant beta-blockers (Fig.
 \@ref(fig:figcofactors)D).
@@
 \@ref(fig:figcofactors)B).
-Cases with cardiac arrest were associated with increased tryptase over 8 ng/ml and this proportion was higher in IVA when compared to other elicitors.
+Cases with cardiac arrest were associated with increased tryptase over eight ng/ml, and this proportion was higher in IVA when compared to other elicitors.
 Loss of consciousness was also associated with increased tryptase levels but only in patients with IVA (Fig.
@@
 
-We compared the Hymenoptera species responsible for triggering IVA according to the severity of the reaction.
-Patients matched according to sex and age had similar reactions to bees and yellow-jackets when severity of a reaction was considered.
+We compared the Hymenoptera species responsible for triggering IVA according to the severity of the reaction and found no differences.
+Patients matched according to sex and age had similar severity of a reaction to known eliciting insects (p = 0.4128).
 <!-- This data, however, is not comparable to the refractory anaphylaxis cases as refractory anaphylaxis seems to be a distinct phenotype.  -->
@@
 
-Physical exercise in the fresh air (e.g. jogging in the park) predisposed to anaphylaxis due to insect stings (`r ttexercise1`).
-But even though the physical exercise was more often associated with IVA than other triggers of anaphylaxis - it was not predictive of the severity of a reaction in these patients, as there were no differences in relation to concomitant exercise (`r ttexercise2`).
+Physical exercise in the fresh air (e.g., jogging in the park) predisposed to anaphylaxis due to insect stings (`r ttexercise1`).
+However, even though the physical exercise was more often associated with IVA than other triggers of anaphylaxis - it was not predictive of the severity of a reaction in these patients, as there were no differences in severity concerning accompanying exercise (`r ttexercise2`).
 
@@
 
-```{r}
+```{r mngmttests}
 ttadren1 <- ANAscore_matched %>% 
@@
 <!-- Antihistaminic drugs, on the other hand, were given more often in cases of IVA when compared to non-IVA. -->
-IVA was most often treated with corticosteroids and antihistamines (significantly more frequently than in other anaphylaxis cases).
-On the other hand adrenaline, beta-2 mimetics and oxygen were given more often in non-IVA.
+Patients with IVA received corticosteroids and antihistamines significantly more frequently than in other anaphylaxis cases.
+On the other hand, adrenaline, beta-2 mimetics, and oxygen were given more often in non-IVA.
 
-We discovered clusters of associated symptoms and therapy modes in IVA patients (association measured useing phi).
-Cardiologic symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria, were treated more similarly than  respiratory or gastroenteral symptoms (Fig.
-\@ref(fig:adrenalineuse)C).
+We discovered clusters of associated symptoms and therapy modes in IVA patients (association measured using phi).
+Cardiologic symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria, were treated more similarly than respiratory or gastrointestinal symptoms (Fig.
+\@ref(fig:adrenalineuse)B).
 Therapy of these symptoms consisted of adrenaline autoinjector (AAI), i.v.
@@
 
-The most differences in the therapy of IVA vs.
-other forms of anaphylaxis was observed in the frequency of inhaled beta2-agonists and antihistamines (probably because these are not available for the IVA patients, but atopic patients use them regularly) (Fig.
-\@ref(fig:adrenalineuse)B).
+The most differences in the therapy of IVA vs. other forms of anaphylaxis was observed in the frequency of inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
+
+```{r }
+pvaladrenprev <- adren_prev$data %>% 
+  spread(key = d_522_adren_agg,value = n) %>% 
+  filter(q_160_ever_react=="reacted previously") %>% 
+  {chisq.test(.[,3:4])$p.value} %>% pval()
+
+pvaladrenprev_no <- adren_prev$data %>% 
+  spread(key = d_522_adren_agg,value = n) %>% 
+  filter(q_160_ever_react=="no prev. ANA") %>% 
+  {chisq.test(.[,3:4])$p.value} %>% pval()
+
+```
+
+Interestingly, adrenaline as a first emergency therapy was given less often in IVA cases when compared to other cases **if patients did not report the previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients who reported previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`).
+
+Similarly, there were no differences in adrenaline use between IVA and non-IVA when severe reactions were taken into consideration (`r adren_severity_sev`).
+However, when we restricted the analysis to mild anaphylaxis cases - non-IVA patients received adrenaline more frequently than IVA (`r adren_severity_mild`).
 
@@
 
-Cardiologic symptoms and hypotensive collapse might be associated with venom anaphylaxis due to the vaso-vagal reflex.
+<!-- Cardiologic symptoms and hypotensive collapse might be associated with venom anaphylaxis due to the vaso-vagal reflex.
 As seen in patients undergoing blood sampling needles (and probably also insect stings) may elicit a hypotensive response due to extreme emotional distress with bradycardia, limiting the blood flow to the brain and causing syncope.
 This mechanism may contribute to the ongoing histamine-induced vasodilatation and may worsen the symptoms of an allergic reaction.
+-->
+
+Surprisingly, detailed information on the seasonality of insect-elicited hypersensitivity reactions is scarce [@old].
+The activity of *Vespula germanica* depends on the climate, and in invaded regions (e.i.
+Australia) it can even extend throughout the whole year [@Spradbery_1992].
+The changing climate in Europe may influence the activity of Hymenoptera in this region in the upcoming years, but in the period from 2007 - 2019 the perennial distribution of IVA has remained stable (data not shown).
+
+We saw differences in IVA frequency in age groups.
+Especially at the age extremes, IVA was less frequent, and the density distribution showed two modes (for ages of 10 and 54).
+Although frequency of insect stings hypersensitivity reactions in children is comparable to food hypersensitivity reactions (12-15% of cases of hypersensitivity reactions admitted to the ER), anaphylaxis is triggered significantly more often by food elicitors in childhood (56% of food hypersensitivity cases vs 5.3% of sting cases seen in the ER) [@Braganza2005].
+Senior patients, on the other hand, suffer from drug-related hypersensitivity more often than insect sting hypersensitivity [@Aurich2019].
+
+IVA was more associated with cardiologic symptoms than other anaphylaxis types.
+Additionally, cardiologic diseases were vital cofactor increasing the risk of a severe reaction in IVA only.
+The body of evidence suggests [@Worm2018; @nittner2015insect; @Ruff2009] that there is an essential link between the cardiovascular system and insect stings as they have been associated with cardiac arrhythmias usually occurring in patients with preexisting heart disease [@Sharma2016].
+
+Kounis syndrome (coronary arterial spasm induced by the release of mast cell mediators) could be another explanation for the increased concomitance of cardiovascular symptoms with IVA ([@CDT1609; @sinkiewicz2008allergic]).
+As the severity of cardiologic reactions was positively associated with the elevated BST (> 8 ng/ml), we theorize that insect venom is unspecifically activating more mast cells at once than food or drug elicitors (potentially due to its composition and subcutaneous route of administration).
+Such rapid activation leads to a more evident cardiovascular spasm (especially if the concomitant cardiovascular disease is present) and in result to more severe anaphylaxis.
+
+We propose to review the currently used reference values of serum tryptase levels.
+In our opinion, we should consider a cut-off value of 8 ng/ml as elevated, when evaluating patients in the context of IVA.
+Baseline serum tryptase levels (BST) of > 8 ng/ml were better predicting the risk of severe anaphylaxis and corresponded with cardiovascular symptom frequency.
+
+Cases of IVA had been treated with adrenaline less often than the age-and-sex-matched cases of anaphylaxis due to other elicitors.
+Moreover, administration of adrenaline did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patient experienced their first episode of IVA.
+One explanation for this fact may be that although symptoms of IVA are more severe (cardiologic symptoms are considered life-threatening), they tend to be more transient.
+In case of food allergy, patients could develop symptoms for an extended time and will eventually be treated with adrenaline, whereas patients with IVA will often be recovering from hypotension or collapse when the emergency medical team arrive at the site (with AAI).
+
+We saw less IVA in patients with concomitant atopic diseases, as these patients more often present with food anaphylaxis [@Tham2019].
 
@@
 
+Based on our results, IVA is a distinctive subtype of anaphylaxis, with specific symptom profile and risk factors.
+Although therapy of IVA should be performed according to the guidelines and should not differ from other types of anaphylaxis, it is unclear why patients with IVA tend to receive adrenaline less often.
+
+When evaluating the risk of future severe episodes - baseline serum tryptase levels over eight ng/ml should be considered.
+
 # Acknowledgements
+We thank all patients, parents, and their children for their support in providing data on the occurrence of anaphylaxis for this study.
+We thank the study personnel for patients counseling and data entry, and we thank the members of The European Anaphylaxis Registry in detail:
+J. Grünhagen (Berlin, Germany),
+K. Beyer (Berlin, Germany),
+A. Möser (Jena, Germany),
+T. Fuchs (Göttingen, Germany),
+F. Ruëff (Munich, Germany),
+B. Wedi (Hanover, Germany),
+F. Friedrichs (Aachen, Germany),
+H. Dickel (Bochum, Germany),
+H. Merk (Aachen, Germany),
+U. Hillen (Essen, Germany),
+A. Bauer (Dresden, Germany),
+N. Wagner (Erlangen, Germany),
+E. Rietschel (Cologne, Germany),
+R. Treudler, S. Aurich (Leipzig, Germany),
+L. Klimek, O. Pfaar (Wiesbaden, Germany),
+N. Reider (Innsbruck, Germany),
+W. Aberer (Graz, Austria),
+A. Köhli (Zurich, Switzerland),
+F. Riffelmann (Schmallenberg, Germany),
+B. Kreft (Halle, Germany),
+K. Nemat (Dresden, Germany),
+T. Kinaciyan (Vienna, Austria),
+R. Brehler (Münster, Germany),
+J. Witte (Hamburg, Germany),
+N. Hunzelmann und I. Huseynow (Cologne, Germany),
+T. Bieber (Bonn, Germany),
+U. Rabe (Treuenbritzen, Germany),
+P. Schmid-Grendelmeier (Zurich, Switzerland),
+W. Brosi (Würzburg, Germany),
+S. Nestoris (Lippe-Lemgo, Germany),
+T. Hawranek (Salzburg, Austria),
+R. Bruns (Greifswald, Germany),
+S. Lehmann (Aachen, Germany),
+L. Lange (Bonn, Germany),
+G. Hansen (Hanover, Germany),
+C. Pföhler (Homburg, Germany),
+E. Varga (Graz, Austria),
+Z. Szepfalusi (Vienna, Austria),
+P. Eng (Aarau, Switzerland),
+P. Eng (Lucerne, Switzerland),
+T. Reese (Rheine, Germany),
+M. Polz (Rüsselsheim, Germany),
+S. Schweitzer-Krantz (Düsseldorf, Germany),
+H. Rebmann (Tübingen, Germany),
+G. Stichtenoth (Lübeck, Germany),
+S. Theis (Schwedt, Germany),
+I. Yildiz (Neumünster, Germany),
+M. Gerstlauer (Augsburg, Germany),
+A. Nordwig (Dresden, Germany),
+R. Schlags (Wangen im Allgäu, Germany),
+I. Neustädter (Fürth, Germany),
+C. Stadlin (Zurich, Switzerland),
+M. Bücheler (Bonn, Germany),
+S. Volkmuth (Velbert, Germany),
+J. Fischer (Tübingen, Germany),
+A. Henschel (Berlin, Germany),
+S. Plank-Habibi (Alzenau, Germany),
+B. Schilling (Passau, Germany),
+A. Kleinheinz (Buxtehude, Germany),
+K. Schäkel (Heidelberg, Germany),
+N. Papadopoulos (Athen, Greece),
+M. Kowalski (Lodz, Poland),
+M. Fernandez-Rivas (Madrid, Spain),
+K. Solarewicz-Madajek (Wroclaw, Poland),
+C. Körner-Rettberg (Bochum, Germany),
+T. Mustakov (Sofia, Bulgaria),
+K. Hartmann (Lübeck, Germany),
+C. Kemen (Hamburg, Germany),
+U. Miehe (Leipzig, Germany),
+C. Ebner (Vienna, Austria),
+S. Haak (Oldenburg, Deutschland),
+V. Cardona (Barcelona, Spain),
+S. Hämmerling (Heidelberg, Germany),
+E. Arroabarren (Pamplona, Spain),
+M. Bilo (Ancona, Italy),
+N. Cabañes Higuero (Toledo, Spain),
+A. Vega Castro (Guadalajara, Spain),
+I. Poziomkowska-Gęsicka (Szczecin, Poland),
+S. Büsing (Osnabrück, Germany),
+U. Klettke (Berlin, Germany),
+C. Virchow (Rostock, Germany),
+G. Christoff (Sofia, Bulgaria),
+U. Jappe (Borstel, Germany),
+K. Breuer (Hamburg, Germany),
+S. Müller and T. Jakob (Freiburg, Germany),
+H. Straube (Darmstadt, Germany),
+C. Vogelberg (Dresden, Germany),
+F. Knöpfel (Norderney, Germany),
+J. Hourihane (Cork, Ireland),
+B. Rogala (Silesia, Poland),
+A. Montoro (Madrid, Spain),
+A. Brandes (Frankfurt/Oder, Germany),
+A. Muraro (Padua, Italy),
+T. Buck und J. Büsselberg (Hanover-Misburg, Germany),
+N. Zimmermann (Potsdam, Germany),
+D. Hernandez (Valencia, Spain),
+P. Minale (Genua, Italy),
+J. Niederwimmer und B. Zahel (Linz, Austria),
+A. Fiocchi (Rome, Italy),
+A. Reissig (Gera, Germany),
+F. Horak (Vienna, Austria),
+N. Klossowiski (Düsseldorf, Germany),
+F. Eitelberger (Wels, Austria),
+H. Ott (Hanover, Germany),
+R. Asero (Milan, Italy),
+S. Pistauer (Sylt/Westerland, Germany),
+M. Geißler (Ribnitz-Damgarten, Germany),
+L. Ensina (Sao Paulo, Brazil),
+A. Plaza Martin (Barcelona, Spain),
+J. Meister (Aue, Germany),
+S. Hompes (Hamburg, Germany),
+S. Stieglitz (Wuppertal, Germany).
 
-<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
 #####
@@
 ```{r flow,fig.width=6,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale for including registry cases in the final analysis.
-B,C,D: Age, sex, and severity distribution was matched in cases in both groups to allow for comparable results between IVA and non-IVA cases.
+B, C, D: Age, sex, and severity distribution was matched in cases in both groups to allow for comparable results between IVA and non-IVA cases.
 Two age-subsets of patients could be recognized based on the density plot of age (B)."}
@@
 ```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occurred.
-Less frequent insect species were grouped together as 'other'.
+Less common insect species grouped as 'other'.
 B: The density distribution of IVA cases to cases elicited by other triggers considering the patient's age.
@@
 A: Proportional presentation of specific reaction symptoms.
-B: High-level overview of involved organ systems and selected cofactors in form of a radar plot.", fig.width=8.5, fig.height = 3.2}
+B: High-level overview of involved organ systems and selected cofactors in the form of a radar plot.", fig.width=8.5, fig.height = 3.2}
 fig_symptoms
@@
 
-```{r adrenalineuse, fig.cap = "Therapy in patients with IVA compared to oter elicitors, cases matched according to sex, age and severity of a reaction.
+```{r adrenalineuse, fig.cap = "Therapy in patients with IVA compared to other elicitors, cases matched according to sex, age, and severity of a reaction.
 A: Proportional use of therapy measures in both anaphylaxis groups.
-B: Heatmap visualizing the associacion of symptoms and corresponding treatment -  presented as a scaled correlation coefficient (phi).", fig.height=10, fig.width = 7}
+B: Heatmap visualizing the association of symptoms and corresponding treatment -  presented as a scaled correlation coefficient (phi).", fig.height=9, fig.width = 7}
 
@@
 
-```{r treatmentmatrixother}
-#matrix_other
-```
+```{r supplement, fig.width= 6, fig.height= 5, fig.cap="Symptoms of anaphylaxis"}
 
-```{r crammersv}
-tests_matched %>%
+g_age <- severity_joined_brown %>%
+    ggbarplot(x = "grouping",
+              y = "n",
+              fill = "d_severity_rm",
+              position = position_fill(reverse = F),
+              color = "d_severity_rm",
+              palette = c("#848484","#1f1f1f"))+
+    facet_grid(.~subset,scale = "free_x",space = "free_x")+
+    labs(fill = "Severity")+
+    theme(axis.title.x =  element_blank(),
+          axis.text.x = element_text(angle=20,hjust = 1),
+          legend.position = "right"
+    )+
+    #annotate("text",x = 1.5, y = 0.5, label = c("*"),
+     #        size =12)+
+    guides(color = F)
+
+cramV <- tests_matched %>%
     mutate(organ = factor(ifelse(grepl(variableName,pattern = "111"),"skin",
@@
     labs(x = "",y = "Cramer's V",color = "")+
-    theme(legend.justification = c(-3,0))
+    theme(legend.justification = c(0,0))
 
+ggarrange(
+  ggarrange(
+  cardiac_masto_plots[[6]]+
+              labs(x = "cardiac arrest"),
+  hypotension2,
+    nrow = 1,
+    ncol = 2
+  ),
+cramV,
+nrow = 2, 
+ncol = 1,
+heights = c(2,1.8)
+)
+```
+
+```{r treatmentmatrixother}
+#matrix_other
 ```
@@
 ```{r sev, fig.cap="Severity of anaphylaxis in subgroups.
-Severity of patients with IVA in two age groups (left), according to elicitor type (center) and according to the responsible insect species (right)"}
+The severity of patients with IVA in two age groups (left), according to elicitor type (center) and according to the responsible insect species (right)"}
 severity_joined_brown %>%
@@
 
-```{r hypotensionkids, fig.cap="Hypotension in IVA of kids"}
-hypotensionPlot
+```{r rfplot}
+rf_ggplot_treatment
 ```
@@
 ```{r}
-ggarrange(dt,
-          dt3 %>%
-ggbarplot(x = "q_610_sit_prior_v5",
-          y = "n",
-          fill = "severity_brown",
-          position = position_fill())+
-  labs(fill="FAAN"))
+adren_prev
 ```
-
-```{r cofactorscomparison}
-#testInsectsbinomialTab$cofactors[,-7] %>% knitr::kable()
-
-testInsectsbinomialTab$cofactors[,-7] %>%
-  tidyr::gather(value ="Proportion",
-         key="Group",
-         4:5) %>% 
-  ggplot(aes(reorder(variableName,-Proportion),Proportion,fill=Group))+
-  geom_bar(stat="identity",position="dodge")+
-  theme(axis.text.x=element_text(angle=90))
-```
-
-```{r severityinsectspecies}
-yj_bee %>%
-  ggbarplot(x = "grouping",
-            y = "n",
-            fill = "d_severity_rm",
-            position = position_fill())
-
-```
-
-```{r acei, fig.cap="Ace inhibitors and cardiologic symptoms"}
-
-ggarrange(
-  plot_ace_cardiacs,
-ggarrange(rdb %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_423_ace,q_116_VAS_v7))+
-  geom_boxplot(),
-rdb %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_423_ace,ANAscore))+
-  geom_boxplot()
-),
-nrow = 2,
-ncol = 1
-)
-
-```
-
-```{r betas, fig.cap = "Beta blockers and cardiologic symptoms"}
-rdb_temp <- rdb
-rdb_temp$q_423_beta <- yesonly(rdb_temp$q_423_beta)
-
-wilcox.test(q_116_VAS_v7~q_423_beta,data = rdb_temp)
-rdb_temp %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_423_beta,q_116_VAS_v7))+
-  geom_violin()
-
-rdb_temp %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-group_by(q_423_beta) %>% 
-  summarise(medianVAS=median(q_116_VAS_v7,na.rm=T),
-            IQRVAS = IQR(q_116_VAS_v7,na.rm=T))
-
-wilcox.test(ANAscore~q_423_beta,data = rdb_temp %>% 
-  filter(d_elicitor_gr5=="insects") 
-)
-
-rdb_temp %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_423_beta,ANAscore))+
-  geom_violin()
-rdb_temp %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
- group_by(q_423_beta) %>% 
-  summarise(medianANAscore=median(ANAscore,na.rm=T),
-            IQRANAscore = IQR(ANAscore,na.rm=T))
-
-ggarrange(
-  plot_beta_cardiacs,
-  ggarrange(
-    rdb_temp %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_423_beta,q_116_VAS_v7))+
-  geom_boxplot(),
-rdb_temp %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-ggplot(aes(q_423_beta,ANAscore))+
-  geom_boxplot()
-  ),
-nrow =2,
-ncol = 1
-)
-
-```
-
-```{r cardiacmasto, fig.cap = "MAsotcytosis and cardiologic symptoms" }
-cardiac_masto_plots[[6]]+
-              labs(x = "cardiac arrest")
-```
-
-```{r tryp, fig.cap="Severity of anaphylaxis is influenced by baseline serum tryptase levels with the cut off of 8 ng/ml in IVA (top row) vs other anaphylaxis cases",fig.width = 7}
-ggarrange(
-rdb %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
-                  "low",
-                  "high")) %>% 
-  filter(!is.na(typ_cat)) %>% 
-  ggplot(aes(typ_cat,q_116_VAS_v7))+
-  geom_boxplot(),
-rdb %>% 
-  filter(d_elicitor_gr5=="insects") %>% 
-  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
-                  "low",
-                  "high")) %>% 
-  filter(!is.na(typ_cat)) %>% 
-  ggplot(aes(typ_cat,ANAscore))+
-  geom_boxplot(),
-rdb %>% 
-  filter(d_elicitor_gr5=="insects",
-         !is.na(d_severity_rm)) %>% 
-  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
-                  "low",
-                  "high"),
-         d_severity_rm = factor(d_severity_rm)) %>% 
-  group_by(d_severity_rm,typ_cat) %>% 
-  filter(!is.na(typ_cat)) %>% 
-    summarize(n = n()) %>% 
-  ggplot(aes(typ_cat,n,fill=d_severity_rm))+
-  geom_bar(stat="identity",position= "fill"),
-rdb %>%
-  filter(d_elicitor_gr5!="insects") %>%
-  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
-                  "low",
-                  "high")) %>%
-  filter(!is.na(typ_cat)) %>%
-  ggplot(aes(typ_cat,q_116_VAS_v7))+
-  geom_boxplot(),
-rdb %>%
-  filter(d_elicitor_gr5!="insects") %>%
-  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
-                  "low",
-                  "high")) %>%
-  filter(!is.na(typ_cat)) %>%
-  ggplot(aes(typ_cat,ANAscore))+
-  geom_boxplot(),
-rdb %>%
-  filter(d_elicitor_gr5!="insects",
-         !is.na(d_severity_rm)) %>%
-  mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
-                  "low",
-                  "high"),
-         d_severity_rm = factor(d_severity_rm)) %>%
-  group_by(d_severity_rm,typ_cat) %>%
-  filter(!is.na(typ_cat)) %>%
-    summarize(n = n()) %>%
-  ggplot(aes(typ_cat,n,fill=d_severity_rm))+
-  geom_bar(stat="identity",position= "fill"),
-ncol =3 , nrow = 2,
-widths = c(1,1.1,2))
-
-```
-
-```{r yjbee, fig.height=8, fig.width = 7, fig.cap="Models of severity and the influence of jellow jacket vs bees"}
-require(broom)
-
-ana_outcome_matched <- function(data = data4,
-         outcome = "d_severity_rmr",
-         matching_var = "q_340_insects",
-         confounders_cat = c("b_sex", "cc_cardio"),
-         confounders_int = c("d_age"),
-         selected_predictors = c(1:4)
-         ){
-  possible_outcomes <- c("d_severity_rm",
-                         "d_severity_rmr",
-                         "severity_brown",
-                         "q_116_VAS_v7",
-                         "ANAscore")
-  predictors <- c(matching_var, confounders_cat,confounders_int)
-  df <- data %>% # first get the data frame
-   filter(!!as.name(matching_var) %in% c("yellow jacket", "bee")) %>% 
-   select(!!!matching_var,
-          !!!outcome,
-          !!!confounders_cat,
-          !!!confounders_int
-          ) %>%
-          {.[complete.cases(.),]}
-  ### we need to match the cases now to the first level of theoutcome variable
-  ref_level <- levels(df[,matching_var] %>% droplevels())
-  matched <- df %>% 
-    mutate(match = ifelse(!!as.name(matching_var)==ref_level[1],T,F),
-           out=ifelse(!!as.name(outcome)=="severe",1,0)) %>% 
-    matchit(formula = as.formula(
-      paste("match ~ ",paste0(confounders_cat),"+", paste0(confounders_int,collapse = " + ")))) %>% 
-    match.data() 
-  ## IFELSE
-  if(outcome %in% possible_outcomes[c(2,3)]){
-    plot <- matched %>% 
-      ggplot(aes(!!as.name(matching_var),
-               fill = !!as.name(outcome)))+
-      geom_bar()+
-      facet_grid(reformulate(confounders_cat[1],confounders_cat[2]))
-
-    fit <- glm(as.formula(paste0(
-      "out ~ ",paste0(predictors[selected_predictors], collapse = " + ")
-    )),data= matched,family= "binomial")
-  } else if (outcome %in% possible_outcomes[c(1,4,5)]){
-     plot <- matched %>% 
-      ggplot(aes(!!as.name(matching_var),
-               y = !!as.name(outcome)))+
-      geom_boxplot()+
-      facet_grid(reformulate(confounders_cat[1],confounders_cat[2]))
-
-    fit <- glm(as.formula(paste0(
-      outcome, " ~ ", paste0(predictors[selected_predictors], collapse = " + "))),
-             data = matched)
-  }
-
-  return(list(mtched_df = matched,
-              ref_level = ref_level,
-              fit = fit,
-              plot = plot)) 
-}
-
-possible_outcomes <- c("d_severity_rm",
-                         "d_severity_rmr",
-                         "severity_brown",
-                         "q_116_VAS_v7",
-                         "ANAscore")
-
-output_severity <- lapply(possible_outcomes,
-                         function(x){
-                         ana_outcome_matched(data = data4,
-                            outcome = x,
-                            matching_var = "q_340_insects",
-                            confounders_cat = c("b_sex", "cc_cardio"),
-                            confounders_int = c("d_age"),
-                            selected_predictors = c(1:4))
-                           }
-)
-
-do_cof_tab <- function(x){
-  output_severity[[x]]$fit %>% summary() %>% {.$coefficients}%>% tidy() %>% as.matrix()
-}
-require(broom)
-temp <- lapply(1:5,do_cof_tab) %>% do.call(what = rbind) %>%
-                          as_tibble() %>%
-  {mutate(.,
-         Estimate = as.numeric(Estimate),
-         Pr...t.. = as.numeric(Pr...t..)
-  )} %>%
-          {cbind(.,outcome = lapply(possible_outcomes,
-                          rep,
-                          nrow(.)/5) %>% unlist)
-          } %>% 
-           filter(.rownames!="(Intercept)")
-gridExtra::grid.arrange(output_severity[[2]]$plot,
-                        output_severity[[3]]$plot,
-                        output_severity[[1]]$plot,
-                        output_severity[[4]]$plot,
-                        output_severity[[5]]$plot,
-                        temp %>% 
-  ggplot(aes(.rownames,Pr...t.., color=outcome))+
-  geom_segment(x=0.5,xend=4.5,y=log10(0.05),yend=log10(0.05))+
-  scale_y_log10()+
-  geom_jitter()+
-  theme(axis.text.x = element_text(angle=40, hjust=1))+
-    annotate("text",x=4,y = c(log10(c(0.0001,0.00001,.000001,.0000001,.00000001))),
-             label = c(1776,1780,1206,234,1780))
-  )#+
-  #geom_line(aes(group=outcome))
-```
-
-```{r modelfitness, fig.cap="Comparison of modelling to different severity scales."}
-
-ggplot(temp,aes(.rownames,Pr...t.., color=outcome))+
-  geom_segment(x=0.5,xend=4.5,y=log10(0.05),yend=log10(0.05))+
-  scale_y_log10()+
-  geom_jitter(width = 0.2)+
-  theme(axis.text.x = element_text(angle=40, hjust=1))+
-    annotate("text",x=4,y = c(0.0001,0.00001,.000001,.0000001,.00000001)*10,
-             label = c("n=1776","n=1780","n=1206","n=234","n=1780"),size = 3)+
-  labs(x = "",y="p value")
-```
-
-```{r tryp2,fig.height=8, fig.width = 7, fig.cap = "tryptase cut off value plot"}
-output_severity <- lapply(possible_outcomes,
-                         function(x){
-                         ana_outcome_matched(data = data4,
-                            outcome = x,
-                            matching_var = "q_340_insects",
-                            confounders_cat = c("tryp_cat", "cc_cardio"),
-                            confounders_int = c("d_age"),
-                            selected_predictors = c(1,2,4))
-                           }
-)
-
-do_cof_tab <- function(x){
-  output_severity[[x]]$fit %>% summary() %>% {.$coefficients}%>% tidy() %>% as.matrix()
-}
-
-temp <- lapply(1:5,do_cof_tab) %>% do.call(what = rbind) %>%
-                          as_tibble() %>%
-  {mutate(.,
-         Estimate = as.numeric(Estimate),
-         Pr...t.. = as.numeric(Pr...t..)
-  )} %>%
-          {cbind(.,outcome = lapply(possible_outcomes,
-                          rep,
-                          nrow(.)/5) %>% unlist)
-          } %>% 
-           filter(.rownames!="(Intercept)")
-gridExtra::grid.arrange(output_severity[[2]]$plot,
-                        output_severity[[3]]$plot,
-                        output_severity[[1]]$plot,
-                        output_severity[[4]]$plot,
-                        output_severity[[5]]$plot,
-                        temp %>% 
-  ggplot(aes(.rownames,Pr...t.., color=outcome))+
-  geom_segment(x=0.5,xend=4.5,y=log10(0.05),yend=log10(0.05))+
-  scale_y_log10()+
-  geom_jitter()+
-  theme(axis.text.x = element_text(angle=40, hjust=1))
-  )
-```
-
-```{r rfplot}
-rf_ggplot_treatment
-```

### 2019-07-23 R/modelfitness.R

@@ -1,0 +1,10 @@
+modelfiteness <- function(){
+  ggplot(temp,aes(.rownames,Pr...t.., color=outcome))+
+  geom_segment(x=0.5,xend=4.5,y=log10(0.05),yend=log10(0.05))+
+  scale_y_log10()+
+  geom_jitter(width = 0.2)+
+  theme(axis.text.x = element_text(angle=40, hjust=1))+
+  annotate("text",x=4,y = c(0.0001,0.00001,.000001,.0000001,.00000001)*10,
+           label = c("n=1776","n=1780","n=1206","n=234","n=1780"),size = 3)+
+  labs(x = "",y="p value")
+}

### 2019-07-23 R/plot_severity_model.R

@@ -1,0 +1,41 @@
+plot_severity <- function(){
+  output_severity <- lapply(possible_outcomes,
+                            function(x){
+                              ana_outcome_matched(data = data4,
+                                                  outcome = x,
+                                                  matching_var = "q_340_insects",
+                                                  confounders_cat = c("tryp_cat", "cc_cardio"),
+                                                  confounders_int = c("d_age"),
+                                                  selected_predictors = c(1,2,4))
+                            }
+  )
+
+
+  do_cof_tab <- function(x){
+    output_severity[[x]]$fit %>% summary() %>% {.$coefficients}%>% tidy() %>% as.matrix()
+  }
+
+  temp <- lapply(1:5,do_cof_tab) %>% do.call(what = rbind) %>%
+    as_tibble() %>%
+    {mutate(.,
+            Estimate = as.numeric(Estimate),
+            Pr...t.. = as.numeric(Pr...t..)
+    )} %>%
+    {cbind(.,outcome = lapply(possible_outcomes,
+                              rep,
+                              nrow(.)/5) %>% unlist)
+    } %>%
+    filter(.rownames!="(Intercept)")
+  gridExtra::grid.arrange(output_severity[[2]]$plot,
+                          output_severity[[3]]$plot,
+                          output_severity[[1]]$plot,
+                          output_severity[[4]]$plot,
+                          output_severity[[5]]$plot,
+                          temp %>%
+                            ggplot(aes(.rownames,Pr...t.., color=outcome))+
+                            geom_segment(x=0.5,xend=4.5,y=log10(0.05),yend=log10(0.05))+
+                            scale_y_log10()+
+                            geom_jitter()+
+                            theme(axis.text.x = element_text(angle=40, hjust=1))
+  )
+}

### 2019-07-23 R/plot_severity_tryptase.R

@@ -1,0 +1,61 @@
+plot_severity_tryptase <- function(){
+  ggarrange(
+    rdb %>%
+      filter(d_elicitor_gr5=="insects") %>%
+      mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                            "low",
+                            "high")) %>%
+      filter(!is.na(typ_cat)) %>%
+      ggplot(aes(typ_cat,q_116_VAS_v7))+
+      geom_boxplot(),
+    rdb %>%
+      filter(d_elicitor_gr5=="insects") %>%
+      mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                            "low",
+                            "high")) %>%
+      filter(!is.na(typ_cat)) %>%
+      ggplot(aes(typ_cat,ANAscore))+
+      geom_boxplot(),
+    rdb %>%
+      filter(d_elicitor_gr5=="insects",
+             !is.na(d_severity_rm)) %>%
+      mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                            "low",
+                            "high"),
+             d_severity_rm = factor(d_severity_rm)) %>%
+      group_by(d_severity_rm,typ_cat) %>%
+      filter(!is.na(typ_cat)) %>%
+      summarize(n = n()) %>%
+      ggplot(aes(typ_cat,n,fill=d_severity_rm))+
+      geom_bar(stat="identity",position= "fill"),
+    rdb %>%
+      filter(d_elicitor_gr5!="insects") %>%
+      mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                            "low",
+                            "high")) %>%
+      filter(!is.na(typ_cat)) %>%
+      ggplot(aes(typ_cat,q_116_VAS_v7))+
+      geom_boxplot(),
+    rdb %>%
+      filter(d_elicitor_gr5!="insects") %>%
+      mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                            "low",
+                            "high")) %>%
+      filter(!is.na(typ_cat)) %>%
+      ggplot(aes(typ_cat,ANAscore))+
+      geom_boxplot(),
+    rdb %>%
+      filter(d_elicitor_gr5!="insects",
+             !is.na(d_severity_rm)) %>%
+      mutate(typ_cat=ifelse(q_212_tryptase_value_v5<8,
+                            "low",
[…]

### 2019-07-23 R/stat.R

@@ -44,6 +44,8 @@ manual_greens <- rev(c("#E5F5E0",
 ####### General Functions #############
 source("R/functions.R")
 
+palBW <- c("#848484","#1f1f1f")
+
 #### Data cleaning ####
 data4 %<>% correctLabels1()
 
@@ -494,6 +496,20 @@ hypotensionPlot <-
          legend.position = "top",
          legend.title = element_blank())
 
+hypotension2 <- ex %>%
+  filter(variableName =="q_114_hypotension_collapse_v5") %>%
+  tidyr::gather(key = "group",
+                value = "Fraction",
+                c("fraq_1",
+                  "fraq_2")) %>%
+  mutate(group = car::recode(group,
+                             "'fraq_1'='IVA';
+                   'fraq_2'='non-IVA'")) %>%
+  ggbarplot(x = "subset",y = "Fraction",
+            fill = "group",
+            position = position_dodge(),
+            palette = rev(palBW))+
+  labs(x = "hypotension", y = "proportion [%]", fill = "")
 
 
 #### Mastocytose patienten  as a subroup #####
@@ -1023,8 +1039,8 @@ lower_panel <- ggpubr::ggarrange(
                  linetype = 2),
 
   ANAscore_matched %>%
-    group_by(b_sex,grouping) %>%
-    ggplot(aes(fill=grouping,x=b_sex))+
+    group_by(b_sex, grouping) %>%
+    ggplot(aes(fill=grouping, x=b_sex))+
     geom_bar(position = "fill")+
     theme_classic()+
     theme(legend.position = "none")+
@@ -2687,3 +2703,132 @@ yj_bee <- rdb %>%
            grouping) %>%
   summarize(n = n())
 
+
+#### Plotting stings among years ####
+
+data %>%
+  filter(d_elicitor_gr5=="insects") %>%
+  mutate(MOR = substr(b_reactiondate,4,5)) %>%
+  mutate(MOR = car::recode(MOR,
+           "'01' = 'Jan';
+                           '02' = 'Feb';
+                           '03' = 'Mar';
+                           '04' = 'Apr';
+                           '05' = 'May';
+                           '06' = 'Jun';
+                           '07' = 'Jul';
+                           '08' = 'Aug';
+                           '09' = 'Sep';
[…]

### 2019-07-24 R/stat.R
fig 1 change to sex and age matched

@@ -1027,7 +1027,7 @@ align = "h",
 labels = c("A","B"))
 
 lower_panel <- ggpubr::ggarrange(
-  ANAscore_matched  %>%
+    age_  %>%
     ggplot(aes(x = d_age, fill = grouping)) +
     geom_density(alpha = 0.5)+
     labs(x = "Age [years]",fill = "elicitor")+

### 2019-07-24 analysis/paper/paper.Rmd
references rformat

@@
 bibliography: references.bib
-csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
+csl: "../templates/expert-review-of-clinical-immunology.csl" # Insert path for the bib-style
 ---
@@
 
-**Highlights:**
-  These are the highlights. 
+<!-- **Highlights:** -->
+<!--   These are the highlights.  -->
 **Document statistics:**
@@
 Mastocytosis was increasing the risk of cardiac arrest in patients undergoing IVA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
-S\@ref(fig:cardiacmasto)).
+S\@ref(fig:supplement)).
 
@@
 Based on the severity and symptom profile, we decided to use a tryptase cut off value of 8 ng/ml (Fig.
-S\@ref(fig:tryp2)) rather than the currently used 11.5 ng/ml.
+S\@ref(fig:figcofactors)B) rather than the currently used 11.5 ng/ml.
 

### 2019-07-24 R/stat.R

@@ -1027,7 +1027,7 @@ align = "h",
 labels = c("A","B"))
 
 lower_panel <- ggpubr::ggarrange(
-    age_  %>%
+    age_sex_matched  %>%
     ggplot(aes(x = d_age, fill = grouping)) +
     geom_density(alpha = 0.5)+
     labs(x = "Age [years]",fill = "elicitor")+
@@ -1038,7 +1038,7 @@ lower_panel <- ggpubr::ggarrange(
     geom_segment(aes(x = 22, xend =22, y = 0, yend = 0.02),
                  linetype = 2),
 
-  ANAscore_matched %>%
+  age_sex_matched %>%
     group_by(b_sex, grouping) %>%
     ggplot(aes(fill=grouping, x=b_sex))+
     geom_bar(position = "fill")+

### 2019-07-24 analysis/paper/paper.Rmd
printout prof worm

@@
 
-The data show that `r {{rdb$q_340_insects=="yellow jacket"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`% of all insect elicited cases were elicited by yellow jackets, followed by bees (`r {{rdb$q_340_insects=="bee"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`%).
+The data showed that `r {{rdb$q_340_insects=="yellow jacket"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`% of all insect elicited cases were elicited by yellow jackets, followed by bees (`r {{rdb$q_340_insects=="bee"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`%).
 The insect venom elicited cases occurred mostly in outdoor places (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(4),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%) patients' homes (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(8),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%) or urban places (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(5),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%).
@@
 
-Recent expert position on the future anaphylaxis research indicates the need for a more precise description of diagnosis, biomarkers, and phenotypes of anaphylaxis [@JimenezRodriguez2018].
+Recent expert position on the future of anaphylaxis research indicates the need for a more precise description of diagnosis, biomarkers, and phenotypes of anaphylaxis [@JimenezRodriguez2018].
 Nevertheless, in order to facilitate precision-medicine approach [@Muraro2017] in the treatment of anaphylaxis - we first need to investigate its subtypes and pathomechanisms better.
@@
 
-This study aimed to identify current patterns of Insect venom anaphylaxis (IVA) in symptoms, cofactors, and management and compare them to other types of anaphylaxis (non-IVA) in a case-control manner.
+This study aimed to identify current patterns of insect venom anaphylaxis (IVA) in regard to symptoms, cofactors, and management and compare them to other types of anaphylaxis (non-IVA) in a case-control manner.
 
@@
 The cases were either assigned to IVA group (if the triggering factor was an insect sting) or non-IVA group (other elicitors of anaphylaxis).
-We compared the frequency of various symptoms, cofactors, known to increase the risk of severe anaphylaxis, [@Worm2018] and management in both groups.
+We compared the frequency of various symptoms, cofactors --- known to increase the risk of severe anaphylaxis, [@Worm2018] and management in both groups.
 
@@
 The results of the propensity score matching are in Fig.
-\ref(fig:flow)B-D.
+\@ref(fig:flow)B-D.
 
-We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find which therapy approaches were best fit for classifying cases either to IVA or non-IVA group and presented the results as mean decrease in Gini.
+We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapy approaches that varied the most between IVA / non-IVA group and presented the results Gini importance [@Strobl2007].
 We also performed an association study of therapeutic interventions and symptoms.
-The resulting phi values are presented in a heatmap with automatic clustering using Ward's D and euclidean distancing method [@heatmaply].
+The resulting phi values are presented in a heatmap with automatic clustering using Ward’s Agglomerative Hierarchical Clustering Method with Euclidean distance [@heatmaply].
 
@@
 IVA showed a specific symptom pattern.
-Patients, who underwent IVA, more often experienced cardiologic symptoms (dizziness, reduced alertness, unconsciousness) than in other elicitors of anaphylaxis and less often showed gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A).
+Patients, who underwent IVA, more often experienced cardiologic symptoms (dizziness, reduced alertness, unconsciousness) than in other elicitors of anaphylaxis and less often showed skin or gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A).
 The difference in severe hypotension frequency was especially prominent in the younger group under 22 (Fig.
@@
 other elicitors (Fig.
-\@ref(fig:sev)D).
-There were no differences in severity of reactions elicited by yellow-jackets and insect species.
+S\@ref(fig:sev)D).
+There were no differences in severity of reactions elicited by yellow-jackets and other insect species.
 
@@
 We observed `r sum(repeated_reaction_in_Reg)` patients with two documented reactions in our registry.
-Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had Insect elicited anaphylaxis and in `r greater_same[1,2]` (`r greater_same[1,2]/sum(greater_same[,2])*100`%) the following reaction was more severe than before.
+Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had Insect elicited anaphylaxis and in `r greater_same[1,2]` (`r (greater_same[1,2]/sum(greater_same[,2])*100 )%>%round(1) `%) the following reaction was more severe than before.
 In `r greater_same[3,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was graded exactly as before.

### 2019-09-09 R/temp.Rmd
Draft v2.0

---
title: "MWD"
output: bookdown::word_document2
---

# Results
This text refers to Fig. \@ref(fig:fig1main). 
We also want to refere here to Fig. \@ref(fig:fig2main).

In some cases we also need supplementary data. Please see Suppl. Fig. S\@ref(fig:fig1supp).

Please note that the 'S' before the reference should optimally NOT be there and ideally one should write:

```
Fig. \@ref(fig:fig1supp) 
```

what would evaluate to Fig. S1.

# Figures

```{r fig1main, fig.cap="First Main Figure"}
plot(1)
```

```{r fig2main, fig.cap="Second Main Figure"}
plot(1)
```

# Supplementary data

```{r fig1supp, fig.cap="This is a supplementary figure and it should be called Fig. S1 and not Fig 3."}
plot(1)
```

### 2019-09-09 analysis/paper/paper.Rmd
Draft v2.0

@@
 ---
-title: "Insect venom anaphylaxis: a case - control study of the European Anaphylaxis Registry"
+title: "Phenotypes and risk assesment of insect venom anaphylaxis: a case - control study of the European Anaphylaxis Registry"
 author:
@@
 **Keywords:**
-  anaphylaxis, adrenaline (epinephrine), beta-blockers, insect venom allergy, yellow-jacket
+  anaphylaxis, adrenaline (epinephrine), beta-blockers, insect venom allergy, Hymenoptera
 
@@
   Word count, figures, tables, references
-
-<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->
 
@@
 Insect-venom elicited anaphylaxis is a common hypersensitivity reaction which may be life-threatening.
-Using the data from the European Anaphylaxis Registry (`r nrow(data)` cases in total) we identified insect-venom elicited anaphylaxis cases (n = `r {data$d_elicitor_gr5=="insects"} %>% sum()`) and analyzed these in comparison to anaphylaxis elicited by other elicitors (n = `r {data$d_elicitor_gr5!="insects"} %>% sum()`).
+Using data from the European Anaphylaxis Registry (`r nrow(data)` cases) we identified `r {data$d_elicitor_gr5=="insects"} %>% sum()` with insect-venom elicited anaphylaxis and analyzed these in comparison to anaphylaxis triggered by other elicitors (n = `r {data$d_elicitor_gr5!="insects"} %>% sum()`).
 
-The data showed that `r {{rdb$q_340_insects=="yellow jacket"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`% of all insect elicited cases were elicited by yellow jackets, followed by bees (`r {{rdb$q_340_insects=="bee"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`%).
-The insect venom elicited cases occurred mostly in outdoor places (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(4),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%) patients' homes (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(8),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%) or urban places (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(5),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%).
+The data show that `r {{rdb$q_340_insects=="yellow jacket"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`% of all insect elicited cases were elicited by yellow jackets, followed by bees (`r {{rdb$q_340_insects=="bee"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`%).
+The venom triggered cases occurred mostly in outdoor places (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(4),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%) patients' homes (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(8),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%) or urban places (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(5),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%).
 
-Skin, gastrointestinal and respiratory symptoms occurred less frequently in insect elicited cases of anaphylaxis, whereas cardiologic symptoms (with hypotension, collapse, and loss of consciousness) were more frequent.
+Regarding the symptoms: skin, gastrointestinal and respiratory manifestations occurred less frequently in insect induced cases of anaphylaxis, but cardiovascular symptoms (with hypotension, collapse, and loss of consciousness) were more frequent.
 Intramuscular adrenaline (as a first-line therapy) was administered significantly less often in insect venom elicited cases (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(d_522_adren_agg) %>% summarize(n=n()) %>% {.[2,2]/(.[1,2])*100} %>% round(1)`% vs `r rdb %>% filter(is.na(q_340_insects)) %>% group_by(d_522_adren_agg) %>% summarize(n=n()) %>% {.[2,2]/(.[1,2])*100} %>% round(1)`, p < 0.0001).
@@
 
-Patients who experienced insect-venom anaphylaxis were older (p < 0.0001), more often had concomitant mastocytosis (p < 0.0001) and cardiologic conditions (p < 0.0001) and females more often had concomitant thyroid diseases and less often suffered from a food allergy or atopic dermatitis.
+Patients who experienced insect-venom anaphylaxis were older than patients with anaphylaxis elicited by other elicitors (p < 0.0001), more often had concomitant mastocytosis (p < 0.0001) and cardiologic conditions (p < 0.0001) and females more often had concomitant thyroid diseases and less often suffered from a food allergy or atopic dermatitis.
 
-Symptoms of insect venom anaphylaxis are distinctively different from other anaphylactic reactions.
-What made us consider the therapy of insect elicited cases of anaphylaxis separately.
-Indeed we observed different therapeutic patterns in insect elicited cases of anaphylaxis (more antihistaminics but fewer corticosteroids, bronchodilators, and surprisingly - adrenaline).
-The presented evidence indicates that the management of insect-venom anaphylaxis may be improved.
+ <!-- MORE DISCUSSION!!!!!-->
+The data show that the management of insect-venom anaphylaxis may be improved.
 Patients with concomitant cardiologic conditions and these with hyperreactive mast cells require intensified prophylactic measures.
@@
 
-Hypersensitivity to insect venom may manifest as a systemic reaction (anaphylaxis) in up to 0.3-7.5% of the adult population [@Bil2008].
+Hypersensitivity to insect venom can manifest as a systemic reaction (anaphylaxis) in up to 0.3-7.5% of the adult population [@Bil2008].
 Insect venom anaphylaxis (IVA) can be fatal, and patients sometimes require lifelong systemic immunotherapy [@Sturm2017].
 
-Recent expert position on the future of anaphylaxis research indicates the need for a more precise description of diagnosis, biomarkers, and phenotypes of anaphylaxis [@JimenezRodriguez2018].
-Nevertheless, in order to facilitate precision-medicine approach [@Muraro2017] in the treatment of anaphylaxis - we first need to investigate its subtypes and pathomechanisms better.
+Recent expert position on the future of anaphylaxis research indicates the need for a more precise description of the diagnosis, biomarkers, and phenotypes of anaphylaxis [@JimenezRodriguez2018].
+Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] in the treatment of anaphylaxis - it is first needed to study its clinical subtypes and pathomechanisms in more detail.
 
@@
 
-This study aimed to identify current patterns of insect venom anaphylaxis (IVA) in regard to symptoms, cofactors, and management and compare them to other types of anaphylaxis (non-IVA) in a case-control manner.
+This study aimed to identify current patterns of insect venom anaphylaxis (IVA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-IVA).
 
@@
 
-We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapy approaches that varied the most between IVA / non-IVA group and presented the results Gini importance [@Strobl2007].
-We also performed an association study of therapeutic interventions and symptoms.
-The resulting phi values are presented in a heatmap with automatic clustering using Ward’s Agglomerative Hierarchical Clustering Method with Euclidean distance [@heatmaply].
-
 <!-- ## Data handling and statistics -->
-The statistical analysis was performed in the R Statistical Package [@R].
+We used the R Statistical Package [@R] for statistical analysis.
 A simple comparison of categorical variables was performed using either Chi^2^ test or Fisher's exact test (where the number of observation in a bin was less than 10), continuous variables were analyzed using Mann-Whitney U test.
@@
 
+We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapeutic approaches that varied the most between IVA / non-IVA group and presented the results Gini importance [@Strobl2007].
+We also performed an association study of therapeutic interventions and symptoms.
+The resulting phi values are presented in a heatmap with automatic clustering using Ward’s Agglomerative Hierarchical Clustering Method with Euclidean distance [@heatmaply].
+
 # Results
 
-## Perennial distribution of anaphylaxis cases by elicitor.
+## Hymenoptera anaphylaxis shows seasonal patterns.
 
-Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was frequently reported from May till October.
-Their proportion of IVA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was as under 1% of cases during winter.
-Nevertheless, `r rare_spring_autumn$n %>% sum` cases of IVA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were also reported in March, April, and November.
+Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
+Their proportion of IVA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was as below 1% of cases during winter.
+Nevertheless, `r rare_spring_autumn$n %>% sum` cases of IVA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were triggered in March, April, and November.
 Yellow-jacket was the most prominent IVA-causing insect followed by bees.
-The IVA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in the south of Europe.
+The IVA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in southern Europe.
 
-IVA was more frequent in adults and seniors when compared to children and young adults.
+IVA was more frequent in adults and seniors than children and young adults.
 (Fig \@ref(fig:insectstime)B).
-When we looked at the density plot of IVA cases according to age - we saw a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B.
+When we analyzed a density plot of IVA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B.
 
-## Symptoms
+## IVA is associated with cardiovascular symptoms
+
 IVA showed a specific symptom pattern.
-Patients, who underwent IVA, more often experienced cardiologic symptoms (dizziness, reduced alertness, unconsciousness) than in other elicitors of anaphylaxis and less often showed skin or gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A).
-The difference in severe hypotension frequency was especially prominent in the younger group under 22 (Fig.
+Patients, who underwent IVA, more often experienced cardiologic symptoms (dizziness, reduced alertness, unconsciousness) than patients with other elicitors inducing anaphylaxis and less often had skin or gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A).
+The difference in severe hypotension frequency was especially prominent in the younger age group (under 22, Fig.
 S\@ref(fig:supplement)).
@@
 
-Severe reactions were more prevalent in older patients in comparison to patients under 22, and in IVA cases vs.
+Severe reactions of IVA were more prevalent in older patients in comparison to patients below 22 years, and in IVA cases vs.
 other elicitors (Fig.
@@
 
-## Repeated reactions
+## Absence of skin symptoms is associated with more severe episodes of IVA
+
+Patients with concomitant mastocytosis had significantly more often anaphylaxis without skin symptoms (namely urticaria and flushing) when compared to patients without diagnosed mastocytosis
+(`r (skin_symptoms$masto_tab$prop[c(2,1)]*100) %>% round(1) %>% paste0(collapse="% vs.
+")`%, `r pval(skin_symptoms$masto_chisq$p.value)`).
+We decided to further investigate cases where mastocytosis was not diagnosed.
+In these patients urticaria or flushing as symptoms of anaphylaxis were present more often in IVA than anaphylaxis triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
+Morover, in patients without concomitant mastocytosis who were also lacking skin symptoms, anaphylaxis was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(2,6)]*100) %>% round(1) %>% paste0(collapse="% in IVA vs.
+")`%, `r pval(skin_symptoms$masto_chisq$p.value)`% in non-IVA).
+We performed a factorial logistic regression modelling and confirmed the significant interaction effect between the presence of skin symptoms and the trigger (eiter IVA or non-IVA) on the severity of anaphylaxis (`r pval(skin_symptoms$severity_fit_skin_symptoms %>% coef %>% .[4,4])`).
+Therefore, non-mastocytic patients presenting without urticara or flushing tended to have more severe anaphylaxis when it was triggered by insects, but not other elicitors of anaphylaxis (Fig.
+\@ref(fig:figsymptoms)C).
+
+Following this observation, we decided to investigate the association of skin symptoms with the tryptase levels in patients who did not have a diagnosis of mastocytosis.
+For this comparison, we removed the cases with BST above 11.5 ng/ml, potentially indicating non-diagnosed mastocytosis.
+Similarly, tryptase levels correlated with the severity of anaphylaxis, were higher in IVA patients, and the effect was prominently visible for IVA patients (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`).
+We did not observe such an interaction in the non-IVA group.
+
+## Nearly one-third of IVA patients experience repeated reactions
 
@@
 
-In general `r treact[1,2] %>% round(1)`% of patients with insect allergy had experienced anaphylaxis in the past.
+In general `r treact[1,2] %>% round(1)`% of patients with insect allergy had experienced venom anaphylaxis in the past.
 If the reaction was elicited by other elcitors previous reactions were more frequently seen (`r treact[2,2]%>% round(1)`%, `r p %>% pval()`).
 We observed `r sum(repeated_reaction_in_Reg)` patients with two documented reactions in our registry.
-Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had Insect elicited anaphylaxis and in `r greater_same[1,2]` (`r (greater_same[1,2]/sum(greater_same[,2])*100 )%>%round(1) `%) the following reaction was more severe than before.
-In `r greater_same[3,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was graded exactly as before.
+Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had insect elicited anaphylaxis and in `r greater_same[1,2]` (`r (greater_same[1,2]/sum(greater_same[,2])*100 )%>%round(1) `%) the following reaction was more severe than before.
+In `r greater_same[3,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was similar in severity.
 
-## Co-factors
+## Increased baseline serum tryptase and concomitant cardiologic conditions increase the risk of severe IVA.
 
@@
 \@ref(fig:figcofactors)).
-Mastocytosis was increasing the risk of cardiac arrest in patients undergoing IVA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
+However, mastocytosis increased the risk of cardiac arrest in patients undergoing IVA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
 S\@ref(fig:supplement)).
 
-Concomitant cardiovascular diseases were associated with a higher risk of severe anaphylaxis when elicited by insects.
-They played an insignificant role in non-IVA cases (Fig.
+Concomitant diseases were more prevalent in IVA than in non-IVA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare()`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were irrelevant in non-IVA cases (Fig.
 \@ref(fig:figcofactors)).
+Interestingly, BST values were increased in patients with concomitant cardiovasular diseases, irrespectively of the reaction severity (Fig.
+S\@ref(trypcardiocur)).
 
-Risk of severe anaphylaxis in patients concomitantly using ACE-I (as well as beta-blockers) could not be independently measured due to coexisting cardiovascular pathologies.
-ACE-I use was, however, more often associated with cardiac arrests in both IVA and anaphylaxis by other triggers (Fig.
-\@ref(fig:figcofactors)C) whereas beta-blockers use was associated with higher severity of anaphylaxis and with cardiovascular symptoms (cardiac arrest, chest pain).
+The risk of severe anaphylaxis in patients concomitantly using ACE-I (as well as beta-blockers) could not be independently measured due to coexisting cardiovascular pathologies.
+ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) and there were no differences between IVA nad non-IVA  (Fig.
+\@ref(fig:figcofactors)C).
+Beta-blocker use was generally associated with higher severity of anaphylaxis and with the onset of cardiovascular symptoms (cardiac arrest, chest pain), but we saw no difference between groups `r summary(beta_fit)[["coefficients"]][3,4] %>% pval()`).
 Surprisingly, arrhythmia was more frequently reported in patients undergoing IVA with concomitant beta-blockers (Fig.
-\@ref(fig:figcofactors)D).
+\@ref(fig:figcofactors)B).
 
-Baseline serum tryptase correlated with the severity of reactions in the Ring and Messmer scale and was more prominently increasing risk of severe anaphylaxis in IVA patients than in other cases (Fig.
-\@ref(fig:figcofactors)B).
-Cases with cardiac arrest were associated with increased tryptase over eight ng/ml, and this proportion was higher in IVA when compared to other elicitors.
-Loss of consciousness was also associated with increased tryptase levels but only in patients with IVA (Fig.
+Baseline serum tryptase (BST) correlated with the severity of reactions (Ring and Messmer scale).
+Moreover, elevated BST was more prominently increasing the risk of severe anaphylaxis in IVA patients than in other cases (Fig.
+\@ref(fig:figcofactors)C).
+
+Cases with cardiac arrest were associated with an increased tryptase above 8 ng/ml, and this proportion was higher in IVA when compared to other elicitors.
+Loss of consciousness was a symptom associated with increased tryptase levels but only in patients with IVA (Fig.
 \@ref(fig:figcofactors)C).
 Based on the severity and symptom profile, we decided to use a tryptase cut off value of 8 ng/ml (Fig.
-S\@ref(fig:figcofactors)B) rather than the currently used 11.5 ng/ml.
+\@ref(fig:figcofactors)B) rather than the currently used 11.5 ng/ml.
 
@@
 
-Physical exercise in the fresh air (e.g., jogging in the park) predisposed to anaphylaxis due to insect stings (`r ttexercise1`).
-However, even though the physical exercise was more often associated with IVA than other triggers of anaphylaxis - it was not predictive of the severity of a reaction in these patients, as there were no differences in severity concerning accompanying exercise (`r ttexercise2`).
+Outdoor physical exercise (e.g., jogging in the park) was more often associated with IVA than other triggers of anaphylaxis (`r ttexercise1`).
+However, it was not associated with the severity of a reaction in these patients.
+(`r ttexercise2`).
 
-## Management
+## Interplay of symptoms and medication
 
@@
 <!-- Antihistaminic drugs, on the other hand, were given more often in cases of IVA when compared to non-IVA. -->
-Patients with IVA received corticosteroids and antihistamines significantly more frequently than in other anaphylaxis cases.
+Patients with IVA received corticosteroids and antihistamines significantly more frequently than patients with anaphylaxis to other elicitors.
 On the other hand, adrenaline, beta-2 mimetics, and oxygen were given more often in non-IVA.
 
-We discovered clusters of associated symptoms and therapy modes in IVA patients (association measured using phi).
+We discovered clusters of symptoms and therapy modes in IVA patients (association measured using phi).
 Cardiologic symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria, were treated more similarly than respiratory or gastrointestinal symptoms (Fig.
 \@ref(fig:adrenalineuse)B).
-Therapy of these symptoms consisted of adrenaline autoinjector (AAI), i.v.
+The treatment of these symptoms consisted of using adrenaline autoinjector (AAI), i.v.
 adrenaline in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists.
@@
 
-The most differences in the therapy of IVA vs. other forms of anaphylaxis was observed in the frequency of inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
+The most noticeable differences in the therapy of IVA vs. other forms of anaphylaxis were observed in the frequency of inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
 
@@
 However, when we restricted the analysis to mild anaphylaxis cases - non-IVA patients received adrenaline more frequently than IVA (`r adren_severity_mild`).
+
+<!-- ## Time of the symptom onset  -->
+
+<!-- # Model for severity prediction -->
+<!-- As a last attempt for the analysis - perform modeling with the use of random forest or regression using the following inputs: Absence of urticaria, BST, short time for the onset of symptoms, older age, male sex?
+Cardiologic conditions.
+-->
 
@@
 
-Surprisingly, detailed information on the seasonality of insect-elicited hypersensitivity reactions is scarce [@old].
+<!-- DONE Top view findings -->
+
+In this study, we identified the prolonged seasonality of IVA, its clinical symptom-profile, and treatment patterns.
+The data unravel phenotypes of IVA, which may contribute to the development of tools incorporating both clinical and biomarker data for predicting the severity of future episodes of anaphylaxis.
+
+<!-- DONE Seasonality -->
+
+Based on our findings, insects are the most probable anaphylaxis elicitor in Europe during Summer-season, with IVA cases extending from early Spring to the end of Autumn.
+Detailed information on the seasonality of insect-elicited hypersensitivity reactions is scarce [@old].
 The activity of *Vespula germanica* depends on the climate, and in invaded regions (e.i.
-Australia) it can even extend throughout the whole year [@Spradbery_1992].
-The changing climate in Europe may influence the activity of Hymenoptera in this region in the upcoming years, but in the period from 2007 - 2019 the perennial distribution of IVA has remained stable (data not shown).
+Australia) it can even extend throughout the year [@Spradbery_1992].
+The changing climate in Europe may influence the activity of Hymenoptera in this region in the upcoming years.
+However, in the period from 2007 - 2019, the perennial ratio of IVA to non-IVA cases has remained unchanged (data not shown).
 
-We saw differences in IVA frequency in age groups.
-Especially at the age extremes, IVA was less frequent, and the density distribution showed two modes (for ages of 10 and 54).
-Although frequency of insect stings hypersensitivity reactions in children is comparable to food hypersensitivity reactions (12-15% of cases of hypersensitivity reactions admitted to the ER), anaphylaxis is triggered significantly more often by food elicitors in childhood (56% of food hypersensitivity cases vs 5.3% of sting cases seen in the ER) [@Braganza2005].
+<!-- Symptoms -->
+<!-- Cardio symptoms phenomenon -->
+IVA was more often associated with cardiologic symptoms non-IVA.
+Previous studies suggest [@Worm2018; @nittner2015insect; @Bonadonna2010] an essential link between the cardiovascular system and insect sting hypersensitivity. 
+IVA has been associated with cardiac arrhythmias usually occurring in patients with preexisting heart disease [@Sharma2016] and Kounis syndrome (coronary arterial spasm induced by the release of mast cell mediators [@CDT1609; @sinkiewicz2008allergic]).
+
+The rate of concomitant cardiovascular diseases was higher in IVA than non-IVA; we observed them be an essential cofactor increasing the *risk of a severe reaction if insects elicited the anaphylaxis*.
+This association was not significant in anaphylaxis elicited by other triggers.
+Notably, cardiac arrest occurred more frequently in patients with the elevated BST (> 8 ng/ml), especially in IVA.
+Nevertheless, the pathomechanism promoting cardiovascular symptoms in IVA requires further investigation.
+
+Of note, cardiologic symptoms like hypotension, collapse, or cardiac arrest grade higher on the Ring and Messmer scale than skin or gastrointestinal symptoms.
+Therefore, since IVA predisposes to cardiologic symptoms, it is likely to be associated with more severe anaphylaxis.
+
+Importantly, the absence of skin symptoms was associated with more severe IVA, which was also after we excluded patients with a known diagnosis of mastocytosis.
+Subsequently, we correlated the severity of anaphylaxis with BST levels and identified an interaction between the absence of skin symptoms and IVA using generalized linear regression.
+
+Our findings indicate that patients with BST of over 8 ng/ml are prone to severe anaphylaxis to insect venom.
+Patients with normal BST in the range of 8-11.4 ng/ml may have indolent systemic mastocytosis or concomitant undiagnosed mast cell activation syndrome (MCAS) [@Valent2019].
+Zanotti et al.
+identified mast cell disorders in 17 out of 22 patients with IVA lacking skin symptoms [@Zanotti2015] and concluded that patients with BST of 7.95 ng/ml and IVA should undergo extensive diagnostic procedures.
+
+Based on our and previous findings [@Ruff2009;@Zanotti2015;@JaraAcevedo2015] we propose to perform a peripheral blood KIT D816V mutation test in cases of BST of above 8 ng/ml and with a history of anaphylaxis presenting without urticaria or flushing.
+
+<!-- Age -->
+
+Regarding the factors increasing risk of IVA - older patients experienced IVA more frequently.
+Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].
+One of the limitations of our registry is the fact that we can only compare cases of anaphylaxis and due to the lack of healthy cohort - cannot draw conclusions on the frequency of a particular type of anaphylaxis in the population.
+
+Emergency room (ER) admission data indicate that the frequency of insect stings hypersensitivity reactions in children is comparable to food hypersensitivity reactions (12-15% of cases of hypersensitivity reactions admitted to the ER), but pediatric anaphylaxis is triggered significantly more often by food elicitors (56% of food hypersensitivity cases vs.
+5.3% of sting cases seen in the ER) [@Braganza2005].
 Senior patients, on the other hand, suffer from drug-related hypersensitivity more often than insect sting hypersensitivity [@Aurich2019].
+<!-- Risk factors - Concomitant disease -->
+Similarly, we saw less IVA in patients with concomitant atopic diseases, as these patients more often present with food anaphylaxis [@Tham2019].
 
-IVA was more associated with cardiologic symptoms than other anaphylaxis types.
-Additionally, cardiologic diseases were vital cofactor increasing the risk of a severe reaction in IVA only.
-The body of evidence suggests [@Worm2018; @nittner2015insect; @Ruff2009] that there is an essential link between the cardiovascular system and insect stings as they have been associated with cardiac arrhythmias usually occurring in patients with preexisting heart disease [@Sharma2016].
+The influence of cardiologic medication could not be isolated from the effect of concomitant cardiologic conditions; therefore, we could not state if ACE-I and beta-blockers increased the severity of anaphylaxis.
+We did observe that there were no significant differences between IVA and non-IVA cases regarding the symptoms and severity of an episode with concomitant use of ACE-I or beta-blockers.
 
-Kounis syndrome (coronary arterial spasm induced by the release of mast cell mediators) could be another explanation for the increased concomitance of cardiovascular symptoms with IVA ([@CDT1609; @sinkiewicz2008allergic]).
-As the severity of cardiologic reactions was positively associated with the elevated BST (> 8 ng/ml), we theorize that insect venom is unspecifically activating more mast cells at once than food or drug elicitors (potentially due to its composition and subcutaneous route of administration).
-Such rapid activation leads to a more evident cardiovascular spasm (especially if the concomitant cardiovascular disease is present) and in result to more severe anaphylaxis.
+<!-- Therapy  -->
+Cases of IVA had been treated with adrenaline less often than the age- sex- and severity-matched cases of non-IVA.
+Moreover, administration of adrenaline did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patient experienced their first episode of IVA.
+The difference between groups was prominent for milder cases of anaphylaxis.
+The reason for this observation is unclear. To our knowledge, this is the only data on the comparative adrenaline usage in a case-controlled group of IVA vs. non-IVA.
 
-We propose to review the currently used reference values of serum tryptase levels.
-In our opinion, we should consider a cut-off value of 8 ng/ml as elevated, when evaluating patients in the context of IVA.
-Baseline serum tryptase levels (BST) of > 8 ng/ml were better predicting the risk of severe anaphylaxis and corresponded with cardiovascular symptom frequency.
+Nevertheless, international guidelines of anaphylaxis state that adrenaline (i.m.) is the first-line agent in all diagnosed cases of anaphylaxis [@Muraro2014].
+Clinicians should not undermine the less severe IVA cases and treat them with adrenaline accordingly.
 
-Cases of IVA had been treated with adrenaline less often than the age-and-sex-matched cases of anaphylaxis due to other elicitors.
-Moreover, administration of adrenaline did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patient experienced their first episode of IVA.
-One explanation for this fact may be that although symptoms of IVA are more severe (cardiologic symptoms are considered life-threatening), they tend to be more transient.
-In case of food allergy, patients could develop symptoms for an extended time and will eventually be treated with adrenaline, whereas patients with IVA will often be recovering from hypotension or collapse when the emergency medical team arrive at the site (with AAI).
+Although there are no absolute contraindications for using adrenaline in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using adrenaline is hypersensitivity reaction with high blood pressure with tachycardia which might be present at the initial phase of IVA, due to a psychologic reflex.
+In theory, these less severe cases of IVA might exhibit some form of stress-related blood pressure increase, but we lack data to confirm or discard this theory.
 
-We saw less IVA in patients with concomitant atopic diseases, as these patients more often present with food anaphylaxis [@Tham2019].
+<!-- Limitations and strengths -->
+Due to the design of the European Anaphylaxis Registry, our analysis was restricted only to cases of anaphylaxis.
+Milder hypersensitivity reactions, as well as healthy controls, are not included in the database.
+Although The European Anaphylaxis Registry is ideal for investigating anaphylaxis subtypes - it might give an incomplete picture of the populational distribution of hypersensitivity reactions and restricts us to only comparing various forms of anaphylaxis.
+
+Nevertheless, because the European Anaphylaxis Registry has until now gathered over 12 000 cases of anaphylaxis - it was possible to perform a case-controlled analysis on a relatively large number of cases and investigate many aspects of IVA.
 
@@
 Based on our results, IVA is a distinctive subtype of anaphylaxis, with specific symptom profile and risk factors.
-Although therapy of IVA should be performed according to the guidelines and should not differ from other types of anaphylaxis, it is unclear why patients with IVA tend to receive adrenaline less often.
+IVA cases should undergo therapy according to the international management guidelines, and adrenaline should be given more often in IVA.
 
-When evaluating the risk of future severe episodes - baseline serum tryptase levels over eight ng/ml should be considered.
+When evaluating the risk of future severe episodes - baseline serum tryptase levels over ng/ml should be considered.
 
@@
 We thank the study personnel for patients counseling and data entry, and we thank the members of The European Anaphylaxis Registry in detail:
-J. Grünhagen (Berlin, Germany),
-K. Beyer (Berlin, Germany),
-A. Möser (Jena, Germany),
-T. Fuchs (Göttingen, Germany),
-F. Ruëff (Munich, Germany),
-B. Wedi (Hanover, Germany),
-F. Friedrichs (Aachen, Germany),
-H. Dickel (Bochum, Germany),
-H. Merk (Aachen, Germany),
-U. Hillen (Essen, Germany),
-A. Bauer (Dresden, Germany),
-N. Wagner (Erlangen, Germany),
-E. Rietschel (Cologne, Germany),
-R. Treudler, S. Aurich (Leipzig, Germany),
-L. Klimek, O. Pfaar (Wiesbaden, Germany),
-N. Reider (Innsbruck, Germany),
-W. Aberer (Graz, Austria),
-A. Köhli (Zurich, Switzerland),
-F. Riffelmann (Schmallenberg, Germany),
-B. Kreft (Halle, Germany),
-K. Nemat (Dresden, Germany),
-T. Kinaciyan (Vienna, Austria),
-R. Brehler (Münster, Germany),
-J. Witte (Hamburg, Germany),
-N. Hunzelmann und I. Huseynow (Cologne, Germany),
-T. Bieber (Bonn, Germany),
-U. Rabe (Treuenbritzen, Germany),
-P. Schmid-Grendelmeier (Zurich, Switzerland),
-W. Brosi (Würzburg, Germany),
-S. Nestoris (Lippe-Lemgo, Germany),
-T. Hawranek (Salzburg, Austria),
-R. Bruns (Greifswald, Germany),
-S. Lehmann (Aachen, Germany),
-L. Lange (Bonn, Germany),
-G. Hansen (Hanover, Germany),
-C. Pföhler (Homburg, Germany),
-E. Varga (Graz, Austria),
-Z. Szepfalusi (Vienna, Austria),
-P. Eng (Aarau, Switzerland),
-P. Eng (Lucerne, Switzerland),
-T. Reese (Rheine, Germany),
-M. Polz (Rüsselsheim, Germany),
-S. Schweitzer-Krantz (Düsseldorf, Germany),
-H. Rebmann (Tübingen, Germany),
-G. Stichtenoth (Lübeck, Germany),
-S. Theis (Schwedt, Germany),
-I. Yildiz (Neumünster, Germany),
-M. Gerstlauer (Augsburg, Germany),
-A. Nordwig (Dresden, Germany),
-R. Schlags (Wangen im Allgäu, Germany),
-I. Neustädter (Fürth, Germany),
-C. Stadlin (Zurich, Switzerland),
-M. Bücheler (Bonn, Germany),
-S. Volkmuth (Velbert, Germany),
-J. Fischer (Tübingen, Germany),
-A. Henschel (Berlin, Germany),
-S. Plank-Habibi (Alzenau, Germany),
-B. Schilling (Passau, Germany),
-A. Kleinheinz (Buxtehude, Germany),
-K. Schäkel (Heidelberg, Germany),
-N. Papadopoulos (Athen, Greece),
-M. Kowalski (Lodz, Poland),
-M. Fernandez-Rivas (Madrid, Spain),
-K. Solarewicz-Madajek (Wroclaw, Poland),
-C. Körner-Rettberg (Bochum, Germany),
-T. Mustakov (Sofia, Bulgaria),
-K. Hartmann (Lübeck, Germany),
-C. Kemen (Hamburg, Germany),
-U. Miehe (Leipzig, Germany),
-C. Ebner (Vienna, Austria),
-S. Haak (Oldenburg, Deutschland),
-V. Cardona (Barcelona, Spain),
-S. Hämmerling (Heidelberg, Germany),
-E. Arroabarren (Pamplona, Spain),
-M. Bilo (Ancona, Italy),
-N. Cabañes Higuero (Toledo, Spain),
-A. Vega Castro (Guadalajara, Spain),
-I. Poziomkowska-Gęsicka (Szczecin, Poland),
-S. Büsing (Osnabrück, Germany),
-U. Klettke (Berlin, Germany),
-C. Virchow (Rostock, Germany),
-G. Christoff (Sofia, Bulgaria),
-U. Jappe (Borstel, Germany),
-K. Breuer (Hamburg, Germany),
-S. Müller and T. Jakob (Freiburg, Germany),
-H. Straube (Darmstadt, Germany),
-C. Vogelberg (Dresden, Germany),
-F. Knöpfel (Norderney, Germany),
-J. Hourihane (Cork, Ireland),
-B. Rogala (Silesia, Poland),
-A. Montoro (Madrid, Spain),
-A. Brandes (Frankfurt/Oder, Germany),
-A. Muraro (Padua, Italy),
-T. Buck und J. Büsselberg (Hanover-Misburg, Germany),
-N. Zimmermann (Potsdam, Germany),
-D. Hernandez (Valencia, Spain),
-P. Minale (Genua, Italy),
-J. Niederwimmer und B. Zahel (Linz, Austria),
-A. Fiocchi (Rome, Italy),
-A. Reissig (Gera, Germany),
-F. Horak (Vienna, Austria),
-N. Klossowiski (Düsseldorf, Germany),
-F. Eitelberger (Wels, Austria),
-H. Ott (Hanover, Germany),
-R. Asero (Milan, Italy),
-S. Pistauer (Sylt/Westerland, Germany),
-M. Geißler (Ribnitz-Damgarten, Germany),
-L. Ensina (Sao Paulo, Brazil),
-A. Plaza Martin (Barcelona, Spain),
-J. Meister (Aue, Germany),
-S. Hompes (Hamburg, Germany),
-S. Stieglitz (Wuppertal, Germany).
 
@@
 
-```{r flow,fig.width=6,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale for including registry cases in the final analysis.
+```{r flow,fig.width=6,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale for case inclusion and exclusion from the final analysis.
 B, C, D: Age, sex, and severity distribution was matched in cases in both groups to allow for comparable results between IVA and non-IVA cases.
@@
 A: Proportional presentation of specific reaction symptoms.
-B: High-level overview of involved organ systems and selected cofactors in the form of a radar plot.", fig.width=8.5, fig.height = 3.2}
-fig_symptoms
+B: High-level overview of involved organ systems and selected cofactors in the form of a radar plot.", fig.width=8.5, fig.height = 6.2}
+  fig_symptoms2
 ```
@@
 B: Proportion of cases elicited by insects or other elicitors (upper panels) according to tryptase levels and cardiologic symptoms.
-", fig.width = 5.6, fig.height = 10}
+", fig.width = 6.7, fig.height = 7.4}
+
 cof_fig
@@
 B: Heatmap visualizing the association of symptoms and corresponding treatment -  presented as a scaled correlation coefficient (phi).", fig.height=9, fig.width = 7}
-
-ggpubr::ggarrange(
-    plotManagement,
-
-  #lower_panel,
-  matrix_venom$p+
-    labs (x = "", y = "")+
-    geom_segment(aes(x = c(8.5),
-                     xend = c(8.5), 
-                     y = c(0.5),
-                     yend = c(39.5)),color = "red")+
-    geom_segment(aes(x = c(0.5),
-                     xend = c(20.5), 
-                     y = c(28.5),
-                     yend = c(28.5)), color = "red" )+
-    geom_segment(aes(x = c(0.5),
-                     xend = c(20.5), 
-                     y = c(34.5),
-                     yend = c(34.5)), color = "red", linetype = 2  ),
-  nrow = 2,
-  heights = c(1.5,3),
-  labels = c("","C")
-)
+fig_adrenalineuse
 
@@
 ```
+
+```{r trypcardiocur, fig.cap="Tryptase levels in patients with concomitant cardiologic diseases. "}
+
+plot_tryptase_in_cardio_cur
+```

### 2019-09-09 R/functions.R

@@ -64,18 +64,18 @@ match_patients <- function(data,
                            grouping_var_level = "insects"
 ){
   o <- data %>%
-    select(b_case_id,!!!grouping_var,!!!matching_vars) %>%
+    dplyr::select(b_case_id,!!!grouping_var,!!!matching_vars) %>%
     {filter(.,complete.cases(.))} %>%
     mutate(grouping = ifelse(grouping==grouping_var_level,1,0)) %>%
     matchit(formula = as.formula(paste0(grouping_var,
                                         "~",
                                         paste0(matching_vars,
                                                collapse = "+"))
-    ),
+                                          ),
     method = "nearest",
     ratio = 1) %>%
     match.data() %>%
-    select(b_case_id) %>% pull()
+    dplyr::select(b_case_id) %>% pull()
 
   if(df == T){
     o <- data[data$b_case_id %in% o,]
@@ -91,6 +91,18 @@ unknown_to_na <- function(vect){
                 0,
                 vect)
   )
+}
+
+compare <- function(x,percent =T,rounding =1){
+  if(percent == T){
+    paste0(round(x[1]*100,rounding),
+           "% vs. ",
+           round(x[2]*100,rounding),"%")
+  } else {
+    paste0(round(x[1],rounding),
+           " vs. ",
+           round(x[2],rounding))
+  }
 }
 
 
@@ -400,7 +412,7 @@ makeTests <- function(groups,rdb){
 
   # Perform chi2 test
   varDF <- rdb[,variableSelectionTab %>% filter(fun=="chi") %>%
-                 select(variableName) %>% pull %>% as.character()]
+                 dplyr::select(variableName) %>% pull %>% as.character()]
   ts <- lapply(varDF, # Check which tables are givin an error with chisq test
                function(x){ # You may implement also fisher tests here later
                  table(x,grouping)
@@ -417,7 +429,7 @@ makeTests <- function(groups,rdb){
   )
   #perform t.tests
   varDF <- rdb[,variableSelectionTab %>% filter(fun=="t.test")
-               %>% select(variableName) %>% pull %>% as.character()]
+               %>% dplyr::select(variableName) %>% pull %>% as.character()]
   #chek if t.test possible
[…]

### 2019-09-09 R/make_flow.R

@@ -43,10 +43,9 @@ nodes <-
                     format(length(which(data4$d_elicitor_gr5=="unkown" &
                                    data4$reaction_type_brown=="anaphylaxis")),
                            big.mark = ",")),
-             paste0("Other known\nelicitors\n",
-                    format(length(which(data4$d_elicitor_gr5!="insects"&
-                                   data4$d_elicitor_gr5!="unkown" &
-                                   data4$reaction_type_brown=="anaphylaxis")),
+             paste0("Control group of\nSex- and age-matched reactions\nto other known elicitors\n",
+                    format(nrow(age_sex_matched %>%
+                                  filter(grouping == "other")),
                            big.mark = ","))
     ),
 
@@ -54,10 +53,10 @@ nodes <-
     #          "grey", "blue"),
     #value = c(3.5, 2.6, 9.4, 2.7),
     shape= "rectangle",
-    width = c(2.5,3,1.8,1.2,1.8,1.2,1.2),
-    x = c(2, 2, 1.1, 3.1,1.1,4.7,3.1),
+    width = c(2.5,3,1.8,1.2,1.8,1.2,2.5),
+    x = c(2, 2, 1.1, 3.1,1.1,4.7,3.6),
     y = c(0,-1,-2,-2,-3,-2,-3),
-    height = c(rep(0.6,4),1,0.6,0.6))
+    height = c(rep(0.6,4),1,0.6,1))
 
 edges <- create_edge_df(from = c(1,2,2,3,4,4),
                         to =   c(2,3,4,5,6,7))

### 2019-09-09 R/stat.R

@@ -141,13 +141,32 @@ data4$route_sc[data4$d_330_drug_group %in%
                  )] <- TRUE
 data4$route_sc[data4$d_elicitor_gr5 == "insects"] <- T
 
+#### d_111_urti_flush
+
+data4 <- data4 %>%
+  mutate(d_111_urti_flush = ifelse(q_111_urticaria=="yes"|
+                                     q_111_erythema_flush_v5=="yes",
+                                   "yes",
+                                   "no"))
+
+##### medium Tryptase between 4 and 8 11.5 ####
+
+data4 <- data4 %>%
+  mutate(tryptase_value_3cat =
+           ifelse(q_212_tryptase_value_v5 < 4,
+                  "low",
+                  ifelse(q_212_tryptase_value_v5 < 8,
+                         "medium",
+                         ifelse(q_212_tryptase_value_v5 <11.5,
+                         "high",
+                         ">11.5"))))
 
 ##### FINAL DATABASE RDB ######
 
 rdb <- data4[data4$reaction_type_brown=="anaphylaxis",]
 countries <-rdb %>%
   filter(d_elicitor_gr5=="insects") %>%
-  select(d_centres_country) %>%
+  dplyr::select(d_centres_country) %>%
   group_by(d_centres_country) %>%
   summarize(n=n())%>% arrange(desc(n))
 
@@ -165,18 +184,15 @@ rdbp$grouping<- grouping
 
 
 
-##### DIAGRAM ##############
-source("R/make_flow.R")
-make_flow()
-
 
 ###Statistic###########
 
 ### Binomial trigger either insects or other
-testInsectsbinomial <- makeTests(groups = "grouping",rdb=rdb) %>% arrange(pval)
+testInsectsbinomial <- makeTests(groups = "grouping",rdb=rdb) %>%
+  arrange(pval)
 
 testInsectsbinomialTab <- testInsectsbinomial %>% #filter(pval<1e-30) %>%
-  select(variableName,counts_1,counts_2,fraq_1,fraq_2,
+  dplyr::select(variableName,counts_1,counts_2,fraq_1,fraq_2,
          pval,
          section) %>%
          {split(.,.$section)}
@@ -190,7 +206,7 @@ tests_matched <- makeTests(groups ="grouping",
                            rdb =age_sex_matched)
 
 tests_matchedTab <- tests_matched %>% filter(pval<1e-30) %>%
[…]

### 2019-09-09 analysis/paper/paper.Rmd
V2.1 - gramarly corrected

@@
 ---
-title: "Phenotypes and risk assesment of insect venom anaphylaxis: a case - control study of the European Anaphylaxis Registry"
+title: "Phenotype and risk assesment of insect venom anaphylaxis: a case - control study of the European Anaphylaxis Registry"
 author:
@@
 
-Hypersensitivity to insect venom can manifest as a systemic reaction (anaphylaxis) in up to 0.3-7.5% of the adult population [@Bil2008].
+Hypersensitivity to insect venom can manifest as a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008].
 Insect venom anaphylaxis (IVA) can be fatal, and patients sometimes require lifelong systemic immunotherapy [@Sturm2017].
@@
 Recent expert position on the future of anaphylaxis research indicates the need for a more precise description of the diagnosis, biomarkers, and phenotypes of anaphylaxis [@JimenezRodriguez2018].
-Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] in the treatment of anaphylaxis - it is first needed to study its clinical subtypes and pathomechanisms in more detail.
+Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] in the treatment of anaphylaxis --- it is first needed to study its clinical subtypes and pathomechanisms in more detail.
 
-Clinical controlled trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition.
+Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition.
 Therefore large registries, gathering clinical data as well as biological samples from patients with a well documented (recent) history of anaphylaxis are a crucial tool in investigating this entity.
@@
 We searched the European Anaphylaxis Registry [@Grabenhenrich2016] database from March 2019 for anaphylaxis cases elicited by insect's venom.
-The flowchart in figure \@ref(fig:flow) represents the detailed case-selection process.
+The flowchart in Fig.
+\@ref(fig:flow) represents the detailed case-selection process.
 
-The final database consisted of `r countries %>% summarise(sum(n))` cases of insect elicited anaphylaxis from `r countries %>% summarize(length(n))` countries.
-Severe reactions were identified based on the definition by NIAID/FAAN [@Sampson2006] or as Ring and Messmer Scale - grades III and IV) and presented with significant hypoxia, hypotension, confusion, collapse and loss of consciousness, or incontinence.
+The final database consisted of `r countries %>% summarise(sum(n))` cases of insect-elicited anaphylaxis reported from allergology centers in `r countries %>% summarize(length(n))` countries.
+We based the diagnosis of anaphylaxis on the definition by NIAID/FAAN [@Sampson2006] and the severity according to Ring and Messmer Scale [@Ring1977].
+Grades III and IV (presenting with significant hypoxia, hypotension, confusion, collapse and loss of consciousness, or incontinence or cardiac arrest) were considered severe.
 
-The cases were either assigned to IVA group (if the triggering factor was an insect sting) or non-IVA group (other elicitors of anaphylaxis).
+The cases were either assigned to IVA group (if the triggering factor was an insect sting) or non-IVA group (if other elicitors triggered anaphylaxis).
 We compared the frequency of various symptoms, cofactors --- known to increase the risk of severe anaphylaxis, [@Worm2018] and management in both groups.
@@
 The results of the propensity score matching are in Fig.
-\@ref(fig:flow)B-D.
+\@ref(fig:flow)B-D and supplementary Fig \@ref(fig:matchingresult).
 
@@
 We used the R Statistical Package [@R] for statistical analysis.
-A simple comparison of categorical variables was performed using either Chi^2^ test or Fisher's exact test (where the number of observation in a bin was less than 10), continuous variables were analyzed using Mann-Whitney U test.
+A simple comparison of categorical variables was performed using either Chi^2^ test or Fisher's exact test (where the number of observation in a bin was less than 10).
+Continuous variables were analyzed using Mann-Whitney U test.
+In case of comparisons with two or more independent variables, we used Factorial ANOVA or Generalized Linear Models.
 We defined statistical significance as α = 0.05.
@@
 
-## Hymenoptera anaphylaxis shows seasonal patterns.
-
-Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
-Their proportion of IVA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was as below 1% of cases during winter.
-Nevertheless, `r rare_spring_autumn$n %>% sum` cases of IVA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were triggered in March, April, and November.
-Yellow-jacket was the most prominent IVA-causing insect followed by bees.
-The IVA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in southern Europe.
-
 IVA was more frequent in adults and seniors than children and young adults.
 (Fig \@ref(fig:insectstime)B).
-When we analyzed a density plot of IVA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B.
+When we analyzed a density plot of IVA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B).
 
@@
 
-Patients with concomitant mastocytosis had significantly more often anaphylaxis without skin symptoms (namely urticaria and flushing) when compared to patients without diagnosed mastocytosis
-(`r (skin_symptoms$masto_tab$prop[c(2,1)]*100) %>% round(1) %>% paste0(collapse="% vs.
+Patients with concomitant mastocytosis had significantly more often anaphylaxis without skin symptoms (namely urticaria and flushing) when compared to patients without diagnosed mastocytosis (`r (skin_symptoms$masto_tab$prop[c(2,1)]*100) %>% round(1) %>% paste0(collapse="% vs.
 ")`%, `r pval(skin_symptoms$masto_chisq$p.value)`).
-We decided to further investigate cases where mastocytosis was not diagnosed.
-In these patients urticaria or flushing as symptoms of anaphylaxis were present more often in IVA than anaphylaxis triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
+We decided to further investigate non-mastocytic patients.
+In these patients urticaria or flushing as symptoms of anaphylaxis were present less often in IVA than anaphylaxis triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
 Morover, in patients without concomitant mastocytosis who were also lacking skin symptoms, anaphylaxis was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(2,6)]*100) %>% round(1) %>% paste0(collapse="% in IVA vs.
@@
 Similarly, tryptase levels correlated with the severity of anaphylaxis, were higher in IVA patients, and the effect was prominently visible for IVA patients (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`).
-We did not observe such an interaction in the non-IVA group.
+We did not observe such an interaction in the non-IVA group (Fig.
+\@ref(fig:figsymptoms)D).
 
@@
 However, mastocytosis increased the risk of cardiac arrest in patients undergoing IVA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
-S\@ref(fig:supplement)).
+S\@ref(fig:supplement)A).
 
-Concomitant diseases were more prevalent in IVA than in non-IVA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare()`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were irrelevant in non-IVA cases (Fig.
+Concomitant cardiovascular diseases were more prevalent in IVA than in non-IVA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare()`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were irrelevant in non-IVA cases (Fig.
 \@ref(fig:figcofactors)).
 Interestingly, BST values were increased in patients with concomitant cardiovasular diseases, irrespectively of the reaction severity (Fig.
-S\@ref(trypcardiocur)).
+S\@ref(fig:trypcardiocur)).
 
 The risk of severe anaphylaxis in patients concomitantly using ACE-I (as well as beta-blockers) could not be independently measured due to coexisting cardiovascular pathologies.
-ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) and there were no differences between IVA nad non-IVA  (Fig.
+ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) and there were no differences between IVA nad non-IVA  (Fig.
 \@ref(fig:figcofactors)C).
@@
 Surprisingly, arrhythmia was more frequently reported in patients undergoing IVA with concomitant beta-blockers (Fig.
-\@ref(fig:figcofactors)B).
+\@ref(fig:figcofactors)C).
 
 Baseline serum tryptase (BST) correlated with the severity of reactions (Ring and Messmer scale).
-Moreover, elevated BST was more prominently increasing the risk of severe anaphylaxis in IVA patients than in other cases (Fig.
-\@ref(fig:figcofactors)C).
-
+Moreover, elevated BST was more prominently increasing the risk of severe anaphylaxis in IVA than in non-IVA (Fig.
+\@ref(fig:figsymptoms)D and Fig.
+\@ref(fig:figcofactors)B).
 Cases with cardiac arrest were associated with an increased tryptase above 8 ng/ml, and this proportion was higher in IVA when compared to other elicitors.
@@
 
+## Hymenoptera anaphylaxis shows seasonal patterns.
+
+Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
+Their proportion of IVA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was as below 1% of cases during winter.
+Nevertheless, `r rare_spring_autumn$n %>% sum` cases of IVA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were triggered in March, April, and November.
+Yellow-jacket was the most prominent IVA-causing insect followed by bees.
+The IVA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in southern Europe.
+
 <!-- ## Time of the symptom onset  -->
@@
 
-<!-- DONE Seasonality -->
-
-Based on our findings, insects are the most probable anaphylaxis elicitor in Europe during Summer-season, with IVA cases extending from early Spring to the end of Autumn.
-Detailed information on the seasonality of insect-elicited hypersensitivity reactions is scarce [@old].
-The activity of *Vespula germanica* depends on the climate, and in invaded regions (e.i.
-Australia) it can even extend throughout the year [@Spradbery_1992].
-The changing climate in Europe may influence the activity of Hymenoptera in this region in the upcoming years.
-However, in the period from 2007 - 2019, the perennial ratio of IVA to non-IVA cases has remained unchanged (data not shown).
-
 <!-- Symptoms -->
@@
 Previous studies suggest [@Worm2018; @nittner2015insect; @Bonadonna2010] an essential link between the cardiovascular system and insect sting hypersensitivity. 
-IVA has been associated with cardiac arrhythmias usually occurring in patients with preexisting heart disease [@Sharma2016] and Kounis syndrome (coronary arterial spasm induced by the release of mast cell mediators [@CDT1609; @sinkiewicz2008allergic]).
+IVA has been associated with Kounis syndrome (coronary arterial spasm induced by the release of mast cell mediators [@CDT1609; @sinkiewicz2008allergic]) and cardiac arrhythmias usually occurring in patients with preexisting heart disease [@Sharma2016].
 
@@
 
-Importantly, the absence of skin symptoms was associated with more severe IVA, which was also after we excluded patients with a known diagnosis of mastocytosis.
+Importantly, the absence of skin symptoms was associated with more severe IVA, which was also true after we excluded patients with a known diagnosis of mastocytosis.
 Subsequently, we correlated the severity of anaphylaxis with BST levels and identified an interaction between the absence of skin symptoms and IVA using generalized linear regression.
@@
 Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].
-One of the limitations of our registry is the fact that we can only compare cases of anaphylaxis and due to the lack of healthy cohort - cannot draw conclusions on the frequency of a particular type of anaphylaxis in the population.
+One of the limitations of our registry is the fact that we can only compare cases of anaphylaxis and due to the lack of healthy cohort --- cannot draw conclusions on the frequency of a particular type of anaphylaxis in the population.
 
@@
 
+<!-- DONE Seasonality -->
+
+Based on our findings, insects are the most probable anaphylaxis elicitor in Europe during Summer-season, with IVA cases extending from early Spring to the end of Autumn.
+Detailed information on the seasonality of insect-elicited hypersensitivity reactions is scarce [@old].
+The activity of *Vespula germanica* depends on the climate, and in invaded regions (e.i.
+Australia) it can even extend throughout the year [@Spradbery_1992].
+The changing climate in Europe may influence the activity of Hymenoptera in this region in the upcoming years.
+However, in the period from 2007 - 2019, the perennial ratio of IVA to non-IVA cases has remained unchanged (data not shown).
+
 <!-- Limitations and strengths -->
@@
 
-When evaluating the risk of future severe episodes - baseline serum tryptase levels over ng/ml should be considered.
+When evaluating the risk of future severe episodes - patients with BST over 8 ng/ml should undergo extensive diagnostic tests to exclude ISM or MCAS and should be provided with two adrenaline autoinjectors.
 
@@
 
-```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occurred.
-Less common insect species grouped as 'other'.
-B: The density distribution of IVA cases to cases elicited by other triggers considering the patient's age.
-C: Geographical differences in the most common elicitors of IVA.
-Countries which reported less than 10 IVA cases were not illustrated in this figure.", fig.height=4, fig.width = 7}
-fig_basic
-```
-
-#####
-
 ```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors.
 A: Proportional presentation of specific reaction symptoms.
-B: High-level overview of involved organ systems and selected cofactors in the form of a radar plot.", fig.width=8.5, fig.height = 6.2}
+B: High-level overview of involved organ systems and selected cofactors in the form of a radar plot.
+C: Interaction effect of IVA and lack of skin symptoms (urticaria or flushing) on the severity of anaphylaxis.
+D: Baseline Serum Tryptase (BST) increases with the severity of anaphylaxis a lot more prominently in IVA.", fig.width=8.5, fig.height = 6.2}
 fig_symptoms2
@@
 B: Proportion of cases elicited by insects or other elicitors (upper panels) according to tryptase levels and cardiologic symptoms.
-", fig.width = 6.7, fig.height = 7.4}
+", fig.width = 7, fig.height = 8}
 
@@
 
+#####
+
+```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occurred.
+Less common insect species grouped as 'other'.
+B: The density distribution of IVA cases to cases elicited by other triggers considering the patient's age.
+C: Geographical differences in the most common elicitors of IVA.
+Countries which reported less than 10 IVA cases were not illustrated in this figure.", fig.height=4, fig.width = 7}
+fig_basic
+```
+
+#####
+
 # Supplementary Figures
 
-```{r supplement, fig.width= 6, fig.height= 5, fig.cap="Symptoms of anaphylaxis"}
+```{r matchingresult, fig.cap="Results of matching the cohort according to sex and age in order to perform a case-controlled study.
+A: The original distribution of IVA and non-IVA cases according to age group and sex.
+B: The distribution of IVA and non-IVA after age and sex matching with the use of MatchIt package for R."}
+matching.result 
+```
+
+```{r supplement, fig.width= 6, fig.height= 5, fig.cap="Symptoms of anaphylaxis.
+A: The association between cardiac arrest and concomitant mastocytosis in IVA and non-IVA.
+B: Hypotension frequency in two age groups of anaphylaxis.
+C: Crammer's V as the measure of association between groups anaphylaxis (IVA vs.
+non-IVA)."}
 
@@
 
+```{r trypcardiocur, fig.cap="Tryptase levels in patients with concomitant cardiologic diseases. "}
+
+plot_tryptase_in_cardio_cur
+```
+
 ```{r treatmentmatrixother}
@@
 
-```{r rfplot}
-rf_ggplot_treatment
+```{r rfplot, fig.cap="Therapy of anaphylaxis. A: Adrenaline use considering if patients had previous anaphylaxis. B: Variable impo"}
+ggarrange(
+  adren_prev+
+    geom_segment(aes(x = 0.5,
+                     xend = 2.5,
+                     y = 0.33,
+                     yend = 0.33),
+                 color="white",
+                 linetype =2),
+  rf_ggplot_treatment,
+  nrow = 1,
+  ncol = 2,
+  labels=c("A","B")
+)
 ```
-
-```{r}
-adren_prev
-```
-
-```{r trypcardiocur, fig.cap="Tryptase levels in patients with concomitant cardiologic diseases. "}
-
-plot_tryptase_in_cardio_cur
-```

### 2019-09-09 R/stat.R

@@ -859,12 +859,12 @@ rdb %>%
 
 
 ANAscore_matched <-
-  match_patients(rdb %>%
+  match_patients(data = rdb %>%
                    filter(d_522_adren_agg %in% c("yes", "no")),
-                 "grouping",
-                 c("ANAscore","d_age"),
+                 grouping_var ="grouping",
+                 matching_vars = c("ANAscore","d_age"),
                  grouping_var_level = "insects",
-                 T)
+                 df = T)
 testANAscoreMatched <- makeTests(groups = "grouping",
                                  rdb=ANAscore_matched) %>%
   #arrange(pval) %>%
@@ -1059,9 +1059,15 @@ lower_panel <- ggpubr::ggarrange(
                  linetype = 2),
 
   age_sex_matched %>%
-    group_by(b_sex, grouping) %>%
+    # count(b_sex, grouping) %>% #spread(b_sex,n) %>% {chisq.test(.[,2:3])}
+    # ggpubr::ggbarplot(
+    #   fill = "grouping",
+    #   y = "n",
+    #   x = "b_sex",
+    # ),
     ggplot(aes(fill=grouping, x=b_sex))+
     geom_bar(position = "fill")+
+    #facet_grid(.~b_sex)+
     theme_classic()+
     theme(legend.position = "none")+
     scale_fill_manual(values = rev(c("#c4c4c4", "#454545")))+
@@ -2184,15 +2190,17 @@ cof_fig<- ggarrange(
         y = "prop",
         facet.by = c("grouping",
                      "d_111_urti_flush"),
-        fill = "tryptase_value_3cat",
-        x = "d_severity_rmr",
+        x = "tryptase_value_3cat",
+        fill = "d_severity_rmr",
         position = position_fill(reverse = T),
         palette = "grey"
       )+
-      labs(x = "severity",
+      theme(axis.text.x = element_text(angle = 90, hjust = 1))+
+      guides(fill= guide_legend(ncol = 2,nrow = 2, byrow = F))+
+      labs(fill = "severity",
            y = "proportion",
-           fill = "BST\n[ng/ml]")+
-      guides(fill= guide_legend(ncol = 2,nrow = 2, byrow = F)),
+           x = "BST [ng/ml]")
+      ,
[…]

### 2019-09-09 analysis/paper/paper.Rmd
V2.2 - figure height fix

@@
 
-```{r rfplot, fig.cap="Therapy of anaphylaxis. A: Adrenaline use considering if patients had previous anaphylaxis. B: Variable impo"}
+```{r rfplot, fig.cap="Therapy of anaphylaxis. A: Adrenaline use considering if patients had previous anaphylaxis. B: Variable impo", fig.height=3.5, fig.width=7.1}
 ggarrange(
@@
   ncol = 2,
-  labels=c("A","B")
+  labels=c("A","B"),
+  widths = c(1,1.3)
 )

### 2019-10-07 analysis/paper/paper.Rmd
IVA to VIA

@@
   - Wojciech Francuzik^1^
+  - Franziska Ruëff^2^ 
   - Sabine Dölle-Bierke^1^
-  - Franziska Ruëff^2^ 
   - Claudia Pföhler^3^
@@
 
-Hypersensitivity to insect venom can manifest as a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008].
-Insect venom anaphylaxis (IVA) can be fatal, and patients sometimes require lifelong systemic immunotherapy [@Sturm2017].
+Hypersensitivity to insect venom is a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008].
+Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong systemic immunotherapy [@Sturm2017].
 
-Recent expert position on the future of anaphylaxis research indicates the need for a more precise description of the diagnosis, biomarkers, and phenotypes of anaphylaxis [@JimenezRodriguez2018].
-Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] in the treatment of anaphylaxis --- it is first needed to study its clinical subtypes and pathomechanisms in more detail.
+The need for a more precise description of the diagnosis, identification of biomarkers, and phenotypes of anaphylaxis is discussed [@JimenezRodriguez2018].
+Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] for the diagnosis of anaphylaxis --- a better understanding of its clinical subtypes is required.
 
-Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition.
-Therefore large registries, gathering clinical data as well as biological samples from patients with a well documented (recent) history of anaphylaxis are a crucial tool in investigating this entity.
+Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and their infrequent and random occurence.
+Therefore registries, gathering clinical data from patients with a well documented (recent) history of anaphylaxis are a crucial tool in investigating this entity.
 
-This study aimed to identify current patterns of insect venom anaphylaxis (IVA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-IVA).
+This study aimed to identify clinical patterns of venom induced anaphylaxis (VIA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA).
 
@@
 
-The final database consisted of `r countries %>% summarise(sum(n))` cases of insect-elicited anaphylaxis reported from allergology centers in `r countries %>% summarize(length(n))` countries.
-We based the diagnosis of anaphylaxis on the definition by NIAID/FAAN [@Sampson2006] and the severity according to Ring and Messmer Scale [@Ring1977].
+The final database consisted of `r countries %>% summarise(sum(n))` cases of insect-elicited anaphylaxis reported from allergy centers in `r countries %>% summarize(length(n))` countries.
+The diagnosis of anaphylaxis was based on the definition by NIAID/FAAN [@Sampson2006] and the severity according to Ring and Messmer Scale [@Ring1977].
 Grades III and IV (presenting with significant hypoxia, hypotension, confusion, collapse and loss of consciousness, or incontinence or cardiac arrest) were considered severe.
 
-The cases were either assigned to IVA group (if the triggering factor was an insect sting) or non-IVA group (if other elicitors triggered anaphylaxis).
+The cases were either assigned to VIA group (if the triggering factor was an insect sting) or non-VIA group (if other elicitors triggered anaphylaxis).
 We compared the frequency of various symptoms, cofactors --- known to increase the risk of severe anaphylaxis, [@Worm2018] and management in both groups.
 
-Due to a large number of documented reactions in the European Anaphylaxis registry - we were able to match the IVA cases to selected non-IVA cases according to sex and age to reduce the comparison bias.
+Due to a large number of documented reactions in the European Anaphylaxis registry - we were able to match the VIA cases to selected non-VIA cases according to sex and age to reduce the comparison bias.
 Subsequently, we compared the management in both groups and matched the control group according to the severity of a reaction.
@@
 
-We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapeutic approaches that varied the most between IVA / non-IVA group and presented the results Gini importance [@Strobl2007].
+We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapeutic approaches that varied the most between VIA / non-VIA group and presented the results Gini importance [@Strobl2007].
 We also performed an association study of therapeutic interventions and symptoms.
@@
 
-IVA was more frequent in adults and seniors than children and young adults.
+VIA was more frequent in adults and seniors than children and young adults.
 (Fig \@ref(fig:insectstime)B).
-When we analyzed a density plot of IVA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B).
+When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B).
 
-## IVA is associated with cardiovascular symptoms
+## VIA is associated with cardiovascular symptoms
 
-IVA showed a specific symptom pattern.
-Patients, who underwent IVA, more often experienced cardiologic symptoms (dizziness, reduced alertness, unconsciousness) than patients with other elicitors inducing anaphylaxis and less often had skin or gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A).
+VIA showed a specific symptom pattern.
+Patients, who underwent VIA, more often experienced cardiologic symptoms (dizziness, reduced alertness, unconsciousness) than patients with other elicitors inducing anaphylaxis and less often had skin or gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A).
 The difference in severe hypotension frequency was especially prominent in the younger age group (under 22, Fig.
@@
 The pattern of organ involvement showed similarities in gastrologic, skin, and respiratory symptoms, and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
-Although, patients undergoing IVA  less often had concomitant atopic diseases (Fig \@ref(fig:figsymptoms)B).
+Although, patients undergoing VIA  less often had concomitant atopic diseases (Fig \@ref(fig:figsymptoms)B).
 
-Severe reactions of IVA were more prevalent in older patients in comparison to patients below 22 years, and in IVA cases vs.
+Severe reactions of VIA were more prevalent in older patients in comparison to patients below 22 years, and in VIA cases vs.
 other elicitors (Fig.
@@
 
-## Absence of skin symptoms is associated with more severe episodes of IVA
+## Absence of skin symptoms is associated with more severe episodes of VIA
 
@@
 We decided to further investigate non-mastocytic patients.
-In these patients urticaria or flushing as symptoms of anaphylaxis were present less often in IVA than anaphylaxis triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
-Morover, in patients without concomitant mastocytosis who were also lacking skin symptoms, anaphylaxis was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(2,6)]*100) %>% round(1) %>% paste0(collapse="% in IVA vs.
-")`%, `r pval(skin_symptoms$masto_chisq$p.value)`% in non-IVA).
-We performed a factorial logistic regression modelling and confirmed the significant interaction effect between the presence of skin symptoms and the trigger (eiter IVA or non-IVA) on the severity of anaphylaxis (`r pval(skin_symptoms$severity_fit_skin_symptoms %>% coef %>% .[4,4])`).
+In these patients urticaria or flushing as symptoms of anaphylaxis were present less often in VIA than anaphylaxis triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
+Morover, in patients without concomitant mastocytosis who were also lacking skin symptoms, anaphylaxis was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(2,6)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
+")`%, `r pval(skin_symptoms$masto_chisq$p.value)`% in non-VIA).
+We performed a factorial logistic regression modelling and confirmed the significant interaction effect between the presence of skin symptoms and the trigger (eiter VIA or non-VIA) on the severity of anaphylaxis (`r pval(skin_symptoms$severity_fit_skin_symptoms %>% coef %>% .[4,4])`).
 Therefore, non-mastocytic patients presenting without urticara or flushing tended to have more severe anaphylaxis when it was triggered by insects, but not other elicitors of anaphylaxis (Fig.
@@
 For this comparison, we removed the cases with BST above 11.5 ng/ml, potentially indicating non-diagnosed mastocytosis.
-Similarly, tryptase levels correlated with the severity of anaphylaxis, were higher in IVA patients, and the effect was prominently visible for IVA patients (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`).
-We did not observe such an interaction in the non-IVA group (Fig.
+Similarly, tryptase levels correlated with the severity of anaphylaxis, were higher in VIA patients, and the effect was prominently visible for VIA patients (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`).
+We did not observe such an interaction in the non-VIA group (Fig.
 \@ref(fig:figsymptoms)D).
 
-## Nearly one-third of IVA patients experience repeated reactions
+## Nearly one-third of VIA patients experience repeated reactions
 
@@
 
-## Increased baseline serum tryptase and concomitant cardiologic conditions increase the risk of severe IVA.
+## Increased baseline serum tryptase and concomitant cardiologic conditions increase the risk of severe VIA.
 
@@
 \@ref(fig:figcofactors)).
-However, mastocytosis increased the risk of cardiac arrest in patients undergoing IVA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
+However, mastocytosis increased the risk of cardiac arrest in patients undergoing VIA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
 S\@ref(fig:supplement)A).
 
-Concomitant cardiovascular diseases were more prevalent in IVA than in non-IVA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare()`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were irrelevant in non-IVA cases (Fig.
+Concomitant cardiovascular diseases were more prevalent in VIA than in non-VIA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare()`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were irrelevant in non-VIA cases (Fig.
 \@ref(fig:figcofactors)).
@@
 The risk of severe anaphylaxis in patients concomitantly using ACE-I (as well as beta-blockers) could not be independently measured due to coexisting cardiovascular pathologies.
-ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) and there were no differences between IVA nad non-IVA  (Fig.
+ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) and there were no differences between VIA nad non-VIA  (Fig.
 \@ref(fig:figcofactors)C).
 Beta-blocker use was generally associated with higher severity of anaphylaxis and with the onset of cardiovascular symptoms (cardiac arrest, chest pain), but we saw no difference between groups `r summary(beta_fit)[["coefficients"]][3,4] %>% pval()`).
-Surprisingly, arrhythmia was more frequently reported in patients undergoing IVA with concomitant beta-blockers (Fig.
+Surprisingly, arrhythmia was more frequently reported in patients undergoing VIA with concomitant beta-blockers (Fig.
 \@ref(fig:figcofactors)C).
@@
 Baseline serum tryptase (BST) correlated with the severity of reactions (Ring and Messmer scale).
-Moreover, elevated BST was more prominently increasing the risk of severe anaphylaxis in IVA than in non-IVA (Fig.
+Moreover, elevated BST was more prominently increasing the risk of severe anaphylaxis in VIA than in non-VIA (Fig.
 \@ref(fig:figsymptoms)D and Fig.
 \@ref(fig:figcofactors)B).
-Cases with cardiac arrest were associated with an increased tryptase above 8 ng/ml, and this proportion was higher in IVA when compared to other elicitors.
-Loss of consciousness was a symptom associated with increased tryptase levels but only in patients with IVA (Fig.
+Cases with cardiac arrest were associated with an increased tryptase above 8 ng/ml, and this proportion was higher in VIA when compared to other elicitors.
+Loss of consciousness was a symptom associated with increased tryptase levels but only in patients with VIA (Fig.
 \@ref(fig:figcofactors)C).
@@
 
-We compared the Hymenoptera species responsible for triggering IVA according to the severity of the reaction and found no differences.
+We compared the Hymenoptera species responsible for triggering VIA according to the severity of the reaction and found no differences.
 Patients matched according to sex and age had similar severity of a reaction to known eliciting insects (p = 0.4128).
@@
 
-Outdoor physical exercise (e.g., jogging in the park) was more often associated with IVA than other triggers of anaphylaxis (`r ttexercise1`).
+Outdoor physical exercise (e.g., jogging in the park) was more often associated with VIA than other triggers of anaphylaxis (`r ttexercise1`).
 However, it was not associated with the severity of a reaction in these patients.
@@
 
-Patients who underwent IVA significantly less often received adrenaline treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
+Patients who underwent VIA significantly less often received adrenaline treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
 After adjusting both groups for similar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)).
-<!-- Antihistaminic drugs, on the other hand, were given more often in cases of IVA when compared to non-IVA. -->
-Patients with IVA received corticosteroids and antihistamines significantly more frequently than patients with anaphylaxis to other elicitors.
-On the other hand, adrenaline, beta-2 mimetics, and oxygen were given more often in non-IVA.
+<!-- Antihistaminic drugs, on the other hand, were given more often in cases of VIA when compared to non-VIA. -->
+Patients with VIA received corticosteroids and antihistamines significantly more frequently than patients with anaphylaxis to other elicitors.
+On the other hand, adrenaline, beta-2 mimetics, and oxygen were given more often in non-VIA.
 
-We discovered clusters of symptoms and therapy modes in IVA patients (association measured using phi).
+We discovered clusters of symptoms and therapy modes in VIA patients (association measured using phi).
 Cardiologic symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria, were treated more similarly than respiratory or gastrointestinal symptoms (Fig.
@@
 
-The most noticeable differences in the therapy of IVA vs. other forms of anaphylaxis were observed in the frequency of inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
+The most noticeable differences in the therapy of VIA vs. other forms of anaphylaxis were observed in the frequency of inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
 
@@
 
-Interestingly, adrenaline as a first emergency therapy was given less often in IVA cases when compared to other cases **if patients did not report the previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients who reported previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`).
+Interestingly, adrenaline as a first emergency therapy was given less often in VIA cases when compared to other cases **if patients did not report the previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients who reported previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`).
 
-Similarly, there were no differences in adrenaline use between IVA and non-IVA when severe reactions were taken into consideration (`r adren_severity_sev`).
-However, when we restricted the analysis to mild anaphylaxis cases - non-IVA patients received adrenaline more frequently than IVA (`r adren_severity_mild`).
+Similarly, there were no differences in adrenaline use between VIA and non-VIA when severe reactions were taken into consideration (`r adren_severity_sev`).
+However, when we restricted the analysis to mild anaphylaxis cases - non-VIA patients received adrenaline more frequently than VIA (`r adren_severity_mild`).
 
@@
 Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
-Their proportion of IVA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was as below 1% of cases during winter.
-Nevertheless, `r rare_spring_autumn$n %>% sum` cases of IVA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were triggered in March, April, and November.
-Yellow-jacket was the most prominent IVA-causing insect followed by bees.
-The IVA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in southern Europe.
+Their proportion of VIA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was as below 1% of cases during winter.
+Nevertheless, `r rare_spring_autumn$n %>% sum` cases of VIA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were triggered in March, April, and November.
+Yellow-jacket was the most prominent VIA-causing insect followed by bees.
+The VIA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in southern Europe.
 
@@
 
-In this study, we identified the prolonged seasonality of IVA, its clinical symptom-profile, and treatment patterns.
-The data unravel phenotypes of IVA, which may contribute to the development of tools incorporating both clinical and biomarker data for predicting the severity of future episodes of anaphylaxis.
+In this study, we identified the prolonged seasonality of VIA, its clinical symptom-profile, and treatment patterns.
+The data unravel phenotypes of VIA, which may contribute to the development of tools incorporating both clinical and biomarker data for predicting the severity of future episodes of anaphylaxis.
 
@@
 <!-- Cardio symptoms phenomenon -->
-IVA was more often associated with cardiologic symptoms non-IVA.
+VIA was more often associated with cardiologic symptoms non-VIA.
 Previous studies suggest [@Worm2018; @nittner2015insect; @Bonadonna2010] an essential link between the cardiovascular system and insect sting hypersensitivity. 
-IVA has been associated with Kounis syndrome (coronary arterial spasm induced by the release of mast cell mediators [@CDT1609; @sinkiewicz2008allergic]) and cardiac arrhythmias usually occurring in patients with preexisting heart disease [@Sharma2016].
+VIA has been associated with Kounis syndrome (coronary arterial spasm induced by the release of mast cell mediators [@CDT1609; @sinkiewicz2008allergic]) and cardiac arrhythmias usually occurring in patients with preexisting heart disease [@Sharma2016].
 
-The rate of concomitant cardiovascular diseases was higher in IVA than non-IVA; we observed them be an essential cofactor increasing the *risk of a severe reaction if insects elicited the anaphylaxis*.
+The rate of concomitant cardiovascular diseases was higher in VIA than non-VIA; we observed them be an essential cofactor increasing the *risk of a severe reaction if insects elicited the anaphylaxis*.
 This association was not significant in anaphylaxis elicited by other triggers.
-Notably, cardiac arrest occurred more frequently in patients with the elevated BST (> 8 ng/ml), especially in IVA.
-Nevertheless, the pathomechanism promoting cardiovascular symptoms in IVA requires further investigation.
+Notably, cardiac arrest occurred more frequently in patients with the elevated BST (> 8 ng/ml), especially in VIA.
+Nevertheless, the pathomechanism promoting cardiovascular symptoms in VIA requires further investigation.
 
 Of note, cardiologic symptoms like hypotension, collapse, or cardiac arrest grade higher on the Ring and Messmer scale than skin or gastrointestinal symptoms.
-Therefore, since IVA predisposes to cardiologic symptoms, it is likely to be associated with more severe anaphylaxis.
+Therefore, since VIA predisposes to cardiologic symptoms, it is likely to be associated with more severe anaphylaxis.
 
-Importantly, the absence of skin symptoms was associated with more severe IVA, which was also true after we excluded patients with a known diagnosis of mastocytosis.
-Subsequently, we correlated the severity of anaphylaxis with BST levels and identified an interaction between the absence of skin symptoms and IVA using generalized linear regression.
+Importantly, the absence of skin symptoms was associated with more severe VIA, which was also true after we excluded patients with a known diagnosis of mastocytosis.
+Subsequently, we correlated the severity of anaphylaxis with BST levels and identified an interaction between the absence of skin symptoms and VIA using generalized linear regression.
 
@@
 Zanotti et al.
-identified mast cell disorders in 17 out of 22 patients with IVA lacking skin symptoms [@Zanotti2015] and concluded that patients with BST of 7.95 ng/ml and IVA should undergo extensive diagnostic procedures.
+identified mast cell disorders in 17 out of 22 patients with VIA lacking skin symptoms [@Zanotti2015] and concluded that patients with BST of 7.95 ng/ml and VIA should undergo extensive diagnostic procedures.
 
@@
 
-Regarding the factors increasing risk of IVA - older patients experienced IVA more frequently.
+Regarding the factors increasing risk of VIA - older patients experienced VIA more frequently.
 Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].
@@
 <!-- Risk factors - Concomitant disease -->
-Similarly, we saw less IVA in patients with concomitant atopic diseases, as these patients more often present with food anaphylaxis [@Tham2019].
+Similarly, we saw less VIA in patients with concomitant atopic diseases, as these patients more often present with food anaphylaxis [@Tham2019].
 
 The influence of cardiologic medication could not be isolated from the effect of concomitant cardiologic conditions; therefore, we could not state if ACE-I and beta-blockers increased the severity of anaphylaxis.
-We did observe that there were no significant differences between IVA and non-IVA cases regarding the symptoms and severity of an episode with concomitant use of ACE-I or beta-blockers.
+We did observe that there were no significant differences between VIA and non-VIA cases regarding the symptoms and severity of an episode with concomitant use of ACE-I or beta-blockers.
 
 <!-- Therapy  -->
-Cases of IVA had been treated with adrenaline less often than the age- sex- and severity-matched cases of non-IVA.
-Moreover, administration of adrenaline did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patient experienced their first episode of IVA.
+Cases of VIA had been treated with adrenaline less often than the age- sex- and severity-matched cases of non-VIA.
+Moreover, administration of adrenaline did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patient experienced their first episode of VIA.
 The difference between groups was prominent for milder cases of anaphylaxis.
-The reason for this observation is unclear. To our knowledge, this is the only data on the comparative adrenaline usage in a case-controlled group of IVA vs. non-IVA.
+The reason for this observation is unclear. To our knowledge, this is the only data on the comparative adrenaline usage in a case-controlled group of VIA vs. non-VIA.
 
 Nevertheless, international guidelines of anaphylaxis state that adrenaline (i.m.) is the first-line agent in all diagnosed cases of anaphylaxis [@Muraro2014].
-Clinicians should not undermine the less severe IVA cases and treat them with adrenaline accordingly.
+Clinicians should not undermine the less severe VIA cases and treat them with adrenaline accordingly.
 
-Although there are no absolute contraindications for using adrenaline in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using adrenaline is hypersensitivity reaction with high blood pressure with tachycardia which might be present at the initial phase of IVA, due to a psychologic reflex.
-In theory, these less severe cases of IVA might exhibit some form of stress-related blood pressure increase, but we lack data to confirm or discard this theory.
+Although there are no absolute contraindications for using adrenaline in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using adrenaline is hypersensitivity reaction with high blood pressure with tachycardia which might be present at the initial phase of VIA, due to a psychologic reflex.
+In theory, these less severe cases of VIA might exhibit some form of stress-related blood pressure increase, but we lack data to confirm or discard this theory.
 
@@
 
-Based on our findings, insects are the most probable anaphylaxis elicitor in Europe during Summer-season, with IVA cases extending from early Spring to the end of Autumn.
+Based on our findings, insects are the most probable anaphylaxis elicitor in Europe during Summer-season, with VIA cases extending from early Spring to the end of Autumn.
 Detailed information on the seasonality of insect-elicited hypersensitivity reactions is scarce [@old].
@@
 The changing climate in Europe may influence the activity of Hymenoptera in this region in the upcoming years.
-However, in the period from 2007 - 2019, the perennial ratio of IVA to non-IVA cases has remained unchanged (data not shown).
+However, in the period from 2007 - 2019, the perennial ratio of VIA to non-VIA cases has remained unchanged (data not shown).
 
@@
 
-Nevertheless, because the European Anaphylaxis Registry has until now gathered over 12 000 cases of anaphylaxis - it was possible to perform a case-controlled analysis on a relatively large number of cases and investigate many aspects of IVA.
+Nevertheless, because the European Anaphylaxis Registry has until now gathered over 12 000 cases of anaphylaxis - it was possible to perform a case-controlled analysis on a relatively large number of cases and investigate many aspects of VIA.
 
@@
 
-Based on our results, IVA is a distinctive subtype of anaphylaxis, with specific symptom profile and risk factors.
-IVA cases should undergo therapy according to the international management guidelines, and adrenaline should be given more often in IVA.
+Based on our results, VIA is a distinctive subtype of anaphylaxis, with specific symptom profile and risk factors.
+VIA cases should undergo therapy according to the international management guidelines, and adrenaline should be given more often in VIA.
 
@@
 ```{r flow,fig.width=6,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale for case inclusion and exclusion from the final analysis.
-B, C, D: Age, sex, and severity distribution was matched in cases in both groups to allow for comparable results between IVA and non-IVA cases.
+B, C, D: Age, sex, and severity distribution was matched in cases in both groups to allow for comparable results between VIA and non-VIA cases.
 Two age-subsets of patients could be recognized based on the density plot of age (B)."}
@@
 
-```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors.
+```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (VIA) compared to other elicitors.
 A: Proportional presentation of specific reaction symptoms.
 B: High-level overview of involved organ systems and selected cofactors in the form of a radar plot.
-C: Interaction effect of IVA and lack of skin symptoms (urticaria or flushing) on the severity of anaphylaxis.
-D: Baseline Serum Tryptase (BST) increases with the severity of anaphylaxis a lot more prominently in IVA.", fig.width=8.5, fig.height = 6.2}
+C: Interaction effect of VIA and lack of skin symptoms (urticaria or flushing) on the severity of anaphylaxis.
+D: Baseline Serum Tryptase (BST) increases with the severity of anaphylaxis a lot more prominently in VIA.", fig.width=8.5, fig.height = 6.2}
 fig_symptoms2
@@
 
-```{r adrenalineuse, fig.cap = "Therapy in patients with IVA compared to other elicitors, cases matched according to sex, age, and severity of a reaction.
+```{r adrenalineuse, fig.cap = "Therapy in patients with VIA compared to other elicitors, cases matched according to sex, age, and severity of a reaction.
 A: Proportional use of therapy measures in both anaphylaxis groups.
@@
 Less common insect species grouped as 'other'.
-B: The density distribution of IVA cases to cases elicited by other triggers considering the patient's age.
-C: Geographical differences in the most common elicitors of IVA.
-Countries which reported less than 10 IVA cases were not illustrated in this figure.", fig.height=4, fig.width = 7}
+B: The density distribution of VIA cases to cases elicited by other triggers considering the patient's age.
+C: Geographical differences in the most common elicitors of VIA.
+Countries which reported less than 10 VIA cases were not illustrated in this figure.", fig.height=4, fig.width = 7}
 fig_basic
@@
 ```{r matchingresult, fig.cap="Results of matching the cohort according to sex and age in order to perform a case-controlled study.
-A: The original distribution of IVA and non-IVA cases according to age group and sex.
-B: The distribution of IVA and non-IVA after age and sex matching with the use of MatchIt package for R."}
+A: The original distribution of VIA and non-VIA cases according to age group and sex.
+B: The distribution of VIA and non-VIA after age and sex matching with the use of MatchIt package for R."}
 matching.result 
@@
 ```{r supplement, fig.width= 6, fig.height= 5, fig.cap="Symptoms of anaphylaxis.
-A: The association between cardiac arrest and concomitant mastocytosis in IVA and non-IVA.
+A: The association between cardiac arrest and concomitant mastocytosis in VIA and non-VIA.
 B: Hypotension frequency in two age groups of anaphylaxis.
-C: Crammer's V as the measure of association between groups anaphylaxis (IVA vs.
-non-IVA)."}
+C: Crammer's V as the measure of association between groups anaphylaxis (VIA vs.
+non-VIA)."}
 
@@
 ```{r sev, fig.cap="Severity of anaphylaxis in subgroups.
-The severity of patients with IVA in two age groups (left), according to elicitor type (center) and according to the responsible insect species (right)"}
+The severity of patients with VIA in two age groups (left), according to elicitor type (center) and according to the responsible insect species (right)"}
 severity_joined_brown %>%

### 2019-10-07 R/stat.R

@@ -2925,7 +2925,7 @@ adren_prev <- ANAscore_matched %>%
             facet.by = "q_160_ever_react",
             palette = palBW)+
   labs(x = "",
-       fill = "adrenaline administered?",
+       fill = "adrenaline\nadministered?",
        y = "proportion [%]")
 
 adren_severity <- ANAscore_matched %>%

### 2019-10-07 analysis/paper/paper.Rmd
V2.5 after corrections from Prof Worm

@@
 
-Patients who experienced insect-venom anaphylaxis were older than patients with anaphylaxis elicited by other elicitors (p < 0.0001), more often had concomitant mastocytosis (p < 0.0001) and cardiologic conditions (p < 0.0001) and females more often had concomitant thyroid diseases and less often suffered from a food allergy or atopic dermatitis.
+Patients who experienced insect-venom anaphylaxis were older than patients with anaphylaxis elicited by other elicitors (p < 0.0001), more often had concomitant mastocytosis (p < 0.0001) and cardiac conditions (p < 0.0001) and females more often had concomitant thyroid diseases and less often suffered from a food allergy or atopic dermatitis.
 
@@
 The data show that the management of insect-venom anaphylaxis may be improved.
-Patients with concomitant cardiologic conditions and these with hyperreactive mast cells require intensified prophylactic measures.
+Patients with concomitant cardiovascular conditions and these with hyperreactive mast cells require intensified prophylactic measures.
 
@@
 
-VIA was more frequent in adults and seniors than children and young adults.
-(Fig \@ref(fig:insectstime)B).
 When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B).
 
-## VIA is associated with cardiovascular symptoms
+## VIA is often associated with cardiovascular symptoms
 
-VIA showed a specific symptom pattern.
-Patients, who underwent VIA, more often experienced cardiologic symptoms (dizziness, reduced alertness, unconsciousness) than patients with other elicitors inducing anaphylaxis and less often had skin or gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A).
+VIA displayed a specific symptom pattern.
+Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, reduced alertness, unconsciousness) than patients with other elicitors inducing anaphylaxis and less often had skin or gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A).
 The difference in severe hypotension frequency was especially prominent in the younger age group (under 22, Fig.
@@
 
-## Increased baseline serum tryptase and concomitant cardiologic conditions increase the risk of severe VIA.
+## Increased baseline serum tryptase and concomitant cardiovascular conditions increase the risk of severe VIA.
 
@@
 The risk of severe anaphylaxis in patients concomitantly using ACE-I (as well as beta-blockers) could not be independently measured due to coexisting cardiovascular pathologies.
-ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) and there were no differences between VIA nad non-VIA  (Fig.
+ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) and there were no differences between VIA and non-VIA  (Fig.
 \@ref(fig:figcofactors)C).
-Beta-blocker use was generally associated with higher severity of anaphylaxis and with the onset of cardiovascular symptoms (cardiac arrest, chest pain), but we saw no difference between groups `r summary(beta_fit)[["coefficients"]][3,4] %>% pval()`).
-Surprisingly, arrhythmia was more frequently reported in patients undergoing VIA with concomitant beta-blockers (Fig.
+Beta-blocker use was generally associated with higher severity of anaphylaxis and with the onset of cardiovascular symptoms (cardiac arrest, chest pain), but showed no difference between groups `r summary(beta_fit)[["coefficients"]][3,4] %>% pval()`).
+Surprisingly, arrhythmia was more frequently reported in patients with VIA and concomitant beta-blockers (Fig.
 \@ref(fig:figcofactors)C).
@@
 Baseline serum tryptase (BST) correlated with the severity of reactions (Ring and Messmer scale).
-Moreover, elevated BST was more prominently increasing the risk of severe anaphylaxis in VIA than in non-VIA (Fig.
+Elevated BST was more prominently increasing the risk of severe anaphylaxis in VIA than in non-VIA (Fig.
 \@ref(fig:figsymptoms)D and Fig.
 \@ref(fig:figcofactors)B).
-Cases with cardiac arrest were associated with an increased tryptase above 8 ng/ml, and this proportion was higher in VIA when compared to other elicitors.
+Cases with cardiac arrest were associated with an increased tryptase above 8 ng/ml.
+This proportion was higher in VIA when compared to other elicitors.
 Loss of consciousness was a symptom associated with increased tryptase levels but only in patients with VIA (Fig.
 \@ref(fig:figcofactors)C).
-Based on the severity and symptom profile, we decided to use a tryptase cut off value of 8 ng/ml (Fig.
-\@ref(fig:figcofactors)B) rather than the currently used 11.5 ng/ml.
+Based on the severity and symptom profile, we defined a tryptase cut-off value of 8 ng/ml (Fig.
+\@ref(fig:figcofactors)B) instead of the currently used 11.5 ng/ml.
 
-We compared the Hymenoptera species responsible for triggering VIA according to the severity of the reaction and found no differences.
+The Hymenoptera species responsible for triggering VIA was independent from  the severity of the reaction.
 Patients matched according to sex and age had similar severity of a reaction to known eliciting insects (p = 0.4128).
@@
 We discovered clusters of symptoms and therapy modes in VIA patients (association measured using phi).
-Cardiologic symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria, were treated more similarly than respiratory or gastrointestinal symptoms (Fig.
+Cardiovascular symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria, were treated more similarly than respiratory or gastrointestinal symptoms (Fig.
 \@ref(fig:adrenalineuse)B).
-The treatment of these symptoms consisted of using adrenaline autoinjector (AAI), i.v.
+The treatment of these symptoms consisted of adrenaline autoinjector (AAI) use, i.v.
 adrenaline in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists.
@@
 
-The most noticeable differences in the therapy of VIA vs. other forms of anaphylaxis were observed in the frequency of inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
+The most intriguing difference in the therapy of VIA vs. other forms of anaphylaxis were regarding the frequency of inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
 
@@
   {chisq.test(.[,3:4])$p.value} %>% pval()
-
 ```
 
-Interestingly, adrenaline as a first emergency therapy was given less often in VIA cases when compared to other cases **if patients did not report the previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients who reported previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`).
+Adrenaline as a first-line treatment was given less often in VIA cases when compared to other cases **if patients did not report the previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients who reported previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`, Suppl.
+Fig.
+\@ref(fig:rfplot)).
 
 Similarly, there were no differences in adrenaline use between VIA and non-VIA when severe reactions were taken into consideration (`r adren_severity_sev`).
-However, when we restricted the analysis to mild anaphylaxis cases - non-VIA patients received adrenaline more frequently than VIA (`r adren_severity_mild`).
+However, when we restricted the analysis to mild anaphylaxis cases --- non-VIA patients received adrenaline more frequently than VIA (`r adren_severity_mild`).
 
-## Hymenoptera anaphylaxis shows seasonal patterns.
+## Hymenoptera anaphylaxis is a seasonal disease.
 
@@
 In this study, we identified the prolonged seasonality of VIA, its clinical symptom-profile, and treatment patterns.
-The data unravel phenotypes of VIA, which may contribute to the development of tools incorporating both clinical and biomarker data for predicting the severity of future episodes of anaphylaxis.
+The data unravel phenotypes of VIA, which may contribute to the development of tools incorporating both clinical <!--- and biomarker--->data for predicting the severity of future episodes of anaphylaxis.
 
@@
 <!-- Cardio symptoms phenomenon -->
-VIA was more often associated with cardiologic symptoms non-VIA.
+VIA was more often associated with cardiovascular symptoms than non-VIA.
 Previous studies suggest [@Worm2018; @nittner2015insect; @Bonadonna2010] an essential link between the cardiovascular system and insect sting hypersensitivity. 
@@
 
-The rate of concomitant cardiovascular diseases was higher in VIA than non-VIA; we observed them be an essential cofactor increasing the *risk of a severe reaction if insects elicited the anaphylaxis*.
+The rate of concomitant cardiovascular diseases was higher in VIA than non-VIA; we observed them be an essential cofactor increasing the risk of a severe reaction **if insects elicited the anaphylaxis**.
 This association was not significant in anaphylaxis elicited by other triggers.
@@
 
-Of note, cardiologic symptoms like hypotension, collapse, or cardiac arrest grade higher on the Ring and Messmer scale than skin or gastrointestinal symptoms.
-Therefore, since VIA predisposes to cardiologic symptoms, it is likely to be associated with more severe anaphylaxis.
+Of note, cardiovascular symptoms like hypotension, collapse, or cardiac arrest grade higher on the Ring and Messmer scale than skin or gastrointestinal symptoms.
+Therefore, since VIA predisposes to cardiovascular symptoms, it is likely to be associated with more severe anaphylaxis.
 
@@
 
-Regarding the factors increasing risk of VIA - older patients experienced VIA more frequently.
+Regarding the factors increasing the risk of VIA - older patients experienced VIA more frequently.
 Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].
@@
 
-The influence of cardiologic medication could not be isolated from the effect of concomitant cardiologic conditions; therefore, we could not state if ACE-I and beta-blockers increased the severity of anaphylaxis.
+The influence of cardiovascular medication could not be isolated from the effect of concomitant cardiovascular conditions; therefore, we could not state if ACE-I and beta-blockers increased the severity of anaphylaxis.
 We did observe that there were no significant differences between VIA and non-VIA cases regarding the symptoms and severity of an episode with concomitant use of ACE-I or beta-blockers.
@@
 
-Based on our results, VIA is a distinctive subtype of anaphylaxis, with specific symptom profile and risk factors.
+Based on our results, VIA is a distinctive subtype of anaphylaxis, with a specific symptom profile and risk factors.
 VIA cases should undergo therapy according to the international management guidelines, and adrenaline should be given more often in VIA.
 
-When evaluating the risk of future severe episodes - patients with BST over 8 ng/ml should undergo extensive diagnostic tests to exclude ISM or MCAS and should be provided with two adrenaline autoinjectors.
+When evaluating the risk of future severe episodes - patients with BST over 8 ng/ml should undergo extensive diagnostic tests to exclude ISM or MCAS and should be provided with two adrenaline autoinjectors for acute self management.
 
@@
 A: Odds ratios of eliciting severe anaphylaxis.
-B: Proportion of cases elicited by insects or other elicitors (upper panels) according to tryptase levels and cardiologic symptoms.
+B: Proportion of cases elicited by insects or other elicitors (upper panels) according to tryptase levels and cardiovascular symptoms.
 ", fig.width = 7, fig.height = 8}
@@
 
-```{r trypcardiocur, fig.cap="Tryptase levels in patients with concomitant cardiologic diseases. "}
+```{r trypcardiocur, fig.cap="Tryptase levels in patients with concomitant cardiovascular diseases. "}
 

### 2019-10-07 R/stat.R

@@ -2953,6 +2953,11 @@ adren_severity_mild <- adren_severity$data %>%
   {chisq.test(.[,3:4])$p.value} %>% pval()
 
 
+ANAscore_matched %>%
+  group_by(d_522_adren_agg, grouping, d_severity_rmr,q_561_hospital_admission_v6) %>%
+  summarize(n())
+
+
 ### Statistical modelling ####
 
 # As a last attempt for the analysis -

### 2019-10-10 analysis/paper/paper.Rmd
changed the Fig. 2 and results secion

@@
 
-^1^ Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,
+^1^ Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Hmuboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,
 
@@
 
-Hypersensitivity to insect venom is a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008].
-Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong systemic immunotherapy [@Sturm2017].
-
-The need for a more precise description of the diagnosis, identification of biomarkers, and phenotypes of anaphylaxis is discussed [@JimenezRodriguez2018].
-Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] for the diagnosis of anaphylaxis --- a better understanding of its clinical subtypes is required.
-
-Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and their infrequent and random occurence.
-Therefore registries, gathering clinical data from patients with a well documented (recent) history of anaphylaxis are a crucial tool in investigating this entity.
-
-This study aimed to identify clinical patterns of venom induced anaphylaxis (VIA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA).
-
 # Methods
@@
 Due to a large number of documented reactions in the European Anaphylaxis registry - we were able to match the VIA cases to selected non-VIA cases according to sex and age to reduce the comparison bias.
+When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B).
 Subsequently, we compared the management in both groups and matched the control group according to the severity of a reaction.
@@
 
-When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B).
-
 ## VIA is often associated with cardiovascular symptoms
@@
 VIA displayed a specific symptom pattern.
-Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, reduced alertness, unconsciousness) than patients with other elicitors inducing anaphylaxis and less often had skin or gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A).
-The difference in severe hypotension frequency was especially prominent in the younger age group (under 22, Fig.
-S\@ref(fig:supplement)).
+Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, hypotension, unconsciousness, reduced alertness) than patients with anaphylaxis due to other elicitors and less often presented with expiratory distress, rhinitis or diarrhoea (Fig.
+\@ref(fig:figsymptoms)A).
 
-The pattern of organ involvement showed similarities in gastrologic, skin, and respiratory symptoms, and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
-Although, patients undergoing VIA  less often had concomitant atopic diseases (Fig \@ref(fig:figsymptoms)B).
-
-Severe reactions of VIA were more prevalent in older patients in comparison to patients below 22 years, and in VIA cases vs.
-other elicitors (Fig.
-S\@ref(fig:sev)D).
-There were no differences in severity of reactions elicited by yellow-jackets and other insect species.
+Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrologic, skin, and respiratory systems, VIA more frequently involved over three organ systems (`r compare(c(tempdf[3:4,">3 organ\nsystems"]))`,<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
+Although, patients undergoing VIA  less often had concomitant atopic diseases  -->
+Fig. \@ref(fig:figsymptoms)B).
 
@@
 \@ref(fig:figsymptoms)D).
+
+## Severity 
+Severe reactions of VIA were more prevalent in older patients in comparison to patients below 22 years, and in VIA cases vs.
+other elicitors (Fig.
+S\@ref(fig:sev)D).
+There were no differences in severity of reactions elicited by yellow-jackets and other insect species.
 

### 2019-10-10 analysis/paper/paper_backupV2.5.Rmd
changed the Fig. 2 and results secion

--- 2019-10-07 analysis/paper/paper.Rmd
@@
 
-^1^ Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,
+  ^1^ Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Hmuboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,
 
@@
 Patients matched according to sex and age had similar severity of a reaction to known eliciting insects (p = 0.4128).
-<!-- This data, however, is not comparable to the refractory anaphylaxis cases as refractory anaphylaxis seems to be a distinct phenotype.  -->
+                                                                                                                                                                                                <!-- This data, however, is not comparable to the refractory anaphylaxis cases as refractory anaphylaxis seems to be a distinct phenotype.
+-->
 
@@
 After adjusting both groups for similar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)).
-<!-- Antihistaminic drugs, on the other hand, were given more often in cases of VIA when compared to non-VIA. -->
+                                                                                                                                                                                                <!-- Antihistaminic drugs, on the other hand, were given more often in cases of VIA when compared to non-VIA.
+-->
                                                                                                                                                                                                 Patients with VIA received corticosteroids and antihistamines significantly more frequently than patients with anaphylaxis to other elicitors.
@@
 adrenaline in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists.
-Corticosteroids, i.v. volume replacement and i.v. β-2 agonists formed another therapy mode.
+                                                                                                                                                                                                Corticosteroids, i.v.
+volume replacement and i.v.
+β-2 agonists formed another therapy mode.
 
-The most intriguing difference in the therapy of VIA vs. other forms of anaphylaxis were regarding the frequency of inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
+                                                                                                                                                                                                The most intriguing difference in the therapy of VIA vs.
+other forms of anaphylaxis were regarding the frequency of inhaled beta2-agonists and antihistamines (Fig.
+\@ref(fig:adrenalineuse)B).
 
@@
                                                                                                                                                                                                 spread(key = d_522_adren_agg,value = n) %>%
-  filter(q_160_ever_react=="no prev. ANA") %>% 
+                                                                                                                                                                                                filter(q_160_ever_react=="no prev.
+ANA") %>%
                                                                                                                                                                                                 {chisq.test(.[,3:4])$p.value} %>% pval()
@@
 The difference between groups was prominent for milder cases of anaphylaxis.
-The reason for this observation is unclear. To our knowledge, this is the only data on the comparative adrenaline usage in a case-controlled group of VIA vs. non-VIA.
+                                                                                                                                                                                                The reason for this observation is unclear.
+To our knowledge, this is the only data on the comparative adrenaline usage in a case-controlled group of VIA vs.
+non-VIA.
 

### 2019-10-10 R/stat.R

@@ -161,6 +161,8 @@ data4 <- data4 %>%
                          "high",
                          ">11.5"))))
 
+data4 <- data4 %>%
+  mutate(d_organ_sum_binom = ifelse(d_organ_sum<4,"1-3","4-5") %>% factor)
 ##### FINAL DATABASE RDB ######
 
 rdb <- data4[data4$reaction_type_brown=="anaphylaxis",]
@@ -402,8 +404,8 @@ iCodeDF %>% filter(difference=="greater")
 #     'q_114_loss_of_consciousness'='loss of consciousness';
 #     'q_112_abdominal_pain'='abdominal pain';
 #     'q_113_wheezing_expiratory_distress_v5'='expiratory distress'"),
-#     group = car::recode(group,"'fraq_1'='IVA';
-#                         'fraq_2'='non-IVA'")) %>%
+#     group = car::recode(group,"'fraq_1'='VIA';
+#                         'fraq_2'='non-VIA'")) %>%
 #                         {mutate(.,variableName = factor(.$variableName,
 #                                                         levels=unique(.$variableName)))} %>%
 #
@@ -425,7 +427,9 @@ temp1 <- lapply(list(quo(q_111),
             quo(q_112),
             quo(q_113),
             quo(q_114),
-            quo(atopy)),
+            quo(d_organ_sum_binom)#,
+            #quo(d_organ_sum)
+            ),
        funtempspider) %>%
   do.call(what = rbind) %>%
   as.data.frame()
@@ -446,23 +450,31 @@ tempdf <- temp %>% data.frame(stringsAsFactors = F) %>%
 {rbind(rep(1,length(.)),rep(0,length(.)),.[2,],temp2 %>% t() %>% {.[2,]})}
 tempdf<- tempdf %>%
   mutate_if(is.character,as.numeric)
-tempdf <- age_sex_matched %>%
-  group_by(grouping) %>%
-  summarize(tryptase = mean(q_212_tryptase_value_v5,na.rm = T)) %>%
-    {rbind(11,0,.[,2])} %>%
-    cbind(tempdf,.)
- radarchart(tempdf,
-             #axistype=1,
-           #seg=5, %>%
-           #plty=1,
-           #title="(axis=1, 5 segments, with specified vlabels)",
-           #vlcex=0.5
-           vlabels = c("skin",
-                      "gastrologic",
-                      "respiratory",
-                      "cardiac",
-                      "atopic",
[…]

### 2019-10-10 R/transform_ntprobnp_dataframe.R

@@ -1,0 +1,23 @@
+names(datED)[14] <- "ntprobnp"
+write.csv(datED,"ntprobnp_results.csv")
+data <- read.csv("ntprobnp_results.csv")
+data$ntprobnp <- as.character(data$ntprobnp) %>% as.numeric()
+require(tidyverse)
+
+data$yob <- data$b_patient_code %>% gsub(pattern = "m-(....)-.*",
+                             replacement = "\\1")
+data$yob <- paste0(data$yob,"-01-01")
+data$yob <- as.Date(data$yob)
+
+data$Datum.der.BE <- as.Date(data$Datum.der.BE, origin = "1900-01-01")
+time_length(difftime(data$Datum.der.BE, data$yob), "years")
+
+
+data %>%
+  ggplot(aes(x = q_114_hypotension_collapse_v5, y = ntprobnp))+
+  geom_boxplot()
+
+data %>%
+    ggplot(aes(x = nach.Müller, y = ntprobnp))+
+  geom_boxplot()
+

### 2019-10-14 analysis/paper/paper.Rmd
corrections to figures and results round 3

@@
 
+Based on the severity and symptom profile, we defined a sub-elevated baseline serum tryptase (BST) cut-off value of above 8 ng/ml (Fig.
+\@ref(fig:figcofactors)B) as being more sensitive indicator compared to the currently used 11.5 ng/ml.
+
 <!-- ## Data handling and statistics -->
@@
 
-## VIA is often associated with cardiovascular symptoms
+## VIA is associated with cardiovascular symptoms
 
@@
 
-## Absence of skin symptoms is associated with more severe episodes of VIA
+Younger patients (under 22) presented even more prominent differences in hypotension symptoms and more frequently reported gastrointestinal symptoms (e.g.
+vomiting) **when the reaction was not triggered by insect venom**.(Fig.
+\@ref(fig:figsymptoms)C-E)
 
-Patients with concomitant mastocytosis had significantly more often anaphylaxis without skin symptoms (namely urticaria and flushing) when compared to patients without diagnosed mastocytosis (`r (skin_symptoms$masto_tab$prop[c(2,1)]*100) %>% round(1) %>% paste0(collapse="% vs.
-")`%, `r pval(skin_symptoms$masto_chisq$p.value)`).
-We decided to further investigate non-mastocytic patients.
-In these patients urticaria or flushing as symptoms of anaphylaxis were present less often in VIA than anaphylaxis triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
-Morover, in patients without concomitant mastocytosis who were also lacking skin symptoms, anaphylaxis was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(2,6)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
+## Absence of skin symptoms during anaphylaxis is associated with more severe episodes of VIA
+
+We found that `r (skin_symptoms$masto_tab$prop[c(2)]*100) %>% round(1)`% of patients with concomitant mastocytosis had anaphylaxis without skin symptoms (i.e.
+urticaria and flushing) which was significantly more compared to patients without diagnosed mastocytosis (`r (skin_symptoms$masto_tab$prop[c(1)]*100) %>% round(1)`%, `r pval(skin_symptoms$masto_chisq$p.value)`) and this observarion was most prominently seen in VIA (Fig.
+\@ref(fig:figskinsymptoms)).
+
+Similarily, in non-mastocytic patients undergoing VIA, urticaria or flushing as symptoms of anaphylaxis were present less often than if anaphylaxis was triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
+Moreover, in this specific subgroup of patients (i.e.
+non-mastocytic and at the same time lacking urticaria or flushing) VIA was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
 ")`%, `r pval(skin_symptoms$masto_chisq$p.value)`% in non-VIA).
-We performed a factorial logistic regression modelling and confirmed the significant interaction effect between the presence of skin symptoms and the trigger (eiter VIA or non-VIA) on the severity of anaphylaxis (`r pval(skin_symptoms$severity_fit_skin_symptoms %>% coef %>% .[4,4])`).
-Therefore, non-mastocytic patients presenting without urticara or flushing tended to have more severe anaphylaxis when it was triggered by insects, but not other elicitors of anaphylaxis (Fig.
-\@ref(fig:figsymptoms)C).
 
-Following this observation, we decided to investigate the association of skin symptoms with the tryptase levels in patients who did not have a diagnosis of mastocytosis.
-For this comparison, we removed the cases with BST above 11.5 ng/ml, potentially indicating non-diagnosed mastocytosis.
-Similarly, tryptase levels correlated with the severity of anaphylaxis, were higher in VIA patients, and the effect was prominently visible for VIA patients (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`).
-We did not observe such an interaction in the non-VIA group (Fig.
-\@ref(fig:figsymptoms)D).
+After performing a factorial logistic regression modelling we confirmed the significant interaction effect between the presence of skin symptoms and the trigger (VIA) on the severity of anaphylaxis (`r pval(skin_symptoms$severity_fit_skin_symptoms %>% coef %>% .[4,4])`).
+To recapitulate, non-mastocytic patients presenting without urticaria or flushing tended to have more severe anaphylaxis when it was triggered by insects, but not other elicitors of anaphylaxis (Fig.
+\@ref(fig:figskinsymptoms)C, and Tab.
+S \@ref(tab:factoriallogit)).
 
-## Severity 
-Severe reactions of VIA were more prevalent in older patients in comparison to patients below 22 years, and in VIA cases vs.
+## Absence of skin symptoms correlates with BST levels and increases the risk of severe anaphylaxis specifically in IVA
+
+Following the above-mentioned observation, we investigated the association of skin symptoms with the tryptase levels in non-masctoscytic patients.
+For this model, we excluded the cases with known mastocytosis and with BST above 11.5 ng/ml, potentially indicating non-diagnosed mastocytosis.
+Similarily, 1) tryptase levels correlated with the severity of anaphylaxis, 2) were higher in VIA patients, and 3) the effect was more prominently visible in VIA (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`).
+We did not observe this effect in the non-VIA group (Fig.
+\@ref(fig:figskinsymptoms)D).
+
+## Sub-elevated baseline serum tryptase and concomitant cardiovascular conditions increase the risk of severe VIA.
+
+We evaluated potential cofactors increasing the risk of severe IVA.
+The factor most prominently associated with an increased risk of severe anaphylaxis was mastocytosis (Fig.
+\@ref(fig:figcofactors)), and it increased the risk of: 1) cardiac arrest and 2) loss of consciousness in patients undergoing VIA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
+S\@ref(fig:supplement)A and Fig.
+\@ref(fig:figcofactors)C).
+
+In line with the above, BST levels also correlated with the severity of anaphylaxis (on the Ring and Messmer scale) and, most importantly, sub-elevated BST was more prominently increasing the risk of severe anaphylaxis in VIA than in non-VIA (Fig.
+\@ref(fig:figsymptoms)D and Fig.
+\@ref(fig:figcofactors)B).
+
+Concomitant cardiovascular diseases were more prevalent in VIA than in non-VIA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare()`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were irrelevant in non-VIA cases (Fig.
+\@ref(fig:figcofactors)).
+Interestingly, BST values were increased in patients with concomitant cardiovasular diseases, irrespectively of the reaction severity (Fig.
+S\@ref(fig:trypcardiocur)).
+
+## Other cofactors of severe reactions (potentially could be shifted to th supplement)
+
+Severe reactions of VIA were more prevalent in patients above 22 years of age, and in VIA cases vs.
 other elicitors (Fig.
 S\@ref(fig:sev)D).
-There were no differences in severity of reactions elicited by yellow-jackets and other insect species.
+There were no differences in severity of reactions elicited by yellow-jackets and other
+insect species (p = 0.4128). .
+
+The effect of using ACE-I (as well as beta-blockers) on the risk of severe anaphylaxis could not be independently measured from the coexisting cardiovascular pathologies.
+ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) and there were no differences between VIA and non-VIA  (Fig.
+\@ref(fig:figcofactors)C).
+Beta-blocker use was generally associated with higher severity of anaphylaxis and with the onset of cardiovascular symptoms (cardiac arrest, chest pain), but showed no difference between groups `r summary(beta_fit)[["coefficients"]][3,4] %>% pval()`).
+Surprisingly, arrhythmia was more frequently reported in patients with VIA and concomitant beta-blockers (Fig.
+\@ref(fig:figcofactors)C).
 
@@
 
-## Increased baseline serum tryptase and concomitant cardiovascular conditions increase the risk of severe VIA.
-
-The factor most prominently associated with an increased risk of severe anaphylaxis was mastocytosis, and there were no differences in elicitor groups (Fig.
-\@ref(fig:figcofactors)).
-However, mastocytosis increased the risk of cardiac arrest in patients undergoing VIA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
-S\@ref(fig:supplement)A).
-
-Concomitant cardiovascular diseases were more prevalent in VIA than in non-VIA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare()`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were irrelevant in non-VIA cases (Fig.
-\@ref(fig:figcofactors)).
-Interestingly, BST values were increased in patients with concomitant cardiovasular diseases, irrespectively of the reaction severity (Fig.
-S\@ref(fig:trypcardiocur)).
-
-The risk of severe anaphylaxis in patients concomitantly using ACE-I (as well as beta-blockers) could not be independently measured due to coexisting cardiovascular pathologies.
-ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) and there were no differences between VIA and non-VIA  (Fig.
-\@ref(fig:figcofactors)C).
-Beta-blocker use was generally associated with higher severity of anaphylaxis and with the onset of cardiovascular symptoms (cardiac arrest, chest pain), but showed no difference between groups `r summary(beta_fit)[["coefficients"]][3,4] %>% pval()`).
-Surprisingly, arrhythmia was more frequently reported in patients with VIA and concomitant beta-blockers (Fig.
-\@ref(fig:figcofactors)C).
-
-Baseline serum tryptase (BST) correlated with the severity of reactions (Ring and Messmer scale).
-Elevated BST was more prominently increasing the risk of severe anaphylaxis in VIA than in non-VIA (Fig.
-\@ref(fig:figsymptoms)D and Fig.
-\@ref(fig:figcofactors)B).
-Cases with cardiac arrest were associated with an increased tryptase above 8 ng/ml.
-This proportion was higher in VIA when compared to other elicitors.
-Loss of consciousness was a symptom associated with increased tryptase levels but only in patients with VIA (Fig.
-\@ref(fig:figcofactors)C).
-Based on the severity and symptom profile, we defined a tryptase cut-off value of 8 ng/ml (Fig.
-\@ref(fig:figcofactors)B) instead of the currently used 11.5 ng/ml.
-
-The Hymenoptera species responsible for triggering VIA was independent from  the severity of the reaction.
-Patients matched according to sex and age had similar severity of a reaction to known eliciting insects (p = 0.4128).
 <!-- This data, however, is not comparable to the refractory anaphylaxis cases as refractory anaphylaxis seems to be a distinct phenotype.  -->
@@
 
-## Interplay of symptoms and medication
+## Therapy: VIA recived adrenaline less often than non-VIA
 
@@
 
-We discovered clusters of symptoms and therapy modes in VIA patients (association measured using phi).
+We discovered clusters of symptoms and therapy modes in VIA patients (association measured using phi coefficient).
 Cardiovascular symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria, were treated more similarly than respiratory or gastrointestinal symptoms (Fig.
@@
 
-The most intriguing difference in the therapy of VIA vs. other forms of anaphylaxis were regarding the frequency of inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
+The most prominent difference in the therapy of VIA vs. other forms of anaphylaxis were regarding the frequency of inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
 
@@
 
-## Hymenoptera anaphylaxis is a seasonal disease.
+## Hymenoptera anaphylaxis is a seasonal disease (may be moved to supplement).
 
@@
 
-In this study, we identified the prolonged seasonality of VIA, its clinical symptom-profile, and treatment patterns.
-The data unravel phenotypes of VIA, which may contribute to the development of tools incorporating both clinical <!--- and biomarker--->data for predicting the severity of future episodes of anaphylaxis.
+In this study, we identified distinctive clinical symptom-profile, and treatment patterns of venom induced anaphlaxis.
+The data unravel phenotypes of VIA, which may contribute to the development of tools incorporating clinical <!--- and biomarker--->data for predicting the severity of future episodes of anaphylaxis.
 
@@
 We thank all patients, parents, and their children for their support in providing data on the occurrence of anaphylaxis for this study.
+We thank Dr.
+rer.
+biol.
+Kristin Franke.
 We thank the study personnel for patients counseling and data entry, and we thank the members of The European Anaphylaxis Registry in detail:
@@
 D: Baseline Serum Tryptase (BST) increases with the severity of anaphylaxis a lot more prominently in VIA.", fig.width=8.5, fig.height = 6.2}
-fig_symptoms2
+fig_symptoms_3
+```
+
+#####
+
+```{r figskinsymptoms, fig.cap="Lack of skin symptoms during anaphylaxis is associated with more severe IVA."}
+
 ```
@@
 ```
+
+```{r factoriallogit, results="asis"}
+stargazer::stargazer(fit_skin_symptoms, type = "html",out = "stargazer_tab.doc")
+
+```

### 2019-10-14 R/make_flow.R

@@ -7,7 +7,7 @@ require(DiagrammeR)
 # Create a node data frame
 nodes <-
   create_node_df(
-    n = 7,
+    n = 8,
     type = "a",
     label= c(paste0("All cases in the\nEuropean Anaphylaxis Registry\n",
                     format(length(data4$b_submitdate),
@@ -15,51 +15,71 @@ nodes <-
              paste0("Formal anaphylaxis definition met\n",
                     format(length(which(data4$reaction_type_brown=="anaphylaxis")),
                            big.mark = ",")),
-             paste0("Insects as elicitors\n",
+             paste0("Sex- and age-matching of venom induced\nanaphylaxis and other known elicitors\n",
+                    format(nrow(age_sex_matched %>%
+                                  filter(grouping == "other")),
+                           big.mark = ",")),
+             paste0("Insects as elicitors\n\n",
                     format(length(which(data4$d_elicitor_gr5== "insects" &
                                    data4$reaction_type_brown=="anaphylaxis")),
                            big.mark = ",")),
-             paste0("Other elicitors\n",
-                    format(length(which(data4$d_elicitor_gr5!= "insects"&
-                                   data4$reaction_type_brown=="anaphylaxis")),
+             paste0("Other elicitors\n(matched control group)\n",
+                    format(nrow(age_sex_matched %>%
+                                  filter(grouping == "other")),
                            big.mark = ",")),
-             paste(paste0("Yellow-jackets = ",
+                    #format(length(which(data4$d_elicitor_gr5!= "insects"&
+                    #               data4$reaction_type_brown=="anaphylaxis")),
+                    #       big.mark = ",")),
+             paste(paste0("yellow-jackets = ",
                           format(length(which(data4$q_340_insects== "yellow jacket"&
                                          data4$reaction_type_brown=="anaphylaxis")),
                                  big.mark = ",")),
-                   paste0("Bees = ",
+                   paste0("bees = ",
                           format(length(which(data4$q_340_insects== "bee"&
                                          data4$reaction_type_brown=="anaphylaxis")),
                                  big.mark = ",")),
-                   paste0("Hornets = ",
+                   paste0("hornets = ",
                           length(which(data4$q_340_insects== "hornet"&
[…]

### 2019-10-14 R/stat.R

@@ -491,6 +491,8 @@ symptTabAdults <- makeTests(rdb=rdb[rdb$d_age>=22,],
                           .$symptoms %>% mutate(subset = "over 22")
 ex <- full_join(symptTabKids,symptTabAdults)
 
+ex %>%
+  filter(variableName=="q_112_vomiting")
 # ggplot(ex %>% filter(pval<1e-10) %>%
 #          tidyr::gather(key = "group",value = "Fraction", c("fraq_1","fraq_2")) %>%
 #          arrange(pval),
@@ -3038,6 +3040,27 @@ skin_symptoms$masto_chisq <-
   matrix(nrow = 2, byrow = T) %>%
   chisq.test()
 
+skin_symptoms$masto_grouping_tab <-
+  age_sex_matched %>%
+  filter(q_111 != "unknown",
+         q_410_masto_cur != "unknown") %>%
+  #dplyr::select(d_111_urti_flush, q_410_masto_cur) %>%
+  #gather("var", "value") %>%
+  count(grouping,d_111_urti_flush, q_410_masto_cur) %>%
+  group_by(q_410_masto_cur,grouping) %>% # now required with changes to dplyr::count()
+  mutate(prop = prop.table(n))
+
+skin_symptoms$masto_grouping_plot <- ggbarplot(skin_symptoms$masto_grouping_tab %>%
+            filter(d_111_urti_flush == "no"),
+          #x = "d_111_urti_flush",
+          y = "prop",
+          x = "q_410_masto_cur",
+          fill = "grouping",
+          palette = palBW,
+          position = position_dodge2())+
+  labs(x = "concomitant mastocytosis",
+       y = "proportion of patients\n without skin symptoms")
+
 skin_symptoms$severe_tab <-
 age_sex_matched %>%
   mutate(grouping = car::recode(grouping, "'insects'='VIA';
@@ -3507,3 +3530,63 @@ common.legend = T,
 heights = c(1,1.4),
 labels = c("A","B")
 )
+
+fig_symptoms_3 <-
+  ggarrange(
+    fig_symptoms,
+    ggpubr::ggarrange(
+      ex %>%
+        filter(variableName=="q_114_hypotension_collapse_v5") %>%
+        dplyr::select(variableName,fraq_1,fraq_2,subset) %>%
+        gather(key = "group", value = "proportion",2:3) %>%
[…]

### 2019-10-15 analysis/paper/paper.Rmd
Figures redo for version 3.0

@@
 ---
-title: "Phenotype and risk assesment of insect venom anaphylaxis: a case - control study of the European Anaphylaxis Registry"
+title: "Phenotype and risk assessment of insect venom anaphylaxis: a case - control study of the European Anaphylaxis Registry"
 author:
@@
 
+Hypersensitivity to insect venom is a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008].
+Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong systemic immunotherapy [@Sturm2017].
+The need for a more precise description of the diagnosis, identification of biomarkers, and phenotypes of anaphylaxis is discussed [@JimenezRodriguez2018].
+Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] for the diagnosis of anaphylaxis --- a better understanding of its clinical subtypes is required.
+
+Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and their infrequent and random occurence.
+Therefore registries, gathering clinical data from patients with a well documented (recent) history of anaphylaxis are a crucial tool in investigating this entity.
+
+This study aimed to identify clinical patterns of venom induced anaphylaxis (VIA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA).
+
 # Methods
 
-We searched the European Anaphylaxis Registry [@Grabenhenrich2016] database from March 2019 for anaphylaxis cases elicited by insect's venom.
+We searched the European Anaphylaxis Registry [@Grabenhenrich2016] (status from March 2019) for anaphylaxis cases elicited by insect's venom.
 The flowchart in Fig.
@@
 
-The final database consisted of `r countries %>% summarise(sum(n))` cases of insect-elicited anaphylaxis reported from allergy centers in `r countries %>% summarize(length(n))` countries.
 The diagnosis of anaphylaxis was based on the definition by NIAID/FAAN [@Sampson2006] and the severity according to Ring and Messmer Scale [@Ring1977].
@@
 
-The cases were either assigned to VIA group (if the triggering factor was an insect sting) or non-VIA group (if other elicitors triggered anaphylaxis).
-We compared the frequency of various symptoms, cofactors --- known to increase the risk of severe anaphylaxis, [@Worm2018] and management in both groups.
-
-Due to a large number of documented reactions in the European Anaphylaxis registry - we were able to match the VIA cases to selected non-VIA cases according to sex and age to reduce the comparison bias.
+Due to a large number of documented reactions in the European Anaphylaxis Registry - we were able to match the VIA with non-VIA cases according to sex and age to reduce the comparison bias. 
 When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B).
@@
 
-Based on the severity and symptom profile, we defined a sub-elevated baseline serum tryptase (BST) cut-off value of above 8 ng/ml (Fig.
-\@ref(fig:figcofactors)B) as being more sensitive indicator compared to the currently used 11.5 ng/ml.
+The final database consisted of `r countries %>% summarise(sum(n))` cases of venom induced anaphylaxis reported from allergy centers in `r countries %>% summarize(length(n))` countries and sex-and-age matched control group.
+We compared the frequency of various symptoms, cofactors --- known to increase the risk of severe anaphylaxis, [@Worm2018] and management in both groups.
+
+Based on the severity and symptom profile, we defined a sub-elevated baseline serum tryptase (BST) cut-off value of 8 - 11.5 ng/ml (Fig. \@ref(fig:figcofactors)B).
 
@@
 We also performed an association study of therapeutic interventions and symptoms.
-The resulting phi values are presented in a heatmap with automatic clustering using Ward’s Agglomerative Hierarchical Clustering Method with Euclidean distance [@heatmaply].
+The resulting phi values are presented in a heatmap with automatic clustering using Ward’s Agglomerative Hierarchical Clustering with Euclidean distances [@heatmaply].
 
@@
 
-<!-- This data, however, is not comparable to the refractory anaphylaxis cases as refractory anaphylaxis seems to be a distinct phenotype.  -->
+## Therapy: VIA recived adrenaline less often than non-VIA
+
+```{r mngmttests}
+ttadren1 <- ANAscore_matched %>% 
+  group_by(d_522_adren_agg,grouping) %>% 
+  summarize(n = n()) %>% 
+  spread(key = grouping, value = n) %>% 
+  {chisq.test(.[,2:3])$p.value} %>% 
+  pval()
+```
+
+Patients who underwent VIA significantly less often received adrenaline treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
+After adjusting both groups for similar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)).
+<!-- Antihistaminic drugs, on the other hand, were given more often in cases of VIA when compared to non-VIA. -->
+```{r }
+pvaladrenprev <- adren_prev$data %>% 
+  spread(key = d_522_adren_agg,value = n) %>% 
+  filter(q_160_ever_react=="reacted previously") %>% 
+  {chisq.test(.[,3:4])$p.value} %>% pval()
+
+pvaladrenprev_no <- adren_prev$data %>% 
+  spread(key = d_522_adren_agg,value = n) %>% 
+  filter(q_160_ever_react=="no prev. ANA") %>% 
+  {chisq.test(.[,3:4])$p.value} %>% pval()
+```
+
+Having a history of anaphylaxis also infuenced the therapy of a current episode.
+Adrenaline as a first-line treatment was given less often in VIA cases when compared to other cases **if patients did not report the previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients who reported previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`, Suppl.
+Fig.
+\@ref(fig:rfplot)).
+Similarly, there were no differences in adrenaline use between VIA and non-VIA when severe reactions were taken into consideration (`r adren_severity_sev`).
+However, when we restricted the analysis to mild anaphylaxis cases --- non-VIA patients received adrenaline more frequently than VIA (`r adren_severity_mild`).
+
+Patients with VIA received corticosteroids and antihistamines significantly more frequently than patients with anaphylaxis to other elicitors.
+On the other hand, adrenaline, beta-2 mimetics, and oxygen were given more often in non-VIA.
+
+We discovered clusters of symptoms and therapy modes in VIA patients (association measured using phi coefficient).
+Cardiovascular symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria, were treated more similarly than respiratory or gastrointestinal symptoms (Fig.
+\@ref(fig:adrenalineuse)B).
+The treatment of these symptoms consisted of adrenaline autoinjector (AAI) use, i.v.
+adrenaline in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists.
+Corticosteroids, i.v. volume replacement and i.v. β-2 agonists formed another therapy mode. 
+The most prominent difference in the therapy of VIA vs. non-VIA were regarging the  inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
+
+## Hymenoptera anaphylaxis is a seasonal disease (may be moved to supplement).
+
+Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
+Their proportion of VIA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was as below 1% of cases during winter.
+Nevertheless, `r rare_spring_autumn$n %>% sum` cases of VIA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were triggered in March, April, and November.
+Yellow-jacket was the most prominent VIA-causing insect followed by bees.
+The VIA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in southern Europe.
 
@@
 (`r ttexercise2`).
-
-## Therapy: VIA recived adrenaline less often than non-VIA
-
-```{r mngmttests}
-ttadren1 <- ANAscore_matched %>% 
-  group_by(d_522_adren_agg,grouping) %>% 
-  summarize(n = n()) %>% 
-  spread(key = grouping, value = n) %>% 
-  {chisq.test(.[,2:3])$p.value} %>% 
-  pval()
-```
-
-Patients who underwent VIA significantly less often received adrenaline treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
-After adjusting both groups for similar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)).
-<!-- Antihistaminic drugs, on the other hand, were given more often in cases of VIA when compared to non-VIA. -->
-Patients with VIA received corticosteroids and antihistamines significantly more frequently than patients with anaphylaxis to other elicitors.
-On the other hand, adrenaline, beta-2 mimetics, and oxygen were given more often in non-VIA.
-
-We discovered clusters of symptoms and therapy modes in VIA patients (association measured using phi coefficient).
-Cardiovascular symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria, were treated more similarly than respiratory or gastrointestinal symptoms (Fig.
-\@ref(fig:adrenalineuse)B).
-The treatment of these symptoms consisted of adrenaline autoinjector (AAI) use, i.v.
-adrenaline in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists.
-Corticosteroids, i.v. volume replacement and i.v. β-2 agonists formed another therapy mode.
-
-The most prominent difference in the therapy of VIA vs. other forms of anaphylaxis were regarding the frequency of inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
-
-```{r }
-pvaladrenprev <- adren_prev$data %>% 
-  spread(key = d_522_adren_agg,value = n) %>% 
-  filter(q_160_ever_react=="reacted previously") %>% 
-  {chisq.test(.[,3:4])$p.value} %>% pval()
-
-pvaladrenprev_no <- adren_prev$data %>% 
-  spread(key = d_522_adren_agg,value = n) %>% 
-  filter(q_160_ever_react=="no prev. ANA") %>% 
-  {chisq.test(.[,3:4])$p.value} %>% pval()
-```
-
-Adrenaline as a first-line treatment was given less often in VIA cases when compared to other cases **if patients did not report the previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients who reported previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`, Suppl.
-Fig.
-\@ref(fig:rfplot)).
-
-Similarly, there were no differences in adrenaline use between VIA and non-VIA when severe reactions were taken into consideration (`r adren_severity_sev`).
-However, when we restricted the analysis to mild anaphylaxis cases --- non-VIA patients received adrenaline more frequently than VIA (`r adren_severity_mild`).
-
-## Hymenoptera anaphylaxis is a seasonal disease (may be moved to supplement).
-
-Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
-Their proportion of VIA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was as below 1% of cases during winter.
-Nevertheless, `r rare_spring_autumn$n %>% sum` cases of VIA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were triggered in March, April, and November.
-Yellow-jacket was the most prominent VIA-causing insect followed by bees.
-The VIA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in southern Europe.
 
@@
 Of note, cardiovascular symptoms like hypotension, collapse, or cardiac arrest grade higher on the Ring and Messmer scale than skin or gastrointestinal symptoms.
-Therefore, since VIA predisposes to cardiovascular symptoms, it is likely to be associated with more severe anaphylaxis.
+Since VIA predisposes to cardiovascular symptoms, it is likely to be associated with more severe anaphylaxis.
 
@@
 biol.
-Kristin Franke.
+Kristin Franke for insight about the manuscript.
 We thank the study personnel for patients counseling and data entry, and we thank the members of The European Anaphylaxis Registry in detail:
@@
 
-```{r flow,fig.width=6,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale for case inclusion and exclusion from the final analysis.
+```{r flow,fig.width=5.2,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale for case inclusion and exclusion from the final analysis.
 B, C, D: Age, sex, and severity distribution was matched in cases in both groups to allow for comparable results between VIA and non-VIA cases.
@@
 ```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (VIA) compared to other elicitors.
-A: Proportional presentation of specific reaction symptoms.
+A: Proportional presentation of specific reaction symptoms in VIA and non-VIA according to cardiovascular (cardio.), gastroenterologic (gastro.) and respirator (resp.) organ systems.
 B: High-level overview of involved organ systems and selected cofactors in the form of a radar plot.
-C: Interaction effect of VIA and lack of skin symptoms (urticaria or flushing) on the severity of anaphylaxis.
+C: difference in symptoms of VIA among patients under 22 and over 22 years of age.
+* denotes significant differences between groups.
+(other figure:Interaction effect of VIA and lack of skin symptoms (urticaria or flushing) on the severity of anaphylaxis.
 D: Baseline Serum Tryptase (BST) increases with the severity of anaphylaxis a lot more prominently in VIA.", fig.width=8.5, fig.height = 6.2}
@@
 
-```{r figskinsymptoms, fig.cap="Lack of skin symptoms during anaphylaxis is associated with more severe IVA."}
-
+```{r figskinsymptoms, fig.cap="Lack of skin symptoms during anaphylaxis is associated with more severe IVA.", fig.height=5}
+fig_skinsymptoms
 ```
@@
 B: Proportion of cases elicited by insects or other elicitors (upper panels) according to tryptase levels and cardiovascular symptoms.
-", fig.width = 7, fig.height = 8}
+", fig.width = 6.6, fig.height = 9}
 
@@
 A: Proportional use of therapy measures in both anaphylaxis groups.
-B: Heatmap visualizing the association of symptoms and corresponding treatment -  presented as a scaled correlation coefficient (phi).", fig.height=9, fig.width = 7}
+B:   C: Heatmap visualizing the association of symptoms and corresponding treatment -  presented as a scaled correlation coefficient (phi).
+* - p value < 0.05 after false discovery rate correction.
+", fig.height=9, fig.width = 7}
 fig_adrenalineuse

### 2019-10-15 R/functions.R

@@ -415,14 +415,14 @@ makeTests <- function(groups,rdb){
                  dplyr::select(variableName) %>% pull %>% as.character()]
   ts <- lapply(varDF, # Check which tables are givin an error with chisq test
                function(x){ # You may implement also fisher tests here later
-                 table(x,grouping)
+                 table(x,grouping)[1:2,] # get rid of the unknown row
                }) %>%
     lapply(function(x){
       length(which(x==0))
     }) %>% lapply(function(x){
-      ifelse(x==0,T,F)
+      ifelse(x==0,T,F) # Gives True output if there are no 0 counts in any table cells
     }) %>% unlist()
-  varDF[,ts]
+ # varDF[,ts]
   #ChiF(varDF[ts],grouping) # put this into the dataframe
   variableSelectionTab %<>% full_join(ChiF(varDF[,ts],grouping),
                                       by = c("variableName"="variables")
@@ -447,9 +447,10 @@ makeTests <- function(groups,rdb){
   }) %>% do.call(what=rbind) %>% as.data.frame() %>%
   {data.frame(variableName=rownames(.),.,stringsAsFactors = F)}
   # Join with the maind DF wit h substituting the missing values
-  variableSelectionTab %<>% full_join( ttsts, by = "variableName") %>%
-    mutate(pval = coalesce(pval.y, pval.x)) %>%
-    dplyr::select(-c(pval.x,pval.y))
+  variableSelectionTab %<>%
+    full_join( ttsts)
+    #mutate(pval = coalesce(pval.y, pval.x)) %>%
+    #dplyr::select(-c(pval.x,pval.y))
   # Perform Kruskall
   varDF <- rdb[,variableSelectionTab %>% filter(fun=="Kruskal-Wallis")
                %>% dplyr::select(variableName) %>% pull %>% as.character()]
@@ -480,9 +481,9 @@ makeTests <- function(groups,rdb){
     {data.frame(variableName=rownames(.),
                 .,
                 stringsAsFactors = F)}
-  variableSelectionTab %<>% full_join(ktest, by = "variableName") %>%
-    mutate(pval = coalesce(pval.y, pval.x)) %>%
-    dplyr::select(-c(pval.x,pval.y))
+  variableSelectionTab %<>% full_join(ktest) #%>%
+    #mutate(pval = coalesce(pval.y, pval.x)) %>%
+    #dplyr::select(-c(pval.x,pval.y))
   return(variableSelectionTab)
 }
 

### 2019-10-15 R/stat.R

@@ -580,7 +580,7 @@ checkVarTab(data = rdb,
 #### The most associated differences
 AssociatedVars <- testInsectsbinomial %>%
   arrange(desc(Cramer)) %>%
-  dplyr::select(-c(2:5,11:16)) %>%
+  dplyr::select(-c(2:5,12:16)) %>%
   filter(Cramer>0.25)
 
 #### Crammer Plot ####
@@ -877,6 +877,7 @@ ANAscore_matched <-
                    filter(d_522_adren_agg %in% c("yes", "no")),
                  grouping_var ="grouping",
                  matching_vars = c("ANAscore","d_age"),
+                 #matching_vars = c("d_severity_rm","d_age"),
                  grouping_var_level = "insects",
                  df = T)
 testANAscoreMatched <- makeTests(groups = "grouping",
@@ -889,6 +890,7 @@ testANAscoreMatched <- makeTests(groups = "grouping",
          {split(.,.$section)}
 
 
+
 countYesManagment <- function(var){
   tbl <- ANAscore_matched %>%
     group_by(grouping,!! sym(var))  %>%
@@ -899,11 +901,22 @@ countYesManagment <- function(var){
              nonVIA = tbl[2,3])
 }
 
-plotManagement <- testANAscoreMatched$management$variableName %>%
-  map(countYesManagment) %>%
-  do.call(what = rbind)  %>%
-  tidyr::gather(key = "grouping", value = "positive",2:3) %>%
-  filter(variable %in% c(
+countYesManagment2 <- function(var){
+  tbl <- ANAscore_matched %>%
+    group_by(grouping,!! sym(var))  %>%
+    summarize(n = n()) %>%
+    tidyr::spread(d_522_adren_agg,n)
+    filter(!!sym(var) == "yes")
+  data.frame(variable = names(tbl)[2],
+             VIA_no = tbl[tbl$grouping=="insects","no"],
+             nonVIA_no = tbl[2,3])
+}
+
+plotManagement <- {df_temp <- testANAscoreMatched$management %>%
+  #map(countYesManagment) %>%
+  #do.call(what = rbind)  %>%
+  #tidyr::gather(key = "grouping", value = "positive",2:3) %>%
+  filter(variableName %in% c(
     "d_522_adren_agg",
     "q_522_antih_iv",
     "q_522_cortico_iv",
@@ -915,7 +928,7 @@ plotManagement <- testANAscoreMatched$management$variableName %>%
   )) %>%
   mutate(
[…]

### 2019-10-16 analysis/paper/paper.Rmd
Abstract and introduction

@@
 # Abstract
-Insect-venom elicited anaphylaxis is a common hypersensitivity reaction which may be life-threatening.
-Using data from the European Anaphylaxis Registry (`r nrow(data)` cases) we identified `r {data$d_elicitor_gr5=="insects"} %>% sum()` with insect-venom elicited anaphylaxis and analyzed these in comparison to anaphylaxis triggered by other elicitors (n = `r {data$d_elicitor_gr5!="insects"} %>% sum()`).
+**Introduction:** Insect-venom elicited anaphylaxis is a common, life-threatening, hypersensitivity reaction.
+It is bound with specific: 1) symptom profile, 2) cofactors and 3) management.
+Identifying the diffeneces in subtypes of anaphylaxis is crucial for future management guidelines and personalized medicine approach.
 
-The data show that `r {{rdb$q_340_insects=="yellow jacket"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`% of all insect elicited cases were elicited by yellow jackets, followed by bees (`r {{rdb$q_340_insects=="bee"} %>% sum(na.rm=T)/sum({!is.na(rdb$q_340_insects)})*100} %>% {round(.,1)}`%).
-The venom triggered cases occurred mostly in outdoor places (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(4),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%) patients' homes (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(8),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%) or urban places (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(q_152_location) %>% summarize(n=n()) %>% {.[c(5),2] %>% sum()/(.[,2] %>% sum)*100} %>% round(1)`%).
+**Methods:** Using data from the European Anaphylaxis Registry (`r nrow(data)` cases) we identified `r length(which(data4$d_elicitor_gr5== "insects" & data4$reaction_type_brown=="anaphylaxis"))` with insect-venom elicited anaphylaxis and analyzed these in comparison to sex-and-age-matched anaphylaxis cases triggered by other elicitors (n = `r nrow(age_sex_matched %>% filter(grouping == "other"))`).
 
-Regarding the symptoms: skin, gastrointestinal and respiratory manifestations occurred less frequently in insect induced cases of anaphylaxis, but cardiovascular symptoms (with hypotension, collapse, and loss of consciousness) were more frequent.
-Intramuscular adrenaline (as a first-line therapy) was administered significantly less often in insect venom elicited cases (`r rdb %>% filter(!is.na(q_340_insects)) %>% group_by(d_522_adren_agg) %>% summarize(n=n()) %>% {.[2,2]/(.[1,2])*100} %>% round(1)`% vs `r rdb %>% filter(is.na(q_340_insects)) %>% group_by(d_522_adren_agg) %>% summarize(n=n()) %>% {.[2,2]/(.[1,2])*100} %>% round(1)`, p < 0.0001).
-The mortality rate in insect anaphylaxis was comparable to other cases (0.295%, p = 0.174).
-
-Patients who experienced insect-venom anaphylaxis were older than patients with anaphylaxis elicited by other elicitors (p < 0.0001), more often had concomitant mastocytosis (p < 0.0001) and cardiac conditions (p < 0.0001) and females more often had concomitant thyroid diseases and less often suffered from a food allergy or atopic dermatitis.
+**Results:** Venom induced anaphylaxis (VIA) more frequently involed over 3 organ systems and was associated with cardiovascular symptoms what was most prominently visible in patients under 22 years of age.
+Absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase and was associated with an increased risk of severe reaction specificaly in VIA.
+Intramuscular or intravenous adrenaline was administered significantly less often in IVA especially without prior history of anaphylaxis.
+Subelevated baseline serum tryptase (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis and this effect was most prominantly visible for IVA.
 
  <!-- MORE DISCUSSION!!!!!-->
-The data show that the management of insect-venom anaphylaxis may be improved.
-Patients with concomitant cardiovascular conditions and these with hyperreactive mast cells require intensified prophylactic measures.
+**Conclusions:** Due to the specific symptom profile, frequently affecting cardiovascular system, patients undergoing IVA are undertreated with adrenaline.
+Lack of urticaria or flushing during previous anaphylaxis (especialy in combination with subelevated baseline serum tryptase levels) is associated with severe reactions.
 
@@
 Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and their infrequent and random occurence.
-Therefore registries, gathering clinical data from patients with a well documented (recent) history of anaphylaxis are a crucial tool in investigating this entity.
+Therefore registries, gathering clinical data from patients with a well documented (recent) history of anaphylaxis are crucial in investigating this entity.
 
@@
 The flowchart in Fig.
-\@ref(fig:flow) represents the detailed case-selection process.
+\@ref(fig:flow)A represents the detailed case-selection process.
 
 The diagnosis of anaphylaxis was based on the definition by NIAID/FAAN [@Sampson2006] and the severity according to Ring and Messmer Scale [@Ring1977].
-Grades III and IV (presenting with significant hypoxia, hypotension, confusion, collapse and loss of consciousness, or incontinence or cardiac arrest) were considered severe.
+Grades III and IV (presenting with significant hypoxia, hypotension, confusion and loss of consciousness, or incontinence or cardiac arrest) were considered severe.
 
@@
 
-Younger patients (under 22) presented even more prominent differences in hypotension symptoms and more frequently reported gastrointestinal symptoms (e.g.
-vomiting) **when the reaction was not triggered by insect venom**.(Fig.
+Younger patients (under 22) presented even more prominent differences in hypotension symptoms and significantly less frequently reported gastrointestinal symptoms (e.g.
+vomiting) when the reaction was triggered by insect venom.(Fig.
 \@ref(fig:figsymptoms)C-E)

### 2019-10-16 additional analyses/Shruti_sample_severity.R

@@ -1,41 +1,0 @@
-#Shruti serum severity anlysis
-d1 <- readxl::read_excel("../../Downloads/Kopie von Seren_ANA_Insekten ab 172_für Shruti.xls")
-
-res <- data %>%
-  filter(data$b_patient_code %in% c("w-1969-l-09-b-05-26 "))
-
-data %>%
-  select(b_patient_code) %>%
-  filter(data$b_patient_code %>% grepl(pattern = "m-1986-t"))
-data$b_patient_code[1:6] %>% as.character()
-data$b_patient_code %<>% gsub(pattern = "[ ]+",replacement = "")
-
-res <- data %>%
-  filter(b_patient_code %in% (d1 %>%
-                                     filter(!is.na(`AR-ID`)) %>%
-                                     select(`AR-ID`) %>% pull())) %>%
-  select(b_patient_code,d_severity_rm,ANAscore)
-res %>%
-  write.csv(file = "file_172_severity.csv")
-
-
-
-check<-  (d1 %>%
-         filter(!is.na(`AR-ID`)) %>%
-         select(`AR-ID`) %>% pull())
-check[which (!(check %in% data$b_patient_code))]
-
-
-##### Second file
-d2 <- readxl::read_excel("../../Downloads/Seren_ANA_Insekten_03.01.2019 für Shruti.xlsx",skip = 1)
-
-res <- data %>%
-  filter(b_patient_code %in% (d2 %>%
-                                filter(!is.na(`AR-ID`),
-                                       `ANA-Code`>98) %>%
-                                select(`AR-ID`) %>% pull())) %>%
-  select(b_patient_code,d_severity_rm,ANAscore)
-res %>%
-  write.csv(file = "file_SEREN_ANA.csv")
-
-

### 2019-10-16 additional analyses/VAS.R

@@ -1,97 +1,0 @@
-#library(plyr)
-require(tidyverse)
-#require(dplyr)
-
-## Summarizes data.
-## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
-##   data: a data frame.
-##   measurevar: the name of a column that contains the variable to be summariezed
-##   groupvars: a vector containing names of columns that contain grouping variables
-##   na.rm: a boolean that indicates whether to ignore NA's
-##   conf.interval: the percent range of the confidence interval (default is 95%)
-summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
-                      conf.interval=.95, .drop=TRUE) {
-  #library(plyr)
-
-  # New version of length which can handle NA's: if na.rm==T, don't count them
-  length2 <- function (x, na.rm=FALSE) {
-    if (na.rm) sum(!is.na(x))
-    else       length(x)
-  }
-
-  # This does the summary. For each group's data frame, return a vector with
-  # N, mean, and sd
-  datac <- ddply(data, groupvars, .drop=.drop,
-                 .fun = function(xx, col) {
-                   c(N    = length2(xx[[col]], na.rm=na.rm),
-                     mean = mean   (xx[[col]], na.rm=na.rm),
-                     sd   = sd     (xx[[col]], na.rm=na.rm)
-                   )
-                 },
-                 measurevar
-  )
-
-  # Rename the "mean" column
-  datac <- rename(datac, c("mean" = measurevar))
-
-  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
-
-  # Confidence interval multiplier for standard error
-  # Calculate t-statistic for confidence interval:
-  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
-  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
-  datac$ci <- datac$se * ciMult
-
-  return(datac)
-}
-
-
-d <- readxl::read_xlsx("analysis/data/raw_data/NORA Konferenz Auswertung Fragebogen.xlsx")
-d %>% str()
-names(d)
-clean <- d %>%
-  select(1:44) %>%
[…]

### 2019-10-16 additional analyses/npprobnp/transform_ntprobnp_dataframe.R

@@ -1,23 +1,0 @@
-names(datED)[14] <- "ntprobnp"
-write.csv(datED,"ntprobnp_results.csv")
-data <- read.csv("ntprobnp_results.csv")
-data$ntprobnp <- as.character(data$ntprobnp) %>% as.numeric()
-require(tidyverse)
-
-data$yob <- data$b_patient_code %>% gsub(pattern = "m-(....)-.*",
-                             replacement = "\\1")
-data$yob <- paste0(data$yob,"-01-01")
-data$yob <- as.Date(data$yob)
-
-data$Datum.der.BE <- as.Date(data$Datum.der.BE, origin = "1900-01-01")
-time_length(difftime(data$Datum.der.BE, data$yob), "years")
-
-
-data %>%
-  ggplot(aes(x = q_114_hypotension_collapse_v5, y = ntprobnp))+
-  geom_boxplot()
-
-data %>%
-    ggplot(aes(x = nach.Müller, y = ntprobnp))+
-  geom_boxplot()
-

### 2019-10-16 additional analyses/test_diagramm.R

@@ -1,56 +1,0 @@
-nodes <-
-  create_node_df(
-    n = 6,
-    type = c("a","b","b","c","d","d"),
-    label= c("insect sting reaction in the ER",
-             "anaphylaxis",
-             "local reaction",
-             "treatment",
-             "acute serum samples\nwithin 2h of reaction",
-             "baseline serum samples\nafter > 2 weeks"
-             ),
-
-    #color = c("red", "green",
-    #          "grey", "blue"),
-    #value = c(3.5, 2.6, 9.4, 2.7),
-    shape = "rectangle",
-    width = c(2,1,1,1,2,2),
-    x = c(0, -.6, .6, -0.6,-0, -0),
-    y = c(0,-0.5,-0.5,  -1,-1.5,-2),
-    height = c(0.3,0.3,0.3,0.3,0.4,0.4))
-
-
-edges <- create_edge_df(from = c(1,1,2,4,5,3),
-                        to =   c(2,3,4,5,6,5),
-                        headport = c(5,0,0,0,0,0))
-# Add the node data frame to the
-# graph object to create a graph
-# with nodes
-gr <- create_graph() %>%
-  add_node_df(node_df = nodes) %>%
-  add_edge_df(edge_df = edges) %>%
-  add_global_graph_attrs(
-    attr = "ortho",
-    value = 1,
-    attr_type = "graph") %>%
-  add_global_graph_attrs(value="black",
-                         attr = "fontcolor",
-                         attr_type = "node") #%>% render_graph()
-  export_graph(file_name = "analysis/figures/flow.png", file_type = "png",
-               title = NULL,
-               width = NULL,
-               height = NULL)
-
-  library(network)
-  library(sna)
-  n <- network(rgraph(6, tprob = 0.2), directed = FALSE)
-
-library(ggnetwork)
-
-  df <- ggnetwork(n, layout = "fruchtermanreingold", cell.jitter = 0.75)
-  df$x <- c(0, -.6, .6, -0.6,-0, -0)
-  ggplot(df,
-         aes(x = x, y = y, xend = xend, yend = yend))+
-   #ggplot(n,aes(x = x, y = y, xend = xend, yend = yend))+
-    geom_edges(color = "grey50") +
-    theme_blank()

### 2019-10-16 analysis/paper/paper.Rmd
Introduction checked

@@
 
+Anaphylaxis is a clinical diagnosis with a variety of trigger factors and clinical presentations.
+Symptom profiles and specific cofactors for venom induced anaphylaxis had previously been analysed in uncontrolled manner in relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].
+
 Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and their infrequent and random occurence.
@@
 
-Based on the severity and symptom profile, we defined a sub-elevated baseline serum tryptase (BST) cut-off value of 8 - 11.5 ng/ml (Fig. \@ref(fig:figcofactors)B).
+Based on the severity and symptom profile, we defined sub-elevated baseline serum tryptase (BST) values as 8 - 11.5 ng/ml (Fig. \@ref(fig:figcofactors)B).
 
@@
 
-We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapeutic approaches that varied the most between VIA / non-VIA group and presented the results Gini importance [@Strobl2007].
-We also performed an association study of therapeutic interventions and symptoms.
+We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapeutic approaches that varied the most between VIA / non-VIA group and presented the results as Gini importance [@Strobl2007].
+We also performed an association analysis of therapeutic interventions and symptoms.
 The resulting phi values are presented in a heatmap with automatic clustering using Ward’s Agglomerative Hierarchical Clustering with Euclidean distances [@heatmaply].

### 2019-10-16 analysis/paper/paper.Rmd
fig 1 from vertical to horizontal

@@
 Due to a large number of documented reactions in the European Anaphylaxis Registry - we were able to match the VIA with non-VIA cases according to sex and age to reduce the comparison bias. 
-When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B).
+When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig.
+\@ref(fig:flow)B).
 Subsequently, we compared the management in both groups and matched the control group according to the severity of a reaction.
@@
 
-```{r flow,fig.width=5.2,fig.height=6, fig.cap=" A) Flow-diagram illustrating the rationale for case inclusion and exclusion from the final analysis.
+```{r flow,fig.width=7,fig.height=4.5, fig.cap=" A) Flow-diagram illustrating the rationale for case inclusion and exclusion from the final analysis.
 B, C, D: Age, sex, and severity distribution was matched in cases in both groups to allow for comparable results between VIA and non-VIA cases.
@@
           lower_panel,
-          ncol = 1,
-          nrow = 2,
-          heights = c(1,0.4),
+          ncol = 2,
+          nrow = 1,
+          widths = c(1,0.5),
           labels = c("A","")

### 2019-10-16 R/stat.R

@@ -1107,24 +1107,26 @@ lower_panel <- ggpubr::ggarrange(
     #facet_grid(.~b_sex)+
     theme_classic()+
     theme(legend.position = "none")+
-    scale_fill_manual(values = rev(c("#c4c4c4", "#454545")))+
+    scale_fill_manual(values = palBW)+
     labs(y ="proportion",x = "Sex",fill = "elicitor"),
 
   ANAscore_matched  %>%
+    mutate(grouping = car::recode(grouping, "'insects'='VIA';'other'='non-VIA'")) %>%
     #select(grouping,d_severity_rm) %>%
     group_by(grouping,d_severity_rm) %>%
     summarize(n = n()) %>%
     ggplot(aes(x = as.factor(d_severity_rm),y=n, fill = grouping)) +
     geom_bar(stat="identity",position="dodge")+
     theme_classic()+
-    labs(x = "Severity grade [R&M]",fill = "R&M",y="cases [n]")+
-    theme(legend.position = "none")+
-    scale_fill_manual(values = rev(c("#c4c4c4", "#454545"))),
+    labs(x = "Severity grade [R&M]",fill = "",y="cases [n]")+
+    theme(legend.position = "bottom")+
+    scale_fill_manual(values = palBW),
 
-common.legend = T,
-legend = "right",
-align = "h",
-ncol = 3,
+common.legend = F,
+#legend = "right",
+align = "v",
+ncol = 1,
+heights = c(1,1,1.5),
 labels = c("B","C","D"))
 
 

### 2019-10-16 analysis/paper/paper.Rmd
corrected figure 3

@@
 Younger patients (under 22) presented even more prominent differences in hypotension symptoms and significantly less frequently reported gastrointestinal symptoms (e.g.
-vomiting) when the reaction was triggered by insect venom.(Fig.
-\@ref(fig:figsymptoms)C-E)
+vomiting) when the reaction was triggered by insect venom (Fig.
+\@ref(fig:figsymptoms)C-E).
 
@@
 We found that `r (skin_symptoms$masto_tab$prop[c(2)]*100) %>% round(1)`% of patients with concomitant mastocytosis had anaphylaxis without skin symptoms (i.e.
-urticaria and flushing) which was significantly more compared to patients without diagnosed mastocytosis (`r (skin_symptoms$masto_tab$prop[c(1)]*100) %>% round(1)`%, `r pval(skin_symptoms$masto_chisq$p.value)`) and this observarion was most prominently seen in VIA (Fig.
-\@ref(fig:figskinsymptoms)).
+urticaria and flushing) which was significantly more compared to patients without diagnosed mastocytosis (`r (skin_symptoms$masto_tab$prop[c(1)]*100) %>% round(1)`%, `r pval(skin_symptoms$masto_chisq$p.value)`) and this observarion was most prominently seen in VIA.
 
@@
 non-mastocytic and at the same time lacking urticaria or flushing) VIA was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
-")`%, `r pval(skin_symptoms$masto_chisq$p.value)`% in non-VIA).
+")`%, `r pval(skin_symptoms$masto_chisq$p.value)`% in non-VIA,  Fig.
+\@ref(fig:figskinsymptoms)B).
 
@@
 
-```{r figskinsymptoms, fig.cap="Lack of skin symptoms during anaphylaxis is associated with more severe IVA.", fig.height=5}
+```{r figskinsymptoms, fig.cap="Lack of skin symptoms (i.e.
+urticara and flushing) during anaphylaxis is associated with more severe IVA.
+A: lack of skin symptoms and mastocytosis in VIA and non-VIA cases.
+B: Lack of skin symptoms according to severirty in both anaphylaxis groups.
+C: Relation of reaction severity according to the elicitor and the absence of skin symptoms in relation to categorized BST values.
+D: Continous values of BST according to severity in both non-VIA and VIA with subgrouping to skin symptoms.", fig.height=5}
 fig_skinsymptoms

### 2019-10-16 R/stat.R

@@ -3069,6 +3069,7 @@ skin_symptoms$masto_chisq <-
 
 skin_symptoms$masto_grouping_tab <-
   age_sex_matched %>%
+  mutate(grouping = car::recode(grouping,"'insects'='VIA'; 'other'='non-VIA'" )) %>%
   filter(q_111 != "unknown",
          q_410_masto_cur != "unknown") %>%
   #dplyr::select(d_111_urti_flush, q_410_masto_cur) %>%
@@ -3086,7 +3087,9 @@ skin_symptoms$masto_grouping_plot <- ggbarplot(skin_symptoms$masto_grouping_tab 
           palette = palBW,
           position = position_dodge2())+
   labs(x = "concomitant mastocytosis",
-       y = "proportion of patients\n without skin symptoms")
+       y = "proportion of patients\n without skin symptoms",
+       fill = "")+
+  annotate("text", x = 1.5, y = 0.5, label = "*", size = 8)
 
 skin_symptoms$severe_tab <-
 age_sex_matched %>%
@@ -3104,22 +3107,31 @@ age_sex_matched %>%
 
 skin_symptoms$severe_urti_flush_plot <-
   skin_symptoms$severe_tab %>%
+  filter(d_111_urti_flush == "no") %>%
   mutate(d_severity_rmr = ifelse(d_severity_rmr == "mild",
                                  "I+II",
                                  "III+IV")) %>%
   ggpubr::ggbarplot(
-    x = "d_111_urti_flush",
-    y = "n",
-    fill = "d_severity_rmr",
-    facet.by = "grouping",
-    position= position_fill(),
+    y = "prop",
+    x = "d_severity_rmr",
+    fill = "grouping",
+    position= position_dodge2(),
     palette = palBW
   ) +
-  labs(x = "urticaria or flush",
-       y = "proportion",
-       fill = "severity [R&M]")
+  labs(y = "proportion",
+       x = "severity [R&M]",
+       title = "non-mastocytic\npatients")+
+  theme(axis.title.y = element_blank(),
+        legend.position = "none")+
+  annotate("text",x = 1.5, y = 0.45, label = "*",
+           size = 8)
 
-
+fit1 <- glm(d_severity_rmr~grouping*d_111_urti_flush,
+    data = age_sex_matched %>%
+      mutate(d_severity_rmr = car::recode(d_severity_rmr,
[…]

### 2019-10-16 analysis/paper/paper.Rmd
gramarly corrrection for v3.0

--- 2019-10-16 analysis/paper/paper.Rmd
@@
 
-^1^ Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Hmuboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,
+^1^ Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,
 
@@
 # Abstract
-**Introduction:** Insect-venom elicited anaphylaxis is a common, life-threatening, hypersensitivity reaction.
-It is bound with specific: 1) symptom profile, 2) cofactors and 3) management.
-Identifying the diffeneces in subtypes of anaphylaxis is crucial for future management guidelines and personalized medicine approach.
+**Introduction:** Insect-venom elicited anaphylaxis is a common, life-threatening hypersensitivity reaction bound with specific: 1) symptom profile, 2) cofactors, and 3) management.
+Identifying the differences in subtypes of anaphylaxis is crucial for future management guidelines and personalized medicine approach.
 
@@
 
-**Results:** Venom induced anaphylaxis (VIA) more frequently involed over 3 organ systems and was associated with cardiovascular symptoms what was most prominently visible in patients under 22 years of age.
-Absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase and was associated with an increased risk of severe reaction specificaly in VIA.
-Intramuscular or intravenous adrenaline was administered significantly less often in IVA especially without prior history of anaphylaxis.
-Subelevated baseline serum tryptase (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis and this effect was most prominantly visible for IVA.
-
- <!-- MORE DISCUSSION!!!!!-->
-**Conclusions:** Due to the specific symptom profile, frequently affecting cardiovascular system, patients undergoing IVA are undertreated with adrenaline.
-Lack of urticaria or flushing during previous anaphylaxis (especialy in combination with subelevated baseline serum tryptase levels) is associated with severe reactions.
+**Results:** Venom induced anaphylaxis (VIA) more frequently involved over three organ systems and was associated with cardiovascular symptoms what was most prominently visible in patients under 22 years of age.
+Absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase and was associated with an increased risk of a severe reaction, specifically in VIA.
+Intramuscular or intravenous adrenaline was administered significantly less often in IVA, especially without prior history of anaphylaxis.
+Subelevated baseline serum tryptase (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis, and this effect was most prominently visible for IVA.
+**Conclusions:** Due to the specific symptom profile, frequently affecting the cardiovascular system, patients undergoing IVA are undertreated with adrenaline.
+Lack of urticaria or flushing during previous anaphylaxis (especially in combination with sub-elevated baseline serum tryptase levels) is associated with severe reactions.
 
@@
 Anaphylaxis is a clinical diagnosis with a variety of trigger factors and clinical presentations.
-Symptom profiles and specific cofactors for venom induced anaphylaxis had previously been analysed in uncontrolled manner in relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].
+Symptom profiles and specific cofactors for venom-induced anaphylaxis had previously been analyzed in an uncontrolled manner in relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].
 
-Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and their infrequent and random occurence.
+Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and their infrequent and random occurrence.
 Therefore registries, gathering clinical data from patients with a well documented (recent) history of anaphylaxis are crucial in investigating this entity.
 
-This study aimed to identify clinical patterns of venom induced anaphylaxis (VIA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA).
+This study aimed to identify clinical patterns of venom-induced anaphylaxis (VIA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA).
 
@@
 The diagnosis of anaphylaxis was based on the definition by NIAID/FAAN [@Sampson2006] and the severity according to Ring and Messmer Scale [@Ring1977].
-Grades III and IV (presenting with significant hypoxia, hypotension, confusion and loss of consciousness, or incontinence or cardiac arrest) were considered severe.
+Grades III and IV (presenting with significant hypoxia, hypotension, confusion, and loss of consciousness, or incontinence or cardiac arrest) were considered severe.
 
@@
 The final database consisted of `r countries %>% summarise(sum(n))` cases of venom induced anaphylaxis reported from allergy centers in `r countries %>% summarize(length(n))` countries and sex-and-age matched control group.
-We compared the frequency of various symptoms, cofactors --- known to increase the risk of severe anaphylaxis, [@Worm2018] and management in both groups.
+We compared the frequency of various symptoms, cofactors --- known to increase the risk of severe anaphylaxis, [@Worm2018], and management in both groups.
 
@@
 We used the R Statistical Package [@R] for statistical analysis.
-A simple comparison of categorical variables was performed using either Chi^2^ test or Fisher's exact test (where the number of observation in a bin was less than 10).
+A simple comparison of categorical variables was performed using either Chi^2^ test or Fisher's exact test (where the number of observations in a bin was less than 10).
 Continuous variables were analyzed using Mann-Whitney U test.
@@
 VIA displayed a specific symptom pattern.
-Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, hypotension, unconsciousness, reduced alertness) than patients with anaphylaxis due to other elicitors and less often presented with expiratory distress, rhinitis or diarrhoea (Fig.
+Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, hypotension, unconsciousness, reduced alertness) than patients with anaphylaxis due to other elicitors and less often presented with expiratory distress, rhinitis or diarrhea (Fig.
 \@ref(fig:figsymptoms)A).
@@
 
-Younger patients (under 22) presented even more prominent differences in hypotension symptoms and significantly less frequently reported gastrointestinal symptoms (e.g.
-vomiting) when the reaction was triggered by insect venom.(Fig.
-\@ref(fig:figsymptoms)C-E)
+Younger patients (under 22) presented even more prominent differences in hypotension symptoms and significantly less frequently reported gastrointestinal symptoms (e.g., vomiting) when the reaction was triggered by insect venom (Fig.
+\@ref(fig:figsymptoms)C-E).
 
@@
 urticaria and flushing) which was significantly more compared to patients without diagnosed mastocytosis (`r (skin_symptoms$masto_tab$prop[c(1)]*100) %>% round(1)`%, `r pval(skin_symptoms$masto_chisq$p.value)`) and this observarion was most prominently seen in VIA (Fig.
-\@ref(fig:figskinsymptoms)).
+\@ref(fig:figskinsymptoms)A).
 
@@
 non-mastocytic and at the same time lacking urticaria or flushing) VIA was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
-")`%, `r pval(skin_symptoms$masto_chisq$p.value)`% in non-VIA).
+")`%, `r pval(skin_symptoms$masto_chisq$p.value)`,  Fig.
+\@ref(fig:figskinsymptoms)B).
 
-After performing a factorial logistic regression modelling we confirmed the significant interaction effect between the presence of skin symptoms and the trigger (VIA) on the severity of anaphylaxis (`r pval(skin_symptoms$severity_fit_skin_symptoms %>% coef %>% .[4,4])`).
-To recapitulate, non-mastocytic patients presenting without urticaria or flushing tended to have more severe anaphylaxis when it was triggered by insects, but not other elicitors of anaphylaxis (Fig.
-\@ref(fig:figskinsymptoms)C, and Tab.
-S \@ref(tab:factoriallogit)).
+After performing a factorial logistic regression modeling, we confirmed the significant interaction effect between the presence of skin symptoms and the trigger (VIA) on the severity of anaphylaxis (`r pval(skin_symptoms$severity_fit_skin_symptoms %>% coef %>% .[4,4])`).
+To recapitualte, non-mastocytic patients presenting without urticaria or flushing tended to have more severe anaphylaxis when triggered by insects, but not other elicitors of anaphylaxis (Fig.
+\@ref(fig:figskinsymptoms)B, and Tab.
+S1 \@ref(tab:factoriallogit)).
 
@@
 
-Following the above-mentioned observation, we investigated the association of skin symptoms with the tryptase levels in non-masctoscytic patients.
+Following the observation, as mentioned earlier, we investigated the association of skin symptoms with the tryptase levels in non-mastocytic patients.
 For this model, we excluded the cases with known mastocytosis and with BST above 11.5 ng/ml, potentially indicating non-diagnosed mastocytosis.
@@
 We did not observe this effect in the non-VIA group (Fig.
-\@ref(fig:figskinsymptoms)D).
+\@ref(fig:figskinsymptoms)C-D).
 
@@
 The factor most prominently associated with an increased risk of severe anaphylaxis was mastocytosis (Fig.
-\@ref(fig:figcofactors)), and it increased the risk of: 1) cardiac arrest and 2) loss of consciousness in patients undergoing VIA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
-S\@ref(fig:supplement)A and Fig.
-\@ref(fig:figcofactors)C).
+\@ref(fig:figcofactors)), and it increased the risk of 1) cardiac arrest and 2) loss of consciousness in patients undergoing VIA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
+\@ref(fig:figcofactors)C and Fig.
+S\@ref(fig:supplement)A).
 
-In line with the above, BST levels also correlated with the severity of anaphylaxis (on the Ring and Messmer scale) and, most importantly, sub-elevated BST was more prominently increasing the risk of severe anaphylaxis in VIA than in non-VIA (Fig.
+In line with the above, BST levels also correlated with the severity of anaphylaxis (on the Ring and Messmer scale) and, most importantly, sub-elevated BST was more prominently associated with increasing the risk of severe anaphylaxis in VIA than in non-VIA (Fig.
 \@ref(fig:figsymptoms)D and Fig.
@@
 
-## Other cofactors of severe reactions (potentially could be shifted to th supplement)
+## Other cofactors of severe reactions (potentially could be moved to the supplement)
 
@@
 There were no differences in severity of reactions elicited by yellow-jackets and other
-insect species (p = 0.4128). .
+insect species (p = 0.4128).
 
@@
 
-## Therapy: VIA recived adrenaline less often than non-VIA
+## Therapy: VIA received adrenaline less often than non-VIA
 
@@
 Patients who underwent VIA significantly less often received adrenaline treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
-After adjusting both groups for similar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)).
+After adjusting both groups for similar age, sex, and severity distribution - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)).
 <!-- Antihistaminic drugs, on the other hand, were given more often in cases of VIA when compared to non-VIA. -->
@@
 
-Having a history of anaphylaxis also infuenced the therapy of a current episode.
+Having a history of anaphylaxis also influenced the therapy of a current episode.
 Adrenaline as a first-line treatment was given less often in VIA cases when compared to other cases **if patients did not report the previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients who reported previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`, Suppl.
@@
 We discovered clusters of symptoms and therapy modes in VIA patients (association measured using phi coefficient).
-Cardiovascular symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria, were treated more similarly than respiratory or gastrointestinal symptoms (Fig.
+Cardiovascular symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria were treated more similarly than respiratory or gastrointestinal symptoms (Fig.
 \@ref(fig:adrenalineuse)B).
@@
 adrenaline in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists.
-Corticosteroids, i.v. volume replacement and i.v. β-2 agonists formed another therapy mode. 
-The most prominent difference in the therapy of VIA vs. non-VIA were regarging the  inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
+Corticosteroids, i.v. volume replacement, and i.v. β-2 agonists formed another therapy mode. 
+The most noticeable difference in the therapy of VIA vs. non-VIA was regarding the inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
 
@@
 Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
-Their proportion of VIA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was as below 1% of cases during winter.
-Nevertheless, `r rare_spring_autumn$n %>% sum` cases of VIA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were triggered in March, April, and November.
+Their proportion of VIA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was below 1% of cases during winter.
+Nevertheless, `r rare_spring_autumn$n %>% sum` cases of VIA (bee -- *Apis mellifera* in Spring; yellow jacket -- *Vespula spp.* in Autumn) were triggered in March, April, and November.
 Yellow-jacket was the most prominent VIA-causing insect followed by bees.
@@
 <!-- Cardiologic symptoms and hypotensive collapse might be associated with venom anaphylaxis due to the vaso-vagal reflex.
-As seen in patients undergoing blood sampling needles (and probably also insect stings) may elicit a hypotensive response due to extreme emotional distress with bradycardia, limiting the blood flow to the brain and causing syncope.
+As seen in patients undergoing blood sampling, needles (and probably also insect stings) may elicit a hypotensive response due to extreme emotional distress with bradycardia, limiting the blood flow to the brain and causing syncope.
 This mechanism may contribute to the ongoing histamine-induced vasodilatation and may worsen the symptoms of an allergic reaction.
@@
 
-In this study, we identified distinctive clinical symptom-profile, and treatment patterns of venom induced anaphlaxis.
+In this study, we identified distinctive clinical symptom-profile and treatment patterns of venom-induced anaphylaxis.
 The data unravel phenotypes of VIA, which may contribute to the development of tools incorporating clinical <!--- and biomarker--->data for predicting the severity of future episodes of anaphylaxis.
@@
 This association was not significant in anaphylaxis elicited by other triggers.
-Notably, cardiac arrest occurred more frequently in patients with the elevated BST (> 8 ng/ml), especially in VIA.
+Notably, cardiac arrest occurred more frequently in patients with elevated BST (> 8 ng/ml), especially in VIA.
 Nevertheless, the pathomechanism promoting cardiovascular symptoms in VIA requires further investigation.
@@
 
-Regarding the factors increasing the risk of VIA - older patients experienced VIA more frequently.
-Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].
-One of the limitations of our registry is the fact that we can only compare cases of anaphylaxis and due to the lack of healthy cohort --- cannot draw conclusions on the frequency of a particular type of anaphylaxis in the population.
+Regarding the factors increasing the risk of VIA - older patients experienced VIA more frequently. Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].
 
@@
 Cases of VIA had been treated with adrenaline less often than the age- sex- and severity-matched cases of non-VIA.
-Moreover, administration of adrenaline did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patient experienced their first episode of VIA.
+Moreover, the administration of adrenaline did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patient experienced their first episode of VIA.
 The difference between groups was prominent for milder cases of anaphylaxis.
@@
 
-Although there are no absolute contraindications for using adrenaline in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using adrenaline is hypersensitivity reaction with high blood pressure with tachycardia which might be present at the initial phase of VIA, due to a psychologic reflex.
+Although there are no absolute contraindications for using adrenaline in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using adrenaline is hypersensitivity reaction with high blood pressure with tachycardia, which might be present at the initial phase of VIA, due to a psychologic reflex.
 In theory, these less severe cases of VIA might exhibit some form of stress-related blood pressure increase, but we lack data to confirm or discard this theory.
@@
 
-Based on our findings, insects are the most probable anaphylaxis elicitor in Europe during Summer-season, with VIA cases extending from early Spring to the end of Autumn.
+Based on our findings, insects are the most probable elicitor of anaphylaxis in Europe during Summer-season, with VIA cases extending from early Spring to the end of Autumn.
 Detailed information on the seasonality of insect-elicited hypersensitivity reactions is scarce [@old].
 The activity of *Vespula germanica* depends on the climate, and in invaded regions (e.i.
-Australia) it can even extend throughout the year [@Spradbery_1992].
+Australia), it can even extend throughout the year [@Spradbery_1992].
 The changing climate in Europe may influence the activity of Hymenoptera in this region in the upcoming years.
@@
 <!-- Limitations and strengths -->
+## Limitations
 Due to the design of the European Anaphylaxis Registry, our analysis was restricted only to cases of anaphylaxis.
@@
 
-When evaluating the risk of future severe episodes - patients with BST over 8 ng/ml should undergo extensive diagnostic tests to exclude ISM or MCAS and should be provided with two adrenaline autoinjectors for acute self management.
+When evaluating the risk of future severe episodes - patients with BST over 8 ng/ml should undergo extensive diagnostic tests to exclude ISM or MCAS and should be provided with two adrenaline autoinjectors for acute self-management.
 
@@
 biol.
-Kristin Franke for insight about the manuscript.
+Kristin Franke for revising the manuscript.
 We thank the study personnel for patients counseling and data entry, and we thank the members of The European Anaphylaxis Registry in detail:
@@
 ```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (VIA) compared to other elicitors.
-A: Proportional presentation of specific reaction symptoms in VIA and non-VIA according to cardiovascular (cardio.), gastroenterologic (gastro.) and respirator (resp.) organ systems.
+A: Proportional presentation of specific reaction symptoms in VIA and non-VIA according to cardiovascular (cardio.), gastroenterologic (gastro.), and respiratory (resp.) organ systems.
 B: High-level overview of involved organ systems and selected cofactors in the form of a radar plot.
 C: difference in symptoms of VIA among patients under 22 and over 22 years of age.
-* denotes significant differences between groups.
-(other figure:Interaction effect of VIA and lack of skin symptoms (urticaria or flushing) on the severity of anaphylaxis.
-D: Baseline Serum Tryptase (BST) increases with the severity of anaphylaxis a lot more prominently in VIA.", fig.width=8.5, fig.height = 6.2}
+* denotes significant differences between groups.", fig.width=8.5, fig.height = 6.2}
 fig_symptoms_3
@@
 
-```{r figskinsymptoms, fig.cap="Lack of skin symptoms during anaphylaxis is associated with more severe IVA.", fig.height=5}
+```{r figskinsymptoms, fig.cap="Lack of skin symptoms (i.e., urticaria and flushing) during anaphylaxis is associated with more severe IVA.
+A: lack of skin symptoms and mastocytosis in VIA and non-VIA cases.
+B: Lack of skin symptoms, according to the severity in both anaphylaxis groups.
+C: Relation of reaction severity according to the elicitor and the absence of skin symptoms concerning categorized BST values.
+D: Continous values of BST according to the severity in both non-VIA and VIA with subgrouping to skin symptoms.", fig.width = 8, fig.height=5.5}
 fig_skinsymptoms
@@
 B:   C: Heatmap visualizing the association of symptoms and corresponding treatment -  presented as a scaled correlation coefficient (phi).
-* - p value < 0.05 after false discovery rate correction.
+* - p-value < 0.05 after false discovery rate correction.
 ", fig.height=9, fig.width = 7}

### 2019-10-16 R/functions.R

@@ -511,8 +511,8 @@ makeDF <- function(data,variables,grouping,outcome){
     grB <- which(grouping==levels(grouping)[2])
     rbind(
       #c(makeOdds(data =data, var1 = x, var2 = outcome),S="all"),
-      c(makeOdds(data =data[grA,], var1 = x, var2 = outcome),S="IVA"),
-      c(makeOdds(data =data[grB,], var1 = x, var2 = outcome),S="non-IVA"),
+      c(makeOdds(data =data[grA,], var1 = x, var2 = outcome),S="VIA"),
+      c(makeOdds(data =data[grB,], var1 = x, var2 = outcome),S="non-VIA"),
       NA
     )
   }) %>% do.call(what=rbind) %>% data.frame()

### 2019-10-16 R/make_all_forest_plots.R

@@ -1,7 +1,6 @@
 ### make all forest plots
 make_all_forest_plots <- function(){
 
-
 test1 <- makeDF(age_sex_matched,
                 c( "q_410_cur",
                    "q_410_asthma_cur",
@@ -76,10 +75,10 @@ forestplot(labeltext = makeTableText(test1),
 dev.off()
 
 png("analysis/figures/figForestfinalrmr.png",
-    height = 640,
-    width = 860,
+    height = 640*2,
+    width = 860*2,
     #res = 300,
-    pointsize = 30
+    pointsize = 30*2
     #units="px"
 )
 forestplot(labeltext = makeTableText(test2)[,1:2],
@@ -119,7 +118,7 @@ makeForestPlot(rdbp[rdbp$d_age<18,],
                testInsectsbinomial %>%
                  filter(section=="cofactors") %>%
                  arrange(pval) %>%
-                 select(variableName) %>% pull() %>% {c(.[c(1,2,6,7,15,18
+                 dplyr::select(variableName) %>% pull() %>% {c(.[c(1,2,6,7,15,18
                  )])},
                "grouping",
                "severity_brown",
@@ -131,7 +130,7 @@ makeForestPlot(rdbp[rdbp$d_age>18,],
                testInsectsbinomial %>%
                  filter(section=="cofactors") %>%
                  arrange(pval) %>%
-                 select(variableName) %>% pull() %>% {c(.[c(1,6,16,41,18,27
+                 dplyr::select(variableName) %>% pull() %>% {c(.[c(1,6,16,41,18,27
                  )])},
                "grouping",
                "severity_brown",

### 2019-10-29 analysis/paper/paper.Rmd
corrections according to Prof Worm

@@
   - Kathrin Scherer Hofmeier^4^
+  - Bettina Wedi^5^
+  - A Bauer^6^ 
+  - N Wagner^7^
+  - Regina Treudler^8^
+  - Thomas Hawranek^9^
+  - Nicos Papadopoulus^10^
+  - Montserat Fernandes-Rivas^11^ 
+  - Jean Marie Reanudin^12^
+  - Victoria Cardona^13^
+  - M. Bilo^14^
+  - Izabela Poziomkowska-Gęsicka^15^
+  - George Christoff^16^
+  - Tilo Jakob^17^
+  - John Hourihane^18^
+  - L Ensina^19^
   - Margitta Worm^1^
@@
 
-^1^ Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,
+^1^ Division of Allergy and Immunology, Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,
 
@@
 # Abstract
-**Introduction:** Insect-venom elicited anaphylaxis is a common, life-threatening hypersensitivity reaction bound with specific: 1) symptom profile, 2) cofactors, and 3) management.
-Identifying the differences in subtypes of anaphylaxis is crucial for future management guidelines and personalized medicine approach.
+**Introduction:** Insect-venom elicited anaphylaxis is a common, potentially life-threatening hypersensitivity reaction associated with specific: 1) symptom profile, 2) cofactors, and 3) management.
+Identifying the differences in subtypes of anaphylaxis is crucial for future management guidelines and the development of a personalized medicine approach.
 
@@
 
-**Results:** Venom induced anaphylaxis (VIA) more frequently involved over three organ systems and was associated with cardiovascular symptoms what was most prominently visible in patients under 22 years of age.
-Absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase and was associated with an increased risk of a severe reaction, specifically in VIA.
-Intramuscular or intravenous adrenaline was administered significantly less often in IVA, especially without prior history of anaphylaxis.
-Subelevated baseline serum tryptase (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis, and this effect was most prominently visible for IVA.
-**Conclusions:** Due to the specific symptom profile, frequently affecting the cardiovascular system, patients undergoing IVA are undertreated with adrenaline.
-Lack of urticaria or flushing during previous anaphylaxis (especially in combination with sub-elevated baseline serum tryptase levels) is associated with severe reactions.
+**Results:** Venom induced anaphylaxis (VIA) more frequently involved more than three organ systems and was associated with cardiovascular symptoms.
+The absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase and was associated with an increased risk of a severe reaction.
+Intramuscular or intravenous adrenaline was administered significantly less often in VIA, in particular in patients without prior history of anaphylaxis.
+Baseline serum tryptase (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis.
+
+**Conclusions:** Due to a specific symptom profile, frequently affecting the cardiovascular system, patients undergoing VIA are undertreated with adrenaline.
+The lack of skin symptoms (i.e.
+urticaria or flushing) during anaphylaxis (especially in combination with baseline serum tryptase levels within upper normal limits) is associated with severe reactions.
 
@@
 Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong systemic immunotherapy [@Sturm2017].
-The need for a more precise description of the diagnosis, identification of biomarkers, and phenotypes of anaphylaxis is discussed [@JimenezRodriguez2018].
+There is a need for a more precise description of the diagnosis, identification of biomarkers, and better definition of phenotypes of anaphylaxis [@JimenezRodriguez2018].
 Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] for the diagnosis of anaphylaxis --- a better understanding of its clinical subtypes is required.
@@
 Anaphylaxis is a clinical diagnosis with a variety of trigger factors and clinical presentations.
-Symptom profiles and specific cofactors for venom-induced anaphylaxis had previously been analyzed in an uncontrolled manner in relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].
+Symptom profiles and specific cofactors for venom-induced anaphylaxis had previously been analyzed in an uncontrolled manner albeit relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].
 
-Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and their infrequent and random occurrence.
+Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and its infrequent and random occurrence.
 Therefore registries, gathering clinical data from patients with a well documented (recent) history of anaphylaxis are crucial in investigating this entity.
 
-This study aimed to identify clinical patterns of venom-induced anaphylaxis (VIA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA).
+This study aimed to identify clinical patterns of venom-induced anaphylaxis (VIA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA) based on the the data from the European Anaphylaxis Registry.
 
@@
 
-Based on the severity and symptom profile, we defined sub-elevated baseline serum tryptase (BST) values as 8 - 11.5 ng/ml (Fig. \@ref(fig:figcofactors)B).
+Based on the severity and symptom profile, and previous reports [@Ruff2009] we defined sub-elevated baseline serum tryptase (BST) values as 8 - 11.5 ng/ml (Fig. \@ref(fig:figcofactors)B).
 
@@
 We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapeutic approaches that varied the most between VIA / non-VIA group and presented the results as Gini importance [@Strobl2007].
-We also performed an association analysis of therapeutic interventions and symptoms.
-The resulting phi values are presented in a heatmap with automatic clustering using Ward’s Agglomerative Hierarchical Clustering with Euclidean distances [@heatmaply].
+Moreover, an association analysis of therapeutic interventions and symptoms was performed.
+The resulting phi values were scaled and presented in a heatmap with automatic clustering using Ward’s Agglomerative Hierarchical Clustering with Euclidean distances [@heatmaply].
 
@@
 
-## VIA is associated with cardiovascular symptoms
+## VIA is more frequently associated with cardiovascular symptoms
 
@@
 
-Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrologic, skin, and respiratory systems, VIA more frequently involved over three organ systems (`r compare(c(tempdf[3:4,">3 organ\nsystems"]))`,<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
+Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrologic, skin, and respiratory systems, VIA more frequently involved more than three organ systems (`r compare(c(tempdf[3:4,">3 organ\nsystems"]))`,<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
 Although, patients undergoing VIA  less often had concomitant atopic diseases  -->
@@
 We found that `r (skin_symptoms$masto_tab$prop[c(2)]*100) %>% round(1)`% of patients with concomitant mastocytosis had anaphylaxis without skin symptoms (i.e.
-urticaria and flushing) which was significantly more compared to patients without diagnosed mastocytosis (`r (skin_symptoms$masto_tab$prop[c(1)]*100) %>% round(1)`%, `r pval(skin_symptoms$masto_chisq$p.value)`) and this observarion was most prominently seen in VIA (Fig.
+urticaria and flushing) which was significantly more frequent compared to patients without diagnosed mastocytosis (`r (skin_symptoms$masto_tab$prop[c(1)]*100) %>% round(1)`%, `r pval(skin_symptoms$masto_chisq$p.value)`).
+This finding was most prominently seen in VIA (Fig.
 \@ref(fig:figskinsymptoms)A).
 
-Similarily, in non-mastocytic patients undergoing VIA, urticaria or flushing as symptoms of anaphylaxis were present less often than if anaphylaxis was triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
+Similarily, in non-mastocytosis patients undergoing VIA skin smptoms(i.e.
+urticaria or flushing) were less often present than if anaphylaxis was triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
 Moreover, in this specific subgroup of patients (i.e.
-non-mastocytic and at the same time lacking urticaria or flushing) VIA was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
+non-mastosis patients lacking skin symptoms) VIA was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
 ")`%, `r pval(skin_symptoms$masto_chisq$p.value)`,  Fig.
@@
 
-After performing a factorial logistic regression modeling, we confirmed the significant interaction effect between the presence of skin symptoms and the trigger (VIA) on the severity of anaphylaxis (`r pval(skin_symptoms$severity_fit_skin_symptoms %>% coef %>% .[4,4])`).
-To recapitualte, non-mastocytic patients presenting without urticaria or flushing tended to have more severe anaphylaxis when triggered by insects, but not other elicitors of anaphylaxis (Fig.
+By applying factorial logistic regression modeling, we confirmed a significant interaction effect between the presence of skin symptoms and insect venom on the severity of anaphylaxis (`r pval(skin_symptoms$severity_fit_skin_symptoms %>% coef %>% .[4,4])`).
+In other words, non-mastosis patients presenting without urticaria or flushing tended to have more severe anaphylaxis when triggered by insects, but not other elicitors of anaphylaxis (Fig.
 \@ref(fig:figskinsymptoms)B, and Tab.
-S1 \@ref(tab:factoriallogit)).
+S1).
 
-## Absence of skin symptoms correlates with BST levels and increases the risk of severe anaphylaxis specifically in IVA
+## Absence of skin symptoms correlates with BST levels and increases the risk of severe anaphylaxis specifically in VIA
 
-Following the observation, as mentioned earlier, we investigated the association of skin symptoms with the tryptase levels in non-mastocytic patients.
+Next, we investigated the association of skin symptoms with the tryptase levels in non-mastocytosis patients.
 For this model, we excluded the cases with known mastocytosis and with BST above 11.5 ng/ml, potentially indicating non-diagnosed mastocytosis.
-Similarily, 1) tryptase levels correlated with the severity of anaphylaxis, 2) were higher in VIA patients, and 3) the effect was more prominently visible in VIA (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`).
-We did not observe this effect in the non-VIA group (Fig.
+Similarily, 1) tryptase levels were higher in VIA patients, 2) correlated with the severity of anaphylaxis, and 3) this effect was significant in VIA (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`) but not in the non-VIA group (Fig.
 \@ref(fig:figskinsymptoms)C-D).
 
-## Sub-elevated baseline serum tryptase and concomitant cardiovascular conditions increase the risk of severe VIA.
+## Baseline serum tryptase over 8 ng/ml and concomitant cardiovascular conditions increase the risk of severe VIA.
 
-We evaluated potential cofactors increasing the risk of severe IVA.
-The factor most prominently associated with an increased risk of severe anaphylaxis was mastocytosis (Fig.
-\@ref(fig:figcofactors)), and it increased the risk of 1) cardiac arrest and 2) loss of consciousness in patients undergoing VIA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
+Cofactors acn promote the onset and increase severity of anaphylaxis. 
+Therefore we evaluated a variety of known cofactors regarding their impact to increase the risk of severe VIA.
+As expected, the cofactor most prominently associated with an increased risk of severe anaphylaxis was mastocytosis (Fig.
+\@ref(fig:figcofactors)).
+Concomitant mastocytosis increased the risk for 1) cardiac arrest and 2) loss of consciousness in patients undergoing VIA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
 \@ref(fig:figcofactors)C and Fig.
@@
 
-In line with the above, BST levels also correlated with the severity of anaphylaxis (on the Ring and Messmer scale) and, most importantly, sub-elevated BST was more prominently associated with increasing the risk of severe anaphylaxis in VIA than in non-VIA (Fig.
+In line with the findings above, BST levels also correlated with the severity of anaphylaxis (on the Ring and Messmer scale) and, most importantly, sub-elevated BST was more prominently associated with increasing the risk of severe anaphylaxis in VIA than in non-VIA (Fig.
 \@ref(fig:figsymptoms)D and Fig.
@@
 
-Concomitant cardiovascular diseases were more prevalent in VIA than in non-VIA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare()`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were irrelevant in non-VIA cases (Fig.
+Concomitant cardiovascular diseases were more prevalent in VIA than in non-VIA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare()`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were not relevant in non-VIA cases (Fig.
 \@ref(fig:figcofactors)).
@@
 
-## Other cofactors of severe reactions (potentially could be moved to the supplement)
+## Other cofactors of severe reactions
 
@@
 other elicitors (Fig.
-S\@ref(fig:sev)D).
+S\@ref(fig:sev)).
 There were no differences in severity of reactions elicited by yellow-jackets and other
@@
 
-The effect of using ACE-I (as well as beta-blockers) on the risk of severe anaphylaxis could not be independently measured from the coexisting cardiovascular pathologies.
-ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) and there were no differences between VIA and non-VIA  (Fig.
+The effect of using ACE-I (as well as beta-blockers) on the risk of severe anaphylaxis correlated with the coexisting cardiovascular pathologies.
+ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) regardless of the elicitor  (Fig.
 \@ref(fig:figcofactors)C).
-Beta-blocker use was generally associated with higher severity of anaphylaxis and with the onset of cardiovascular symptoms (cardiac arrest, chest pain), but showed no difference between groups `r summary(beta_fit)[["coefficients"]][3,4] %>% pval()`).
+Beta-blocker use was associated with a higher severity of anaphylaxis and with the onset of cardiovascular symptoms (cardiac arrest, chest pain), but was comparable between `r summary(beta_fit)[["coefficients"]][3,4] %>% pval()`).
 Surprisingly, arrhythmia was more frequently reported in patients with VIA and concomitant beta-blockers (Fig.
@@
 
-## Nearly one-third of VIA patients experience repeated reactions
+## One-third of VIA patients experience repeated reactions
 
@@
 
-In general `r treact[1,2] %>% round(1)`% of patients with insect allergy had experienced venom anaphylaxis in the past.
-If the reaction was elicited by other elcitors previous reactions were more frequently seen (`r treact[2,2]%>% round(1)`%, `r p %>% pval()`).
-We observed `r sum(repeated_reaction_in_Reg)` patients with two documented reactions in our registry.
-Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had insect elicited anaphylaxis and in `r greater_same[1,2]` (`r (greater_same[1,2]/sum(greater_same[,2])*100 )%>%round(1) `%) the following reaction was more severe than before.
-In `r greater_same[3,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was similar in severity.
+`r treact[1,2] %>% round(1)`% of patients with insect allergy had experienced venom anaphylaxis in the past.
+If the reaction was elicited by other elcitors (i.e.
+non-VIA) --- previous reactions were more frequently seen (`r treact[2,2]%>% round(1)`%, `r p %>% pval()`).
+We observed `r sum(repeated_reaction_in_Reg)` patients with at least two fully-documented reactions.
+Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had insect elicited anaphylaxis and in `r greater_same[1,2]` of them (`r (greater_same[1,2]/sum(greater_same[,2])*100 )%>%round(1) `%) the following reaction was more severe than before.
+In `r greater_same[3,2]` out of the `r greater_same[1,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was similar in severity.
 
-## Therapy: VIA received adrenaline less often than non-VIA
+## VIA patients receive adrenaline less often than non-VIA
 
@@
 Patients who underwent VIA significantly less often received adrenaline treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
-After adjusting both groups for similar age, sex, and severity distribution - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)).
+After adjusting both groups for similar age, sex, and severity distribution - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)B).
 <!-- Antihistaminic drugs, on the other hand, were given more often in cases of VIA when compared to non-VIA. -->
@@
 
-Having a history of anaphylaxis also influenced the therapy of a current episode.
-Adrenaline as a first-line treatment was given less often in VIA cases when compared to other cases **if patients did not report the previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients who reported previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`, Suppl.
-Fig.
-\@ref(fig:rfplot)).
-Similarly, there were no differences in adrenaline use between VIA and non-VIA when severe reactions were taken into consideration (`r adren_severity_sev`).
+A positive history of anaphylaxis influenced the therapy of a current episode as well.
+Adrenaline as a first-line treatment was given less often in VIA cases when compared to other cases **if patients did not report a previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients reporting previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`, Fig.
+\@ref(fig:adrenalineuse)B).
+Similarly, there were no differences in the adrenaline use between VIA and non-VIA when only severe reactions were taken into consideration (`r adren_severity_sev`).
 However, when we restricted the analysis to mild anaphylaxis cases --- non-VIA patients received adrenaline more frequently than VIA (`r adren_severity_mild`).
@@
 Patients with VIA received corticosteroids and antihistamines significantly more frequently than patients with anaphylaxis to other elicitors.
-On the other hand, adrenaline, beta-2 mimetics, and oxygen were given more often in non-VIA.
+On the other hand, adrenaline, beta-2 mimetics, and oxygen were given more often to patients suffering from non-VIA  (Fig.
+\@ref(fig:adrenalineuse)A, Fig.
+S\@ref(fig:rfplot)).
 
-We discovered clusters of symptoms and therapy modes in VIA patients (association measured using phi coefficient).
-Cardiovascular symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria were treated more similarly than respiratory or gastrointestinal symptoms (Fig.
-\@ref(fig:adrenalineuse)B).
-The treatment of these symptoms consisted of adrenaline autoinjector (AAI) use, i.v.
+Next, we asked wether specific symptom clusters and treatment profiles could be identified within our cohort (association measured using phi coefficient).
+We found that patients dysplaing cardiovascular symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria were treated differently than patients with respiratory or gastrointestinal symptoms (Fig.
+\@ref(fig:adrenalineuse)C).
+The treatment of the former symptoms consisted of adrenaline autoinjector (AAI) use, i.v.
 adrenaline in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists.
-Corticosteroids, i.v. volume replacement, and i.v. β-2 agonists formed another therapy mode. 
-The most noticeable difference in the therapy of VIA vs. non-VIA was regarding the inhaled beta2-agonists and antihistamines (Fig. \@ref(fig:adrenalineuse)B).
+Corticosteroids, i.v.
+volume replacement, and i.v.
+β-2 agonists formed another therapy mode.
 
@@
 
-Outdoor physical exercise (e.g., jogging in the park) was more often associated with VIA than other triggers of anaphylaxis (`r ttexercise1`).
+<!-- Outdoor physical exercise (e.g., jogging in the park) was more often associated with VIA than other triggers of anaphylaxis (`r ttexercise1`).
 However, it was not associated with the severity of a reaction in these patients.
 (`r ttexercise2`).
-
-<!-- ## Time of the symptom onset  -->
-
-<!-- # Model for severity prediction -->
-<!-- As a last attempt for the analysis - perform modeling with the use of random forest or regression using the following inputs: Absence of urticaria, BST, short time for the onset of symptoms, older age, male sex?
-Cardiologic conditions.
 -->
-
-<!-- ### Mastocytosis -->
-
-<!-- ### Older Age -->
-
-<!-- ### ACE - I -->
-
-<!-- ### previous reactions (severe) -->
-
-<!-- ### Male sex -->
-
-<!-- ### Beta blockers and CARDIOSELECTIVE BLOCKERS -->
-
-<!-- ### Evaluate the pediatric population separately -->
-
-<!-- ### Number of SAR during VIT compare to other hypos -->
-<!-- #### Compare bee sit to wasp sit for AEs -->
-<!-- #### correlate the tryptase value with SIT types and severity -->
 
@@
 
-In this study, we identified distinctive clinical symptom-profile and treatment patterns of venom-induced anaphylaxis.
-The data unravel phenotypes of VIA, which may contribute to the development of tools incorporating clinical <!--- and biomarker--->data for predicting the severity of future episodes of anaphylaxis.
+In this study, we identified distinct symptom-profile and treatment patterns of venom-induced anaphylaxis.
+The data unravel phenotypes of VIA, which may support to the development of tools incorporating clinical <!--- and biomarker--->data for predicting the severity of future episodes of anaphylaxis.
 
@@
 VIA was more often associated with cardiovascular symptoms than non-VIA.
-Previous studies suggest [@Worm2018; @nittner2015insect; @Bonadonna2010] an essential link between the cardiovascular system and insect sting hypersensitivity. 
+Previous studies suggest an essential link between the cardiovascular system and insect sting hypersensitivity [@Worm2018; @nittner2015insect; @Bonadonna2010]. 
 VIA has been associated with Kounis syndrome (coronary arterial spasm induced by the release of mast cell mediators [@CDT1609; @sinkiewicz2008allergic]) and cardiac arrhythmias usually occurring in patients with preexisting heart disease [@Sharma2016].
 
-The rate of concomitant cardiovascular diseases was higher in VIA than non-VIA; we observed them be an essential cofactor increasing the risk of a severe reaction **if insects elicited the anaphylaxis**.
+The rate of concomitant cardiovascular diseases was higher in VIA than non-VIA.
+They are an essential cofactor increasing the risk of a severe reaction **if insects elicited the anaphylaxis**.
 This association was not significant in anaphylaxis elicited by other triggers.
@@
 
-Of note, cardiovascular symptoms like hypotension, collapse, or cardiac arrest grade higher on the Ring and Messmer scale than skin or gastrointestinal symptoms.
-Since VIA predisposes to cardiovascular symptoms, it is likely to be associated with more severe anaphylaxis.
+As the onset of cardiovascular symptoms like hypotension, collapse, or cardiac arrest lead to a higher grade on the Ring and Messmer scale than skin or gastrointestinal symptoms, VIA (being associated with cardiovascular symptoms) is likely to be associated with more severe anaphylaxis.
 
-Importantly, the absence of skin symptoms was associated with more severe VIA, which was also true after we excluded patients with a known diagnosis of mastocytosis.
-Subsequently, we correlated the severity of anaphylaxis with BST levels and identified an interaction between the absence of skin symptoms and VIA using generalized linear regression.
+Importantly, the absence of skin symptoms was associated with more severe VIA, which was still present after excluding patients with a known diagnosis of mastocytosis.
+Subsequently, the corellation of BST levels with severity of anaphylaxis lead us to identify an interaction between the absence of skin symptoms and VIA using generalized linear regression.
 
@@
 Zanotti et al.
-identified mast cell disorders in 17 out of 22 patients with VIA lacking skin symptoms [@Zanotti2015] and concluded that patients with BST of 7.95 ng/ml and VIA should undergo extensive diagnostic procedures.
+identified mast cell disorders in 17 out of 22 patients with VIA lacking skin symptoms  and concluded that patients with BST of 7.95 ng/ml and VIA should undergo extensive diagnostic procedures [@Zanotti2015].
+We recently identified that elederly patient undergoing anaphylaxis without concomitant skin symptoms tended to have more severe reactions [@Aurich2019]
 
-Based on our and previous findings [@Ruff2009;@Zanotti2015;@JaraAcevedo2015] we propose to perform a peripheral blood KIT D816V mutation test in cases of BST of above 8 ng/ml and with a history of anaphylaxis presenting without urticaria or flushing.
+Based on these and previous findings [@Ruff2009;@Zanotti2015;@JaraAcevedo2015] we propose to perform a peripheral blood KIT D816V mutation test in cases of BST of above 8 ng/ml and with a history of anaphylaxis presenting without urticaria or flushing.
 
@@
 
-Regarding the factors increasing the risk of VIA - older patients experienced VIA more frequently. Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].
+Regarding the factors increasing the risk of VIA - adult patients experienced VIA more frequently. Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].
 
@@
 <!-- Risk factors - Concomitant disease -->
-Similarly, we saw less VIA in patients with concomitant atopic diseases, as these patients more often present with food anaphylaxis [@Tham2019].
+Similarly, we observed less VIA in patients with concomitant atopic diseases, as these patients more often present with food anaphylaxis [@Tham2019].
 
-The influence of cardiovascular medication could not be isolated from the effect of concomitant cardiovascular conditions; therefore, we could not state if ACE-I and beta-blockers increased the severity of anaphylaxis.
-We did observe that there were no significant differences between VIA and non-VIA cases regarding the symptoms and severity of an episode with concomitant use of ACE-I or beta-blockers.
+The role of cardiovascular medication can not be isolated from the effect of concomitant cardiovascular conditions; therefore, we can not state wether ACE-I and beta-blockers increase the severity of anaphylaxis.
+However, we did observe that there were no significant differences between VIA and non-VIA cases regarding the symptoms and severity of an episode with concomitant use of ACE-I or beta-blockers.
 
@@
 
-Although there are no absolute contraindications for using adrenaline in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using adrenaline is hypersensitivity reaction with high blood pressure with tachycardia, which might be present at the initial phase of VIA, due to a psychologic reflex.
-In theory, these less severe cases of VIA might exhibit some form of stress-related blood pressure increase, but we lack data to confirm or discard this theory.
+Although there are no absolute contraindications for using adrenaline in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using adrenaline is a hypersensitivity reaction presenting with high blood pressure and tachycardia, which may be present at the initial phase of VIA, due to a psychologic reflex.
+In theory, these less severe cases of VIA may display some form of stress-related blood pressure increase.
+As we lack data to confirm or discard this theory future analysis of this question is of great clinical value.
 
@@
 
-Based on our findings, insects are the most probable elicitor of anaphylaxis in Europe during Summer-season, with VIA cases extending from early Spring to the end of Autumn.
+Based on our findings, insects are the most probable elicitor of anaphylaxis in Europe during summer-season, with VIA cases extending from early spring to the end of autumn.
 Detailed information on the seasonality of insect-elicited hypersensitivity reactions is scarce [@old].
@@
 Milder hypersensitivity reactions, as well as healthy controls, are not included in the database.
-Although The European Anaphylaxis Registry is ideal for investigating anaphylaxis subtypes - it might give an incomplete picture of the populational distribution of hypersensitivity reactions and restricts us to only comparing various forms of anaphylaxis.
+Although The European Anaphylaxis Registry is ideal for investigating anaphylaxis subtypes - it might give an incomplete perception of the populational distribution of hypersensitivity reactions and restricts us to only comparing various forms of anaphylaxis.
 
-Nevertheless, because the European Anaphylaxis Registry has until now gathered over 12 000 cases of anaphylaxis - it was possible to perform a case-controlled analysis on a relatively large number of cases and investigate many aspects of VIA.
+Nevertheless, because the European Anaphylaxis Registry has until now gathered over 12,000 cases of anaphylaxis - it was possible to perform a case-controlled analysis on a relatively large number of cases and investigate many aspects of VIA.
 
@@
 We thank all patients, parents, and their children for their support in providing data on the occurrence of anaphylaxis for this study.
-We thank Dr.
-rer.
-biol.
-Kristin Franke for revising the manuscript.
 We thank the study personnel for patients counseling and data entry, and we thank the members of The European Anaphylaxis Registry in detail:
@@
 
-```{r figskinsymptoms, fig.cap="Lack of skin symptoms (i.e., urticaria and flushing) during anaphylaxis is associated with more severe IVA.
+```{r figskinsymptoms, fig.cap="Lack of skin symptoms (i.e., urticaria and flushing) during anaphylaxis is associated with more severe VIA.
 A: lack of skin symptoms and mastocytosis in VIA and non-VIA cases.
@@
   cardiac_masto_plots[[6]]+
-              labs(x = "cardiac arrest"),
-  hypotension2,
-    nrow = 1,
-    ncol = 2
+              labs(x = "cardiac arrest",
+                   fill = "mastocytosis")#,
+  #hypotension2,
+  #  nrow = 1,
+  #  ncol = 2
   ),

### 2019-10-29 R/stat.R

@@ -2143,8 +2143,23 @@ fig_symptoms <- ggarrange(
              group,
              "'fraq_1' = 'VIA';
              'fraq_2' = 'non-VIA'"
-           )),
+           )) %>%
+           filter(!is.na(variableName)),
          aes(variableName,Fraction, fill = group))+
+    geom_rect(fill = c("#E4E4E4"),
+              xmin = c(-Inf),
+              xmax = c(4.5),
+              ymin = -Inf,
+              ymax = Inf,
+              alpha = 0.1,
+              color = "white")+
+    geom_rect(fill = c("#E4E4E4"),
+              xmin = c(8.5),
+              xmax = c(Inf),
+              ymin = -Inf,
+              ymax = Inf,
+              alpha = 0.1,
+              color = "white")+
     geom_bar(stat = "identity", position = "dodge",na.rm = T)+
     theme_classic()+
     theme(axis.text.x = element_text(angle = 30,hjust =1, size = 15),
@@ -2157,9 +2172,9 @@ fig_symptoms <- ggarrange(
     annotate("text", x = 1:11,
              y = c(0.48,0.46,.32,.22,.30,.11,.18,.14,.28,.14,.11),
              label = "*", size = 8)+
-    geom_segment(aes(x = 1, xend = 4, y = .52,yend = .52), linetype =1 )+
-    geom_segment(aes(x = 5, xend = 8, y = .52,yend = .52), linetype =1 )+
-    geom_segment(aes(x = 9, xend = 11, y = .52,yend = .52), linetype =1 )+
+    #geom_segment(aes(x = 1, xend = 4, y = .52,yend = .52), linetype =1 )+
+    #geom_segment(aes(x = 5, xend = 8, y = .52,yend = .52), linetype =1 )+
+    #geom_segment(aes(x = 9, xend = 11, y = .52,yend = .52), linetype =1 )+
     annotate("text", x = c(2.5,6.5,10),y = .54,label = c("cardio.","gastro.","resp."))+
     scale_y_continuous(labels = scales::percent_format(accuracy = 1)),
 
@@ -3120,7 +3135,7 @@ skin_symptoms$severe_urti_flush_plot <-
   ) +
   labs(y = "proportion",
        x = "severity [R&M]",
-       title = "non-mastocytic\npatients")+
+       title = "non-mastocytosis\npatients")+
   theme(axis.title.y = element_blank(),
         legend.position = "none")+
   annotate("text",x = 1.5, y = 0.45, label = "*",
[…]

### 2019-10-29 analysis/paper/paper.Rmd
corrections according to Magda

@@
 Patients who underwent VIA significantly less often received adrenaline treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
-After adjusting both groups for similar age, sex, and severity distribution - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)B).
+After adjusting both groups for similar severity - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)B).
 <!-- Antihistaminic drugs, on the other hand, were given more often in cases of VIA when compared to non-VIA. -->
@@
 Similarly, there were no differences in the adrenaline use between VIA and non-VIA when only severe reactions were taken into consideration (`r adren_severity_sev`).
-However, when we restricted the analysis to mild anaphylaxis cases --- non-VIA patients received adrenaline more frequently than VIA (`r adren_severity_mild`).
+However, when we restricted the analysis to moderate anaphylaxis cases --- non-VIA patients received adrenaline more frequently than VIA (`r adren_severity_mild`).
 

### 2019-10-29 R/make_flow.R

@@ -15,10 +15,7 @@ nodes <-
              paste0("Formal anaphylaxis definition met\n",
                     format(length(which(data4$reaction_type_brown=="anaphylaxis")),
                            big.mark = ",")),
-             paste0("Sex- and age-matching of venom induced\nanaphylaxis and other known elicitors\n",
-                    format(nrow(age_sex_matched %>%
-                                  filter(grouping == "other")),
-                           big.mark = ",")),
+             paste0("Sex- and age-matching of venom induced\nanaphylaxis and other known elicitors\n"),
              paste0("Insects as elicitors\n\n",
                     format(length(which(data4$d_elicitor_gr5== "insects" &
                                    data4$reaction_type_brown=="anaphylaxis")),

### 2019-12-05 analysis/paper/paper.Rmd
Treudler and Cardona

@@
 ---
-title: "Phenotype and risk assessment of insect venom anaphylaxis: a case - control study of the European Anaphylaxis Registry"
+title: "Phenotype and risk factors of venom induced anaphylaxis: a case - control study of the European Anaphylaxis Registry"
 author:
@@
 
+**Funding/ supporting grants:**
+
+**Conflict of interest:**
+
+R. Treudler reports grants and personal fees from Sanofi-Genzyme, ALK-Abello, Takeda,  Novartis, grants from Hautnetz Leipzig and Fraunhofer-IZI Leipzig, outside the submitted work.
+V. Cardona  reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LET, Thermofisher and Stallergens outside the submitted work.
+
+**List of abbreviations:**
+- VIA - Venom induced anaphylaxis
+- BST -Baseline Serum Tryptase
+
 ```{r setup, echo = FALSE,warning=FALSE,include=F}
@@
 ###### SETUP
-
+pacman::p_load(stargazer)
 ```
@@
 # Abstract
-**Introduction:** Insect-venom elicited anaphylaxis is a common, potentially life-threatening hypersensitivity reaction associated with specific: 1) symptom profile, 2) cofactors, and 3) management.
+**Introduction:** Venom induced anaphylaxis (VIA) is a common, potentially life-threatening hypersensitivity reaction associated with specific: 1) symptom profile, 2) cofactors, and 3) management.
 Identifying the differences in subtypes of anaphylaxis is crucial for future management guidelines and the development of a personalized medicine approach.
 
-**Methods:** Using data from the European Anaphylaxis Registry (`r nrow(data)` cases) we identified `r length(which(data4$d_elicitor_gr5== "insects" & data4$reaction_type_brown=="anaphylaxis"))` with insect-venom elicited anaphylaxis and analyzed these in comparison to sex-and-age-matched anaphylaxis cases triggered by other elicitors (n = `r nrow(age_sex_matched %>% filter(grouping == "other"))`).
+**Aim:** This study aimed to evaluate the phenotype and risk factors of venom induced anaphylaxis in comparison to other elicitors of anaphylaxis.
+
+**Methods:** Using data from the European Anaphylaxis Registry (`r nrow(data)` cases) we identified `r length(which(data4$d_elicitor_gr5== "insects" & data4$reaction_type_brown=="anaphylaxis"))` with VIA and analyzed these in comparison to sex-and-age-matched anaphylaxis cases triggered by other elicitors (non-VIA n = `r nrow(age_sex_matched %>% filter(grouping == "other"))`).
 
 **Results:** Venom induced anaphylaxis (VIA) more frequently involved more than three organ systems and was associated with cardiovascular symptoms.
-The absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase and was associated with an increased risk of a severe reaction.
+The absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase (BST) and was associated with an increased risk of a severe reaction.
 Intramuscular or intravenous adrenaline was administered significantly less often in VIA, in particular in patients without prior history of anaphylaxis.
-Baseline serum tryptase (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis.
+BST (8-11.5 ng/ml) was more frequently associated with severe anaphylaxis compared to BST values of < 8 ng/ml.
 
-**Conclusions:** Due to a specific symptom profile, frequently affecting the cardiovascular system, patients undergoing VIA are undertreated with adrenaline.
+**Conclusions:** Regarding the fact that VIA frequently affects the cardiovascular system, VIA patients were undertreated with adrenaline.
 The lack of skin symptoms (i.e.
-urticaria or flushing) during anaphylaxis (especially in combination with baseline serum tryptase levels within upper normal limits) is associated with severe reactions.
+urticaria or flushing) during anaphylaxis (especially in combination with BST levels within upper normal limits) is associated with severe reactions.
 
@@
 
-Hypersensitivity to insect venom is a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008].
-Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong systemic immunotherapy [@Sturm2017].
+Hypersensitivity to insect venom presents as a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008].
+Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong venom immunotherapy [@Sturm2017].
 There is a need for a more precise description of the diagnosis, identification of biomarkers, and better definition of phenotypes of anaphylaxis [@JimenezRodriguez2018].
@@
 Anaphylaxis is a clinical diagnosis with a variety of trigger factors and clinical presentations.
-Symptom profiles and specific cofactors for venom-induced anaphylaxis had previously been analyzed in an uncontrolled manner albeit relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].
+Symptom profiles and specific cofactors for venom-induced anaphylaxis (VIA) had previously been analyzed in an uncontrolled manner albeit relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].
 
 Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and its infrequent and random occurrence.
-Therefore registries, gathering clinical data from patients with a well documented (recent) history of anaphylaxis are crucial in investigating this entity.
+Therefore registries, gathering clinical data from patients with a well-documented (recent) history of anaphylaxis are crucial in investigating this entity.
 
-This study aimed to identify clinical patterns of venom-induced anaphylaxis (VIA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA) based on the the data from the European Anaphylaxis Registry.
+This study aimed to identify clinical patterns of VIA regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA) based on the data from the European Anaphylaxis Registry.
 
@@
 VIA displayed a specific symptom pattern.
-Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, hypotension, unconsciousness, reduced alertness) than patients with anaphylaxis due to other elicitors and less often presented with expiratory distress, rhinitis or diarrhea (Fig.
+Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, hypotension, unconsciousness, reduced alertness) than patients with anaphylaxis due to other elicitors and less often presented with respiratory distress, rhinitis or diarrhea (Fig.
 \@ref(fig:figsymptoms)A).
 
-Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrologic, skin, and respiratory systems, VIA more frequently involved more than three organ systems (`r compare(c(tempdf[3:4,">3 organ\nsystems"]))`,<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
+Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrointestinal, skin, and respiratory systems, VIA more frequently involved more than three organ systems (`r compare(c(tempdf[3:4,">3 organ\nsystems"]))`,<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
 Although, patients undergoing VIA  less often had concomitant atopic diseases  -->
@@
 
-Similarily, in non-mastocytosis patients undergoing VIA skin smptoms(i.e.
+Similarly, in non-mastocytosis patients undergoing VIA, skin symptoms (i.e.
 urticaria or flushing) were less often present than if anaphylaxis was triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
 Moreover, in this specific subgroup of patients (i.e.
-non-mastosis patients lacking skin symptoms) VIA was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
+non-mastocytosis patients lacking skin symptoms) VIA was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
 ")`%, `r pval(skin_symptoms$masto_chisq$p.value)`,  Fig.
@@
 By applying factorial logistic regression modeling, we confirmed a significant interaction effect between the presence of skin symptoms and insect venom on the severity of anaphylaxis (`r pval(skin_symptoms$severity_fit_skin_symptoms %>% coef %>% .[4,4])`).
-In other words, non-mastosis patients presenting without urticaria or flushing tended to have more severe anaphylaxis when triggered by insects, but not other elicitors of anaphylaxis (Fig.
+In other words, non-mastocytosis patients presenting without urticaria or flushing tended to have more severe anaphylaxis when triggered by insects.
+(Fig.
 \@ref(fig:figskinsymptoms)B, and Tab.
@@
 For this model, we excluded the cases with known mastocytosis and with BST above 11.5 ng/ml, potentially indicating non-diagnosed mastocytosis.
-Similarily, 1) tryptase levels were higher in VIA patients, 2) correlated with the severity of anaphylaxis, and 3) this effect was significant in VIA (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`) but not in the non-VIA group (Fig.
+Similarly, 1) tryptase levels were higher in VIA patients, 2) correlated with the severity of anaphylaxis, and 3) this effect was significant in VIA (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`) but not in the non-VIA group (Fig.
 \@ref(fig:figskinsymptoms)C-D).
 
-## Baseline serum tryptase over 8 ng/ml and concomitant cardiovascular conditions increase the risk of severe VIA.
+## BST over 8 ng/ml and concomitant cardiovascular conditions increase the risk of severe VIA
 
-Cofactors acn promote the onset and increase severity of anaphylaxis. 
-Therefore we evaluated a variety of known cofactors regarding their impact to increase the risk of severe VIA.
-As expected, the cofactor most prominently associated with an increased risk of severe anaphylaxis was mastocytosis (Fig.
+The cofactor most prominently associated with an increased risk of severe anaphylaxis was mastocytosis (Fig.
 \@ref(fig:figcofactors)).
@@
 Severe reactions of VIA were more prevalent in patients above 22 years of age, and in VIA cases vs.
-other elicitors (Fig.
+non-VIA cases (Fig.
 S\@ref(fig:sev)).
@@
 \@ref(fig:figcofactors)C).
-Beta-blocker use was associated with a higher severity of anaphylaxis and with the onset of cardiovascular symptoms (cardiac arrest, chest pain), but was comparable between `r summary(beta_fit)[["coefficients"]][3,4] %>% pval()`).
+Beta-blocker use was associated with a higher severity of anaphylaxis and with the onset of cardiovascular symptoms (cardiac arrest, chest pain), but was comparable between both VIA and non-VIA`r summary(beta_fit)[["coefficients"]][3,4] %>% pval()`).
 Surprisingly, arrhythmia was more frequently reported in patients with VIA and concomitant beta-blockers (Fig.
@@
 `r treact[1,2] %>% round(1)`% of patients with insect allergy had experienced venom anaphylaxis in the past.
-If the reaction was elicited by other elcitors (i.e.
+If the reaction was elicited by other elicitors (i.e.
 non-VIA) --- previous reactions were more frequently seen (`r treact[2,2]%>% round(1)`%, `r p %>% pval()`).
@@
 Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had insect elicited anaphylaxis and in `r greater_same[1,2]` of them (`r (greater_same[1,2]/sum(greater_same[,2])*100 )%>%round(1) `%) the following reaction was more severe than before.
-In `r greater_same[3,2]` out of the `r greater_same[1,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was similar in severity.
+In `r greater_same[3,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was similar in severity.
 
@@
 
-Next, we asked wether specific symptom clusters and treatment profiles could be identified within our cohort (association measured using phi coefficient).
-We found that patients dysplaing cardiovascular symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria were treated differently than patients with respiratory or gastrointestinal symptoms (Fig.
+Next, we asked whether specific symptom clusters and treatment profiles could be identified within our cohort (association measured using phi coefficient).
+We found that patients displaying cardiovascular symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria were treated differently than patients with respiratory or gastrointestinal symptoms (Fig.
 \@ref(fig:adrenalineuse)C).
@@
 
-## Hymenoptera anaphylaxis is a seasonal disease (may be moved to supplement).
+## Venom anaphylaxis is a seasonal disease (may be moved to supplement).
 
@@
 In this study, we identified distinct symptom-profile and treatment patterns of venom-induced anaphylaxis.
-The data unravel phenotypes of VIA, which may support to the development of tools incorporating clinical <!--- and biomarker--->data for predicting the severity of future episodes of anaphylaxis.
+The data unravel phenotypes of VIA, which may support the development of tools incorporating clinical <!--- and biomarker--->data for predicting the severity of future episodes of anaphylaxis.
 
 <!-- Symptoms -->
+
+Cofactors acn promote the onset and increase severity of anaphylaxis. 
+Therefore we evaluated a variety of known cofactors regarding their impact to increase the risk of severe VIA. As expected, t
+
 <!-- Cardio symptoms phenomenon -->
@@
 The rate of concomitant cardiovascular diseases was higher in VIA than non-VIA.
-They are an essential cofactor increasing the risk of a severe reaction **if insects elicited the anaphylaxis**.
+They are an essential cofactor increasing the risk of a severe reaction **if hymenoptera elicited the anaphylaxis**.
 This association was not significant in anaphylaxis elicited by other triggers.
@@
 Importantly, the absence of skin symptoms was associated with more severe VIA, which was still present after excluding patients with a known diagnosis of mastocytosis.
-Subsequently, the corellation of BST levels with severity of anaphylaxis lead us to identify an interaction between the absence of skin symptoms and VIA using generalized linear regression.
+Subsequently, the correlation of BST levels with severity of anaphylaxis lead us to identify an interaction between the absence of skin symptoms and VIA using generalized linear regression.
 
@@
 identified mast cell disorders in 17 out of 22 patients with VIA lacking skin symptoms  and concluded that patients with BST of 7.95 ng/ml and VIA should undergo extensive diagnostic procedures [@Zanotti2015].
-We recently identified that elederly patient undergoing anaphylaxis without concomitant skin symptoms tended to have more severe reactions [@Aurich2019]
+We recently identified that elderly patient undergoing anaphylaxis without concomitant skin symptoms tended to have more severe reactions [@Aurich2019]
 
@@
 
-Regarding the factors increasing the risk of VIA - adult patients experienced VIA more frequently. Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].
+Adult patients experienced VIA more frequently. Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].
 
@@
 <!-- Risk factors - Concomitant disease -->
-Similarly, we observed less VIA in patients with concomitant atopic diseases, as these patients more often present with food anaphylaxis [@Tham2019].
+Similarly, we observed less VIA in patients with concomitant atopic diseases (Fig. ) , as these patients more often present with food anaphylaxis [@Tham2019].
 
-The role of cardiovascular medication can not be isolated from the effect of concomitant cardiovascular conditions; therefore, we can not state wether ACE-I and beta-blockers increase the severity of anaphylaxis.
+The role of cardiovascular medication cannot be isolated from the effect of concomitant cardiovascular conditions; therefore, we cannot state whether ACE-I and beta-blockers increase the severity of anaphylaxis.
 However, we did observe that there were no significant differences between VIA and non-VIA cases regarding the symptoms and severity of an episode with concomitant use of ACE-I or beta-blockers.
@@
 Cases of VIA had been treated with adrenaline less often than the age- sex- and severity-matched cases of non-VIA.
-Moreover, the administration of adrenaline did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patient experienced their first episode of VIA.
+Moreover, the administration of adrenaline did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patient experienced their first episode of VIA (in comparison to non-VIA).
 The difference between groups was prominent for milder cases of anaphylaxis.
-The reason for this observation is unclear. To our knowledge, this is the only data on the comparative adrenaline usage in a case-controlled group of VIA vs. non-VIA.
+The reason for this observation is unclear.
+One explanation could be that emergency team more often attributed the IVA symptoms to anxiety, whereas in non-IVA, they were more often suspecting anaphylaxis.
+To our knowledge, this is the only data on the comparative adrenaline usage in a case-controlled group of VIA vs.
+non-VIA.
 
@@
 We thank the study personnel for patients counseling and data entry, and we thank the members of The European Anaphylaxis Registry in detail:
+Dr. Stephanie Aurich (Leipzig)
 
@@
 
-```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (VIA) compared to other elicitors.
+```{r figsymptoms, fig.cap="Symptoms of venom induced anaphylaxis (VIA) compared to other elicitors.
 A: Proportional presentation of specific reaction symptoms in VIA and non-VIA according to cardiovascular (cardio.), gastroenterologic (gastro.), and respiratory (resp.) organ systems.
@@
 
+```{r sup_cofactors, fig.cap="Cofactors of anaphylaxis."}
+age_sex_matched %>% group_by(grouping,atopy) %>% dplyr::count() %>% 
+  filter(!is.na(atopy)) %>% 
+  ggpubr::ggbarplot(
+    x = "atopy",
+    y = "n",
+    fill = "grouping",
+    palette = palBW,
+    position = position_fill()
+  )+
+  annotate("text", x = 1.5,y = .5,label = "*",size= 9)+
+
+```
+
 ```{r trypcardiocur, fig.cap="Tryptase levels in patients with concomitant cardiovascular diseases. "}

### 2019-12-05 R/functions.R

@@ -651,7 +651,7 @@ funtempspider <- function(var){
     mutate(variable = deparse(var)) %>%
     group_by(grouping) %>%
     nest() %>%
-    mutate(prop = map(data,function(x){
+    mutate(prop = purrr::map(data,function(x){
       x$n/sum(x$n)
     })
     ) %>% unnest() %>%

### 2019-12-05 R/ntprobnp analysis.R

@@ -1,0 +1,65 @@
+require(tidyverse)
+df <-read_csv("ntprobnp_results.csv")
+glimpse(df)
+df$ntprobnp[28] <- NA
+df$ntprobnp <- as.numeric(df$ntprobnp)
+ggpubr::ggscatter(df,
+       x = "d_severity_rm",
+       y = "ntprobnp")
+ggpubr::ggscatter(df,
+                  x = "q_114_hypotension_collapse_v5",
+                  y = "ntprobnp")
+ggpubr::ggscatter(df,
+                  x = "q_212_tryptase_value_v5",
+                  y = "ntprobnp")
+load("data.R")
+sum(data$b_patient_code %in% df$b_patient_code)
+
+df$b_patient_code
+sum(data$b_patient_code %>% as.character() %>% gsub(pattern=" ",replacement="") %in% df$b_patient_code)
+
+fdf <- data %>%
+  mutate(b_patient_code = data$b_patient_code %>% as.character() %>% gsub(pattern=" ",replacement="")) %>%
+  filter(b_patient_code %in% df$b_patient_code) %>%
+  select(b_patient_code,d_age,q_410_ad_cur,q_410_cardio_cur,q_114_loss_of_consciousness,q_111_urticaria,ANAscore,) %>%
+  full_join(df,by="b_patient_code")
+
+ggpubr::ggscatter(fdf,
+                  x = "d_age",
+                  y = "ntprobnp")
+
+ggpubr::ggscatter(fdf,
+                  x = "d_age",
+                  y = "ntprobnp",
+                  color = "q_410_cardio_cur")
+
+ggpubr::ggscatter(fdf,
+                  x = "q_212_tryptase_value_v5",
+                  y = "ntprobnp",
+                  color = "q_410_cardio_cur")
+
+
+ggpubr::ggscatter(fdf,
+                  x = "d_age",
+                  y = "ntprobnp",
+                  color = "q_114_hypotension_collapse_v5",
+                  facet.by = "d_severity_rm")
+
+
+ggpubr::ggscatter(fdf,
+                  x = "d_age",
+                  y = "ntprobnp",
+                  color = "q_114_loss_of_consciousness",
+                  facet.by = "d_severity_rm")
+
+ggpubr::ggscatter(fdf,
+                  x = "q_212_tryptase_value_v5",
+                  y = "ntprobnp",
+                  color = "q_111_urticaria",
+                  facet.by = "d_severity_rm")
+
[…]

### 2019-12-05 R/stat.R

@@ -1,20 +1,28 @@
 #####Setup####
-require(dplyr)
-require(magrittr)
-require(ggplot2)
-require(forestplot)
-require(summarytools)
-require(vcd)
-require(MatchIt)
-require(tidyr)
-library(ggradar)
-library(tibble)
-library(scales)
-require(ggpubr)
-require(rlang)
-library(forcats)
-require(fmsb)
-require(purrr)
+
+if (!require("pacman")) install.packages("pacman")
+pacman::p_load(dplyr,magrittr,ggplot2,forestplot,summarytools,
+               vcd,MatchIt,tidyr,ggradar,tibble,scales,ggpubr,rlang,
+               forcats,fmsb,purrr,DiagrammeR,DiagrammeRsvg,rsvg,
+               DescTools,optmatch,randomForest,ggfortify,heatmaply,verification)
+install.packages("questionr")
+#
+# require(dplyr)
+# require(magrittr)
+# require(ggplot2)
+# require(forestplot)
+# require(summarytools)
+# require(vcd)
+# require(MatchIt)
+# require(tidyr)
+# library(ggradar)
+# library(tibble)
+# library(scales)
+# require(ggpubr)
+# require(rlang)
+# library(forcats)
+# require(fmsb)
+# require(purrr)
 
 
 #### GET THE DATA ########
@@ -444,10 +452,10 @@ temp2 <- temp1 %>%
 
 temp <- temp %>% t()
 
-colnames(temp) <- temp[5,]
+colnames(temp) <- temp[4,]
 
 tempdf <- temp %>% data.frame(stringsAsFactors = F) %>%
-{rbind(rep(1,length(.)),rep(0,length(.)),.[2,],temp2 %>% t() %>% {.[2,]})}
+{rbind(rep(1,length(.)),rep(0,length(.)),.[5,],temp2 %>% t() %>% {.[5,]})}
 tempdf<- tempdf %>%
   mutate_if(is.character,as.numeric)
 # tempdf <- age_sex_matched %>%
@@ -715,17 +723,21 @@ F1$data[["dt"]]  <-
   })) %>%
   tidyr::unnest()
 
+
+# this introduces an error ! nest needs now cols!
 F1$data[["dt2"]] <- F1$data$dt %>%
   group_by(b_patient_code) %>%
-  tidyr::nest() %>%
-  mutate(compare_r = map(data,function(df){
+  nest() %>%
+  mutate(compare_r = purrr::map(data,function(df){
[…]

### 2019-12-11 analysis/paper/paper.Rmd
corrections coauthors done

@@
 title: "Phenotype and risk factors of venom induced anaphylaxis: a case - control study of the European Anaphylaxis Registry"
-author:
-  - Wojciech Francuzik^1^
-  - Franziska Ruëff^2^ 
-  - Sabine Dölle-Bierke^1^
-  - Claudia Pföhler^3^
-  - Kathrin Scherer Hofmeier^4^
-  - Bettina Wedi^5^
-  - A Bauer^6^ 
-  - N Wagner^7^
-  - Regina Treudler^8^
-  - Thomas Hawranek^9^
-  - Nicos Papadopoulus^10^
-  - Montserat Fernandes-Rivas^11^ 
-  - Jean Marie Reanudin^12^
-  - Victoria Cardona^13^
-  - M. Bilo^14^
-  - Izabela Poziomkowska-Gęsicka^15^
-  - George Christoff^16^
-  - Tilo Jakob^17^
-  - John Hourihane^18^
-  - L Ensina^19^
-  - Margitta Worm^1^
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-^1^ Division of Allergy and Immunology, Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,
+Wojciech Francuzik^1^,
+Franziska Ruëff^2^,
+Andrea Bauer^3^,.
+Maria Beatrice Bilò ^4,5^,
+Victoria Cardona^6^,
+George Christoff^7^,
+Sabine Dölle-Bierke^8^,
+Luis Ensina^9^,
+Montserat Fernandes-Rivas^7,11^,
+Thomas Hawranek^12^,
+Johnathan O'B Hourihane^13^,
+Tilo Jakob^14,15^,
+Nicos G. Papadopoulus^16,17^,
+Claudia Pföhler^18^,
+Iwona Poziomkowska-Gęsicka^19^,
+Xavier Van der Brempt^20^,
+Kathrin Scherer Hofmeier^21^,
+Regina Treudler^22^,
+Nicola Wagner^23^,
+Bettina Wedi^24^,
+Margitta Worm^1^
 
-^2^ Department of Dermatology and Allergology, Klinikum der Universität München, Germany
+^1^ Division of Allergy and Immunology, Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Germany
 
-^3^ Department of Dermatology, Saarland University Hospital, Homburg/Saar, Germany
+^2^ Department of Dermatology and Allergy, Ludwig-Maximilian University, Munich, Germany
 
-^4^ Department of Dermatology, University Hospital Basel, Switzerland
+^3^ University Allergy Center, University Hospital Carl Gustav Carus, Technical University Dresden, Dresden, German
+
+^4^ Allergy Unit, Department of Internal Medicine, University Hospital Ospedali Riuniti di Ancona, Ancona, Italy
+
+^5^ Department of Clinical and Molecular Sciences, Polytechnic University of Marche, Ancona, Italy
+
+^6^ Allergy Section, Department of Internal Medicine, Hospital Vall d'Hebron, Barcelona, Spain
+
+^7^ ARADyAL Research Network
+
+^8^  Faculty of Public Health, Medical University-Sofia, Sofia, Bulgaria
+
+^9^ Allergy Out-patient Department, Acibadem CityClinic, Tokuda Medical Centre, Sofia, Bulgaria
+
+^10^ Division of Allergy, Clinical Immunology and Rheumatology, Department of Pediatrics, Federal University of São Paulo, São Paulo, Brazil
+
+^11^ Department of Allergy, Hospital Clinico San Carlos, Universidad Complutense, IdISSC, Madrid, Spain
+
+^12^ Department of Dermatology, University Hospital Salzburg, Paracelsus Medical University, Salzburg, Austria
+
+^13^ DM University College Cork and Cork University Hospital, Cork, Ireland
+
+^14^ Department of Dermatology and Allergology, University Medical Center Giessen and Marburg, Justus-Liebig University Gießen, Gießen, Germany
+
+^15^ Allergy Research Group, Medical Center, University of Freiburg, Freiburg, Germany
+
+^16^ Allergy Department, 2nd Pediatric Clinic, National and Kapodistrian University of Athens, Athens, Greece
+
+^17^ Division of Infection, Immunity & Respiratory Medicine, University of Manchester, Manchester, UK
+
+^18^ Department of Dermatology, Saarland University Medical Center, Homburg/Saar, Germany
+
+^19^ Clinical Allergology Department, Pomeranian Medical University in Szczecin, Szczecin, Poland
+
+^20^ Allergy Vigilance Network, Nancy, France
+
+^21^ Division of Allergy, Department of Dermatology, University Hospital Basel, University of Basel, Basel, Switzerland
+
+^22^ Department of Dermatology, Venereology and Allergology, Leipzig Interdisciplinary Allergy Center (LICA)-Comprehensive Allergy Center, University Hospital, Leipzig, Germany
+
+^23^ 1Department of Dermatology, University Hospital Erlangen, Friedrich-Alexander-Universität Erlangen-Nürnberg (FAU), Erlangen, Germany
+
+^24^ Department of Dermatology and Allergy, Comprehensive Allergy Center, Hannover Medical School, Hannover, Germany
 
@@
 
-**Funding/ supporting grants:**
-
 **Conflict of interest:**
 
+A. Bauer reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LETI, Thermofisher, and Stallergens outside the submitted work
+N. Wagner reports personal fees from ALK outside the submitted work.
 R. Treudler reports grants and personal fees from Sanofi-Genzyme, ALK-Abello, Takeda,  Novartis, grants from Hautnetz Leipzig and Fraunhofer-IZI Leipzig, outside the submitted work.
 V. Cardona  reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LET, Thermofisher and Stallergens outside the submitted work. 
+M. B. Bilò reports personal fees from ALK outside the submitted work.
+K. Scherer reports personal fees from Allergopharma, Sanofi-Aventis, and Shire outside submitted work. 
+Franziska Ruëff reports personal fees outside the submitted work from from ALK-Abelló, Allergopharma, Bencard, Boehringer Ingelheim, Bristol Myers Squibb, Circassia, Dermira, DST, LEO Pharma, Lilly, Dr.
+Gerhard Mann chem.-pharm.
+Fabrik GmbH, Mylan, Novartis, Pfizer, Thermo Fisher Scientific and UCB
+Claudia Pföhler performed clinical studies for Allergy Therapeutics and received speaker honoraria and
+travel support from Bencard, Novartis and ALK.
 
@@
 - VIA - Venom induced anaphylaxis
-- BST -Baseline Serum Tryptase
+- BST - Baseline Serum Tryptase
 
@@
 
+pacman::p_load("tidyvese")
+
 devtools::load_all('.',quiet = T) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
@@
 **Introduction:** Venom induced anaphylaxis (VIA) is a common, potentially life-threatening hypersensitivity reaction associated with specific: 1) symptom profile, 2) cofactors, and 3) management.
-Identifying the differences in subtypes of anaphylaxis is crucial for future management guidelines and the development of a personalized medicine approach.
+Identifying the differences in phenotypes of anaphylaxis is crucial for future management guidelines and the development of a personalized medicine approach.
 
@@
 Intramuscular or intravenous adrenaline was administered significantly less often in VIA, in particular in patients without prior history of anaphylaxis.
-BST (8-11.5 ng/ml) was more frequently associated with severe anaphylaxis compared to BST values of < 8 ng/ml.
+Baseline serum tryptase within upper normal range (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis.
 
@@
 The lack of skin symptoms (i.e.
-urticaria or flushing) during anaphylaxis (especially in combination with BST levels within upper normal limits) is associated with severe reactions.
+urticaria or flushing) during anaphylaxis and BST levels within upper normal limits were associated with severe reactions.
 
@@
 Hypersensitivity to insect venom presents as a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008].
-Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong venom immunotherapy [@Sturm2017].
-There is a need for a more precise description of the diagnosis, identification of biomarkers, and better definition of phenotypes of anaphylaxis [@JimenezRodriguez2018].
-Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] for the diagnosis of anaphylaxis --- a better understanding of its clinical subtypes is required.
+Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong specific immunotherapy [@Sturm2017].
+There is a need for a more precise identification of biomarkers, and better definition of phenotypes of anaphylaxis [@JimenezRodriguez2018].
+Also, in order to facilitate a precision-medicine approach [@Muraro2017] for the diagnosis of anaphylaxis, a better understanding of its clinical phenotypes is required.
 
@@
 
-We searched the European Anaphylaxis Registry [@Grabenhenrich2016] (status from March 2019) for anaphylaxis cases elicited by insect's venom.
+We searched the European Anaphylaxis Registry [@Grabenhenrich2016] (status from March 2019) for anaphylaxis cases elicited by insect venom.
 The flowchart in Fig.
@@
 
-The diagnosis of anaphylaxis was based on the definition by NIAID/FAAN [@Sampson2006] and the severity according to Ring and Messmer Scale [@Ring1977].
+The diagnosis of anaphylaxis was based on the definition by NIAID/FAAN [@Sampson2006] and the severity according to the Ring and Messmer Scale [@Ring1977].
 Grades III and IV (presenting with significant hypoxia, hypotension, confusion, and loss of consciousness, or incontinence or cardiac arrest) were considered severe.
+Mastocytosis patients were defined as having a documented diagnosis of mastocytosis prior to the reaction.
+The Registry is designed for reporting cases of moderate to severe anaphylaxis (Ring and Messmer grades II - IV).
 
@@
 Next, we investigated the association of skin symptoms with the tryptase levels in non-mastocytosis patients.
-For this model, we excluded the cases with known mastocytosis and with BST above 11.5 ng/ml, potentially indicating non-diagnosed mastocytosis.
+For this model, we excluded the cases with known mastocytosis and with BST above 11.5 ng/ml, potentially indicating non-diagnosed mast cell activation disorders.
 Similarly, 1) tryptase levels were higher in VIA patients, 2) correlated with the severity of anaphylaxis, and 3) this effect was significant in VIA (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`) but not in the non-VIA group (Fig.
@@
 \@ref(fig:figcofactors)).
-Concomitant mastocytosis increased the risk for 1) cardiac arrest and 2) loss of consciousness in patients undergoing VIA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig.
+Concomitant mastocytosis increased the risk for 1) cardiac arrest and 2) loss of consciousness in patients undergoing VIA significantly more than in patients undergoing anaphylaxis due to other elicitors (Fig.
 \@ref(fig:figcofactors)C and Fig.
@@
 
-The effect of using ACE-I (as well as beta-blockers) on the risk of severe anaphylaxis correlated with the coexisting cardiovascular pathologies.
+The effect of using ACE-I (as well as beta-blockers) on the risk of severe anaphylaxis correlated with the coexisting cardiovascular diseases.
 ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) regardless of the elicitor  (Fig.
@@
 ttadren1 <- ANAscore_matched %>% 
-  group_by(d_522_adren_agg,grouping) %>% 
+  group_by(d_520_adren1,grouping) %>% 
   summarize(n = n()) %>% 
@@
 
+We evaluated adrenaline use (administered by any route from patients themselves and medical professionals) in both ambulatory and emergency room settings.   
 Patients who underwent VIA significantly less often received adrenaline treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
@@
 pvaladrenprev <- adren_prev$data %>% 
-  spread(key = d_522_adren_agg,value = n) %>% 
+  spread(key = d_520_adren1,value = n) %>% 
   filter(q_160_ever_react=="reacted previously") %>% 
@@
 pvaladrenprev_no <- adren_prev$data %>% 
-  spread(key = d_522_adren_agg,value = n) %>% 
+  spread(key = d_520_adren1,value = n) %>% 
   filter(q_160_ever_react=="no prev. ANA") %>% 
@@
 However, when we restricted the analysis to moderate anaphylaxis cases --- non-VIA patients received adrenaline more frequently than VIA (`r adren_severity_mild`).
+Presence of skin symptoms during these mild reactions also was associated with lower fraction of adrenaline treated patients (Fig.
+\@ref(fig:rfplot)).
 
@@
 β-2 agonists formed another therapy mode.
-
-## Venom anaphylaxis is a seasonal disease (may be moved to supplement).
-
-Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
-Their proportion of VIA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was below 1% of cases during winter.
-Nevertheless, `r rare_spring_autumn$n %>% sum` cases of VIA (bee -- *Apis mellifera* in Spring; yellow jacket -- *Vespula spp.* in Autumn) were triggered in March, April, and November.
-Yellow-jacket was the most prominent VIA-causing insect followed by bees.
-The VIA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in southern Europe.
 
@@
 
-<!-- Outdoor physical exercise (e.g., jogging in the park) was more often associated with VIA than other triggers of anaphylaxis (`r ttexercise1`).
+<!-- Outdoor physical exercise (e.g., jogging in the park) was more often associated with VIA than other elicitors of anaphylaxis (`r ttexercise1`).
 However, it was not associated with the severity of a reaction in these patients.
@@
 # Discussion
+
+The important function of the register especially in anaphylaxis where targetes studies are not possible.
 
@@
 In this study, we identified distinct symptom-profile and treatment patterns of venom-induced anaphylaxis.
-The data unravel phenotypes of VIA, which may support the development of tools incorporating clinical <!--- and biomarker--->data for predicting the severity of future episodes of anaphylaxis.
+The data unraveled phenotypes of VIA, which may support the development of tools incorporating clinical <!--- and biomarker--->data for predicting the severity of future episodes of anaphylaxis.
 
@@
 They are an essential cofactor increasing the risk of a severe reaction **if hymenoptera elicited the anaphylaxis**.
-This association was not significant in anaphylaxis elicited by other triggers.
+This association was not significant in anaphylaxis elicited by other elicitors.
 Notably, cardiac arrest occurred more frequently in patients with elevated BST (> 8 ng/ml), especially in VIA.
@@
 Milder hypersensitivity reactions, as well as healthy controls, are not included in the database.
-Although The European Anaphylaxis Registry is ideal for investigating anaphylaxis subtypes - it might give an incomplete perception of the populational distribution of hypersensitivity reactions and restricts us to only comparing various forms of anaphylaxis.
+Although The European Anaphylaxis Registry is ideal for investigating anaphylaxis phenotypes - it might give an incomplete perception of the populational distribution of hypersensitivity reactions and restricts us to only comparing various forms of anaphylaxis.
 
@@
 
-Based on our results, VIA is a distinctive subtype of anaphylaxis, with a specific symptom profile and risk factors.
+Based on our results, VIA is a distinctive phenotype of anaphylaxis, with a specific symptom profile and risk factors.
 VIA cases should undergo therapy according to the international management guidelines, and adrenaline should be given more often in VIA.
@@
 We thank the study personnel for patients counseling and data entry, and we thank the members of The European Anaphylaxis Registry in detail:
-Dr. Stephanie Aurich (Leipzig)
+Dr. S. Aurich (Leipzig), 
+V. Mahler, 
+Macarena Gonzales Knop (LMU Munich),
+I Manolaraki,
+D. Mitsias,
+E. Manoussakis,
+N. Douladiris,
+P. Xepapadaki (Athens, Greece), 
+Dr. R. Saternus, N. Krecké (Homburg, Germany)
+`r readxl::read_excel(path = "../data/raw_data/Co-Authors insect manuscript Oct-2019 (1).xlsx")[,1] %>% filter(!is.na(.)) %>% pull() %>% paste0(collapse = ", ")`.
 
@@
 C: Relation of reaction severity according to the elicitor and the absence of skin symptoms concerning categorized BST values.
-D: Continous values of BST according to the severity in both non-VIA and VIA with subgrouping to skin symptoms.", fig.width = 8, fig.height=5.5}
+D: Continous values of BST according to the severity in both non-VIA and VIA with subgrouping to skin symptoms.", fig.width = 8.5, fig.height=5.5}
 fig_skinsymptoms
@@
 
+# Supplementary Figures
+
+**Insect venom anaphylaxis is a seasonal disease. **
+
+Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
+Their proportion of VIA to anaphylaxis cases elicited by other elicitors during the summer seasons reached 60% and was below 1% of cases during winter.
+Nevertheless, `r rare_spring_autumn$n %>% sum` cases of VIA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were triggered in March, April, and November.
+Yellow-jacket was the most prominent VIA-causing insect followed by bees.
+The VIA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in southern Europe.
+
 ```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occurred.
 Less common insect species grouped as 'other'.
-B: The density distribution of VIA cases to cases elicited by other triggers considering the patient's age.
+B: The density distribution of VIA cases to cases elicited by other elicitors considering the patient's age.
 C: Geographical differences in the most common elicitors of VIA.
@@
 
-# Supplementary Figures
-
 ```{r matchingresult, fig.cap="Results of matching the cohort according to sex and age in order to perform a case-controlled study.
@@
 
-```{r supplement, fig.width= 6, fig.height= 5, fig.cap="Symptoms of anaphylaxis.
+#####
+
+```{r supplement, fig.width= 7, fig.height= 2.5, fig.cap="Symptoms of anaphylaxis.
 A: The association between cardiac arrest and concomitant mastocytosis in VIA and non-VIA.
@@
            section == "symptoms",
-           Cramer > 0.11,
-           !(variableName %in% c("q_114","q_143_fatal_adre_v5", "q_112_abdominal_pain")))%>%
+           Cramer > 0.10,
+           !(variableName %in% c("q_114",
+                                 "q_143_fatal_adre_v5",
+                                 "q_115_hot_sweat_tremble_v6",
+                                 "q_112_abdominal_pain",
+                                 "q_142_fatal_treatment_v5")))%>%
     mutate(variableName = car::recode(variableName,
@@
 cramV,
-nrow = 2, 
-ncol = 1,
-heights = c(2,1.8)
+nrow = 1, 
+ncol = 2,
+widths = c(1,1.8)
 )
@@
 
-```{r sup_cofactors, fig.cap="Cofactors of anaphylaxis."}
-age_sex_matched %>% group_by(grouping,atopy) %>% dplyr::count() %>% 
-  filter(!is.na(atopy)) %>% 
-  ggpubr::ggbarplot(
-    x = "atopy",
-    y = "n",
-    fill = "grouping",
-    palette = palBW,
-    position = position_fill()
-  )+
-  annotate("text", x = 1.5,y = .5,label = "*",size= 9)+
-
-```
+<!-- # ```{r supcofactors, fig.cap="Cofactors of anaphylaxis."}
+# age_sex_matched %>% group_by(grouping,atopy) %>% dplyr::count() %>% 
+#   filter(!is.na(atopy)) %>% 
+#   ggpubr::ggbarplot(
+#     x = "atopy",
+#     y = "n",
+#     fill = "grouping",
+#     palette = palBW,
+#     position = position_fill()
+#   )+
+#   annotate("text", x = 1.5,y = .5,label = "*",size= 9) ``` -->
 
@@
 
-```{r rfplot, fig.cap="Therapy of anaphylaxis. A: Adrenaline use considering if patients had previous anaphylaxis. B: Variable impo", fig.height=3.5, fig.width=7.1}
+```{r rfplot, fig.cap="Therapy of anaphylaxis.
+A: Patients who presented with skin symptoms and IVA less often recived adrenaline then if skin symptoms were absent during the reaction.
+B: Variable importance in the unsupervised classification between VIA and non-VIA using random Forest classifier.", fig.height=3.5, fig.width=7.1}
 ggarrange(
-  adren_prev+
-    geom_segment(aes(x = 0.5,
-                     xend = 2.5,
-                     y = 0.33,
-                     yend = 0.33),
-                 color="white",
-                 linetype =2),
+  # adren_prev+
+  #   geom_segment(aes(x = 0.5,
+  #                    xend = 2.5,
+  #                    y = 0.33,
+  #                    yend = 0.33),
+  #                color="white",
+  #                linetype =2),
+  skin_sympt_adrenaline,
   rf_ggplot_treatment,
@@
 
+#####
+
 ```{r factoriallogit, results="asis"}

### 2019-12-11 R/stat.R

@@ -5,7 +5,7 @@ pacman::p_load(dplyr,magrittr,ggplot2,forestplot,summarytools,
                vcd,MatchIt,tidyr,ggradar,tibble,scales,ggpubr,rlang,
                forcats,fmsb,purrr,DiagrammeR,DiagrammeRsvg,rsvg,
                DescTools,optmatch,randomForest,ggfortify,heatmaply,verification)
-install.packages("questionr")
+#install.packages("questionr")
 #
 # require(dplyr)
 # require(magrittr)
@@ -171,6 +171,7 @@ data4 <- data4 %>%
 
 data4 <- data4 %>%
   mutate(d_organ_sum_binom = ifelse(d_organ_sum<4,"1-3","4-5") %>% factor)
+
 ##### FINAL DATABASE RDB ######
 
 rdb <- data4[data4$reaction_type_brown=="anaphylaxis",]
@@ -179,6 +180,7 @@ countries <-rdb %>%
   dplyr::select(d_centres_country) %>%
   group_by(d_centres_country) %>%
   summarize(n=n())%>% arrange(desc(n))
+rdb <- rdb %>% filter(!q_340_insects%in%c("horse fly","mosquito"))
 
 #### Grouping variable insects vs no insects #####
 grouping <- ifelse(rdb$d_elicitor_gr5=="insects",
@@ -761,46 +763,46 @@ F2<- list(rationale = "We saw that patients who were treated for anaphylaxis aft
 
 F2$plot[["RM"]] <- rdb %>%
   filter(!is.na(d_severity_rm),
-         !is.na(d_522_adren_agg),
+         !is.na(d_520_adren1),
          !is.na(grouping),
-         d_522_adren_agg!="unknown") %>%
+         d_520_adren1!="unknown") %>%
   group_by(d_severity_rm,grouping) %>%
-  summarize(n =mean(d_522_adren_agg=="yes")) %>%
+  summarize(n =mean(d_520_adren1=="yes")) %>%
   ggplot(aes(d_severity_rm,n,fill = grouping))+
   geom_bar(stat="identity",position="dodge")
 
 
 F2$test <- rdb %>%
   filter(!is.na(d_severity_rm),
-         !is.na(d_522_adren_agg),
+         !is.na(d_520_adren1),
          !is.na(grouping),
-         d_522_adren_agg!="unknown") %>%
-  {glm(d_522_adren_agg~grouping*d_severity_rm,family = "binomial",data =.)} %>%
+         d_520_adren1!="unknown") %>%
[…]

### 2019-12-19 analysis/paper/paper.Rmd
xavier corrections

@@
 **Keywords:**
-  anaphylaxis, adrenaline (epinephrine), beta-blockers, insect venom allergy, Hymenoptera
+  anaphylaxis, epinephrine (adrenaline), beta-blockers, insect venom allergy, Hymenoptera
 
@@
 K. Scherer reports personal fees from Allergopharma, Sanofi-Aventis, and Shire outside submitted work. 
-Franziska Ruëff reports personal fees outside the submitted work from from ALK-Abelló, Allergopharma, Bencard, Boehringer Ingelheim, Bristol Myers Squibb, Circassia, Dermira, DST, LEO Pharma, Lilly, Dr.
+Franziska Ruëff reports personal fees outside the submitted work from ALK-Abelló, Allergopharma, Bencard, Boehringer Ingelheim, Bristol Myers Squibb, Circassia, Dermira, DST, LEO Pharma, Lilly, Dr.
 Gerhard Mann chem.-pharm.
@@
 The absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase (BST) and was associated with an increased risk of a severe reaction.
-Intramuscular or intravenous adrenaline was administered significantly less often in VIA, in particular in patients without prior history of anaphylaxis.
+Intramuscular or intravenous epinephrine was administered significantly less often in VIA, in particular in patients without prior history of anaphylaxis.
 Baseline serum tryptase within upper normal range (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis.
 
-**Conclusions:** Regarding the fact that VIA frequently affects the cardiovascular system, VIA patients were undertreated with adrenaline.
+**Conclusions:** Regarding the fact that VIA frequently affects the cardiovascular system, VIA patients were undertreated with epinephrine.
 The lack of skin symptoms (i.e.
@@
 
-Anaphylaxis is a clinical diagnosis with a variety of trigger factors and clinical presentations.
-Symptom profiles and specific cofactors for venom-induced anaphylaxis (VIA) had previously been analyzed in an uncontrolled manner albeit relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].
+Anaphylaxis is a clinical diagnosis with a variety of triggering factors and clinical presentations.
+Symptom profiles and specific cofactors for venom-induced anaphylaxis (VIA) had previously been analyzed in an uncontrolled manner albeit in relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].
 
-Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and its infrequent and random occurrence.
+Controlled clinical trials in anaphylaxis are difficult to conduct due to the acuteness of this life-threatening condition and its infrequent and random occurrence.
 Therefore registries, gathering clinical data from patients with a well-documented (recent) history of anaphylaxis are crucial in investigating this entity.
@@
 
-We searched the European Anaphylaxis Registry [@Grabenhenrich2016] (status from March 2019) for anaphylaxis cases elicited by insect venom.
+We searched the European Anaphylaxis Registry [@Grabenhenrich2016] (status until March 2019) for anaphylaxis cases elicited by insect venom.
 The flowchart in Fig.
@@
 
-Due to a large number of documented reactions in the European Anaphylaxis Registry - we were able to match the VIA with non-VIA cases according to sex and age to reduce the comparison bias. 
-When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig.
+Due to the large number of documented reactions in the European Anaphylaxis Registry, we were able to match the VIA with non-VIA cases according to sex and age to reduce the comparison bias. 
+When we analyzed a density plot of VIA cases according to age, we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig.
 \@ref(fig:flow)B).
@@
 Propensity score matching has been performed using the "MatchIt" package for R [@matchit].
-The results of the propensity score matching are in Fig.
+The results of the propensity score matching are illustrated in Fig.
 \@ref(fig:flow)B-D and supplementary Fig \@ref(fig:matchingresult).
 
-The final database consisted of `r countries %>% summarise(sum(n))` cases of venom induced anaphylaxis reported from allergy centers in `r countries %>% summarize(length(n))` countries and sex-and-age matched control group.
+The final database included `r countries %>% summarise(sum(n))` cases of venom induced anaphylaxis reported from allergy centers in `r countries %>% summarize(length(n))` countries and sex-and-age matched control group.
 We compared the frequency of various symptoms, cofactors --- known to increase the risk of severe anaphylaxis, [@Worm2018], and management in both groups.
 
-Based on the severity and symptom profile, and previous reports [@Ruff2009] we defined sub-elevated baseline serum tryptase (BST) values as 8 - 11.5 ng/ml (Fig. \@ref(fig:figcofactors)B).
+Based on the severity and symptom profile, and previous reports [@Ruff2009] we defined sub-elevated baseline serum tryptase (BST) values as 8 - 11.5 ng/ml (Fig. \@ref(fig:figskinsymptoms)B).
 
-<!-- ## Data handling and statistics -->
+<!-- ## Data handling and statitics -->
 We used the R Statistical Package [@R] for statistical analysis.
@@
 
-We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapeutic approaches that varied the most between VIA / non-VIA group and presented the results as Gini importance [@Strobl2007].
+We developed a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapeutic approaches that varied the most between VIA / non-VIA group and presented the results as Gini importance [@Strobl2007].
 Moreover, an association analysis of therapeutic interventions and symptoms was performed.
@@
 \@ref(fig:figcofactors)).
-Interestingly, BST values were increased in patients with concomitant cardiovasular diseases, irrespectively of the reaction severity (Fig.
+Interestingly, BST values were increased in patients with concomitant cardiovascular diseases, irrespectively of the reaction severity (Fig.
 S\@ref(fig:trypcardiocur)).
@@
 
-## VIA patients receive adrenaline less often than non-VIA
+## VIA patients receive epinephrine less often than non-VIA
 
@@
 
-We evaluated adrenaline use (administered by any route from patients themselves and medical professionals) in both ambulatory and emergency room settings.   
-Patients who underwent VIA significantly less often received adrenaline treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
-After adjusting both groups for similar severity - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)B).
+We evaluated epinephrine use (administered by any route from patients themselves and medical professionals) in both ambulatory and emergency room settings.   
+Patients who underwent VIA significantly less often received epinephrine treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
+After adjusting both groups for similar severity - the difference in epinephrine use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)B).
 <!-- Antihistaminic drugs, on the other hand, were given more often in cases of VIA when compared to non-VIA. -->
@@
 A positive history of anaphylaxis influenced the therapy of a current episode as well.
-Adrenaline as a first-line treatment was given less often in VIA cases when compared to other cases **if patients did not report a previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients reporting previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`, Fig.
+Epinephrine as a first-line treatment was given less often in VIA cases when compared to other cases **if patients did not report a previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients reporting previous reactions, there was no difference in epinephrine therapy (`r pvaladrenprev`, Fig.
 \@ref(fig:adrenalineuse)B).
-Similarly, there were no differences in the adrenaline use between VIA and non-VIA when only severe reactions were taken into consideration (`r adren_severity_sev`).
-However, when we restricted the analysis to moderate anaphylaxis cases --- non-VIA patients received adrenaline more frequently than VIA (`r adren_severity_mild`).
-Presence of skin symptoms during these mild reactions also was associated with lower fraction of adrenaline treated patients (Fig.
+Similarly, there were no differences in the epinephrine use between VIA and non-VIA when only severe reactions were taken into consideration (`r adren_severity_sev`).
+However, when we restricted the analysis to moderate anaphylaxis cases --- non-VIA patients received epinephrine more frequently than VIA (`r adren_severity_mild`).
+Presence of skin symptoms during these mild reactions also was associated with lower fraction of epinephrine treated patients (Fig.
 \@ref(fig:rfplot)).
@@
 Patients with VIA received corticosteroids and antihistamines significantly more frequently than patients with anaphylaxis to other elicitors.
-On the other hand, adrenaline, beta-2 mimetics, and oxygen were given more often to patients suffering from non-VIA  (Fig.
+On the other hand, epinephrine, beta-2 mimetics, and oxygen were given more often to patients suffering from non-VIA  (Fig.
 \@ref(fig:adrenalineuse)A, Fig.
@@
 \@ref(fig:adrenalineuse)C).
-The treatment of the former symptoms consisted of adrenaline autoinjector (AAI) use, i.v.
-adrenaline in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists.
+The treatment of the former symptoms consisted of epinephrine autoinjector (AAI) use, i.v.
+epinephrine in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists.
 Corticosteroids, i.v.
@@
 
-The important function of the register especially in anaphylaxis where targetes studies are not possible.
-
 <!-- Cardiologic symptoms and hypotensive collapse might be associated with venom anaphylaxis due to the vaso-vagal reflex.
@@
 
-Our findings indicate that patients with BST of over 8 ng/ml are prone to severe anaphylaxis to insect venom.
+Our findings indicate that patients with BST above 8 ng/ml are prone to severe anaphylaxis to insect venom.
 Patients with normal BST in the range of 8-11.4 ng/ml may have indolent systemic mastocytosis or concomitant undiagnosed mast cell activation syndrome (MCAS) [@Valent2019].
 Zanotti et al.
-identified mast cell disorders in 17 out of 22 patients with VIA lacking skin symptoms  and concluded that patients with BST of 7.95 ng/ml and VIA should undergo extensive diagnostic procedures [@Zanotti2015].
+identified mast cell disorders in 17 out of 22 patients with VIA lacking skin symptoms  and concluded that patients with BST above 7.95 ng/ml and VIA should undergo extensive diagnostic procedures [@Zanotti2015].
 We recently identified that elderly patient undergoing anaphylaxis without concomitant skin symptoms tended to have more severe reactions [@Aurich2019]
@@
 Adult patients experienced VIA more frequently. Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016]. 
-
 Emergency room (ER) admission data indicate that the frequency of insect stings hypersensitivity reactions in children is comparable to food hypersensitivity reactions (12-15% of cases of hypersensitivity reactions admitted to the ER), but pediatric anaphylaxis is triggered significantly more often by food elicitors (56% of food hypersensitivity cases vs.
@@
 <!-- Risk factors - Concomitant disease -->
-Similarly, we observed less VIA in patients with concomitant atopic diseases (Fig. ) , as these patients more often present with food anaphylaxis [@Tham2019].
+Similarly, we observed less VIA in patients with concomitant atopic diseases (Fig. \@ref(fig:supplement)) , as these patients more often present with food anaphylaxis [@Tham2019].
 
@@
 <!-- Therapy  -->
-Cases of VIA had been treated with adrenaline less often than the age- sex- and severity-matched cases of non-VIA.
-Moreover, the administration of adrenaline did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patient experienced their first episode of VIA (in comparison to non-VIA).
+Cases of VIA had been treated with epinephrine less often than the age- sex- and severity-matched cases of non-VIA.
+Moreover, the administration of epinephrine did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patients experienced their first episode of VIA (in comparison to non-VIA).
 The difference between groups was prominent for milder cases of anaphylaxis.
 The reason for this observation is unclear.
-One explanation could be that emergency team more often attributed the IVA symptoms to anxiety, whereas in non-IVA, they were more often suspecting anaphylaxis.
-To our knowledge, this is the only data on the comparative adrenaline usage in a case-controlled group of VIA vs.
+One explanation could be that emergency team more often attributed the VIA symptoms to anxiety, whereas in non-VIA, they were more often suspecting anaphylaxis.
+To our knowledge, this is the only data on the comparative epinephrine usage in a case-controlled group of VIA vs.
 non-VIA.
 
-Nevertheless, international guidelines of anaphylaxis state that adrenaline (i.m.) is the first-line agent in all diagnosed cases of anaphylaxis [@Muraro2014].
-Clinicians should not undermine the less severe VIA cases and treat them with adrenaline accordingly.
+Nevertheless, international guidelines of anaphylaxis state that epinephrine (i.m.) is the first-line agent in all diagnosed cases of anaphylaxis [@Muraro2014].
+Clinicians should not undermine the less severe VIA cases and treat them with epinephrine accordingly.
 
-Although there are no absolute contraindications for using adrenaline in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using adrenaline is a hypersensitivity reaction presenting with high blood pressure and tachycardia, which may be present at the initial phase of VIA, due to a psychologic reflex.
+Although there are no absolute contraindications for using epinephrine in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using epinephrine is a hypersensitivity reaction presenting with high blood pressure and tachycardia, which may be present at the initial phase of VIA, due to a psychologic reflex.
 In theory, these less severe cases of VIA may display some form of stress-related blood pressure increase.
@@
 Nevertheless, because the European Anaphylaxis Registry has until now gathered over 12,000 cases of anaphylaxis - it was possible to perform a case-controlled analysis on a relatively large number of cases and investigate many aspects of VIA.
+It is worth underlining the important function of international registries especially in diseases where targeted studies are not possible.
 
@@
 Based on our results, VIA is a distinctive phenotype of anaphylaxis, with a specific symptom profile and risk factors.
-VIA cases should undergo therapy according to the international management guidelines, and adrenaline should be given more often in VIA.
+VIA cases should undergo therapy according to the international management guidelines, and epinephrine should be given more often in VIA.
 
-When evaluating the risk of future severe episodes - patients with BST over 8 ng/ml should undergo extensive diagnostic tests to exclude ISM or MCAS and should be provided with two adrenaline autoinjectors for acute self-management.
+When evaluating the risk of future severe episodes - patients with BST above 8 ng/ml should undergo extensive diagnostic tests to exclude ISM or MCAS and should be provided with two epinephrine autoinjectors for acute self-management.
 
 # Acknowledgements
+
+```{r formatack}
+temp <- readxl::read_excel(path = "../data/raw_data/Co-Authors insect manuscript Oct-2019 (1).xlsx")[,1] %>% filter(!is.na(.)) %>% pull() 
+temp2 <- c(temp, 
+"V. Mahler, (Erlangen, Germany)", 
+"D. Mitsias, (Athens, Greece)",
+"E. Manoussakis, (Athens, Greece)",
+"N. Douladiris, (Athens, Greece)",
+"P. Xepapadaki, (Athens, Greece)", 
+"R. Saternus, (Homburg, Germany)",
+"N. Krecké (Homburg, Germany)",
+"J-M. Renaudin, (Nancy, France)",
+"S. Tscheiller, (Nancy, France)"
+) 
+temp3 <- gsub(temp2, pattern = ".*?\\. (..).*",replacement = "\\1")  %>% 
+  order()
+
+```
+
 We thank all patients, parents, and their children for their support in providing data on the occurrence of anaphylaxis for this study.
 We thank the study personnel for patients counseling and data entry, and we thank the members of The European Anaphylaxis Registry in detail:
-Dr. S. Aurich (Leipzig), 
-V. Mahler, 
-Macarena Gonzales Knop (LMU Munich),
-I Manolaraki,
-D. Mitsias,
-E. Manoussakis,
-N. Douladiris,
-P. Xepapadaki (Athens, Greece), 
-Dr. R. Saternus, N. Krecké (Homburg, Germany)
-`r readxl::read_excel(path = "../data/raw_data/Co-Authors insect manuscript Oct-2019 (1).xlsx")[,1] %>% filter(!is.na(.)) %>% pull() %>% paste0(collapse = ", ")`.
+
+`temp2[temp3] %>% paste0(collapse = ", ")`.
 
@@
 Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
-Their proportion of VIA to anaphylaxis cases elicited by other elicitors during the summer seasons reached 60% and was below 1% of cases during winter.
+The proportion of VIA to anaphylaxis cases elicited by other elicitors during the summer seasons reached 60% and was below 1% of cases during winter.
 Nevertheless, `r rare_spring_autumn$n %>% sum` cases of VIA (bee -- *Apis mellifera* in spring; yellow jacket -- *Vespula spp.* in autumn) were triggered in March, April, and November.
@@
 
-```{r supplement, fig.width= 7, fig.height= 2.5, fig.cap="Symptoms of anaphylaxis.
+```{r supplement, fig.width= 7, fig.height= 5, fig.cap="Symptoms of anaphylaxis.
 A: The association between cardiac arrest and concomitant mastocytosis in VIA and non-VIA.
@@
 C: Crammer's V as the measure of association between groups anaphylaxis (VIA vs.
-non-VIA)."}
+non-VIA).
+Higher values indicate stronger assotiation with IVA."}
 
@@
               labs(x = "cardiac arrest",
-                   fill = "mastocytosis")#,
+                   fill = "mastocytosis"),
   #hypotension2,
@@
   #  ncol = 2
-  ),
+age_sex_matched %>% group_by(grouping,atopy) %>% dplyr::count() %>%
+  filter(!is.na(atopy)) %>%
+  ggpubr::ggbarplot(
+    x = "atopy",
+    y = "n",
+    fill = "grouping",
+    palette = palBW,
+    position = position_fill()
+  )+
+  annotate("text", x = 1.5,y = .5,label = "*",size= 9)),
 cramV,
-nrow = 1, 
-ncol = 2,
-widths = c(1,1.8)
+nrow = 2, 
+ncol = 1
+#widths = c(1,1.8)
 )
@@
 ```{r rfplot, fig.cap="Therapy of anaphylaxis.
-A: Patients who presented with skin symptoms and IVA less often recived adrenaline then if skin symptoms were absent during the reaction.
+A: Patients who presented with skin symptoms and VIA less often received epinephrine than if skin symptoms were absent during the reaction.
 B: Variable importance in the unsupervised classification between VIA and non-VIA using random Forest classifier.", fig.height=3.5, fig.width=7.1}

### 2019-12-19 R/stat.R

@@ -3388,7 +3388,7 @@ plot_tryptase_in_cardio_cur <- age_sex_matched %>%
     position = position_fill(reverse = T),
     palette = "grey"
   )+
-  labs(x = "concomitant\ncardiovascualr disease",
+  labs(x = "concomitant\ncardiovascular disease",
        fill = "baseline\nserum\ntryptase\ncategory",
        y = "proportion")+
   theme(legend.position = "right")

### 2019-12-19 analysis/paper/paper.Rmd
Refs and absolute numbers

@@
 bibliography: references.bib
-csl: "../templates/expert-review-of-clinical-immunology.csl" # Insert path for the bib-style
+csl: "../templates/american-medical-association.csl" # Insert path for the bib-style
 ---
@@
 **Document statistics:**
-  Word count, figures, tables, references
+2838 words, 5 figures, 0 tables, 29 references
 
@@
 **List of abbreviations:**
-- VIA - Venom induced anaphylaxis
-- BST - Baseline Serum Tryptase
+
+- VIA - venom induced anaphylaxis
+- BST - baseline Serum Tryptase
+- EAI - epinephrine autoinjector
+- MCAS - mast cell activation syndrome
+- ER - emergency room
 
@@
 
+#####
+
+# Key points
+
+**Question:** What is the phenotype of insect venom anaphylaxis?
+
+**Findings:** This case-control study showed that venom induced anaphylaxis significantly more often presented with cardiovascular symptoms.
+Secondly, severe cases of venom induced anaphylaxis were significantly more often associated with the lack of skin symptoms upon anaphylaxis and showed a significant interaction with the levels of baseline serum tryptase (in range from 8 - 11 ng/ml).
+
+**Meaning:** Allegologists should consider intesified prophylaxis (and diagnostic) measures in patients with baseline tryptase levels of above 8 and a history of insect venom anaphylaxis.
+
+#####
+
 # Abstract
-**Introduction:** Venom induced anaphylaxis (VIA) is a common, potentially life-threatening hypersensitivity reaction associated with specific: 1) symptom profile, 2) cofactors, and 3) management.
+
+**Importance:** Venom induced anaphylaxis is a common, potentially life-threatening hypersensitivity reaction associated with specific: 1) symptom profile, 2) cofactors, and 3) management.
 Identifying the differences in phenotypes of anaphylaxis is crucial for future management guidelines and the development of a personalized medicine approach.
 
-**Aim:** This study aimed to evaluate the phenotype and risk factors of venom induced anaphylaxis in comparison to other elicitors of anaphylaxis.
+**Objective:** This study aimed to evaluate the phenotype and risk factors of venom induced anaphylaxis in comparison to other elicitors of anaphylaxis.
 
-**Methods:** Using data from the European Anaphylaxis Registry (`r nrow(data)` cases) we identified `r length(which(data4$d_elicitor_gr5== "insects" & data4$reaction_type_brown=="anaphylaxis"))` with VIA and analyzed these in comparison to sex-and-age-matched anaphylaxis cases triggered by other elicitors (non-VIA n = `r nrow(age_sex_matched %>% filter(grouping == "other"))`).
+**Design:** Using data from the European Anaphylaxis Registry (`r nrow(data)` cases) we identified `r length(which(data4$d_elicitor_gr5== "insects" & data4$reaction_type_brown=="anaphylaxis"))` with venom induced anaphylaxis and analyzed these in comparison to sex-and-age-matched anaphylaxis cases triggered by other elicitors (non-VIA n = `r nrow(age_sex_matched %>% filter(grouping == "other"))`).
 
-**Results:** Venom induced anaphylaxis (VIA) more frequently involved more than three organ systems and was associated with cardiovascular symptoms.
-The absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase (BST) and was associated with an increased risk of a severe reaction.
-Intramuscular or intravenous epinephrine was administered significantly less often in VIA, in particular in patients without prior history of anaphylaxis.
+**Setting:** Multicenter, international, prospective, case-control study.
+
+**Participants:** All patients with the diagnosis of moderate to severe anaphylaxis (according to Ring and Messmer grade II-IV) who gave an informed consent to provide their clinical data for the Anaphylaxis Registry were included in the study.
+Subsequently, we matched the group of patients presenting with venom anaphylaxis to a sex and age matched group of patients who presented with anaphylaxis to other elicitors (i.e.
+food, drugs).
+
+**Main Outcome(s) and Measure(s):** This study measured severity of anaphylaxis on the Ring and Messmer scale.
+Hypotheses regarding the comparisons between groups were formulated during and after data collection.
+
+**Results:** Venom induced anaphylaxis more frequently involved more than three organ systems and was associated with cardiovascular symptoms.
+The absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase and was associated with an increased risk of a severe reaction.
+Intramuscular or intravenous epinephrine was administered significantly less often in venom induced anaphylaxis, in particular in patients without prior history of anaphylaxis.
 Baseline serum tryptase within upper normal range (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis.
 
-**Conclusions:** Regarding the fact that VIA frequently affects the cardiovascular system, VIA patients were undertreated with epinephrine.
+**Conclusions and Relevance:** Regarding the fact that venom induced anaphylaxis frequently affects the cardiovascular system, patients were undertreated with epinephrine.
 The lack of skin symptoms (i.e.
-urticaria or flushing) during anaphylaxis and BST levels within upper normal limits were associated with severe reactions.
+urticaria or flushing) during anaphylaxis and baseline serum tryptase levels within upper normal limits were associated with severe reactions.
+Patients with serum tryptase of above 8 ng/ml and a history of venom anaphylaxis may require additional prophylaxis.
+
+#####
 
@@
 
-Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrointestinal, skin, and respiratory systems, VIA more frequently involved more than three organ systems (`r compare(c(tempdf[3:4,">3 organ\nsystems"]))`,<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
+Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrointestinal, skin, and respiratory systems, VIA more frequently involved more than three organ systems (`r compare(x = funtempspider(quo(d_organ_sum_binom))[c(2,5),3] %>% as.numeric(),y = c(tempdf[3:4,">3 organ\nsystems"]),type = "both")`,<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
 Although, patients undergoing VIA  less often had concomitant atopic diseases  -->
@@
 
-We found that `r (skin_symptoms$masto_tab$prop[c(2)]*100) %>% round(1)`% of patients with concomitant mastocytosis had anaphylaxis without skin symptoms (i.e.
-urticaria and flushing) which was significantly more frequent compared to patients without diagnosed mastocytosis (`r (skin_symptoms$masto_tab$prop[c(1)]*100) %>% round(1)`%, `r pval(skin_symptoms$masto_chisq$p.value)`).
+We found that `r skin_symptoms$masto_tab[2,3]` (`r (skin_symptoms$masto_tab$prop[c(2)]*100) %>% round(1)`%) of patients with concomitant mastocytosis had anaphylaxis without skin symptoms (i.e.
+urticaria and flushing) which was significantly more frequent compared to patients without diagnosed mastocytosis (`r skin_symptoms$masto_tab[1,3]`; `r (skin_symptoms$masto_tab$prop[c(1)]*100) %>% round(1)`%, `r pval(skin_symptoms$masto_chisq$p.value)`).
 This finding was most prominently seen in VIA (Fig.
@@
 Similarly, in non-mastocytosis patients undergoing VIA, skin symptoms (i.e.
-urticaria or flushing) were less often present than if anaphylaxis was triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
+urticaria or flushing) were less often present than if anaphylaxis was triggered by other elicitors (`r skin_symptoms$tab[2,3]`; `r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r skin_symptoms$tab[4,3]`; `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
 Moreover, in this specific subgroup of patients (i.e.
-non-mastocytosis patients lacking skin symptoms) VIA was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
-")`%, `r pval(skin_symptoms$masto_chisq$p.value)`,  Fig.
+non-mastocytosis patients lacking skin symptoms) VIA was significantly more frequently severe (`r skin_symptoms$severe_tab[c(6),4]`,`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
+")`%, `r skin_symptoms$severe_tab[c(2),4]`; `r pval(skin_symptoms$masto_chisq$p.value)`,  Fig.
 \@ref(fig:figskinsymptoms)B).
@@
 
-Concomitant cardiovascular diseases were more prevalent in VIA than in non-VIA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare()`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were not relevant in non-VIA cases (Fig.
+Concomitant cardiovascular diseases were more prevalent in VIA than in non-VIA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare(x = (age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% {.[c(2,5),3]} %>% pull()), type = "both")`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were not relevant in non-VIA cases (Fig.
 \@ref(fig:figcofactors)).
@@
 The effect of using ACE-I (as well as beta-blockers) on the risk of severe anaphylaxis correlated with the coexisting cardiovascular diseases.
-ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) regardless of the elicitor  (Fig.
+ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare( x = ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(n) %>% pull() %>% rev(), type = "both")`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) regardless of the elicitor  (Fig.
 \@ref(fig:figcofactors)C).
@@
 
-`r treact[1,2] %>% round(1)`% of patients with insect allergy had experienced venom anaphylaxis in the past.
+`r table(rdb$grouping,rdb$q_160_ever_react)[,1:2][1,2]` (`r treact[1,2] %>% round(1)`%) of patients with insect allergy had experienced venom anaphylaxis in the past.
 If the reaction was elicited by other elicitors (i.e.
-non-VIA) --- previous reactions were more frequently seen (`r treact[2,2]%>% round(1)`%, `r p %>% pval()`).
+non-VIA) --- previous reactions were more frequently seen `r table(rdb$grouping,rdb$q_160_ever_react)[,1:2][2,2]` (`r treact[2,2]%>% round(1)`%, `r p %>% pval()`).
 We observed `r sum(repeated_reaction_in_Reg)` patients with at least two fully-documented reactions.
@@
 \@ref(fig:adrenalineuse)C).
-The treatment of the former symptoms consisted of epinephrine autoinjector (AAI) use, i.v.
+The treatment of the former symptoms consisted of epinephrine autoinjector (EAI) use, i.v.
 epinephrine in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists.
@@
 
-When evaluating the risk of future severe episodes - patients with BST above 8 ng/ml should undergo extensive diagnostic tests to exclude ISM or MCAS and should be provided with two epinephrine autoinjectors for acute self-management.
+When evaluating the risk of future severe episodes - patients with BST above 8 ng/ml should undergo extensive diagnostic tests to exclude indolent systemic mastocytosis or MCAS and should be provided with two EAIs for acute self-management.
 
@@
 
-`temp2[temp3] %>% paste0(collapse = ", ")`.
+`r temp2[temp3] %>% paste0(collapse = ", ")`.
 
@@
 
-# Figures
+# Figure legends
 
@@
 
-# Supplementary Figures
+# Supplementary Figures (online-only material)
 
@@
 
-Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
+VIA in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October.
 The proportion of VIA to anaphylaxis cases elicited by other elicitors during the summer seasons reached 60% and was below 1% of cases during winter.

### 2019-12-19 R/functions.R

@@ -93,15 +93,19 @@ unknown_to_na <- function(vect){
   )
 }
 
-compare <- function(x,percent =T,rounding =1){
-  if(percent == T){
+compare <- function(x,y = NULL,type ="percent",rounding =1){
+  if(type == "percent"){
     paste0(round(x[1]*100,rounding),
            "% vs. ",
            round(x[2]*100,rounding),"%")
-  } else {
+  } else if (type == "absolute"){
     paste0(round(x[1],rounding),
            " vs. ",
            round(x[2],rounding))
+  } else if (type == "both"){
+    paste0(round(x[1],rounding), " (",round(y[1]*100,rounding),"%)",
+           " vs. ",
+           round(x[2],rounding)," (",round(y[2]*100,rounding),"%)")
   }
 }
 

### 2019-12-19 R/stat.R

@@ -1126,11 +1126,13 @@ lower_panel <- ggpubr::ggarrange(
     labs(y ="proportion",x = "Sex",fill = "elicitor"),
 
   ANAscore_matched  %>%
-    mutate(grouping = car::recode(grouping, "'insects'='VIA';'other'='non-VIA'")) %>%
+    #filter(severity_rmr)
+    mutate(grouping = car::recode(grouping, "'insects'='VIA';'other'='non-VIA'"),
+           d_severity_rmr = car::recode(d_severity_rmr, "'mild'='2';'severe'='3 & 4'")) %>%
     #select(grouping,d_severity_rm) %>%
-    group_by(grouping,d_severity_rm) %>%
+    group_by(grouping,d_severity_rmr) %>%
     summarize(n = n()) %>%
-    ggplot(aes(x = as.factor(d_severity_rm),y=n, fill = grouping)) +
+    ggplot(aes(x = as.factor(d_severity_rmr),y=n, fill = grouping)) +
     geom_bar(stat="identity",position="dodge")+
     theme_classic()+
     labs(x = "Severity grade [R&M]",fill = "",y="cases [n]")+
@@ -3463,7 +3465,7 @@ rffit <- randomForest(d_severity_rmr~
                         q_114_hypotension_collapse_v5
                         ,
              data = d %>% filter(subsets==1) %>%
-               dplyr:::select(-subsets),
+               dplyr::select(-subsets),
              importance = T,
              mtry = 3,
              ntree = 500,

### 2019-12-19 analysis/paper/paper.Rmd
gramarly corrections

@@
 ---
-title: "Phenotype and risk factors of venom induced anaphylaxis: a case - control study of the European Anaphylaxis Registry"
+title: "Phenotype and risk factors of venom-induced anaphylaxis: a case - control study of the European Anaphylaxis Registry"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-- VIA - venom induced anaphylaxis
+- VIA - venom-induced anaphylaxis
 - BST - baseline Serum Tryptase
@@
 
-**Findings:** This case-control study showed that venom induced anaphylaxis significantly more often presented with cardiovascular symptoms.
-Secondly, severe cases of venom induced anaphylaxis were significantly more often associated with the lack of skin symptoms upon anaphylaxis and showed a significant interaction with the levels of baseline serum tryptase (in range from 8 - 11 ng/ml).
+**Findings:** This case-control study showed that venom-induced anaphylaxis significantly more often presented with cardiovascular symptoms.
+Secondly, severe cases of venom-induced anaphylaxis were significantly more often associated with the lack of skin symptoms upon anaphylaxis and showed a significant interaction with the levels of baseline serum tryptase (in range from 8 - 11 ng/ml).
 
@@
 
-**Importance:** Venom induced anaphylaxis is a common, potentially life-threatening hypersensitivity reaction associated with specific: 1) symptom profile, 2) cofactors, and 3) management.
+**Importance:** Venom-induced anaphylaxis is a common, potentially life-threatening hypersensitivity reaction associated with specific: 1) symptom profile, 2) cofactors, and 3) management.
 Identifying the differences in phenotypes of anaphylaxis is crucial for future management guidelines and the development of a personalized medicine approach.
 
-**Objective:** This study aimed to evaluate the phenotype and risk factors of venom induced anaphylaxis in comparison to other elicitors of anaphylaxis.
+**Objective:** This study aimed to evaluate the phenotype and risk factors of venom-induced anaphylaxis in comparison to other elicitors of anaphylaxis.
 
-**Design:** Using data from the European Anaphylaxis Registry (`r nrow(data)` cases) we identified `r length(which(data4$d_elicitor_gr5== "insects" & data4$reaction_type_brown=="anaphylaxis"))` with venom induced anaphylaxis and analyzed these in comparison to sex-and-age-matched anaphylaxis cases triggered by other elicitors (non-VIA n = `r nrow(age_sex_matched %>% filter(grouping == "other"))`).
+**Design:** Using data from the European Anaphylaxis Registry (`r nrow(data)` cases) we identified `r length(which(data4$d_elicitor_gr5== "insects" & data4$reaction_type_brown=="anaphylaxis"))` with venom-induced anaphylaxis and analyzed these in comparison to sex- and age-matched anaphylaxis cases triggered by other elicitors (non-VIA n = `r nrow(age_sex_matched %>% filter(grouping == "other"))`).
 
@@
 
-**Participants:** All patients with the diagnosis of moderate to severe anaphylaxis (according to Ring and Messmer grade II-IV) who gave an informed consent to provide their clinical data for the Anaphylaxis Registry were included in the study.
-Subsequently, we matched the group of patients presenting with venom anaphylaxis to a sex and age matched group of patients who presented with anaphylaxis to other elicitors (i.e.
-food, drugs).
+**Participants:** All patients with the diagnosis of moderate to severe anaphylaxis (according to Ring and Messmer grade II-IV) who gave informed consent to provide their clinical data for the Anaphylaxis Registry were included in the study.
+Subsequently, we matched the group of patients presenting with venom anaphylaxis to sex- and age-matched group of patients who presented with anaphylaxis to other elicitors (i.e., food, drugs).
 
-**Main Outcome(s) and Measure(s):** This study measured severity of anaphylaxis on the Ring and Messmer scale.
+**Main Outcome(s) and Measure(s):** This study measured the severity of anaphylaxis on the Ring and Messmer scale.
 Hypotheses regarding the comparisons between groups were formulated during and after data collection.
 
-**Results:** Venom induced anaphylaxis more frequently involved more than three organ systems and was associated with cardiovascular symptoms.
+**Results:** Venom-induced anaphylaxis more frequently involved more than three organ systems and was associated with cardiovascular symptoms.
 The absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase and was associated with an increased risk of a severe reaction.
-Intramuscular or intravenous epinephrine was administered significantly less often in venom induced anaphylaxis, in particular in patients without prior history of anaphylaxis.
-Baseline serum tryptase within upper normal range (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis.
+Intramuscular or intravenous epinephrine was administered significantly less often in venom-induced anaphylaxis, in particular in patients without prior history of anaphylaxis.
+Baseline serum tryptase within the upper normal range (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis.
 
-**Conclusions and Relevance:** Regarding the fact that venom induced anaphylaxis frequently affects the cardiovascular system, patients were undertreated with epinephrine.
-The lack of skin symptoms (i.e.
-urticaria or flushing) during anaphylaxis and baseline serum tryptase levels within upper normal limits were associated with severe reactions.
+**Conclusions and Relevance:** Regarding the fact that venom-induced anaphylaxis frequently affects the cardiovascular system, patients were undertreated with epinephrine.
+The lack of skin symptoms (i.e., urticaria or flushing) during anaphylaxis and baseline serum tryptase levels within the upper normal limits were associated with severe reactions.
 Patients with serum tryptase of above 8 ng/ml and a history of venom anaphylaxis may require additional prophylaxis.
@@
 Hypersensitivity to insect venom presents as a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008].
-Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong specific immunotherapy [@Sturm2017].
-There is a need for a more precise identification of biomarkers, and better definition of phenotypes of anaphylaxis [@JimenezRodriguez2018].
+Venom-induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong specific immunotherapy [@Sturm2017].
+There is a need for more precise identification of biomarkers, and better definition of phenotypes of anaphylaxis [@JimenezRodriguez2018].
 Also, in order to facilitate a precision-medicine approach [@Muraro2017] for the diagnosis of anaphylaxis, a better understanding of its clinical phenotypes is required.
@@
 Anaphylaxis is a clinical diagnosis with a variety of triggering factors and clinical presentations.
-Symptom profiles and specific cofactors for venom-induced anaphylaxis (VIA) had previously been analyzed in an uncontrolled manner albeit in relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].
+Symptom profiles and specific cofactors for venom-induced anaphylaxis (VIA) had previously been analyzed in an uncontrolled manner, albeit in relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].
 
@@
 Mastocytosis patients were defined as having a documented diagnosis of mastocytosis prior to the reaction.
-The Registry is designed for reporting cases of moderate to severe anaphylaxis (Ring and Messmer grades II - IV).
+The Registry is designed for reporting cases of moderate to severe anaphylaxis (Ring and Messmer grades II-IV).
 
-Due to the large number of documented reactions in the European Anaphylaxis Registry, we were able to match the VIA with non-VIA cases according to sex and age to reduce the comparison bias. 
+Due to a large number of documented reactions in the European Anaphylaxis Registry, we were able to match the VIA with non-VIA cases according to sex and age to reduce the comparison bias. 
 When we analyzed a density plot of VIA cases according to age, we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig.
@@
 
-The final database included `r countries %>% summarise(sum(n))` cases of venom induced anaphylaxis reported from allergy centers in `r countries %>% summarize(length(n))` countries and sex-and-age matched control group.
+The final database included `r countries %>% summarise(sum(n))` cases of venom-induced anaphylaxis reported from allergy centers in `r countries %>% summarize(length(n))` countries and sex- and age-matched control group.
 We compared the frequency of various symptoms, cofactors --- known to increase the risk of severe anaphylaxis, [@Worm2018], and management in both groups.
 
-Based on the severity and symptom profile, and previous reports [@Ruff2009] we defined sub-elevated baseline serum tryptase (BST) values as 8 - 11.5 ng/ml (Fig. \@ref(fig:figskinsymptoms)B).
+Based on the severity and symptom profile and the previous reports [@Ruff2009], we defined sub-elevated baseline serum tryptase (BST) values as 8 - 11.5 ng/ml (Fig. \@ref(fig:figskinsymptoms)B).
 
@@
 We used the R Statistical Package [@R] for statistical analysis.
-A simple comparison of categorical variables was performed using either Chi^2^ test or Fisher's exact test (where the number of observations in a bin was less than 10).
-Continuous variables were analyzed using Mann-Whitney U test.
+A simple comparison of categorical variables was performed using either the Chi^2^ test or Fisher's exact test (where the number of observations in a bin was less than 10).
+Continuous variables were analyzed using the Mann-Whitney U test.
 In case of comparisons with two or more independent variables, we used Factorial ANOVA or Generalized Linear Models.
@@
 
-We developed a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapeutic approaches that varied the most between VIA / non-VIA group and presented the results as Gini importance [@Strobl2007].
-Moreover, an association analysis of therapeutic interventions and symptoms was performed.
+We developed a Random Forest classifier (using the "randomForest" package for R [@rf]) in order to find therapeutic approaches that varied the most between VIA / non-VIA group and presented the results as Gini importance [@Strobl2007].
+Moreover, association analysis of therapeutic interventions and symptoms was performed.
 The resulting phi values were scaled and presented in a heatmap with automatic clustering using Ward’s Agglomerative Hierarchical Clustering with Euclidean distances [@heatmaply].
@@
 
-We found that `r skin_symptoms$masto_tab[2,3]` (`r (skin_symptoms$masto_tab$prop[c(2)]*100) %>% round(1)`%) of patients with concomitant mastocytosis had anaphylaxis without skin symptoms (i.e.
-urticaria and flushing) which was significantly more frequent compared to patients without diagnosed mastocytosis (`r skin_symptoms$masto_tab[1,3]`; `r (skin_symptoms$masto_tab$prop[c(1)]*100) %>% round(1)`%, `r pval(skin_symptoms$masto_chisq$p.value)`).
+We found that `r skin_symptoms$masto_tab[2,3]` (`r (skin_symptoms$masto_tab$prop[c(2)]*100) %>% round(1)`%) of patients with concomitant mastocytosis had anaphylaxis without skin symptoms (i.e., urticaria and flushing), which was significantly more frequent compared to patients without diagnosed mastocytosis (`r skin_symptoms$masto_tab[1,3]`; `r (skin_symptoms$masto_tab$prop[c(1)]*100) %>% round(1)`%, `r pval(skin_symptoms$masto_chisq$p.value)`).
 This finding was most prominently seen in VIA (Fig.
@@
 
-Similarly, in non-mastocytosis patients undergoing VIA, skin symptoms (i.e.
-urticaria or flushing) were less often present than if anaphylaxis was triggered by other elicitors (`r skin_symptoms$tab[2,3]`; `r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r skin_symptoms$tab[4,3]`; `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
-Moreover, in this specific subgroup of patients (i.e.
-non-mastocytosis patients lacking skin symptoms) VIA was significantly more frequently severe (`r skin_symptoms$severe_tab[c(6),4]`,`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
+Similarly, in non-mastocytosis patients undergoing VIA, skin symptoms (i.e., urticaria or flushing) were less often present than if anaphylaxis was triggered by other elicitors (`r skin_symptoms$tab[2,3]`; `r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs.
+`r skin_symptoms$tab[4,3]`; `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`).
+Moreover, in this specific subgroup of patients (i.e., non-mastocytosis patients lacking skin symptoms) VIA was significantly more frequently severe (`r skin_symptoms$severe_tab[c(6),4]`,`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs.
 ")`%, `r skin_symptoms$severe_tab[c(2),4]`; `r pval(skin_symptoms$masto_chisq$p.value)`,  Fig.
@@
 
-Concomitant cardiovascular diseases were more prevalent in VIA than in non-VIA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare(x = (age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% {.[c(2,5),3]} %>% pull()), type = "both")`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were not relevant in non-VIA cases (Fig.
+Concomitant cardiovascular diseases were more prevalent in VIA than in non-VIA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare(x = (age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% {.[c(2,5),3]} %>% pull()), type = "both")`) and were associated with higher risk of severe anaphylaxis when elicited by insects but were not relevant in non-VIA cases (Fig.
 \@ref(fig:figcofactors)).
@@
 
-The effect of using ACE-I (as well as beta-blockers) on the risk of severe anaphylaxis correlated with the coexisting cardiovascular diseases.
+The effect of using ACE-I (as well as beta-blockers) on the risk of severe anaphylaxis correlated with coexisting cardiovascular diseases.
 ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare( x = ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(n) %>% pull() %>% rev(), type = "both")`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) regardless of the elicitor  (Fig.
@@
 `r table(rdb$grouping,rdb$q_160_ever_react)[,1:2][1,2]` (`r treact[1,2] %>% round(1)`%) of patients with insect allergy had experienced venom anaphylaxis in the past.
-If the reaction was elicited by other elicitors (i.e.
-non-VIA) --- previous reactions were more frequently seen `r table(rdb$grouping,rdb$q_160_ever_react)[,1:2][2,2]` (`r treact[2,2]%>% round(1)`%, `r p %>% pval()`).
+If the reaction was elicited by other elicitors (i.e., non-VIA) --- previous reactions were more frequently seen (`r table(rdb$grouping,rdb$q_160_ever_react)[,1:2][2,2]`; `r treact[2,2]%>% round(1)`%, `r p %>% pval()`).
 We observed `r sum(repeated_reaction_in_Reg)` patients with at least two fully-documented reactions.
-Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had insect elicited anaphylaxis and in `r greater_same[1,2]` of them (`r (greater_same[1,2]/sum(greater_same[,2])*100 )%>%round(1) `%) the following reaction was more severe than before.
-In `r greater_same[3,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was similar in severity.
+Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had insect elicited anaphylaxis and in `r greater_same[1,2]` of them (`r (greater_same[1,2]/sum(greater_same[,2])*100 )%>%round(1) `%), the following reaction was more severe than before.
+In `r greater_same[3,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases, the reaction was similar in severity.
 
@@
 We evaluated epinephrine use (administered by any route from patients themselves and medical professionals) in both ambulatory and emergency room settings.   
-Patients who underwent VIA significantly less often received epinephrine treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`).
+Patients who underwent VIA significantly less often received epinephrine treatment than in other anaphylaxis cases (`r ANAscore_matched %>% group_by(d_520_adren1,grouping) %>%   summarize(n = n()) %>% {.[3,3]}`; 26.9% vs.
+`r ANAscore_matched %>% group_by(d_520_adren1,grouping) %>%   summarize(n = n()) %>% {.[4,3]}`; 34.6%, `r test_adrenuse1$p.value %>% pval()`).
 After adjusting both groups for similar severity - the difference in epinephrine use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)B).
@@
 However, when we restricted the analysis to moderate anaphylaxis cases --- non-VIA patients received epinephrine more frequently than VIA (`r adren_severity_mild`).
-Presence of skin symptoms during these mild reactions also was associated with lower fraction of epinephrine treated patients (Fig.
+The presence of skin symptoms during these mild reactions also was associated with a lower fraction of epinephrine treated patients (Fig.
 \@ref(fig:rfplot)).
@@
 
-Cofactors acn promote the onset and increase severity of anaphylaxis. 
+Cofactors can promote the onset and increase the severity of anaphylaxis. 
 Therefore we evaluated a variety of known cofactors regarding their impact to increase the risk of severe VIA. As expected, t
@@
 The rate of concomitant cardiovascular diseases was higher in VIA than non-VIA.
-They are an essential cofactor increasing the risk of a severe reaction **if hymenoptera elicited the anaphylaxis**.
+They are an essential cofactor increasing the risk of a severe reaction **if Hymenoptera elicited the anaphylaxis**.
 This association was not significant in anaphylaxis elicited by other elicitors.
@@
 Importantly, the absence of skin symptoms was associated with more severe VIA, which was still present after excluding patients with a known diagnosis of mastocytosis.
-Subsequently, the correlation of BST levels with severity of anaphylaxis lead us to identify an interaction between the absence of skin symptoms and VIA using generalized linear regression.
+Subsequently, the correlation of BST levels with the severity of anaphylaxis lead us to identify an interaction between the absence of skin symptoms and VIA using generalized linear regression.
 
@@
 In theory, these less severe cases of VIA may display some form of stress-related blood pressure increase.
-As we lack data to confirm or discard this theory future analysis of this question is of great clinical value.
+As we lack data to confirm or discard this theory, future analysis of this question is of great clinical value.
 
@@
 Nevertheless, because the European Anaphylaxis Registry has until now gathered over 12,000 cases of anaphylaxis - it was possible to perform a case-controlled analysis on a relatively large number of cases and investigate many aspects of VIA.
-It is worth underlining the important function of international registries especially in diseases where targeted studies are not possible.
+It is worth underlining the important function of international registries, especially in diseases where targeted studies are not possible.
 
@@
 
-# Acknowledgements
+# Acknowledgments
 
@@
 
-```{r figsymptoms, fig.cap="Symptoms of venom induced anaphylaxis (VIA) compared to other elicitors.
+```{r figsymptoms, fig.cap="Symptoms of venom-induced anaphylaxis (VIA) compared to other elicitors.
 A: Proportional presentation of specific reaction symptoms in VIA and non-VIA according to cardiovascular (cardio.), gastroenterologic (gastro.), and respiratory (resp.) organ systems.
@@
 non-VIA).
-Higher values indicate stronger assotiation with IVA."}
+Higher values indicate stronger association with IVA."}
 
@@
 A: Patients who presented with skin symptoms and VIA less often received epinephrine than if skin symptoms were absent during the reaction.
-B: Variable importance in the unsupervised classification between VIA and non-VIA using random Forest classifier.", fig.height=3.5, fig.width=7.1}
+B: Variable importance in the unsupervised classification between VIA and non-VIA using Random Forest classifier.", fig.height=3.5, fig.width=7.1}
 ggarrange(

### 2019-12-19 analysis/paper/paper.Rmd
minor correction in fig e 3

@@
 ---
-title: "Phenotype and risk factors of venom-induced anaphylaxis: a case - control study of the European Anaphylaxis Registry"
+title: "Phenotype and risk factors of venom-induced anaphylaxis: a case-control study of the European Anaphylaxis Registry"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
   #  ncol = 2
-age_sex_matched %>% group_by(grouping,atopy) %>% dplyr::count() %>%
-  filter(!is.na(atopy)) %>%
+age_sex_matched %>% filter(grouping=="insects") %>% 
+  mutate(age_g = ifelse(d_age <22,"<22",">21")) %>% 
+  filter(q_114_hypotension_collapse_v5!="unknown") %>% 
+  group_by(age_g,q_114_hypotension_collapse_v5) %>% dplyr::count() %>%
   ggpubr::ggbarplot(
-    x = "atopy",
+    fill = "q_114_hypotension_collapse_v5",
     y = "n",
-    fill = "grouping",
+    x = "age_g",
     palette = palBW,
@@
   )+
+  labs(x = "age group [years]", fill = "hypotension")+
   annotate("text", x = 1.5,y = .5,label = "*",size= 9)),

### 2020-03-27 R/stat rev comm.Rmd
after review round 1 JACI

---
title: "Point-by-point reply to Reviewers' comments on the manusctipt JACI-D-20-00147"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    #bookdown::html_document2:
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../analysis/templates/template.docx" # Insert path for the DOCX file
bibliography: "../analysis/paper/references.bib"
csl: "../analysis/templates/american-medical-association.csl" # Insert path for the bib-style
---

## Reviewer #1

**The authors have important observations on the concomitant use of beta-blockers and ACE inhibitors and mores severe anaphylaxis; possibly a profound biostatistical analysis of this aspect could help to find  out whether this is just due to the "premorbidity of this patients" or an effect of these drugs.**

This is a very important question, but very difficult to analyze as patients with relevant cardiovascular diseases (mainly hypertension) are taking these drugs according to the current guidelines [@Mallat2013].

We were previously able to demonstrate the influence of ACE inhibitors and beta-blockers on the severity of anaphylaxis in a mouse model and found evidence that beta-blocker and ACE-I increased the activation of human mast cells in vitro, whereas these drugs had a lower effect when used alone [@Nassiri2015].

Taking all these findings together with this analysis we do not anticipate that premorbidity is the primary cause for our observations, but can, of course, be an additional exacerbating factor.

In another study regarding the general risk factors of anaphylaxis, we have also identified a correlation between ACE-I and beta-blockers use and an increased risk of severe anaphylaxis [@Worm2018]. Also there it was not possible to dissect the effect of concomitant cardiologic conditions from the effects of drugs. Therefore a different experimental approach is needed to address this question in humans.

Nevertheless, due to our interest in this topic and to partially answer the reviewer’s question - we have re-analyzed the effects of cardiologic drugs on the severity of anaphylaxis. Patients who were taking cardiologic drugs (i.e ASA, ACE-I, and beta-blockers) more often had severe anaphylaxis if they were taking these drugs in combination, when compared to monotherapy. This is in concordance with the above-mentioned observations.

```{r, warning =F,echo =FALSE, include = F}
knitr::opts_chunk$set(echo = F, warning = F,message = F)
pacman::p_load("rprojroot")
root <- rprojroot::is_r_package
R <- root$make_fix_file()

source(R("R/stat.R"))
cardio_drugs <- age_sex_matched %>%
  filter(!is.na(d_severity_rm),
         d_severity_rm!="1") %>%
  rowwise() %>%
  mutate(ace_beta= ifelse(q_423_beta=="yes"&q_423_ace=="yes"&q_423_asa=="yes",
                          "ACE-ASA-Beta",
                          ifelse(q_423_beta=="yes"&q_423_ace=="yes"&q_423_asa=="no",
                                 "ACE-Beta",
                                 ifelse(q_423_beta=="no"&q_423_ace=="yes"&q_423_asa=="yes",
                                        "ACE-ASA",
                                        ifelse(q_423_beta=="yes"&q_423_ace=="no"&q_423_asa=="yes",
                                               "ASA-Beta",
                                               ifelse(q_423_beta=="no"&q_423_ace=="yes"&q_423_asa=="no",
                                                      "ACE",
                                                      ifelse(q_423_beta=="no"&q_423_ace=="no"&q_423_asa=="yes",
                                                             "ASA",
                                                             ifelse(q_423_beta=="yes"&q_423_ace=="no"&q_423_asa=="no",
                                                                    "Beta",
                                                                     "none"
                                                                    ))))))) %>%
           factor(levels=c("none","ASA","ACE","Beta","ASA-Beta","ACE-Beta","ACE-ASA","ACE-ASA-Beta"))
         ) %>%
  group_by(ace_beta,d_severity_rm,grouping) %>%
  count() %>%
  ungroup() %>%
  filter(!is.na(ace_beta)) %>%
  mutate(d_severity_rm = factor(d_severity_rm))

cardio_drugs_df <- cardio_drugs %>%
  group_by(ace_beta,grouping) %>%
  summarise(sums = sum(n))
cardio_drugs_df$y <- 0.85
cardio_drugs_df$x <- lapply(1:8,rep,2) %>% unlist()
```

```{r fig.cap="Combination of cardiologic drugs and the severity of anaphylxis"}
ggbarplot(cardio_drugs,
          y ="n",
            fill ="d_severity_rm",
            x = "ace_beta",
            position = position_fill(),
            palette = "Greys",
            facet.by = "grouping")+
    labs(x = "Concomitant cardiologic drugs", y = "proportion", fill = "Severity [R&M]")+
  theme(axis.text.x = element_text(angle = 50, hjust =1)) +
  geom_text(data =cardio_drugs_df,aes(y =y, label = sums),inherit.aes = T,angle = 90)

```

**Are there any data available on the latency time between sting and symptoms in correlation with the severity of the reaction?**

We thank the reviewer for this interesting question. Indeed we did not provide data on time between elicitor exposure and the onset of anaphylaxis symptoms. We have analyzed the data and it is now mentioned in the text.

```{r}  
df_sev_time <- age_sex_matched %>%
    filter(!is.na(q_120_time_between_v4)) %>%
    mutate(time_r = ifelse(q_120_time_between_v4=="10","<10",">10")) %>%
    group_by(time_r,d_severity_rmr,grouping) %>%
    count() %>% 
  #ggbarplot(fill = "d_severity_rmr",
  #          x="time_r",
  #          facet.by = "grouping", 
  #          y = "n",
  #          position= position_fill())
  pivot_wider( names_from = d_severity_rmr,names_prefix = "grade_",values_from = n)

sev_time<-list()
sev_time$insects <- df_sev_time %>%
  filter(grouping =="insects") %>%
  {.[,c(3:4)]}  %>% as.matrix() %>% prop.table(margin = 1)
#chisq.test()
sev_time$insects_p <- df_sev_time %>%
  filter(grouping =="insects") %>%
  {.[,c(3:4)]}  %>% as.matrix() %>% chisq.test() %>% {.$p.value} %>% round(4)

sev_time$others <- df_sev_time %>%
  filter(grouping =="other") %>%
  {.[,c(3:4)]} %>% as.matrix() %>%  prop.table(margin = 1)
#  chisq.test()

sev_time$others_p <- df_sev_time %>%
  filter(grouping =="other") %>%
  {.[,c(3:4)]}  %>% as.matrix() %>% chisq.test() %>% {.$p.value} %>% round(4)

```

*Patients with insect venom anaphylaxis had more often severe anaphylaxis if the reaction occured in the first 10 minutes after exposure to venom (`r sev_time$insects[1,2] %>% {.*100} %>% round(2)`% of Ring and Messmer grade III or IV cases) then when the reaction occured after 10 minutes post exposure (`r sev_time$insects[2,2] %>% {.*100} %>% round(2)`% p = `r sev_time$insects_p`).*

```{r, fig.cap="Time between exposure to an elicitor and first anaphylaxis symptoms. Patients undergoing anaphylaxis due to insects more often had a quick reaction during the first 10 minutes, whereas patients with non-VIA reacted more often after 10 minutes. In the Insect venom anaphylaxis, we saw a decrease in severity in these patients that reacted after 10 minutes."}
pacman::p_load(ggmosaic)

sum_time_df <- age_sex_matched %>% 
    mutate(d_severity_rm = d_severity_rm %>%factor()) %>% 
  filter(d_severity_rm!=1,
         !is.na(q_120_time_between_v4),
         grouping =="insects"
         ) %>%
    group_by(q_120_time_between_v4) %>% 
    count()
sum_time_df$y = 0.8
sum_time_df$x = 1:5

  age_sex_matched %>% 
    mutate(d_severity_rm = d_severity_rm %>%factor()) %>% 
  filter(d_severity_rm!=1,
         !is.na(q_120_time_between_v4),
         grouping =="insects"
         ) %>%
    group_by(q_120_time_between_v4,d_severity_rm) %>% 
    count() %>% 
    ggpubr::ggbarplot(
      x = "q_120_time_between_v4",
      fill = "d_severity_rm",
      y = "n",
      position = position_fill(),
      palette = "Greys"
    )+
  geom_text(data =sum_time_df,aes(y =y, label = n,x = x),inherit.aes = T,angle = 90)+
  theme_classic()+
    theme(legend.position = "top")+
    labs(x = "Time between exposure and anaphylaxis [min]", y = "proportion", fill = "Severity [R&M]")
```

**In the supplementary material (also very interesting) the authors mention in Fig 1 other insects. It would be good to mention briefly in the text what insects these were (imported fire ants? others?)**

```{r}
n_unknown_insects <- data[which(data$q_340_insects=="insects"&data$d_centres_country!="not EU"),
                          c("d_390_other_elicitor_from_free","d_centres_country")] %>% 
  nrow()

per_unk_insects <- (n_unknown_insects/nrow(age_sex_matched %>% filter(grouping =="insects"))*100) %>% round(2)
```

Indeed the mentioned "other insects" were fire ants reported from the center in Brazil. We included this information in the figure description. 7 were mosquitos and 5 horse flies. The rest of the other insects' species were not be identified by history (n = `r n_unknown_insects`, `r per_unk_insects`% of all insect cases).

**References: there are 2 newer publications focussing on similar topics, possibly they could be added:**

**1. Risk factors and indicators of severe systemic insect sting reactions. Stoevesandt J, Sturm GJ, Bonadonna P, Oude Elberink JNG, Trautmann A. Allergy. 2019 Jun 13. doi: 10.1111/all.13945**

**2. Risk factors for severe systemic sting reactions in wasp (Vespula spp.) and honeybee (Apis mellifera) venom allergic patients. Fehr D, Micaletto S, Moehr T, Schmid-Grendelmeier P. Clin Transl Allergy. 2019 Oct 11;9:54.**

We have added these publications in the discussion.

## REVIEWER #2
**MAJOR COMMENTS:
The study "Phenotype and risk factors of VIA: a case-control study of the European Anaphylaxis registry" evaluates in a meaningful number of patients including children and adults with VIA and non-VIA. The paper is a descriptive clinical study evaluating the severity of anaphylaxis, involved organs, levels of BST and treatment models and comparing these parameters between VIA and non-VIA.
The study is a clinical descriptive study which does not bring a new concept or mechanistic innovative finding. My opinion is that the study is more appropriate for a journal with a clinical perspective as JACIP, etc.
The major strength of the study is the high number of the participants, It is well designed as a clinical and observational study,**

Major Concerns:
**1. The method part includes a very short information about propensity score matching done by MatchIT package for R and the eFig 2 shows different ratio, it needs more information to understand the propensity score.**

We agree that the propensity score matching has not been in detail explained in our methods. We added additional information to address this issue in the methods:

*"Cases were matched according to sex, age, and reaction severity to reduce the comparison bias by propensity score matching. The propensity score is a statistical approach to quantify the similarity between two unrelated cases. Propensity scores were calculated using the "MatchIt" package for R [@matchit]. MatchIt uses logistic regression to reduce the bias due to multiple confounding variables (i.e. sex and age) by weighing them and choosing cases with minimal differences in both groups."*

The supplementary figure caption also has been changed to:

"Results of matching the cohort according to sex and age to perform a case-controlled study. A: The original distribution of VIA and non-VIA cases according to age group and sex. The distribution of VIA and non-VIA cases in age groups is uneven. B: The distribution of VIA and non-VIA after age and sex matching with the use of MatchIt package for R. The ratio of VIA to non-VIA cases is approaching 50% indicating more balanced matching according to sex and age variables."

**2. Some of the patients registered were previously diagnosed as mastocytosis, what was the baseline tryptase level at these patients and which type of mastocytosis was the diagnosis and which criteria were positive for these pts?**

Thank you for this suggestion. The levels of BST are illustrated in the figure below. We mention the data in the text and added this figure to the supplementary material.

```{r BST, fig.cap="Levels of baseline serum tryptase in patients with VIA and non-VIA. A significant difference in BST between patients with concomitant mastocytosis and other patients (***). There was no significant difference between anaphylaxis elicited by insects and other elicitors (NS). Tested by two-way ANOVA."}

BST_masto$plot

```

When patients present in a participating allergy center after the reaction we ask if they had a previous diagnosis of mastocytosis. We do not include further questions about the classification of mastocytosis nor consider specific diagnosis criteria in our questionaire.

We agree with the reviewer that additional information regarding concomitant mastocytosis would provide data for novel analysis opportunities, however, this data is not available in the Anaphylaxis Registry.

We now mention in the text that the diagnosis of mastocytosis was based on the patients' history.

**4. Fig 2B shows the odd ratio of different factors for a severe reaction. Surprisingly the concomitant asthma is not significant for both VIA and non-VIA. How argue the authors this finding?**

Figure 2B illustrates the involvement of organ systems during VIA and non-VIA cases. We suppose that the reviewer meant figure 4A.

Although concomitant asthma has been previously suggested as a potential factor increasing the risk of anaphylaxis [@Niggemann2014] we have failed to confirm this observation in our data. We previously found that the influence of asthma varied greatly depending on which severity grading was used in a statistical model [@Worm2018]. Nevertheless, there are reports of asthma being a significant risk factor for fatal food anaphylaxis [@Smith2015]. We also recently showed asthma to be more often present in refractory anaphylaxis cases [@Francuzik2019]. It seems that asthma plays a role in life-threatening anaphylaxis, but its effect is insignificant when comparing mild-to-moderate vs. severe anaphylaxis.

```{r, fig.cap="Concomitant asthma and anaphylaxis severity. Number of observations per bar is reported vertically."}
age_sex_matched %>% 
  filter (q_410_asthma_cur !="unknown",
          !is.na(d_severity_rm),
          d_severity_rm!=1) %>%
  mutate(d_severity_rm = d_severity_rm %>% factor()) %>% 
  group_by(q_410_asthma_cur,grouping,d_severity_rm) %>% 
  count() %>% 
  ggbarplot(x = "q_410_asthma_cur",
      fill = "d_severity_rm",
      y = "n",
      position = position_fill(),
      palette = "Greys",
      facet.by = "grouping"
    )+
  geom_text(data =age_sex_matched %>% 
  filter (q_410_asthma_cur !="unknown") %>%
  group_by(q_410_asthma_cur,grouping) %>% 
  count() %>% cbind (y = rep(0.7,4)) ,
  aes(y =y, label = n ),inherit.aes = T,angle = 90)+
  theme_classic()+
    theme(legend.position = "top")+
    labs(x = "Concomitant asthma", y = "proportion", fill = "Severity [R&M]")
```

**5. Fig 2 b shows more cardiac involvement in VIA than non-VIA and this important finding is not mentioned in the results part line 168-170.**

Thank you for your comment, we added the information about the cardiac involvement accordingly.

*"Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrointestinal, skin, and respiratory systems, VIA more frequently involved more than three organ systems (`r compare(x = funtempspider(quo(d_organ_sum_binom))[c(2,5),3] %>% as.numeric(),y = c(tempdf[3:4,">3 organ\nsystems"]),type = "both")`, `r pval(p_organ$p.value)`), and predominantly involved cardiovascular system (`r compare(x = funtempspider(quo(q_114))[c(2,5),3] %>% as.numeric(),y = c(tempdf[3:4,"cardiac"]),type = "both")` `r pval(p_card$p.value)`"*

**6. Fig 4 B and C compare the proportion of pts with cardiac arrest, arrhythmia, etc between VIA and non-VIA and between with beta-blockers and without beta-blockers. The pts in the VIA group using beta-blockers or BST higher than 8 have more cardiac symptoms, is this finding statistically significant?**

The significance of these comparisons is difficult to show in the figures. These relationships are multivariable and there often are interactions between variables that are significant.

Using factorial logistic regression we can confirm that beta-blockers are significantly more frequently taken in patients that presented with cardiac arrest upon anaphylaxis, and this effect was diminished when anaphylaxis was elicited by insects.

There were no significant differences in arrhythmia occurrence in regards to beta-blockers use.
Intake of beta-blockers was associated with chest pain and angina symptoms upon anaphylaxis, independently of the elicitor.

```{r, include =F}
glm1 <- glm(q_114_cardiac_arrest~grouping*q_423_beta,
    data = age_sex_matched %>%
    filter(!is.na(q_423_beta),
           q_423_beta!="unknown",
           q_114_cardiac_arrest != "unknown",
           !is.na(grouping)) %>%
    mutate(grouping = car::recode(grouping,
                                  "'insects'='VIA';'other'='non-VIA'")),
    family = "binomial") %>% summary()

age_sex_matched %>%
    filter(!is.na(q_423_beta),
           q_423_beta!="unknown",
           q_114_cardiac_arrest != "unknown",
           !is.na(grouping)) %>%
    mutate(grouping = car::recode(grouping,
                                  "'insects'='VIA';'other'='non-VIA'")) %>% 
ggplot(aes(fill = q_114_cardiac_arrest, x = q_423_beta))+
  geom_mosaic(aes(x = product(q_423_beta),fill = factor(q_114_cardiac_arrest))
    )+
  facet_grid(.~grouping,scales ="free")+
  theme_classic()+
    theme(legend.position = "top")+
    labs(x = "Time between exposure and anaphylaxis [min]", y = "Severity [R&M]", fill = "Severity [R&M]")+
    scale_fill_brewer(palette = "YlOrRd")

```

```{r, include = F}
glm2 <- glm(q_114_palpitations_cardiac_arrythmia_v5~grouping*q_423_beta,
    data = age_sex_matched %>%
    filter(!is.na(q_423_beta),
           q_423_beta!="unknown",
           q_114_palpitations_cardiac_arrythmia_v5 != "unknown",
           !is.na(grouping)) %>%
    mutate(grouping = car::recode(grouping,
                                  "'insects'='VIA';'other'='non-VIA'")),
    family = "binomial") %>% summary()
```

```{r, include = F}
glm3 <- glm(q_114_chest_pain_angina_v5~grouping*q_423_beta,
    data = age_sex_matched %>%
    filter(!is.na(q_423_beta),
           q_423_beta!="unknown",
           q_114_chest_pain_angina_v5 != "unknown",
           !is.na(grouping)) %>%
    mutate(grouping = car::recode(grouping,
                                  "'insects'='VIA';'other'='non-VIA'")),
    family = "binomial") %>% summary()
```

```{r, include = F}
glm4 <-glm(q_114_cardiac_arrest~grouping*q_423_ace,
    data = age_sex_matched %>%
    filter(!is.na(q_423_ace),
           q_423_ace!="unknown",
           q_114_cardiac_arrest != "unknown",
           !is.na(grouping)) %>%
    mutate(grouping = car::recode(grouping,
                                  "'insects'='VIA';'other'='non-VIA'")),
    family = "binomial") %>% summary()
```

```{r, include = F}

glm6 <- glm(q_114_cardiac_arrest~grouping*tryp_cat,
    data =    age_sex_matched %>%
    filter(!is.na(d_severity_rmr),
           !is.na(tryp_cat)) %>%
    mutate(d_severity_rmr = d_severity_rmr %>%
             factor(labels = c("II","III+IV"))) %>%
#    group_by(grouping,q_114_cardiac_arrest,tryp_cat) %>%
#    summarize(n = n()) %>%
#      ungroup() %>%
      mutate(grouping = car::recode(grouping,
                                    "'insects' = 'VIA';'other'='non-VIA'")),
    family = "binomial") %>% summary()

glm7 <- glm(q_114_loss_of_consciousness~grouping*tryp_cat,
    data =    age_sex_matched %>%
    filter(!is.na(d_severity_rmr),
           !is.na(tryp_cat)) %>%
    mutate(d_severity_rmr = d_severity_rmr %>%
             factor(labels = c("II","III+IV"))) %>%
#    group_by(grouping,q_114_cardiac_arrest,tryp_cat) %>%
#    summarize(n = n()) %>%
#      ungroup() %>%
      mutate(grouping = car::recode(grouping,
                                    "'insects' = 'VIA';'other'='non-VIA'")),
    family = "binomial") %>% summary()

 age_sex_matched %>%
    filter(!is.na(q_423_beta),
           q_423_beta!="unknown",
            q_114_hypotension_collapse_v5 != "unknown",
           !is.na(grouping)) %>%
    mutate(grouping = car::recode(grouping,
                                  "'insects'='VIA';'other'='non-VIA'")) %>%
    group_by(q_423_beta,
              q_114_hypotension_collapse_v5,
             grouping) %>%
    summarize(n=n()) %>%
    ggbarplot(x = "q_423_beta",
              fill = "q_114_hypotension_collapse_v5",
              y = "n",
              facet.by = "grouping",
              position = position_fill(),
              palette = c("#848484","#1f1f1f"))+
    labs(y = "proportion",
         fill = "Beta blockers")
```

```{r include = F}
glm(q_114_loss_of_consciousness~grouping*q_423_beta,
    data = age_sex_matched %>%
    filter(!is.na(q_423_beta),
           q_423_beta!="unknown",
           q_114_loss_of_consciousness != "unknown",
           !is.na(grouping)) %>%
    mutate(grouping = car::recode(grouping,
                                  "'insects'='VIA';'other'='non-VIA'")),
    family = "binomial") %>% summary()

 age_sex_matched %>%
    filter(!is.na(q_423_beta),
           q_423_beta!="unknown",
            q_114_loss_of_consciousness != "unknown",
           !is.na(grouping)) %>%
    mutate(grouping = car::recode(grouping,
                                  "'insects'='VIA';'other'='non-VIA'")) %>%
    group_by(q_423_beta,
              q_114_loss_of_consciousness,
             grouping) %>%
    summarize(n=n()) %>%
    ggbarplot(x = "q_423_beta",
              fill = "q_114_loss_of_consciousness",
              y = "n",
              facet.by = "grouping",
              position = position_fill(),
              palette = c("#848484","#1f1f1f"))+
    labs(y = "proportion",
         fill = "Beta blockers")
```

```{r}
ttt <- glm1$coefficients %>% 
  rbind(glm2$coefficients) %>% 
  rbind(glm3$coefficients) %>%
  rbind(glm4$coefficients) %>%
  rbind(glm6$coefficients) %>%
  rbind(glm7$coefficients) %>% data.frame() %>% 
  tibble::rownames_to_column() %>% 
  mutate(`Pr...z..` = round(`Pr...z..`,4))

df22 <- data.frame(graph = c("B1","B2","B3","B4","C2","C3"),
  symptom = c("cardiac arrest",
                             "arrythmia",
                             "chest pain angina",
                             "cardiac arrest",
                             "cardiac arrest",
                             "loss of consciousness"),
                 IVA = c("0",
                         "↑",
                         "0",
                         "↓",
                         "0",
                         "↑"),
                 Betablockers = c("↑",
                                  "0",
                                  "↑",
                                  "",
                                  "",
                                  ""),
                 ACEI = c("",
                          "",
                          "",
                          "↑",
                          "",
                          ""),
                 Tryptase = c("",
                              "",
                              "",
                              "",
                              "↑",
                              "0"),
                 interaction = c("+",
                                 "0",
                                 "0",
                                 "0",
                                 "0",
                                 "+")

                 )
df22 %>% knitr::kable()
```

**MINOR COMMENTS: Line 147 mention BST values in Fig 3B but the figure should be 3C and D.**

Thank you for pointing that out. It has been corrected.

## REVIEWER #3

**MAJOR COMMENTS:**

**This report focuses on a sub-population of the European Anaphylaxis Registry and attempts to define phenotypes of insect sting anaphylaxis (VIA). This is therefore a study of severe insect sting anaphylaxis (VIA). The exclusion of patients with mild-to-moderate VIA provides a selected dataset and limits the generalizability of the results. The authors should point out at many points in the introduction, methods, results, and discussion, that this is a study limited to severe VIA and does not represent the full spectrum of VIA. There is a lack of depth in the references.**

We are afraid that a misunderstanding occurred. We do consider in this report mild, moderate and severe cases from grades II-IV according to Ring and Messmer. We decided to group the grades of 4 and 3 together as severe vs. mild-to-moderate cases of grade II. We arrived at a distribution that was very suited for statistical analysis as the two formed groups formed a nearly 50-50 split in VIA and non-VIA groups. Please inspect the figure 1D for details.

Due to the reviewer's comment, we have revised the line 130 in the methods section as it seems to be unclear that we also compared the severe cases (grades III+IV) to mild-to-moderate cases of grade II.

The method section now reads: 
*"The diagnosis of anaphylaxis was based on the definition by NIAID/FAAN [@Sampson2006] and the severity according to the Ring and Messmer Scale [@Ring1977]. Reactions of grade II were considered mild and grades III and IV (presenting with significant hypoxia, hypotension, confusion, and loss of consciousness, or incontinence or cardiac arrest) were considered severe."*

**MINOR COMMENTS:**

**1. Line 117 Some mention should be made here, and in the discussion, of Stoevesandt et al, who reported on the risk factors for severe anaphylaxis to insect stings, and specifically identified the absence of urticaria as a major finding. (Stoevesandt J, Hain J, Kerstan A, Trautmann A. Over- and underestimated parameters in severe Hymenoptera venom-induced anaphylaxis: Cardiovascular medication and absence of urticaria/angioedema. J Allergy Clin Immunol. 2012;130:698-704.)**

Thank you for pointing that out, we added this relevant reference in our discussion and introduction sections.

**Line 134 The registry was designed for reporting of grades II - IV, but this study excluded grade II. This selection bias weakens the analysis.**

As answered above we used cases of grades II to IV in this study and apologize if this was not quite clear.

**Line 143-144 What was the distribution of cases in the 11 countries? Did one or two centers predominate?**

```{r}
most <- age_sex_matched %>%
  group_by(d_centres_country) %>% 
  count() %>% 
  mutate(percent = (n/sum(.$n)*100) %>% round(2)) %>% 
  arrange(desc(n)) %>% {.[1:30,1]} %>% pull()

age_sex_matched %>%
  mutate(d_centres_country = forcats::fct_other(d_centres_country,keep = most)) %>% 
  group_by(d_centres_country) %>% 
  count() %>% 
  mutate(percent = (n/sum(.$n)*100) %>% round(2)) %>% 
  arrange(desc(n)) %>% 
  ggplot(aes(reorder(d_centres_country,n),n)) + 
           geom_histogram(stat = "identity")+
  theme(axis.text.x = element_text(angle=70,hjust = 1))+
  labs(x = "")

```

As illustrated in the figure above - most of the cases came from the centers in Germany. This is because the Anaphylaxis Registry was established in Germany and during the first years of its existence it only gathered cases in German centers. Subsequently, it was broadened to include german-speaking countries (Switzerland and Austria), other European countries, and now it enters the phase when centers from other continents are gradually starting to collaborate in this great effort.

**Line 170 There is no p-value here, and no detail in the figure.**

Thank you for pointing that out. We added the information about p value and included the stars in the figure.

```{r}
tempdf  %>%
  {.[,c(2,1,5,3,4)]} %>%
    rownames_to_column( var = "group" ) %>%
    mutate_at(vars(-group),funs(rescale)) %>%
    filter(group%in%c("VIA","non-VIA")) %>%
    ggradar(axis.label.offset = c(1.1),
            axis.label.size = 5,
            group.point.size = 3,
            legend.text.size = 7,
            grid.label.size = 0)+
    theme(legend.position = "none")+
    scale_color_manual(values = c("#1f1f1f", "#848484"))+
  annotate(geom = "text",y = 0.3076874,x = -0.951624024,label = "*",angle = 70,size =6)+
  annotate(geom = "text",y = -0.7587985,x = 0.5584042,label = "*",angle = 20,size =6)
```

**Line 180-184 The frequency of hives reported here is in grade III/IV reactors (a selected population). Here, as in many other places, it would be preferred to refer to "patients undergoing severe VIA" so as to distinguish this population with severe anaphylaxis from the broader spectrum of patients who present with VIA.**

This comment is related to #1 and is no longer valid. Please see our answer above.

**Line 226 The fact that 28% had a prior anaphylactic reaction to a sting is worthy of discussion. Is this a typical example of the failure of emergency and primary health care personnel to advise patients with sting anaphylaxis about the importance of allergy consultation for testing and preventative VIT?**

Thank you for this comment. 
We agree that this rate is high and suggests suboptimal adherence to the therapy guidelines.  We added this observation into the discussion.

*IVA patients had a documented history of anaphylaxis in 28% of the cases, and systemic immunotherapy has not been initiated in these patients. These data underline the utmost importance to recommend SIT to all patients who experienced VIA.*

```{r include = F}
age_sex_matched %>% 
   #filter(q_160_ever_react =="yes") %>% 
   filter(grouping == "insects") %>% 
  filter(d_year_reaction>2005) %>% 
   group_by(d_year_reaction) %>% 
   count() %>% 
  full_join( age_sex_matched %>% 
   filter(q_160_ever_react =="yes") %>% 
   filter(d_severity_rm!="1",
          grouping == "insects") %>% 
  filter(d_year_reaction>2005) %>% 
   group_by(d_year_reaction) %>% 
   count() ,
by = "d_year_reaction"
) %>% 
  mutate(percent = n.y/n.x*100) %>% 
  ggpubr::ggline(
    x = "d_year_reaction",
    y = "percent"
  )+
  labs(x = "Year",
       title = "Fraction of patients that undergo VIA\nas a repeated reaction")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1
                                   ))

```

**Line 230 Note that the 10% frequency of more severe reaction was in patients selected for severe (grade III/IV) reactions. This does not address the frequency of more severe reactions in milder systemic reactors.**

This comment is related to #1 and is no longer valid. Please see our answer above.

**Line 247 These were "moderate" not "mild" reactions, but they were milder than the severe (comparator) reactions. Mild anaphylactic reactions were excluded from the study.**

This comment is related to #1 and is no longer valid. Please see our answer above.

**Line 277 It is not clear what is meant by "at the onset"**

Changed the wording.

**Line 277 The authors make an important point about the inherent selection bias of this study. This is specific to the use of the Ring and Messmer scale, in comparison with the Mueller grading scale, or the scale of Simon Brown (Brown SG. Clinical features and severity grading of anaphylaxis. J Allergy Clin Immunol. 2004;114:371-6.) The Ring scale emphasizes hypoxia as an element of the Grade 3 reaction and shock as the definition of the Grade 4 reaction. In other grading systems, both are considered severe, whereas moderate anaphylaxis might include patients with dyspnea, throat tightness, and moderate hypotensive symptoms. Thus the current study would be more likely to include patients with shock and hypoxia (and fewer cases of respiratory and cutaneous symptoms), than if it were to structure its definition of "moderate and severe anaphylaxis" on a different scale. It would be of interest to include an analysis of the Grade 2 reactors to provide another comparator with the phenotypes described in the current analysis.**

This comment is related to #1 and is no longer valid. Please see our answer above.

**Lines 281-285 The discussion should include the reference to previously published studies that have reported this finding. This study adds to our knowledge by the strength of numbers that permits comparison of those with and without mastocytosis.**

The references have been added.

**Lines 281-285 These are important phenotypic correlations, showing the relation between BST levels, absence of urticaria, and severity of VIA. However, we must recall that the data are based on a population that excluded the majority of milder systemic reactors. The average allergist sees patients of all kinds with no study exclusions. It would be very useful to know how this paradigm fits in the patients with Grade II reactions.**

This comment is related to #1 and is no longer valid. Please see our answer above.

**Line 294 I agree, but perhaps the discussion should indicate the sensitivity of the test in these patients? Patients with a relatively low mast cell burden may have negative blood c-kit.**

We added the following text to the discussion:

*"Previous studies showed 92% sensitivity of this test in patients with Hymenoptera anaphylaxis, presenting without skin symptoms and with tryptase under 20 ng/ml."*

**Line 316 One likely explanation for the under-use of epinephrine in VIA is the absence of urticaria in many cases, especially those with a very severe reaction. Most physicians believe that "it is not anaphylaxis if there are no hives".**

Thank you for this comment. Indeed we also believe that this is true. 
We added the following sentence to the discussion:

*"A second possibility could be that many physicians fail to diagnose anaphylaxis when no skin symptoms are present."*

**Line 321 Do the authors mean "underestimate" rather than "undermine"?**

Yes, thank you.

**Line 326 The observations of Smith et al would contradict the assumption that the tachycardia is psychological. (Smith PL, Kagey-Sobotka A, Bleecker ER, Traystman R, Kaplan AP, Gralnick H, et al. Physiologic manifestations of human anaphylaxis. J Clin Invest. 1980;66:1072-80.)**

Thank you for this important comment, we fully agree with the Reviewer. We have changed the discussion in this paragraph accordingly.

*"Nevertheless, the three exceptionally well-documented cases of anaphylaxis upon sting challenge showed that the initial transient increase in blood pressure should not be interpreted as a contraindication to adrenaline and it could be safely given even if the heart rate was above 120 beats per minute."*

**Lines 351-356 Any recommendation for management of VIA should include the indication for appropriate testing and immunotherapy.**

Thank you for this suggestion. We changed the conclusions to read:

*"All cases should undergo appropriate allergological testing and indication for SIT should be evaluated. Patients with BST above 8 ng/ml should undergo extensive diagnostic tests to exclude indolent systemic mastocytosis or MCAS and should be provided with two EAIs for acute self-management."*

<!-- ## COMMENTS FROM THE EDITORIAL OFFICE: -->
<!-- ** Change the colour of the title & headings to black -->
<!-- ** Remove the date on the title page -->
<!-- ** Please include a disclosure statement for all of the authors. If appropriate, you may state: "The rest of the authors declare that they have no relevant conflicts of interest." -->
<!-- ** Move the Disclosure statement to title page -->
<!-- ** Please follow the JACI standard format for author names in references, according to Vancouver style. Examples are given in the Instructions for Authors. List all author names when there are six or fewer; when there are seven or more, list the first six and add et al. -->
<!-- ** Upload each figure in the manuscript separately and place all of the figure legends together in a section after the references -->
<!-- ** Upload each figure in the Online Repository separately and place all of the figure legends together in a section in the Online Repository text file. -->
<!-- ** Include Table S1 in the Online Repository text file -->
<!-- Your revision must include the following items: (1) point-by-point responses to the Editor and reviewer comments, (2) a marked copy of your revision showing the changes made, and (3) a clean (unmarked) copy of your revised manuscript. If your manuscript has any figures, tables, or Online Repository material in separate files, please be sure these are included in the revision as well. For further information regarding formatting of these elements, please consult the Guidelines for Submitting a Revision (found on the Editorial Manager homepage by clicking on Instructions for Authors). To avoid a delay in a final decision on your manuscript, please follow these instructions carefully. -->
<!-- If you have questions or encounter difficulties in submitting your revised manuscript, please contact James Galipeau at jgalipeau@origineditorial.com, providing the title and manuscript number. -->
<!-- Include interactive data visualizations in your publication and let your readers interact and engage more closely with your research. Follow the instructions here: https://www.elsevier.com/authors/author-services/data-visualization to find out about available data visualization options and how to include them with your article. -->
<!-- Sincerely, -->
<!-- The Editors -->
<!-- The Journal of Allergy and Clinical Immunology -->
<!-- The JACI is the #1 most-cited allergy/immunology journal, with an impact factor of 14.110. -->

## References

### 2020-03-27 analysis/paper/paper.Rmd
after review round 1 JACI

--- 2019-12-19 analysis/paper/paper.Rmd
@@
 ---
-title: "Phenotype and risk factors of venom-induced anaphylaxis: a case - control study of the European Anaphylaxis Registry"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+title: "Phenotype and risk factors of venom-induced anaphylaxis: a case-control study of the European Anaphylaxis Registry"
 output: 
@@
 bibliography: references.bib
-csl: "../templates/american-medical-association.csl" # Insert path for the bib-style
+csl: "../templates/vancouver.csl" # Insert path for the bib-style
 ---
@@
 Margitta Worm^1^
+
+**Conflict of interest:**
+
+A. Bauer reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LETI, Thermofisher, and Stallergens outside the submitted work
+N. Wagner reports personal fees from ALK outside the submitted work.
+R. Treudler reports grants and personal fees from Sanofi-Genzyme, ALK-Abello, Takeda,  Novartis, grants from Hautnetz Leipzig and Fraunhofer-IZI Leipzig, outside the submitted work.
+V. Cardona  reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LET, Thermofisher and Stallergens outside the submitted work. 
+M. B. Bilò reports personal fees from ALK outside the submitted work.
+K. Scherer reports personal fees from Allergopharma, Sanofi-Aventis, and Shire outside submitted work. 
+Franziska Ruëff reports personal fees outside the submitted work from ALK-Abelló, Allergopharma, Bencard, Boehringer Ingelheim, Bristol Myers Squibb, Circassia, Dermira, DST, LEO Pharma, Lilly, Dr.
+Gerhard Mann chem.-pharm.
+Fabrik GmbH, Mylan, Novartis, Pfizer, Thermo Fisher Scientific and UCB
+Claudia Pföhler performed clinical studies for Allergy Therapeutics and received speaker honoraria and
+travel support from Bencard, Novartis and ALK. The rest of the authors declare that they have no relevant conflicts of interest.
+
+**Affiliations:**
 
@@
 2838 words, 5 figures, 0 tables, 29 references
-
-**Conflict of interest:**
-
-A. Bauer reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LETI, Thermofisher, and Stallergens outside the submitted work
-N. Wagner reports personal fees from ALK outside the submitted work.
-R. Treudler reports grants and personal fees from Sanofi-Genzyme, ALK-Abello, Takeda,  Novartis, grants from Hautnetz Leipzig and Fraunhofer-IZI Leipzig, outside the submitted work.
-V. Cardona  reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LET, Thermofisher and Stallergens outside the submitted work. 
-M. B. Bilò reports personal fees from ALK outside the submitted work.
-K. Scherer reports personal fees from Allergopharma, Sanofi-Aventis, and Shire outside submitted work. 
-Franziska Ruëff reports personal fees outside the submitted work from ALK-Abelló, Allergopharma, Bencard, Boehringer Ingelheim, Bristol Myers Squibb, Circassia, Dermira, DST, LEO Pharma, Lilly, Dr.
-Gerhard Mann chem.-pharm.
-Fabrik GmbH, Mylan, Novartis, Pfizer, Thermo Fisher Scientific and UCB
-Claudia Pföhler performed clinical studies for Allergy Therapeutics and received speaker honoraria and
-travel support from Bencard, Novartis and ALK.
 
@@
 The diagnosis of anaphylaxis was based on the definition by NIAID/FAAN [@Sampson2006] and the severity according to the Ring and Messmer Scale [@Ring1977].
-Grades III and IV (presenting with significant hypoxia, hypotension, confusion, and loss of consciousness, or incontinence or cardiac arrest) were considered severe.
-Mastocytosis patients were defined as having a documented diagnosis of mastocytosis prior to the reaction.
+Reactions of grade II were considered mild and grades III and IV (presenting with significant hypoxia, hypotension, confusion, and loss of consciousness, or incontinence or cardiac arrest) were considered severe.
+Mastocytosis patients were defined as having a documented diagnosis of mastocytosis in medical history prior to the reaction.
 The Registry is designed for reporting cases of moderate to severe anaphylaxis (Ring and Messmer grades II-IV).
 
-Due to a large number of documented reactions in the European Anaphylaxis Registry, we were able to match the VIA with non-VIA cases according to sex and age to reduce the comparison bias. 
+Due to a large number of documented reactions in the European Anaphylaxis Registry, we were able to match the VIA with non-VIA cases according to sex and age. 
 When we analyzed a density plot of VIA cases according to age, we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig.
@@
 Subsequently, we compared the management in both groups and matched the control group according to the severity of a reaction.
-Propensity score matching has been performed using the "MatchIt" package for R [@matchit].
+
+Cases were matched according to sex, age, and reaction severity in order to reduce the comparison bias by propensity score matching.
+Propensity score is a statistical approach to quantify the simmilarity between two unrelated cases.
+Propensity scores were calculated using the "MatchIt" package for R [@matchit].
+MatchIt uses logistic regression to reduce the bias due to multiple confounding variables (i.e.
+sex and age) by weighing them and choosing cases with minimal differences in both groups.
 The results of the propensity score matching are illustrated in Fig.
@@
 
-Based on the severity and symptom profile and the previous reports [@Ruff2009], we defined sub-elevated baseline serum tryptase (BST) values as 8 - 11.5 ng/ml (Fig. \@ref(fig:figskinsymptoms)B).
+Based on the severity and symptom profile and the previous reports [@Ruff2009], we defined sub-elevated baseline serum tryptase (BST) values as 8 - 11.5 ng/ml (Fig. \@ref(fig:figskinsymptoms)C-D).
 
@@
 
-Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrointestinal, skin, and respiratory systems, VIA more frequently involved more than three organ systems (`r compare(x = funtempspider(quo(d_organ_sum_binom))[c(2,5),3] %>% as.numeric(),y = c(tempdf[3:4,">3 organ\nsystems"]),type = "both")`,<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
+Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrointestinal, skin, and respiratory systems, VIA more frequently involved more than three organ systems (`r compare(x = funtempspider(quo(d_organ_sum_binom))[c(2,5),3] %>% as.numeric(),y = c(tempdf[3:4,">3 organ\nsystems"]),type = "both")`, `r pval(p_organ$p.value)`), and predominantly involved cardiovascular system (`r compare(x = funtempspider(quo(q_114))[c(2,5),3] %>% as.numeric(),y = c(tempdf[3:4,"cardiac"]),type = "both")` `r pval(p_card$p.value)`<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml).
 Although, patients undergoing VIA  less often had concomitant atopic diseases  -->
@@
 
+BST levels were significantly higher in patients with a prior diagnosis of mastocytosis \@ref(fig:BST). 
 Next, we investigated the association of skin symptoms with the tryptase levels in non-mastocytosis patients.
@@
 \@ref(fig:figcofactors)C).
+
+```{r}  
+df_sev_time <- age_sex_matched %>%
+    filter(!is.na(q_120_time_between_v4)) %>%
+    mutate(time_r = ifelse(q_120_time_between_v4=="10","<10",">10")) %>%
+    group_by(time_r,d_severity_rmr,grouping) %>%
+    count() %>% 
+  #ggbarplot(fill = "d_severity_rmr",
+  #          x="time_r",
+  #          facet.by = "grouping", 
+  #          y = "n",
+  #          position= position_fill())
+  pivot_wider( names_from = d_severity_rmr,names_prefix = "grade_",values_from = n)
+
+sev_time<-list()
+sev_time$insects <- df_sev_time %>%
+  filter(grouping =="insects") %>%
+  {.[,c(3:4)]}  %>% as.matrix() %>% prop.table(margin = 1)
+#chisq.test()
+sev_time$insects_p <- df_sev_time %>%
+  filter(grouping =="insects") %>%
+  {.[,c(3:4)]}  %>% as.matrix() %>% chisq.test() %>% {.$p.value} %>% round(4)
+
+sev_time$others <- df_sev_time %>%
+  filter(grouping =="other") %>%
+  {.[,c(3:4)]} %>% as.matrix() %>%  prop.table(margin = 1)
+#  chisq.test()
+
+sev_time$others_p <- df_sev_time %>%
+  filter(grouping =="other") %>%
+  {.[,c(3:4)]}  %>% as.matrix() %>% chisq.test() %>% {.$p.value} %>% round(4)
+
+```
+
+IVA was more often severe if the reaction occured in the first 10 minutes after exposure to venom (`r sev_time$insects[1,2] %>% {.*100} %>% round(2)`% of Ring and Messmer grade III or IV cases) then when the reaction occured after 10 minutes post exposure (`r sev_time$insects[2,2] %>% {.*100} %>% round(2)`% p = `r sev_time$insects_p`).
 
@@
 
-Cofactors can promote the onset and increase the severity of anaphylaxis. 
-Therefore we evaluated a variety of known cofactors regarding their impact to increase the risk of severe VIA. As expected, t
+<!-- Cofactors can promote the onset and increase the severity of anaphylaxis.  -->
+<!-- Therefore we evaluated a variety of known cofactors regarding their impact to increase the risk of severe VIA. As expected, t -->
 
@@
 The rate of concomitant cardiovascular diseases was higher in VIA than non-VIA.
-They are an essential cofactor increasing the risk of a severe reaction **if Hymenoptera elicited the anaphylaxis**.
+They are an essential cofactor increasing the risk of a severe reaction if Hymenoptera elicited the anaphylaxis.
 This association was not significant in anaphylaxis elicited by other elicitors.
@@
 
-As the onset of cardiovascular symptoms like hypotension, collapse, or cardiac arrest lead to a higher grade on the Ring and Messmer scale than skin or gastrointestinal symptoms, VIA (being associated with cardiovascular symptoms) is likely to be associated with more severe anaphylaxis.
+As cardiovascular symptoms like hypotension, collapse, or cardiac arrest lead to a higher grade on the Ring and Messmer scale than skin or gastrointestinal symptoms, VIA (being associated with cardiovascular symptoms) is likely to be associated with more severe anaphylaxis.
 
 Importantly, the absence of skin symptoms was associated with more severe VIA, which was still present after excluding patients with a known diagnosis of mastocytosis.
-Subsequently, the correlation of BST levels with the severity of anaphylaxis lead us to identify an interaction between the absence of skin symptoms and VIA using generalized linear regression.
+Previous studies also observed this phenomenon [@Stoevesandt2012;@Fehr2019].
+Subsequently, the correlation of BST levels with the severity of anaphylaxis lead us to identify an interaction between the absence of skin symptoms and VIA using generalized linearf regression.
 
@@
 identified mast cell disorders in 17 out of 22 patients with VIA lacking skin symptoms  and concluded that patients with BST above 7.95 ng/ml and VIA should undergo extensive diagnostic procedures [@Zanotti2015].
-We recently identified that elderly patient undergoing anaphylaxis without concomitant skin symptoms tended to have more severe reactions [@Aurich2019]
+We recently identified that elderly patient undergoing anaphylaxis without concomitant skin symptoms tended to have more severe reactions [@Aurich2019].
+Our finding are in corcondance with a recent retrospective study from Fehr et al.
+[@Fehr2019] who identified lack of skin symptoms as a risk factor for severe VIA.
 
 Based on these and previous findings [@Ruff2009;@Zanotti2015;@JaraAcevedo2015] we propose to perform a peripheral blood KIT D816V mutation test in cases of BST of above 8 ng/ml and with a history of anaphylaxis presenting without urticaria or flushing.
+Previous studies showed 92% sensitivity of this test in patients with hymenoptera anaphylaxis, presenting without skin symptoms and with tryptase under 20 ng/ml [@Kristensen2017].
 
@@
 
-Adult patients experienced VIA more frequently. Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016]. 
+Age is an important risk factor for severe anaphylaxis [@Stoevesandt2019].
+Adult patients experienced VIA more frequently.
+Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].
 Emergency room (ER) admission data indicate that the frequency of insect stings hypersensitivity reactions in children is comparable to food hypersensitivity reactions (12-15% of cases of hypersensitivity reactions admitted to the ER), but pediatric anaphylaxis is triggered significantly more often by food elicitors (56% of food hypersensitivity cases vs.
@@
 One explanation could be that emergency team more often attributed the VIA symptoms to anxiety, whereas in non-VIA, they were more often suspecting anaphylaxis.
+A second possibility could be that many physicians fail to diagnose anaphylaxis when no skin symptoms are present.
 To our knowledge, this is the only data on the comparative epinephrine usage in a case-controlled group of VIA vs.
@@
 Nevertheless, international guidelines of anaphylaxis state that epinephrine (i.m.) is the first-line agent in all diagnosed cases of anaphylaxis [@Muraro2014].
-Clinicians should not undermine the less severe VIA cases and treat them with epinephrine accordingly.
+Clinicians should not underestimate the less severe VIA cases and treat them with epinephrine accordingly.
 
-Although there are no absolute contraindications for using epinephrine in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using epinephrine is a hypersensitivity reaction presenting with high blood pressure and tachycardia, which may be present at the initial phase of VIA, due to a psychologic reflex.
-In theory, these less severe cases of VIA may display some form of stress-related blood pressure increase.
-As we lack data to confirm or discard this theory, future analysis of this question is of great clinical value.
+Although there are no absolute contraindications for using epinephrine in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using epinephrine is a hypersensitivity reaction presenting with high blood pressure and tachycardia, which may be present at the initial phase of VIA.
+Nevertheless, the three exceptionally well documented cases of anaphylaxis upon sting challenge showed that the initial transient increase in blood presure should not be interpreted as a contraindication to andrenaline and it could be safely given even if the heart rate was above 120 beats per minute [@Smith1980].
 
+IVA patients had a documented history of anaphylaxis in 28% of the cases, and systemic immunotherapy has not been initiated in these patients, what is against latest management guidelines, although this fraction may be slowly decreasing it is of utmost importance to recommend SIT to all patients who experienced VIA.
 <!-- DONE Seasonality -->
@@
 
-When evaluating the risk of future severe episodes - patients with BST above 8 ng/ml should undergo extensive diagnostic tests to exclude indolent systemic mastocytosis or MCAS and should be provided with two EAIs for acute self-management.
+All cases should undergo appropriate allergologic testing and indication for SIT should be evaluated.
+Patients with BST above 8 ng/ml should undergo extensive diagnostic tests to exclude indolent systemic mastocytosis or MCAS and should be provided with two EAIs for acute self-management.
 
@@
 Two age-subsets of patients could be recognized based on the density plot of age (B)."}
-ggarrange(ggplot()+
+
+fig1<-ggarrange(ggplot()+
             background_image(png::readPNG("../figures/flow.png")),
@@
 )
+
+fig1
+```
+
+```{r,include=F}
+tiff("fig1.tiff", width = 7, height = 4.5,units = 'in', res = 300)
+fig1
+dev.off()
 ```
@@
 
+```{r,include=F}
+tiff("fig2.tiff", width = 8.5, height = 6.2,units = 'in', res = 300)
+fig_symptoms_3
+dev.off()
+```
+
 #####
@@
 
+```{r,include=F}
+tiff("fig3.tiff", width = 8.5, height = 5.5,units = 'in', res = 300)
+fig_skinsymptoms
+dev.off()
+```
+
 #####
@@
 cof_fig
+```
+
+```{r,include=F}
+tiff("fig4.tiff", width = 6.6, height = 9,units = 'in', res = 300)
+cof_fig
+dev.off()
 ```
@@
 
+```{r,include=F}
+tiff("fig5.tiff", width = 7, height = 9,units = 'in', res = 300)
+fig_adrenalineuse
+dev.off()
+```
+
 #####
@@
 C: Geographical differences in the most common elicitors of VIA.
-Countries which reported less than 10 VIA cases were not illustrated in this figure.", fig.height=4, fig.width = 7}
+Countries which reported less than 10 VIA cases were not illustrated in this figure.
+Fire ants and insectsthat could not be identified formed the 'other' group.", fig.height=4, fig.width = 7}
 fig_basic
+```
+
+```{r,include=F}
+tiff("figS1.tiff", width = 7, height = 4.5,units = 'in', res = 300)
+fig_basic
+dev.off()
 ```
@@
 A: The original distribution of VIA and non-VIA cases according to age group and sex.
-B: The distribution of VIA and non-VIA after age and sex matching with the use of MatchIt package for R."}
+Please note the uneaven distribution of VIA and non-VIA cases in age groups.
+B: The distribution of VIA and non-VIA after age and sex matching with the use of MatchIt package for R.
+Please notice how the ratio of VIA to non-VIA cases is approaching 50% indicating balanced matching according to sex and age variables."}
+
 matching.result 
+```
+
+```{r,include=F}
+tiff("figS2.tiff", width = 7, height = 5,units = 'in', res = 300)
+matching.result 
+dev.off()
 ```
@@
 
-ggarrange(
+figS3<-ggarrange(
   ggarrange(
@@
 )
+
+figS3
+```
+
+```{r,include=F}
+tiff("figS3.tiff", width = 7, height = 5,units = 'in', res = 300)
+figS3
+dev.off()
 ```
@@
 
+```{r,include=F}
+tiff("figS4.tiff", width = 7, height = 4,units = 'in', res = 300)
+plot_tryptase_in_cardio_cur
+dev.off()
+```
+
 ```{r treatmentmatrixother}
@@
 The severity of patients with VIA in two age groups (left), according to elicitor type (center) and according to the responsible insect species (right)"}
-severity_joined_brown %>%
+figS5 <- severity_joined_brown %>%
     ggbarplot(x = "grouping",
@@
 
+```{r,include=F}
+tiff("figS5.tiff", width = 7, height = 4,units = 'in', res = 300)
+figS5
+dev.off()
+```
+
 ```{r rfplot, fig.cap="Therapy of anaphylaxis.
@@
 B: Variable importance in the unsupervised classification between VIA and non-VIA using Random Forest classifier.", fig.height=3.5, fig.width=7.1}
-ggarrange(
+figS6 <- ggarrange(
   # adren_prev+
@@
 
+```{r,include=F}
+tiff("figS6.tiff", width = 7., height = 3.5,units = 'in', res = 300)
+figS6
+dev.off()
+```
+
+```{r BST, fig.cap="Levels of baseline serum tryptase in patinent with VIA and non-VIA.
+Significant difference in BST between patients with concomitant mastocytosis and other patients (***).
+There was no significant difference between anaphylaxis elicited by insects and other elicitors (NS).
+Tested by two way ANOVA."}
+
+BST_masto$plot
+
+```
+
+```{r,include=F}
+tiff("figS7.tiff", width = 7, height = 4,units = 'in', res = 300)
+BST_masto$plot
+dev.off()
+```
+
 #####

### 2020-03-27 R/make_all_forest_plots.R

@@ -1,3 +1,6 @@
+root <- rprojroot::is_r_package
+R <- root$make_fix_file()
+
 ### make all forest plots
 make_all_forest_plots <- function(){
 
@@ -54,7 +57,7 @@ test2 <- test2[-c(2:4,20:25),]
 #or_plot(test1)
 
 
-png("analysis/figures/figForestfinalb.png",
+png(R("analysis/figures/figForestfinalb.png"),
     height = 800,
     width = 1000,
     #res = 300,
@@ -74,13 +77,14 @@ forestplot(labeltext = makeTableText(test1),
            lwd.zero = 4)
 dev.off()
 
-png("analysis/figures/figForestfinalrmr.png",
-    height = 640*2,
-    width = 860*2,
+png(R("analysis/figures/figForestfinalrmr.png"),
+    height = 600*3,
+    width = 860*3,
     #res = 300,
-    pointsize = 30*2
+    pointsize = 30*3
     #units="px"
 )
+
 forestplot(labeltext = makeTableText(test2)[,1:2],
            mean = test2[,"mean"] %>% unlist,
            lower = test2[,"lower"] %>% unlist,
@@ -90,9 +94,9 @@ forestplot(labeltext = makeTableText(test2)[,1:2],
            clip=c(0.2,4),
            xlog=TRUE,
            col=fpColors(box="#1f1f1f",line="gray", summary="royalblue"),
-           lwd.xaxis = 5,
-           lwd.ci = 4,
-           lwd.zero = 4,
+           lwd.xaxis = 5*4,
+           lwd.ci = 4*4,
+           lwd.zero = 4*4,
            hrzl_lines = T,
            grid = T,
            txt_gp = fpTxtGp(ticks = gpar(cex = 0.8)))
@@ -101,7 +105,7 @@ dev.off()
 
 
 
-png("analysis/figures/figForest.png")
+png("analysis/figures/figForest.png" %>% R)
 makeForestPlot(age_sex_matched,
                c( "q_410_masto_cur",
                   "q_410_asthma_cur",
@@ -113,7 +117,7 @@ makeForestPlot(age_sex_matched,
                "d_severity_rmr")
 dev.off()
 
-png("analysis/figures/kidsForest.png")
+png("analysis/figures/kidsForest.png" %>% R)
 makeForestPlot(rdbp[rdbp$d_age<18,],
                testInsectsbinomial %>%
                  filter(section=="cofactors") %>%
[…]

### 2020-03-27 R/make_flow.R

@@ -1,4 +1,6 @@
 ### mAKE THE DIAGRAMM FOR THE fLOW CHART
+root <- rprojroot::is_r_package
+R <- root$make_fix_file()
 
 make_flow <- function(){
 require(DiagrammeRsvg)
@@ -93,7 +95,7 @@ create_graph() %>%
   add_global_graph_attrs(value="white",
                          attr = "fillcolor",
                          attr_type = "node") %>% #render_graph()
-  export_graph(file_name = "analysis/figures/flow.png", file_type = "png",
+  export_graph(file_name = "analysis/figures/flow.png" %>% R, file_type = "png",
                title = NULL,
                width = 1000,
                height = 900)

### 2020-03-27 R/stat.R

@@ -5,6 +5,10 @@ pacman::p_load(dplyr,magrittr,ggplot2,forestplot,summarytools,
                vcd,MatchIt,tidyr,ggradar,tibble,scales,ggpubr,rlang,
                forcats,fmsb,purrr,DiagrammeR,DiagrammeRsvg,rsvg,
                DescTools,optmatch,randomForest,ggfortify,heatmaply,verification)
+pacman::p_load("rprojroot")
+root <- rprojroot::is_r_package
+R <- root$make_fix_file()
+
 #install.packages("questionr")
 #
 # require(dplyr)
@@ -34,7 +38,8 @@ pacman::p_load(dplyr,magrittr,ggplot2,forestplot,summarytools,
 #                            trim_values = T)
 
 #load('../RefractoryAnaOrg/analysis/data/raw_data/data4.R')
-load("data.R") # here the anascore is done using the severity_analysis R script
+#load("data.R") # here the anascore is done using the severity_analysis R script
+load(R("data.R")) # here the anascore is done using the severity_analysis R script
 
 # update the anascore according to nora consensus
 #source("R/anascore_vas.R")
@@ -50,7 +55,7 @@ manual_greens <- rev(c("#E5F5E0",
                        "#005A32"))
 
 ####### General Functions #############
-source("R/functions.R")
+source(R("R/functions.R"))
 
 palBW <- c("#848484","#1f1f1f")
 
@@ -94,7 +99,10 @@ data4$d_organ_sum %<>% car::recode("'none' = '0';
                                    'four' = '4';
                                    'unknown' = NA") %>% as.numeric()
 data4$q_116_VAS_v7 %<>% as.numeric()
-data4$q_120_time_between_v4%<>% as.numeric()
+data4 %<>%
+  mutate(q_120_time_between_v4 = as.character(q_120_time_between_v4)) %>%
+  mutate(q_120_time_between_v4 = ifelse(q_120_time_between_v4 %in%c("4h","more","4h-5"),"240",q_120_time_between_v4) %>% as.numeric()) #%>%
+  #dplyr::select(q_120_time_between_v4) %>% summary()
 data4$q_131_biphasic_time_v4 %<>% car::recode("'b_version<4' = NA;
                                               'no biphasic reaction' = NA;
[…]

### 2020-03-27 analysis/paper/paper.Rmd
small correction for JACI

@@
 bibliography: references.bib
-csl: "../templates/vancouver.csl" # Insert path for the bib-style
+csl: "../templates/the-journal-of-allergy-and-clinical-immunology.csl" # Insert path for the bib-style
 ---
 
-Wojciech Francuzik^1^,
-Franziska Ruëff^2^,
-Andrea Bauer^3^,.
-Maria Beatrice Bilò ^4,5^,
-Victoria Cardona^6^,
-George Christoff^7^,
-Sabine Dölle-Bierke^8^,
-Luis Ensina^9^,
-Montserat Fernandes-Rivas^7,11^,
-Thomas Hawranek^12^,
-Johnathan O'B Hourihane^13^,
-Tilo Jakob^14,15^,
-Nicos G. Papadopoulus^16,17^,
-Claudia Pföhler^18^,
-Iwona Poziomkowska-Gęsicka^19^,
-Xavier Van der Brempt^20^,
-Kathrin Scherer Hofmeier^21^,
-Regina Treudler^22^,
-Nicola Wagner^23^,
-Bettina Wedi^24^,
-Margitta Worm^1^
+Wojciech Francuzik, MD ^1^, Franziska Ruëff, MD ^2^ , Andrea Bauer, MD ^3^ , Maria Beatrice Bilò, MD
+^4,5^ , Victoria Cardona, MD, PhD ^6,7^ , George Christoff, MD, PhD ^8,9^ , Sabine Dölle-Bierke, PhD ^1^ , Luis
+Ensina, MD, PhD ^10^ , Montserat Fernandes-Rivas, MD, PhD ^7,11^ , Thomas Hawranek, MD ^12^ ,
+Johnathan O’B Hourihane, MD, PhD ^13^ , Thilo Jakob, MD ^14,15^ , Nicos G. Papadopoulus, MD
+PhD ^16,17^ , Claudia Pföhler MD ^18^ , Iwona Poziomkowska-Gęsicka, MD, PhD ^19^ , Xavier Van der
+Brempt, MD ^20^ , Kathrin Scherer Hofmeier, MD ^21^ , Regina Treudler, MD ^226 , Nicola Wagner, MD ^23^ ,
 
-**Conflict of interest:**
-
-A. Bauer reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LETI, Thermofisher, and Stallergens outside the submitted work
-N. Wagner reports personal fees from ALK outside the submitted work.
-R. Treudler reports grants and personal fees from Sanofi-Genzyme, ALK-Abello, Takeda,  Novartis, grants from Hautnetz Leipzig and Fraunhofer-IZI Leipzig, outside the submitted work.
-V. Cardona  reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LET, Thermofisher and Stallergens outside the submitted work. 
-M. B. Bilò reports personal fees from ALK outside the submitted work.
-K. Scherer reports personal fees from Allergopharma, Sanofi-Aventis, and Shire outside submitted work. 
-Franziska Ruëff reports personal fees outside the submitted work from ALK-Abelló, Allergopharma, Bencard, Boehringer Ingelheim, Bristol Myers Squibb, Circassia, Dermira, DST, LEO Pharma, Lilly, Dr.
-Gerhard Mann chem.-pharm.
-Fabrik GmbH, Mylan, Novartis, Pfizer, Thermo Fisher Scientific and UCB
-Claudia Pföhler performed clinical studies for Allergy Therapeutics and received speaker honoraria and
-travel support from Bencard, Novartis and ALK. The rest of the authors declare that they have no relevant conflicts of interest.
-
-**Affiliations:**
+Bettina Wedi, MD ^24^ , Margitta Worm, MD ^1^
 
@@
 **Corresponding author: **
-  Prof. Dr. med. M. Worm margitta.worm@charite.de
+  Prof. Dr. med. M. Worm, margitta.worm@charite.de,
 Phone: +49 30 450 529 005;      Fax: +49 30 450 529 902
 
-**Keywords:**
-  anaphylaxis, epinephrine (adrenaline), beta-blockers, insect venom allergy, Hymenoptera
+**Conflict of interest:**
+
+A. Bauer reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LETI, Thermofisher, and Stallergens outside the submitted work
+N. Wagner reports personal fees from ALK outside the submitted work.
+R. Treudler reports grants and personal fees from Sanofi-Genzyme, ALK-Abello, Takeda,  Novartis, grants from Hautnetz Leipzig and Fraunhofer-IZI Leipzig, outside the submitted work.
+V. Cardona  reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LET, Thermofisher and Stallergens outside the submitted work. 
+M. B. Bilò reports personal fees from ALK outside the submitted work.
+K. Scherer reports personal fees from Allergopharma, Sanofi-Aventis, and Shire outside submitted work. 
+Franziska Ruëff reports personal fees outside the submitted work from ALK-Abelló, Allergopharma, Bencard, Boehringer Ingelheim, Bristol Myers Squibb, Circassia, Dermira, DST, LEO Pharma, Lilly, Dr.
+Gerhard Mann chem.-pharm.
+Fabrik GmbH, Mylan, Novartis, Pfizer, Thermo Fisher Scientific and UCB
+Claudia Pföhler performed clinical studies for Allergy Therapeutics and received speaker honoraria and
+travel support from Bencard, Novartis and ALK. The rest of the authors declare that they have no relevant conflicts of interest.
 
@@
 **Document statistics:**
-2838 words, 5 figures, 0 tables, 29 references
+3136 words, 5 figures, 0 tables, 34 references
 
@@
 
-# Key points
+<!-- # Key points  -->
 
-**Question:** What is the phenotype of insect venom anaphylaxis?
+<!-- **Question:** What is the phenotype of insect venom anaphylaxis? -->
 
-**Findings:** This case-control study showed that venom-induced anaphylaxis significantly more often presented with cardiovascular symptoms.
+<!-- **Findings:** This case-control study showed that venom-induced anaphylaxis significantly more often presented with cardiovascular symptoms.
 Secondly, severe cases of venom-induced anaphylaxis were significantly more often associated with the lack of skin symptoms upon anaphylaxis and showed a significant interaction with the levels of baseline serum tryptase (in range from 8 - 11 ng/ml).
+-->
 
-**Meaning:** Allegologists should consider intesified prophylaxis (and diagnostic) measures in patients with baseline tryptase levels of above 8 and a history of insect venom anaphylaxis.
+<!-- **Meaning:** Allegologists should consider intesified prophylaxis (and diagnostic) measures in patients with baseline tryptase levels of above 8 and a history of insect venom anaphylaxis.  -->
 
-#####
+<!-- ##### -->
 
@@
 
-**Importance:** Venom-induced anaphylaxis is a common, potentially life-threatening hypersensitivity reaction associated with specific: 1) symptom profile, 2) cofactors, and 3) management.
-Identifying the differences in phenotypes of anaphylaxis is crucial for future management guidelines and the development of a personalized medicine approach.
+**Background:** Venom-induced anaphylaxis is a common, potentially life-threatening
+hypersensitivity reaction associated with specific: 1) symptom profile, 2) cofactors, and 3)
+management. Identifying the differences in phenotypes of anaphylaxis is crucial for future
+management guidelines and the development of a personalized medicine approach.
 
-**Objective:** This study aimed to evaluate the phenotype and risk factors of venom-induced anaphylaxis in comparison to other elicitors of anaphylaxis.
+**Objective:** This study aimed to evaluate the phenotype and risk factors of venom-induced
+anaphylaxis.
 
-**Design:** Using data from the European Anaphylaxis Registry (`r nrow(data)` cases) we identified `r length(which(data4$d_elicitor_gr5== "insects" & data4$reaction_type_brown=="anaphylaxis"))` with venom-induced anaphylaxis and analyzed these in comparison to sex- and age-matched anaphylaxis cases triggered by other elicitors (non-VIA n = `r nrow(age_sex_matched %>% filter(grouping == "other"))`).
+**Methods:** Using data from the European Anaphylaxis Registry (12874 cases) we identified
+3612 with venom-induced anaphylaxis and analyzed these in comparison to sex- and age-
+matched anaphylaxis cases triggered by other elicitors (non-VIA n = 3605). All patients
+with anaphylaxis (Ring and Messmer grade II-IV) who gave informed consent were
+included in the study. Subsequently, we matched the group of patients presenting with
+venom anaphylaxis to sex- and age-matched group of patients who presented with
+anaphylaxis to other elicitors (i.e., food, drugs).
 
-**Setting:** Multicenter, international, prospective, case-control study.
+**Results:** Venom-induced anaphylaxis more frequently involved more than three organ
+systems and was associated with cardiovascular symptoms. The absence of skin symptoms
+during anaphylaxis correlated with baseline serum tryptase and was associated with an
+increased risk of a severe reaction. Intramuscular or intravenous epinephrine was
+administered significantly less often in venom-induced anaphylaxis, in particular in
+patients without prior history of anaphylaxis. Baseline serum tryptase within the upper
+normal range (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis.
 
-**Participants:** All patients with the diagnosis of moderate to severe anaphylaxis (according to Ring and Messmer grade II-IV) who gave informed consent to provide their clinical data for the Anaphylaxis Registry were included in the study.
-Subsequently, we matched the group of patients presenting with venom anaphylaxis to sex- and age-matched group of patients who presented with anaphylaxis to other elicitors (i.e., food, drugs).
+**Conclusion:** Regarding the fact that venom-induced anaphylaxis frequently affects the
+cardiovascular system, patients were undertreated with epinephrine. The lack of skin
+symptoms (i.e., urticaria or flushing) during anaphylaxis and baseline serum tryptase levels
+within the upper normal limits were associated with severe reactions. Patients with serum
+tryptase of above 8 ng/ml and a history of venom anaphylaxis may require additional
+prophylaxis.
 
-**Main Outcome(s) and Measure(s):** This study measured the severity of anaphylaxis on the Ring and Messmer scale.
-Hypotheses regarding the comparisons between groups were formulated during and after data collection.
+# Clinical Implications
 
-**Results:** Venom-induced anaphylaxis more frequently involved more than three organ systems and was associated with cardiovascular symptoms.
-The absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase and was associated with an increased risk of a severe reaction.
-Intramuscular or intravenous epinephrine was administered significantly less often in venom-induced anaphylaxis, in particular in patients without prior history of anaphylaxis.
-Baseline serum tryptase within the upper normal range (8-11.5 ng-ml) was more frequently associated with severe anaphylaxis.
+Allergologists should consider intensified prophylactic (and diagnostic) measures in patients with baseline serum tryptase of above eight ng/ml and a history of insect venom anaphylaxis without skin involvement.
 
-**Conclusions and Relevance:** Regarding the fact that venom-induced anaphylaxis frequently affects the cardiovascular system, patients were undertreated with epinephrine.
-The lack of skin symptoms (i.e., urticaria or flushing) during anaphylaxis and baseline serum tryptase levels within the upper normal limits were associated with severe reactions.
-Patients with serum tryptase of above 8 ng/ml and a history of venom anaphylaxis may require additional prophylaxis.
+# Capsule Summary
+
+Venom-induced anaphylaxis significantly more often presented with cardiovascular symptoms.
+Severe cases more often showed lack of skin involvement and were associated with higher levels of baseline serum tryptase (in range from 8 - 11 ng/ml).
+
+**Keywords:**
+  anaphylaxis, epinephrine (adrenaline), beta-blockers, insect venom allergy, Hymenoptera
+
+**List of abbreviations:**
+* VIA - venom-induced anaphylaxis
+* BST - baseline serum tryptase
+* EAI - epinephrine auto-injector
+* MCAS - mast cell activation syndrome
+* ER - emergency room
 
@@
 
-BST levels were significantly higher in patients with a prior diagnosis of mastocytosis \@ref(fig:BST). 
-Next, we investigated the association of skin symptoms with the tryptase levels in non-mastocytosis patients.
+BST levels were significantly higher in patients with a prior diagnosis of mastocytosis (eFig. \@ref(fig:BST)).
+We investigated the association of skin symptoms with the tryptase levels in non-mastocytosis patients.
 For this model, we excluded the cases with known mastocytosis and with BST above 11.5 ng/ml, potentially indicating non-diagnosed mast cell activation disorders.
@@
 Concomitant mastocytosis increased the risk for 1) cardiac arrest and 2) loss of consciousness in patients undergoing VIA significantly more than in patients undergoing anaphylaxis due to other elicitors (Fig.
-\@ref(fig:figcofactors)C and Fig.
-S\@ref(fig:supplement)A).
+\@ref(fig:figcofactors)C and eFig.
+\@ref(fig:supplement)A).
 
@@
 \@ref(fig:figcofactors)).
-Interestingly, BST values were increased in patients with concomitant cardiovascular diseases, irrespectively of the reaction severity (Fig.
-S\@ref(fig:trypcardiocur)).
+Interestingly, BST values were increased in patients with concomitant cardiovascular diseases, irrespectively of the reaction severity (eFig.
+\@ref(fig:trypcardiocur)).
 
@@
 Severe reactions of VIA were more prevalent in patients above 22 years of age, and in VIA cases vs.
-non-VIA cases (Fig.
-S\@ref(fig:sev)).
+non-VIA cases (eFig.
+\@ref(fig:sev)).
 There were no differences in severity of reactions elicited by yellow-jackets and other
@@
 
-IVA was more often severe if the reaction occured in the first 10 minutes after exposure to venom (`r sev_time$insects[1,2] %>% {.*100} %>% round(2)`% of Ring and Messmer grade III or IV cases) then when the reaction occured after 10 minutes post exposure (`r sev_time$insects[2,2] %>% {.*100} %>% round(2)`% p = `r sev_time$insects_p`).
+IVA was more often severe if the reaction occured in the first 10 minutes after exposure to venom (`r sev_time$insects[1,2] %>% {.*100} %>% round(2)`% were severe cases) then when the reaction occured after 10 minutes post exposure (`r sev_time$insects[2,2] %>% {.*100} %>% round(2)`% were severe cases, p = `r sev_time$insects_p`).
 
@@
 However, when we restricted the analysis to moderate anaphylaxis cases --- non-VIA patients received epinephrine more frequently than VIA (`r adren_severity_mild`).
-The presence of skin symptoms during these mild reactions also was associated with a lower fraction of epinephrine treated patients (Fig.
+The presence of skin symptoms during these mild reactions also was associated with a lower fraction of epinephrine treated patients (eFig.
 \@ref(fig:rfplot)).
@@
 On the other hand, epinephrine, beta-2 mimetics, and oxygen were given more often to patients suffering from non-VIA  (Fig.
-\@ref(fig:adrenalineuse)A, Fig.
-S\@ref(fig:rfplot)).
+\@ref(fig:adrenalineuse)A, eFig.
+\@ref(fig:rfplot)).
 
@@
 <!-- Risk factors - Concomitant disease -->
-Similarly, we observed less VIA in patients with concomitant atopic diseases (Fig. \@ref(fig:supplement)) , as these patients more often present with food anaphylaxis [@Tham2019].
+Similarly, we observed less VIA in patients with concomitant atopic diseases (eFig. \@ref(fig:supplement)) , as these patients more often present with food anaphylaxis [@Tham2019].
 
@@
 
+The European Anaphylaxis Registry was supported by the Network for Online-Registration
+of Anaphylaxis NORA e.
+V.
 We thank all patients, parents, and their children for their support in providing data on the occurrence of anaphylaxis for this study.
@@
 ```{r,include=F}
-tiff("figS1.tiff", width = 7, height = 4.5,units = 'in', res = 300)
+tiff("efig6.tiff", width = 7, height = 4.5,units = 'in', res = 300)
 fig_basic
@@
 ```{r,include=F}
-tiff("figS2.tiff", width = 7, height = 5,units = 'in', res = 300)
+tiff("efig7.tiff", width = 7, height = 5,units = 'in', res = 300)
 matching.result 
@@
 ```{r,include=F}
-tiff("figS3.tiff", width = 7, height = 5,units = 'in', res = 300)
+tiff("efig8.tiff", width = 7, height = 5,units = 'in', res = 300)
 figS3
@@
 ```{r,include=F}
-tiff("figS4.tiff", width = 7, height = 4,units = 'in', res = 300)
+tiff("efig9.tiff", width = 7, height = 4,units = 'in', res = 300)
 plot_tryptase_in_cardio_cur
@@
     guides(color = F)
+figS5
 ```
@@
 ```{r,include=F}
-tiff("figS5.tiff", width = 7, height = 4,units = 'in', res = 300)
+tiff("efig10.tiff", width = 7, height = 4,units = 'in', res = 300)
 figS5
@@
 )
+figS6
 ```
@@
 ```{r,include=F}
-tiff("figS6.tiff", width = 7., height = 3.5,units = 'in', res = 300)
+tiff("efig11.tiff", width = 7., height = 3.5,units = 'in', res = 300)
 figS6
@@
 ```{r,include=F}
-tiff("figS7.tiff", width = 7, height = 4,units = 'in', res = 300)
+tiff("efig12.tiff", width = 7, height = 4,units = 'in', res = 300)
 BST_masto$plot

### 2020-05-27 analysis/paper/paper.Rmd
revision round 2

@@
     reference_docx: "../templates/template.docx" # Insert path for the DOCX file
-bibliography: references.bib
+bibliography: "references.bib"
 csl: "../templates/the-journal-of-allergy-and-clinical-immunology.csl" # Insert path for the bib-style
@@
 
-pacman::p_load("tidyvese")
+pacman::p_load("tidyverse")
 
@@
 
-**Conclusion:** Regarding the fact that venom-induced anaphylaxis frequently affects the
-cardiovascular system, patients were undertreated with epinephrine. The lack of skin
-symptoms (i.e., urticaria or flushing) during anaphylaxis and baseline serum tryptase levels
-within the upper normal limits were associated with severe reactions. Patients with serum
-tryptase of above 8 ng/ml and a history of venom anaphylaxis may require additional
-prophylaxis.
+**Conclusion:** Using a large cohort of VIA cases, we have validated that patients 
+with intermediate baseline serum tryptase levels (8 - 11 ng/ml) and without skin 
+involvement have higher risk of severe VIA. Patients receiving beta-blockers or 
+ACE-I had a higher risk of developing severe cardiovascular symptoms (including cardiac arrest) 
+in VIA and non-VIA cases. Patients undergoing VIA received epinephrine less frequently than non-VIA cases.
 
@@
 
-Importantly, the absence of skin symptoms was associated with more severe VIA, which was still present after excluding patients with a known diagnosis of mastocytosis.
+Importantly, the absence of skin symptoms was associated with more severe VIA, which was still present after excluding patients with a known diagnosis of mastocytosis (although in non-mastocytosis cases the difference between groups was small and the clinical relevance of this needs cautious evaluation).
 Previous studies also observed this phenomenon [@Stoevesandt2012;@Fehr2019].
-Subsequently, the correlation of BST levels with the severity of anaphylaxis lead us to identify an interaction between the absence of skin symptoms and VIA using generalized linearf regression.
+Subsequently, the correlation of BST levels with the severity of anaphylaxis lead us to identify an interaction between the absence of skin symptoms and VIA using generalized linear regression.
 
@@
 Based on our results, VIA is a distinctive phenotype of anaphylaxis, with a specific symptom profile and risk factors.
+Using a large cohort of VIA cases compared to a sex and age matched non-VIA cases, we have validated that patients with intermediate baseline serum tryptase levels (8 - 11 ng/ml) and without skin involvement have higher risk of severe VIA.
+Similarily, patients receiving beta-blockers or ACE-I had a higher risk of developing severe cardiovascular symptoms (including cardiac arrest) in VIA and non-VIA cases.
+Patients undergoing VIA received epinephrine less frequently than non-VIA cases.
+
 VIA cases should undergo therapy according to the international management guidelines, and epinephrine should be given more often in VIA.
-
-All cases should undergo appropriate allergologic testing and indication for SIT should be evaluated.
+All cases should undergo appropriate allergologic testing and indication for SIT should be evaluated along with patient education regarding the risk of future anaphylaxis.
 Patients with BST above 8 ng/ml should undergo extensive diagnostic tests to exclude indolent systemic mastocytosis or MCAS and should be provided with two EAIs for acute self-management.

### 2020-05-27 R/response2_cardio_epinephrine.R

@@ -1,0 +1,14 @@
+# age_sex_matched %>%
+#   mutate(grouping = car::recode(grouping, "'insects' = 'VIA';'other'='non-VIA'")) %>%
+#   group_by(grouping,q_114,d_522_adren_agg) %>%
+#   count() %>%
+#   filter(!is.na(q_114),
+#          !is.na(d_522_adren_agg),
+#          q_114!="unknown",
+#          d_522_adren_agg!="unknown")%>%
+# ggplot(aes(x = q_114,y = n, fill = d_522_adren_agg))+
+#   facet_grid(~grouping)+
+#   geom_bar(stat = "identity") +
+#   labs(x = "Cardiologic symptoms",
+#        y = "Number of cases",
+#        fill = "Epinephrine Tx")
